### 2022-11-21 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@lgrcia}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2022-11-24 simple/latex/ms.tex
add simple paper and some results csv

% Define document class
\documentclass{aastex631}
\usepackage{float}

\newcommand\twostep{two-step\xspace}
\newcommand\full{full-fledged\xspace}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@lgrcia}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section*{Introduction}
\label{sec:intro}

Exoplanets, planets outside our solar system, are discovered at an ever-increasing rate. Beyond the study of their inner structure and atmosphere, they give a unique glimpse to extrasolar systems formation, dynamics, as well as being probes to understand exoplanets' host stars. This is particularly true for systems whose orbital plane is aligned with our line of sight, leading to observable planetary transits. Although these transit signals can be seen in the apparent flux received from their stars (light curves), they are often mixed with other astrophysical and instrumental signals, together referred to as \textit{correlated noise}. If they can be disentangled from these nuisance signals, transits offer a powerful way to detect exoplanets. However, disentangling these signals comes with many challenges.
\bigskip\\
The simplest way to find periodic transit signals is to first clean a light curve from nuisance signals before performing the search. Indeed, widely used transit-search algorithms (namely Box-Least-Square algorithms, \cite{}) are capable of detecting transits on flat light curves only containing white noise. This strategy is widely adopted by the community, both using physically-motivated systematic models like \cite{everest1, everest2}, or empirical filtering techniques, such as the ones described and implemented in \cite{wotan}. However, when systematics and variability start to resemble transits, this cleaning step (often referred to as \textit{detrending}) strongly alters the searched signals. In such cases, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noises, and asses the likelihood of the model to the data on a wide parameter space, an approach largely avoided due to its untractable nature. Nonetheless, \citealt{kovacs2016} face the question by asking: \textit{Periodic transit and variability search with simultaneous systematics filtering: Is it worth it?}. This study discards the general benefit of using a full-fledged approach, it fails at exploring the light-curves characteristics for which it becomes necessary. While it might only represent a handful of systems, this cases are extremely valuable for the exoplanetary science community. First, variability may be associated with star-spots that can be probed with the help of planetary transits. A better understanding of these structures benefit both the study of stellar atmospheres and their concerning impact on planetary atmosphere retrievals. Second, the growing interest of the community for ultra-cool dwarf stars comes with observations featuring enhanced red noise, stellar variability and lower transit SNR. Hence, these hidden system are pristine.
\bigskip\\
In this paper we present nuance, an algorithm using linear models and gaussian processes to simultaneously search for transits while modeling correlated noises in a tractable way. In \autoref{issues}, we describe the issues inherent to the two-step approach described earlier, and study the parameter space for which commonly used filtering techniques alter transit signals to the point of no detection. In section 2 we describe the tractable full-fledged approach on which nuance is based, and its implementation in an open-source Python package. In section 3, we asses the performance of nuance by computing the recovery rate of planetary transits injected into simulated light curves. In section 4, we use nuance in a case study, to detect a known planet observed by the T E S S but undetected by the TESS pipeline due to its light curve characteristics. In section 5, we use nuance to search for transits in a list of .... Finally, we conclude this paper by providing avenues for the improvement of nuance, from its advantage to search for transits in ground-based observations, to its potential for the detection of multi-planetary systems affected by transit-times variations.

\newpage
\section{The effect of correlated noise detrending on transit search}\label{issues}
Two sources of correlated noise particularly justify the need for a detrending step before searching for transits: systematic noises, such as telescope pointing errors or detector ... and stellar variability, such as pulsations, or starspots rotations. In this section, we discuss the impact of such correlated noises on transits detectability, by studying the signal-to-noise (SNR) of a unique event, reduced to the simplified expression (\citealt{pont2006}, Equation 12):

\begin{equation}\label{eq:snr}
  SNR= \frac{df}{\sqrt{\frac{\sigma_w^2}{n} + \frac{\sigma_r^2}{N_{tr}}}}
\end{equation}

where $df$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits (unity here since we consider single transits), and $\sigma_w$ and $\sigma_c$ are the white and correlated noise standard deviations, the subscript.
\bigskip\\

\subsection{Instrumental noises}

The study from \citealt{pont2006} highlight the importance of accounting for correlated noise ($\sigma_r$) when estimating the yield of transit search campaigns. To demonstrate its point, we simulate a unique transit signal (using the empirical analytical model from \citealt{protopapas}) and compute its SNR using \autoref{eq:snr}, both in the presence and absence of correlated noise (\autoref{fig:issue1}).

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=8.5cm]{../figures/first_issue.pdf}
        \caption{This is a pretty visualization of the Mandelbrot set.}
        \label{fig:issue1}
    \end{centering}
\end{figure}

As illustrated in \autoref{fig:issue1}, the presence of correlated noise fundamentally limits transit detectability. This issue rapidly motivated the development of systematics detrending algorithms such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}), or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}). Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal. We note that, except for \textsc{TFA}, these algorithms are mostly applied to space-based continuous observations, that provide continuous stellar baselines and mostly reproducible systematic signals. This is not the case for the vast majority of ground-based observations, in addition subject to periodic daytime interruptions and varying atmospheric extinction.

\subsection{Stellar variability}

Instrumental signals have the benefit to be shared among light curves of the same field, and strongly correlated with measurements from the experimental setup (like detector's temperature, pointing error, sky background or airmass). In opposition, stellar variability is generally unknown and harder to correlate with simultaneous measurements. This gave rise to two types of treatments to reconstruct and remove stellar variability to perform a simpler transit search. One is physically-motivated and make use of Gaussian Processes (e.g. \cite{k2sc}). The other is empirical and make use of filtering algorithms (a wide variety being described in \cite{wotan}). In either case, the complete light curve is modeled while ignoring the presence of transits. In this section, we will study in details the effect of both approaches on transit search, by studying the degradation of a unique transit SNR for a wide variety of stellar variability characteristics.
\bigskip\\
Throughout this paper, we simulate stellar variability (and later model it) thanks to Gaussian Processes (GP), employing the physically-motivated stochastically-driven damped simple harmonic oscillator kernel (SHO) presented in \citealt{celerite1} (see Annexe ...). We generate stellar variability thanks to the \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}} Python package, providing an implententation of the scalable GP method from \citealt{celerite2} powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{../figures/issue2.pdf}
        \caption{This is a pretty visualization of the Mandelbrot set.}
        \label{fig:issue2}
    \end{centering}
\end{figure}

In \autoref{fig:issue2}, we simulate a light curve featuring a single transit (again using the model from \citealt{protopapas}), in addition with stellar variability of different timescales and amplitudes. In case \textit{a} (purple in \autoref{fig:issue2}), we model and detrend these signals using the widely-adopted Tukey's biweight filtering method presented in \citealt{tukey} and its implementatation from \texttt{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan} (case a. in purple). In case \textit{b} (red in \autoref{fig:issue2}) we employ the same GP model used to simulate stellar variability to reconstruct it, and remove it from light curves ignoring the presence of a potential transits (as it is the case in unknown transits search). Once transits cleaned from stellar variability, we asses their remaining SNR using \autoref{eq:snr}.
\bigskip\\
\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and suggests that this degradation is strongly dependant on the stellar variability characteristics encountered. In order to explore the parameter space for which detrending is the most problematic, we simulate light curves including a single transit and a much wider range of correlated noise characteristics defined with two parameters: $\delta_v$, the relative amplitude of the variability against transit depth; and $\tau_v$ the relative timescale of the variability with respect to the transit duration. To follow this parametrization, we simulate the photometric variability using a GP with an SHO kernel with hyperparameters \footnote{}:

\begin{equation}\label{eq:params}
    \omega = \frac{\pi}{\tau_v\tau}, \quad 
    \sigma = \delta_v \frac{\delta^2}{4}  \quad  \text{and}  \quad  
    Q = 10
\end{equation}

with $\delta$ and $\tau$ the depth and duration of the simulated transit, similar in all light curves. We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability. The expressions of $\omega$ and $\sigma$ given in \autoref{eq:params} correspond to a periodic signal with a period half that of the transit duration, and a peak to peak amplitude two times that of the transit depth, so that a varibility with $(\tau_v=1, \delta_v=1)$ strongly resembles the simulated transit signal. Then, as in \autoref{fig:issue2}, we reconstruct the variability using a biweight filter with a window length of $3\tau$ (three times the transit duration, an optimal value according to \citealt{wotan}). We then subtract the variability from each light-curve, estimate the resulting transit depth and compute its SNR. \autoref{fig:snr_detrend} shows the remaining SNR values of transits computed this way, after each variability signals with random $(\tau_v, \delta_v)$ have been detrended.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[height=10cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
        \caption{This is a pretty visualization of the Mandelbrot set.}
        \label{fig:snr_detrend}
    \end{centering}
\end{figure}

\section{\texttt{nuance}}

\bibliography{bib}

\end{document}

### 2022-11-24 simple/lib/utils.py

@@ -1,0 +1,39 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import jax, jaxopt
+jax.config.update("jax_enable_x64", True)
+import jax.numpy as jnp
+from tinygp import kernels, GaussianProcess
+import matplotlib.pyplot as plt
+
+
+def white_red(time, flux, error, plot=False, mean=0.):
+    init = {
+        "log_diag": 2*np.log(np.mean(error)),
+        "log_scale": np.log(1), # 5 mins
+        "log_amp": np.log(np.std(flux))
+    }
+
+    def build_gp(params):
+        kernel = jnp.exp(params["log_amp"]) * kernels.ExpSquared(jnp.exp(params["log_scale"]))
+        return GaussianProcess(kernel, time, diag=jnp.exp(params["log_diag"]), mean=mean)
+
+    @jax.jit
+    def loss(params):
+        gp = build_gp(params)
+        return -gp.log_probability(flux)
+    
+    opt = jaxopt.ScipyMinimize(fun=loss)
+    params = opt.run(init).params  
+    w = np.sqrt(np.exp(params['log_diag']))
+    
+    if plot:
+        gp = build_gp(params)
+        _, cond = gp.condition(y=flux, X_test=time)
+        mu = cond.mean
+        plt.plot(time, flux, ".", c="0.8")
+        plt.plot(time, mu, c="k")
+        noise = w
+        plt.fill_between(time, mu + noise, mu - noise, color="C0", alpha=0.3)
+
+    return w, np.sqrt(np.exp(params['log_amp']))

### 2022-11-24 simple/scripts/plot_cleaning_SNR.py

@@ -1,0 +1,139 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+# In[66]:
+
+
+import pandas as pd
+import numpy as np
+import yaml
+from scipy.stats import binned_statistic_2d
+import sys
+sys.path.append("../workflows/synthetic_injection_recovery/scripts")
+from make_lightcurve import make_lc
+from make_params import make
+import matplotlib.pyplot as plt
+from matplotlib.patches import ConnectionPatch
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+
+df = pd.read_csv("../workflows/cleaning_snr/data/simu1/results.csv")
+
+
+# In[ ]:
+
+
+n = 3 # number of small plots
+scale = 1.2
+W = n+1
+H = n+1
+
+fig = plt.figure(None, (1.6*W*scale, H*1.2*scale))
+
+def I(i, j):
+    return int(j*W + i) + 1
+
+for i in range(H*W):
+    ax = plt.subplot(H, W, i+1)
+    
+## Main plot
+main = plt.subplot(H, W, (I(0, 1), I(W-2, H-1)))
+
+bins=(30, 30)
+var, amp, snr = df.values.T.astype(float)
+stats = binned_statistic_2d(var, amp, snr, bins=bins)
+im = main.imshow(
+    stats.statistic.T, 
+    origin="lower",
+    extent=(
+        stats.x_edge.min(), 
+        stats.x_edge.max(), 
+        stats.y_edge.min(), 
+        stats.y_edge.max()
+    ),
+    aspect="auto",
+    vmax=15
+)
+
+main.set_ylabel(r"$\delta_v \sim \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$", fontsize=14)
+main.set_xlabel(r"$\tau_v \sim \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$", fontsize=14)
+params = yaml.full_load(open("../workflows/cleaning_snr/data/simu1/params/0.yaml", "r"))
+original_snr = params["depth"]/np.sqrt(params["error"]**2/n)
+
+main.set_title(f"Detrending effect on SNR\n", loc="left")
[…]

### 2022-11-24 simple/scripts/plot_issue_1.py

@@ -1,0 +1,121 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import jax, jaxopt
+jax.config.update("jax_enable_x64", True)
+import jax.numpy as jnp
+from tinygp import kernels, GaussianProcess
+import matplotlib.pyplot as plt
+import tinygp
+
+def white_red(time, flux, error, plot=False, mean=0.):
+    init = {
+        "log_diag": 2*np.log(np.mean(error)),
+        "log_scale": np.log(1), # 5 mins
+        "log_amp": np.log(np.std(flux))
+    }
+
+    def build_gp(params):
+        kernel = jnp.exp(params["log_amp"]) * kernels.ExpSquared(jnp.exp(params["log_scale"]))
+        return GaussianProcess(kernel, time, diag=jnp.exp(params["log_diag"]), mean=mean)
+
+    @jax.jit
+    def loss(params):
+        gp = build_gp(params)
+        return -gp.log_probability(flux)
+    
+    opt = jaxopt.ScipyMinimize(fun=loss)
+    params = opt.run(init).params  
+    w = np.sqrt(np.exp(params['log_diag']))
+    
+    if plot:
+        gp = build_gp(params)
+        _, cond = gp.condition(y=flux, X_test=time)
+        mu = cond.mean
+        plt.plot(time, flux, ".", c="0.8")
+        plt.plot(time, mu, c="k")
+        noise = w
+        plt.fill_between(time, mu + noise, mu - noise, color="C0", alpha=0.3)
+
+    return w, np.sqrt(np.exp(params['log_amp']))
+
+# Simulation
+# ------------
+from nuance.utils import transit
+np.random.seed(43)
+
+fluxes = []
+snrs = []
+
+# transit
+time = np.linspace(0, 3, 1000)
+dur = 0.1
+depth = 0.8e-2
+n = dur/np.median(np.diff(time))
+signal = transit(time, time.mean(), dur, depth, c=10) + 1.
+
+# white noise
+wsigma = 1.5e-3
+white_noise = np.random.normal(0, wsigma, len(time))
+flux = white_noise + signal
+
+# No systematics
+# --------------
+
+w, r = white_red(time, flux-signal, wsigma, plot=True)
+SNR = depth/np.sqrt((w**2)/n + r**2)
+
[…]

### 2022-11-24 simple/scripts/plot_issue_2.py

@@ -1,0 +1,144 @@
+from wotan import flatten
+import numpy as np
+from nuance.utils import transit
+import tinygp
+import jax
+import matplotlib.pyplot as plt
+jax.config.update("jax_enable_x64", True)
+import sys
+sys.path.append("./lib")
+sys.path.append("../lib")
+from utils import white_red
+
+seed = 42
+
+# Simulate light curve
+# --------------------
+np.random.seed(seed)
+# transit
+time = np.linspace(0, 3, 1000)
+dur = 0.1
+depth = 0.8e-2
+n = dur/np.median(np.diff(time))
+signal = transit(time, time.mean(), dur, depth, c=10) + 1.
+
+# white noise
+wsigma = 1.5e-3
+white_noise = np.random.normal(0, wsigma, len(time))
+flux = white_noise + signal
+
+# Simulate variability
+# --------------------
+lightcurves = []
+variabilities = []
+gps = []
+cond_gps = []
+
+periods = [1000., 1., 0.2, 0.1, 0.04]
+amps = [0., 1e-2, 1e-2, 1e-2, 5e-3]
+
+for i, (period, amp) in enumerate(zip(periods, amps)):
+    Q = 10.
+    w0 = 1/period
+    S0 = amp
+
+    var_kernel = tinygp.kernels.quasisep.SHO(w0, Q, S0)
+    gp = tinygp.GaussianProcess(var_kernel, time, mean=0.)
+    cond_gp = tinygp.GaussianProcess(var_kernel, time, mean=0., diag=wsigma**2)
+    variability = gp.sample(jax.random.PRNGKey(seed))
+    variability -= np.median(variability)
+    flux = white_noise + signal + variability
+    lightcurves.append(flux)
+    variabilities.append(variability)
+    gps.append(gp)
+    cond_gps.append(cond_gp)
+
+# Trends and SNR
+# -------------
+wotan_trend_snr = []
+gp_trend_snr = []
+
+cut = 200
+intransit = signal < (1. - depth/2)
+
+for i, (flux, var, gp) in enumerate(zip(lightcurves, variabilities, cond_gps)):
+     # WOTAN
+    flatten_flux, flatten_trend = flatten(time, flux, window_length=3*dur, return_trend=True)
[…]

### 2022-11-24 workflows/cleaning_snr/lib/utils.py

@@ -1,0 +1,44 @@
+import numpy as np
+
+duration = 0.04
+depth = 1e-2
+error = 5e-4
+time = np.linspace(0, duration*70, 2000)
+
+def white_red(time, flux, error, plot=False, mean=0.):
+    import numpy as np
+    import matplotlib.pyplot as plt
+    import jax, jaxopt
+    jax.config.update("jax_enable_x64", True)
+    import jax.numpy as jnp
+    from tinygp import kernels, GaussianProcess
+
+    init = {
+        "log_diag": 2*np.log(np.mean(error)),
+        "log_scale": np.log(1), # 5 mins
+        "log_amp": np.log(np.std(flux))
+    }
+
+    def build_gp(params):
+        kernel = jnp.exp(params["log_amp"]) * kernels.ExpSquared(jnp.exp(params["log_scale"]))
+        return GaussianProcess(kernel, time, diag=jnp.exp(params["log_diag"]), mean=mean)
+
+    @jax.jit
+    def loss(params):
+        gp = build_gp(params)
+        return -gp.log_probability(flux)
+    
+    opt = jaxopt.ScipyMinimize(fun=loss)
+    params = opt.run(init).params  
+    w = np.sqrt(np.exp(params['log_diag']))
+    
+    if plot:
+        gp = build_gp(params)
+        _, cond = gp.condition(y=flux, X_test=time)
+        mu = cond.mean
+        plt.plot(time, flux, ".", c="0.8")
+        plt.plot(time, mu, c="k")
+        noise = w
+        plt.fill_between(time, mu + noise, mu - noise, color="C0", alpha=0.3)
+
+    return w, np.sqrt(np.exp(params['log_amp']))

### 2022-11-24 workflows/cleaning_snr/scripts/aggregate_results.py

@@ -1,0 +1,18 @@
+import pandas as pd
+import yaml
+import numpy as np
+
+results = {
+    "snr": [],
+    "relative_duration": [],
+    "relative_depth": [],
+}
+
+for r in snakemake.input:
+    result = yaml.full_load(open(r, "r"))
+    results["relative_depth"].append(result['relative_depth'])
+    results["relative_duration"].append(result['relative_duration'])
+    results["snr"].append(result["snr"])
+
+df = pd.DataFrame(results)
+df.to_csv(snakemake.output[0], index=False)

### 2022-11-24 workflows/cleaning_snr/scripts/clean.py

@@ -1,0 +1,24 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import jax
+jax.config.update("jax_enable_x64", True)
+import yaml
+from wotan import flatten
+
+params = yaml.full_load(open(snakemake.input[0], "r"))
+time, flux, error = np.load(snakemake.input[1])
+
+intransit = np.abs(time - params['t0']) < params['duration']
+n_tr = params['duration']/np.median(np.diff(time))
+
+flatten_flux = flatten(time, flux + 1., window_length=3*params["duration"], return_trend=False)
+new_depth = np.max([0, np.median(flatten_flux[~intransit]) - np.median(flatten_flux[intransit])])
+w = params["error"]
+r = np.max([0, np.std(flatten_flux[~intransit])-w])
+snr = new_depth/np.sqrt((w**2)/n_tr + r**2)
+result = {
+    "relative_duration" : float(params["relative_duration"]),
+    "relative_depth" : float(params["relative_depth"]),
+    "snr": float(snr)
+}
+yaml.safe_dump(result, open(snakemake.output[0], "w"))

### 2022-11-24 workflows/cleaning_snr/scripts/make_lightcurve.py

@@ -1,0 +1,30 @@
+# make_lightcurve.py
+import tinygp
+import jax
+import numpy as np
+import yaml
+from nuance.utils import transit
+jax.config.update("jax_enable_x64", True)
+import sys
+sys.path.append("./lib")
+sys.path.append("../lib")
+from utils import time
+
+def make_lc(params, seed):
+    
+    kernel = tinygp.kernels.quasisep.SHO(params['omega'], params['quality'], sigma=params["sigma"])
+    gp = tinygp.GaussianProcess(kernel, time, diag=params['error']**2)
+
+    signal = transit(time, time.mean(), params["duration"], params["depth"], c=10)
+    y = gp.sample(jax.random.PRNGKey(seed)) + signal
+    
+    error = np.ones_like(y)*params['error']
+    
+    return time, y, error
+
+if __name__=="__main__":
+    
+    seed = int(snakemake.wildcards.seed)
+    params = yaml.full_load(open(snakemake.input[0], "r"))
+    time, y, error = make_lc(params, seed)
+    np.save(snakemake.output[0], np.array([time, y, error]))

### 2022-11-24 workflows/cleaning_snr/scripts/make_params.py

@@ -1,0 +1,47 @@
+# generate_params.py
+import numpy as np
+import yaml
+import sys
+sys.path.append("./lib")
+sys.path.append("../lib")
+from utils import time, duration, depth, error
+
+argv = sys.argv
+
+def make(seed=None, amp=None, var=None):
+    if seed is None:
+        if amp is None:
+            seed = np.random.randint(1, 1e5)
+            np.random.seed(seed)
+            amp = np.random.uniform(0.1, 10)
+            var = np.random.uniform(0.1, 5)
+        
+    params = {
+        "t0" : float(time.mean()),
+        "duration" : duration,
+        "depth" : depth,
+        "error" : error,
+    }
+
+    if seed == -1:
+        return make(var=1000, amp=0.1)
+
+    params.update({
+        "relative_duration": float(var),
+        "relative_depth": float(amp),
+        "omega" : float(2*np.pi/(2*params["duration"] * var)),
+        "quality" : float(np.random.uniform(10, 100)),
+        "sigma" : float(params["depth"] * (amp/2)**2),
+        "seed" : int(seed),
+    })
+    
+    return params
+
+if __name__=="__main__":
+    if len(argv) > 1:
+        seed = int(argv[1])
+    else:
+        seed = None
+        
+    params = make(seed)
+    yaml.safe_dump(params, open(f"data/params/{seed}.yaml", "w"))

### 2022-11-24 workflows/cleaning_snr/scripts/plot_snr_cleaning.py

@@ -1,0 +1,128 @@
+import pandas as pd
+import numpy as np
+import yaml
+from scipy.stats import binned_statistic_2d
+import sys
+sys.path.append("./lib")
+from make_lightcurve import make_lc
+from make_params import make
+import matplotlib.pyplot as plt
+from matplotlib.patches import ConnectionPatch
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+from utils import depth as true_depth
+from utils import error
+
+df = pd.read_csv(snakemake.input[0])
+
+n = 3 # number of small plots
+scale = 1.2
+W = n+1
+H = n+1
+
+fig = plt.figure(None, (1.6*W*scale, H*1.2*scale))
+
+def I(i, j):
+    return int(j*W + i) + 1
+
+for i in range(H*W):
+    ax = plt.subplot(H, W, i+1)
+    
+## Main plot
+main = plt.subplot(H, W, (I(0, 1), I(W-2, H-1)))
+
+bins=(30, 30)
+snr, var, amp = df.values.T.astype(float)
+stats = binned_statistic_2d(var, amp, snr, bins=bins)
+im = main.imshow(
+    stats.statistic.T, 
+    origin="lower",
+    extent=(
+        stats.x_edge.min(), 
+        stats.x_edge.max(), 
+        stats.y_edge.min(), 
+        stats.y_edge.max()
+    ),
+    aspect="auto",
+    vmax=15
+)
+
+main.set_ylabel(r"$\delta_v \sim \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$", fontsize=14)
+main.set_xlabel(r"$\tau_v \sim \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$", fontsize=14)
+params = true_depth
+original_snr = true_depth/np.sqrt(np.mean(error)**2/n)
+
+main.set_title(f"Detrending effect on SNR\n", loc="left")
+main.text(0, stats.y_edge.max()+0.4, f"original SNR: {original_snr:.2f}", color="0.3", va="center")
+
+# removing some axes
+# ------------------
+ax = plt.subplot(H, W, I(W-1, 0))
+plt.axis('off')
+
+# light curves examples
[…]

### 2022-11-25 simple/latex/ms.tex
cleaner plots, including results

@@
 % Title
-\title{An open source scientific article}
+\title{\texttt{nuance}: Detection of planetary transits in the presence of correlated noise}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    We present \texttt{nuance}, an algorithm to simultaneously search for transits while modeling correlated noise in a tractable way, including instrumental signals and stellar photometric variability.
 \end{abstract}
@@
 \bigskip\\
-In this paper we present nuance, an algorithm using linear models and gaussian processes to simultaneously search for transits while modeling correlated noises in a tractable way.
+In this paper we present nuance, an algorithm using linear models and gaussian processes to simultaneously search for transits while modeling correlated noise in a tractable way, including instrumental signals and stellar photometric variability.
 In \autoref{issues}, we describe the issues inherent to the two-step approach described earlier, and study the parameter space for which commonly used filtering techniques alter transit signals to the point of no detection.
@@
 \newpage
-\section{The effect of correlated noise detrending on transit search}\label{issues}
-Two sources of correlated noise particularly justify the need for a detrending step before searching for transits: systematic noises, such as telescope pointing errors or detector ...
-and stellar variability, such as pulsations, or starspots rotations.
-In this section, we discuss the impact of such correlated noises on transits detectability, by studying the signal-to-noise (SNR) of a unique event, reduced to the simplified expression (\citealt{pont2006}, Equation 12):
+\section{The issue with detrending}\label{issues}
+Two sources of correlated noise particularly justify the need for a detrending step before searching for transits: systematic noise (such as telescope pointing errors) and stellar variability (induced by  pulsations or starspots).
+In this section, we discuss the impact of such correlated noise and its detrending on transits detectability.
+For simplicity, we will model transit signals using an analytic empirical model described in \citealt{protopapas} (see Annexe).
+\bigskip\\
+
+\subsection{The effect of correlated noise on transits detectability}
+
+To study transits detectability, we will focus on the signal-to-noise (SNR) of a unique event, reduced to the simplified expression (\citealt{pont2006}, Equation 12):
 
 \begin{equation}\label{eq:snr}
-  SNR= \frac{df}{\sqrt{\frac{\sigma_w^2}{n} + \frac{\sigma_r^2}{N_{tr}}}}
+  SNR= \frac{df}{\sqrt{\frac{\sigma_w^2}{n} + \frac{\sigma_c^2}{N_{tr}}}}
 \end{equation}
 
-where $df$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits (unity here since we consider single transits), and $\sigma_w$ and $\sigma_c$ are the white and correlated noise standard deviations, the subscript.
-\bigskip\\
-
-\subsection{Instrumental noises}
-
-The study from \citealt{pont2006} highlight the importance of accounting for correlated noise ($\sigma_r$) when estimating the yield of transit search campaigns.
-To demonstrate its point, we simulate a unique transit signal (using the empirical analytical model from \citealt{protopapas}) and compute its SNR using \autoref{eq:snr}, both in the presence and absence of correlated noise (\autoref{fig:issue1}).
+where $df$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits (unity here since we consider a single transit), and $\sigma_w$ and $\sigma_c$ are the white and correlated noise standard deviations.
+In this study, the authors highlight the importance of accounting for correlated noise ($\sigma_r$) when estimating the yield of transit search campaigns.
+To demonstrate this point, we simulate a unique transit signal and compute its SNR using \autoref{eq:snr}, both in the presence and absence of correlated noise (\autoref{fig:issue1}).
 
@@
     \begin{centering}
-        \includegraphics[width=8.5cm]{../figures/first_issue.pdf}
-        \caption{This is a pretty visualization of the Mandelbrot set.}
+        \includegraphics[width=8.5cm]{../figures/issue1.pdf}
+        \caption{Illustration of the effect of correlated noise on a single transit singal-to-noise (SNR).
+A 1 hour transit of depth 1\% is simulated on top of white noise during a 24 hours observing window with an exposure time of 1 minute (top).
+Then, in the bottom plot, correlated noise is added to the transit signal and simulated using a Gaussian Process (GP) with a Matèrn-32 kernel of length-scale 1 hour and amplitude 0.2\%.
+The SNR on the right of each light curve is computed using \autoref{eq:snr}.}
         \label{fig:issue1}
@@
 
-As illustrated in \autoref{fig:issue1}, the presence of correlated noise fundamentally limits transit detectability.
+As illustrated in \autoref{fig:issue1}, the presence of correlated noise strongly decreases transit signals SNR, fundamentally limiting their detection.
 This issue rapidly motivated the development of systematics detrending algorithms such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}), or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}).
-Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal.
+Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal (in practice it does).
 We note that, except for \textsc{TFA}, these algorithms are mostly applied to space-based continuous observations, that provide continuous stellar baselines and mostly reproducible systematic signals.
@@
 
-\subsection{Stellar variability}
+\subsection{The effect of detrending on transits detectability}
 
-Instrumental signals have the benefit to be shared among light curves of the same field, and strongly correlated with measurements from the experimental setup (like detector's temperature, pointing error, sky background or airmass).
+Instrumental signals have the benefit to be shared among light curves of stars observed with the same instrument, strongly correlated with measurements from the experimental setup (like detector's temperature, pointing error, sky background or airmass), hence we will make the strong assumption that their detrending based on an incomplete model of the light curve, one that ignore transit signals (because unknowns), do not affect the latter.
 In opposition, stellar variability is generally unknown and harder to correlate with simultaneous measurements.
-This gave rise to two types of treatments to reconstruct and remove stellar variability to perform a simpler transit search.
+This gave rise to two types of treatments in order to reconstruct and detrend stellar variability, and perform transit search on \textit{flattened} light curves.
 One is physically-motivated and make use of Gaussian Processes (e.g.
@@
 The other is empirical and make use of filtering algorithms (a wide variety being described in \cite{wotan}).
-In either case, the complete light curve is modeled while ignoring the presence of transits.
-In this section, we will study in details the effect of both approaches on transit search, by studying the degradation of a unique transit SNR for a wide variety of stellar variability characteristics.
+In this section, we will study in details the effect of both approaches on transit detectability, by studying the degradation of a unique transit SNR for a wide variety of stellar variability characteristics.
 \bigskip\\
@@
 In \autoref{fig:issue2}, we simulate a light curve featuring a single transit (again using the model from \citealt{protopapas}), in addition with stellar variability of different timescales and amplitudes.
-In case \textit{a} (purple in \autoref{fig:issue2}), we model and detrend these signals using the widely-adopted Tukey's biweight filtering method presented in \citealt{tukey} and its implementatation from \texttt{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan} (case a.
-in purple).
-In case \textit{b} (red in \autoref{fig:issue2}) we employ the same GP model used to simulate stellar variability to reconstruct it, and remove it from light curves ignoring the presence of a potential transits (as it is the case in unknown transits search).
+In case \textit{a} (purple in \autoref{fig:issue2}), we model and detrend these signals using the widely-adopted Tukey's biweight filtering method presented in \citealt{tukey} and its implementatation from \texttt{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}.
+In case \textit{b} (red in \autoref{fig:issue2}) we employ the same GP model used to simulate stellar variability to reconstruct and remove it from light curves ignoring the presence of potential transits (the case of unknown transits search).
 Once transits cleaned from stellar variability, we asses their remaining SNR using \autoref{eq:snr}.
@@
 \autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and suggests that this degradation is strongly dependant on the stellar variability characteristics encountered.
-In order to explore the parameter space for which detrending is the most problematic, we simulate light curves including a single transit and a much wider range of correlated noise characteristics defined with two parameters: $\delta_v$, the relative amplitude of the variability against transit depth; and $\tau_v$ the relative timescale of the variability with respect to the transit duration.
+In order to explore the parameter space for which detrending is the most problematic, we simulate light curves including a single transit and a much wider range of correlated noise characteristics defined with two parameters: $\tau_v$ the relative timescale of the variability with respect to the transit duration; and $\delta_v$, the relative amplitude of the variability against transit depth, such that:
+
+\begin{equation}
+    \tau_v \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}} \quad \text{and} \quad 
+    \delta_v \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}} 
+\end{equation}
+
 To follow this parametrization, we simulate the photometric variability using a GP with an SHO kernel with hyperparameters \footnote{}:
@@
 \begin{equation}\label{eq:params}
-    \omega = \frac{\pi}{\tau_v\tau}, \quad 
+    \omega = \frac{\pi}{\tau_v\tau} \quad 
     \sigma = \delta_v \frac{\delta^2}{4} \quad  \text{and}  \quad  
@@
 with $\delta$ and $\tau$ the depth and duration of the simulated transit, similar in all light curves.
-We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability.
-The expressions of $\omega$ and $\sigma$ given in \autoref{eq:params} correspond to a periodic signal with a period half that of the transit duration, and a peak to peak amplitude two times that of the transit depth, so that a varibility with $(\tau_v=1, \delta_v=1)$ strongly resembles the simulated transit signal.
-Then, as in \autoref{fig:issue2}, we reconstruct the variability using a biweight filter with a window length of $3\tau$ (three times the transit duration, an optimal value according to \citealt{wotan}).
-We then subtract the variability from each light-curve, estimate the resulting transit depth and compute its SNR.
+We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability signals.
+For $\tau_v=1$ and $\delta_v=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:params} correspond to a periodic signal with a period half that of the transit duration, and a peak to peak amplitude two times that of the transit depth, i.e.
+strongly resembling the simulated transit signal.
+As in \autoref{fig:issue2}, we reconstruct the variability using a biweight filter with a window length of $3\tau$ (three times the transit duration, an optimal value according to \citealt{wotan}).
+We then subtract the variability from each light-curve, estimate the resulting transit depth (using the in-transit minimum flux) and compute its SNR.
 \autoref{fig:snr_detrend} shows the remaining SNR values of transits computed this way, after each variability signals with random $(\tau_v, \delta_v)$ have been detrended.
+\bigskip\\
+\autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau_v, \delta_v)$ parameter space for which detrending degrades transit SNR to the point of no detection ($SNR < 6$).
+While this region might only represent a fraction of existing exoplanetary systems, we argue that these systems are of great value for the field of exoplanetary science.
+First, as variability is often linked to starspots co-rotation with their star, systems whose  with the stellar photo in is likely to be observed on stars whose equatorial plane lines up with our line of sight, following recent studies on the dynamical evolution of planetary systems \cite{}, hence being most likely to be transited.
+As a direct consequence, these same starspots are more likely to be occulted by planetary companions, events whose measurement benefits both the study of stellar atmospheres and their concerning impact on planetary atmospheric retrievals \cite{}.
+Finally, the growing interest of the community for ultra-cool dwarf stars comes with observations done in the infrared, featuring enhanced red noise and variability, leading to lower transit SNR \cite{}.
+Hence, commonly-used detrending technique (such as the biweight filter studied in this section) make transit-search blind to those pristine systems, motivating the need for a more informed transit search algorithm properly accounting for correlated noise.
+In the next section we present such an algorithm: \texttt{nuance}.
 
@@
 
+\section{Injection-recovery on simulated data}
+
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[height=12cm]{../../workflows/synthetic_injection_recovery/figures/final_result.pdf}
+        \caption{This is a pretty visualization of the Mandelbrot set.}
+        \label{fig:simu}
+    \end{centering}
+\end{figure}
+
 \bibliography{bib}

### 2022-11-25 simple/scripts/plot_cleaning_SNR.py

@@ -1,139 +1,0 @@
-#!/usr/bin/env python
-# coding: utf-8
-
-# In[66]:
-
-
-import pandas as pd
-import numpy as np
-import yaml
-from scipy.stats import binned_statistic_2d
-import sys
-sys.path.append("../workflows/synthetic_injection_recovery/scripts")
-from make_lightcurve import make_lc
-from make_params import make
-import matplotlib.pyplot as plt
-from matplotlib.patches import ConnectionPatch
-import matplotlib.pyplot as plt
-from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-
-df = pd.read_csv("../workflows/cleaning_snr/data/simu1/results.csv")
-
-
-# In[ ]:
-
-
-n = 3 # number of small plots
-scale = 1.2
-W = n+1
-H = n+1
-
-fig = plt.figure(None, (1.6*W*scale, H*1.2*scale))
-
-def I(i, j):
-    return int(j*W + i) + 1
-
-for i in range(H*W):
-    ax = plt.subplot(H, W, i+1)
-    
-## Main plot
-main = plt.subplot(H, W, (I(0, 1), I(W-2, H-1)))
-
-bins=(30, 30)
-var, amp, snr = df.values.T.astype(float)
-stats = binned_statistic_2d(var, amp, snr, bins=bins)
-im = main.imshow(
-    stats.statistic.T, 
-    origin="lower",
-    extent=(
-        stats.x_edge.min(), 
-        stats.x_edge.max(), 
-        stats.y_edge.min(), 
-        stats.y_edge.max()
-    ),
-    aspect="auto",
-    vmax=15
-)
-
-main.set_ylabel(r"$\delta_v \sim \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$", fontsize=14)
-main.set_xlabel(r"$\tau_v \sim \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$", fontsize=14)
-params = yaml.full_load(open("../workflows/cleaning_snr/data/simu1/params/0.yaml", "r"))
-original_snr = params["depth"]/np.sqrt(params["error"]**2/n)
-
-main.set_title(f"Detrending effect on SNR\n", loc="left")
[…]

### 2022-11-25 simple/scripts/plot_issue_1.py

@@ -1,42 +1,9 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import jax, jaxopt
+import jax
 jax.config.update("jax_enable_x64", True)
-import jax.numpy as jnp
-from tinygp import kernels, GaussianProcess
 import matplotlib.pyplot as plt
 import tinygp
-
-def white_red(time, flux, error, plot=False, mean=0.):
-    init = {
-        "log_diag": 2*np.log(np.mean(error)),
-        "log_scale": np.log(1), # 5 mins
-        "log_amp": np.log(np.std(flux))
-    }
-
-    def build_gp(params):
-        kernel = jnp.exp(params["log_amp"]) * kernels.ExpSquared(jnp.exp(params["log_scale"]))
-        return GaussianProcess(kernel, time, diag=jnp.exp(params["log_diag"]), mean=mean)
-
-    @jax.jit
-    def loss(params):
-        gp = build_gp(params)
-        return -gp.log_probability(flux)
-    
-    opt = jaxopt.ScipyMinimize(fun=loss)
-    params = opt.run(init).params  
-    w = np.sqrt(np.exp(params['log_diag']))
-    
-    if plot:
-        gp = build_gp(params)
-        _, cond = gp.condition(y=flux, X_test=time)
-        mu = cond.mean
-        plt.plot(time, flux, ".", c="0.8")
-        plt.plot(time, mu, c="k")
-        noise = w
-        plt.fill_between(time, mu + noise, mu - noise, color="C0", alpha=0.3)
-
-    return w, np.sqrt(np.exp(params['log_amp']))
 
 # Simulation
 # ------------
@@ -47,21 +14,23 @@ fluxes = []
 snrs = []
 
 # transit
-time = np.linspace(0, 3, 1000)
-dur = 0.1
-depth = 0.8e-2
+dur = 1/24
+depth = 1e-2
+total_time = 1
+time = np.arange(0, total_time, 1/60/24)
 n = dur/np.median(np.diff(time))
-signal = transit(time, time.mean(), dur, depth, c=10) + 1.
+signal = transit(time, time.mean(), dur, depth) + 1.
 
 # white noise
-wsigma = 1.5e-3
-white_noise = np.random.normal(0, wsigma, len(time))
+w = 1.5e-3
[…]

### 2022-11-25 workflows/cleaning_snr/scripts/plot_snr_cleaning.py

@@ -20,7 +20,7 @@ scale = 1.2
 W = n+1
 H = n+1
 
-fig = plt.figure(None, (1.6*W*scale, H*1.2*scale))
+fig = plt.figure(None, (1.5*W*scale, H*1.2*scale))
 
 def I(i, j):
     return int(j*W + i) + 1
@@ -44,16 +44,16 @@ im = main.imshow(
         stats.y_edge.max()
     ),
     aspect="auto",
-    vmax=15
+    vmax=5
 )
 
-main.set_ylabel(r"$\delta_v \sim \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$", fontsize=14)
-main.set_xlabel(r"$\tau_v \sim \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$", fontsize=14)
+main.set_ylabel(r"$\delta_v \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$", fontsize=14)
+main.set_xlabel(r"$\tau_v \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$", fontsize=14)
 params = true_depth
 original_snr = true_depth/np.sqrt(np.mean(error)**2/n)
 
-main.set_title(f"Detrending effect on SNR\n", loc="left")
-main.text(0, stats.y_edge.max()+0.4, f"original SNR: {original_snr:.2f}", color="0.3", va="center")
+main.set_title(f"Detrending effect on SNR", loc="left")
+main.text(0.1, stats.y_edge.max()+0.26, f"original SNR: {original_snr:.2f}", color="0.3", va="center")
 
 # removing some axes
 # ------------------
@@ -107,20 +107,12 @@ for i, v in enumerate(var):
     #fig.add_artist(con)
     #main.plot(v, a, ".", c=pcolor, ms=7, alpha=0.5)
     
-fig.tight_layout(pad=1.5)
-axins = inset_axes(
-    main,
-    width="30%",
-    height="5%",
-    loc="upper right",
-    bbox_to_anchor=(0., 0., 1, 1.08),
-    bbox_transform=main.transAxes,
-    borderpad=0,
-)
-cb = fig.colorbar(im, cax=axins, orientation="horizontal", ticks=[1, 5, 10, 15])
-cb.set_label("SNR", labelpad=-10, x=-.15)
[…]

### 2022-11-25 workflows/synthetic_injection_recovery/scripts/plot_final.py

@@ -1,0 +1,125 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy.stats import binned_statistic_2d
+from matplotlib import patches
+
+nuance_main, wotan_main, nuance_zoom, wotan_zoom = snakemake.input
+
+ylabel = r"$\delta_v \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$"
+xlabel = r"$\tau_v \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$"
+
+fig = plt.figure(None, (7, 6.5))
+
+bins = (30,30)
+
+# wtls
+# ----
+main = plt.subplot(2, 2, 1)
+main.set_ylabel(ylabel, fontsize=14)
+df = pd.read_csv(wotan_main)
+detected, _var, _amp = df.values.T.astype(float)
+stats = binned_statistic_2d(_var, _amp, detected, bins=bins)
+im = main.imshow(
+    stats.statistic.T, 
+    origin="lower",
+    extent=(
+        stats.x_edge.min(), 
+        stats.x_edge.max(), 
+        stats.y_edge.min(), 
+        stats.y_edge.max()
+    ),
+    aspect="auto"
+)
+main.set_title("wõtan + TLS", loc="left")
+
+
+zoomw = plt.subplot(2, 2, 2)
+bins = (30,30)
+df = pd.read_csv(wotan_zoom)
+detected, _var, _amp = df.values.T.astype(float)
+stats = binned_statistic_2d(_var, _amp, detected, bins=bins)
+zoomw.imshow(
+    stats.statistic.T, 
+    origin="lower",
+    extent=(
+        stats.x_edge.min(), 
+        stats.x_edge.max(), 
+        stats.y_edge.min(), 
+        stats.y_edge.max()
+    ),
+    aspect="auto"
+)
+
+a = (0.5, 1) # xy of lower left box
+b = (10, 5) # xy of upper right box
+pcolor="0.6"
+p = patches.Polygon([a, (a[0], b[1]), (b[0], b[1]), (b[0], a[1])], fill=False, ec=pcolor)
+main.add_patch(p)
[…]

### 2023-01-17 simple/latex/ms.tex
real dataset injections + nuance formalism

@@
 \usepackage{float}
+\usepackage{amsmath}
+\usepackage{hyperref}
+\usepackage{bm}
 
@@
 
-% Author list
-\author{@lgrcia}
+% Authorz list
+\author{Lionel J. Garcia, Daniel Foreman-Mackey, Fran Pozuelos}
 
@@
 \begin{abstract}
-    We present \texttt{nuance}, an algorithm to simultaneously search for transits while modeling correlated noise in a tractable way, including instrumental signals and stellar photometric variability.
+    We present \texttt{nuance}, an algorithm to search for planetary transits in light curves featuring correlated noises, such as instrumental signals and stellar photometric variability.
+Simpler approaches consist in searching for transits in light curves cleaned from correlated noise, where only transit signals would remain.
+However, we show that commonly used detrending techniques strongly affect transits signal-to-noise, up to the point of no detection when the noise characteristics is close to that of the transits.
+Focusing on stellar variability, we explore the parameter space for which this degradation occur, and quantify its effect on variety of cases.
 \end{abstract}
@@
 This is particularly true for systems whose orbital plane is aligned with our line of sight, leading to observable planetary transits.
-Although these transit signals can be seen in the apparent flux received from their stars (light curves), they are often mixed with other astrophysical and instrumental signals, together referred to as \textit{correlated noise}.
+Although these transit signals can be seen in the apparent flux received from their stars, they are often mixed with other astrophysical and instrumental signals.
+% together referred to as \textit{correlated noise}. 
 If they can be disentangled from these nuisance signals, transits offer a powerful way to detect exoplanets.
 However, disentangling these signals comes with many challenges.
-\bigskip\\
-The simplest way to find periodic transit signals is to first clean a light curve from nuisance signals before performing the search.
-Indeed, widely used transit-search algorithms (namely Box-Least-Square algorithms, \cite{}) are capable of detecting transits on flat light curves only containing white noise.
-This strategy is widely adopted by the community, both using physically-motivated systematic models like \cite{everest1, everest2}, or empirical filtering techniques, such as the ones described and implemented in \cite{wotan}.
-However, when systematics and variability start to resemble transits, this cleaning step (often referred to as \textit{detrending}) strongly alters the searched signals.
+Due to their nature we will refer to these signal as correlated noises\\
+Widely used transit-search algorithms (Box-Least-Square algorithms \cite{bls}) are capable of detecting transits on light curves only containing transits and white noise. Hence, 
+the simplest way to find periodic transit signals is to first clean a light curve from nuisance signals before performing the search.
+This strategy is widely adopted by the community, both using physically-motivated systematics models like \cite{everest1, everest2}, or empirical filtering techniques, such as the ones described and implemented in \cite{wotan}.
+However, when correlated noises start resembling transits, this cleaning step (often referred to as \textit{detrending}) strongly degrades their detectability.
 In such cases, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noises, and asses the likelihood of the model to the data on a wide parameter space, an approach largely avoided due to its untractable nature.
-Nonetheless, \citealt{kovacs2016} face the question by asking: \textit{Periodic transit and variability search with simultaneous systematics filtering: Is it worth it?}.
-This study discards the general benefit of using a full-fledged approach, it fails at exploring the light-curves characteristics for which it becomes necessary.
-While it might only represent a handful of systems, this cases are extremely valuable for the exoplanetary science community.
+Nonetheless, \citealt{kovacs2016} ask the question: \textit{Periodic transit and variability search with simultaneous systematics filtering: Is it worth it?}.
+The latter study discards the general benefit of using a full-fledged approach, but it fails at exploring the light-curves characteristics for which it becomes necessary.\\
+
+% While it might only represent a handful of systems, this cases are extremely valuable for the exoplanetary science community.
 First, variability may be associated with star-spots that can be probed with the help of planetary transits.
@@
 Second, the growing interest of the community for ultra-cool dwarf stars comes with observations featuring enhanced red noise, stellar variability and lower transit SNR.
-Hence, these hidden system are pristine.
-\bigskip\\
-In this paper we present nuance, an algorithm using linear models and gaussian processes to simultaneously search for transits while modeling correlated noise in a tractable way, including instrumental signals and stellar photometric variability.
-In \autoref{issues}, we describe the issues inherent to the two-step approach described earlier, and study the parameter space for which commonly used filtering techniques alter transit signals to the point of no detection.
-In section 2 we describe the tractable full-fledged approach on which nuance is based, and its implementation in an open-source Python package.
-In section 3, we asses the performance of nuance by computing the recovery rate of planetary transits injected into simulated light curves.
+Hence, these hidden system are pristine.$
+We present \texttt{nuance}, an algorithm using linear models and Gaussian processes (GP) to simultaneously search for transits while modeling correlated noise in a tractable way, such as instrumental signals and stellar photometric variability.
+In \autoref{issues}, we describe the issues inherent to the two-step approach described earlier, and study the parameter space for which commonly used detrending techniques degrade transit signals to the point of no detection.
+In \autoref{nuance} we describe the tractable full-fledged approach of \texttt{nuance}, and its implementation in an open-source Python package.
+In \autoref{simu} we asses the performance of \texttt{nuance} by computing the recovery of planetary transits injected into simulated light curves.
 In section 4, we use nuance in a case study, to detect a known planet observed by the T E S S but undetected by the TESS pipeline due to its light curve characteristics.
@@
 \newpage
-\section{The issue with detrending}\label{issues}
-Two sources of correlated noise particularly justify the need for a detrending step before searching for transits: systematic noise (such as telescope pointing errors) and stellar variability (induced by  pulsations or starspots).
-In this section, we discuss the impact of such correlated noise and its detrending on transits detectability.
-For simplicity, we will model transit signals using an analytic empirical model described in \citealt{protopapas} (see Annexe).
-\bigskip\\
+\section{The issue with coorelated noise and its detrending}\label{issues}
+Two sources of correlated noise particularly justify the need for a detrending step before searching for transits: instrumental noises (such as telescope pointing errors) and stellar variability (induced by  pulsations or starspots).
+In this section, we discuss the impact of such correlated noises and their detrending on transits detectability.
+For simplicity, we will model transit signals using an analytic empirical model described in \citealt{protopapas} (see Annexe).\\
 
@@
 where $df$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits (unity here since we consider a single transit), and $\sigma_w$ and $\sigma_c$ are the white and correlated noise standard deviations.
-In this study, the authors highlight the importance of accounting for correlated noise ($\sigma_r$) when estimating the yield of transit search campaigns.
-To demonstrate this point, we simulate a unique transit signal and compute its SNR using \autoref{eq:snr}, both in the presence and absence of correlated noise (\autoref{fig:issue1}).
+To show the impact of correlated noise on transit detectability, we simulate a unique transit signal and compute its SNR using \autoref{eq:snr}, both in the absence and presence of correlated noise (\autoref{fig:issue1}).
 
@@
 
-As illustrated in \autoref{fig:issue1}, the presence of correlated noise strongly decreases transit signals SNR, fundamentally limiting their detection.
+As illustrated in \autoref{fig:issue1}, the presence of correlated noise strongly decreases the transit signal SNR, limiting its detection.
 This issue rapidly motivated the development of systematics detrending algorithms such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}), or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}).
-Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal (in practice it does).
+Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal.
 We note that, except for \textsc{TFA}, these algorithms are mostly applied to space-based continuous observations, that provide continuous stellar baselines and mostly reproducible systematic signals.
@@
 The other is empirical and make use of filtering algorithms (a wide variety being described in \cite{wotan}).
-In this section, we will study in details the effect of both approaches on transit detectability, by studying the degradation of a unique transit SNR for a wide variety of stellar variability characteristics.
-\bigskip\\
+In this section, we will study in details the effect of both approaches on transit detectability, by studying the degradation of a unique transit SNR for a wide variety of stellar variability characteristics.\\
+
 Throughout this paper, we simulate stellar variability (and later model it) thanks to Gaussian Processes (GP), employing the physically-motivated stochastically-driven damped simple harmonic oscillator kernel (SHO) presented in \citealt{celerite1} (see Annexe ...).
@@
         \includegraphics[width=\linewidth]{../figures/issue2.pdf}
-        \caption{This is a pretty visualization of the Mandelbrot set.}
+        \caption{}
         \label{fig:issue2}
@@
 In case \textit{b} (red in \autoref{fig:issue2}) we employ the same GP model used to simulate stellar variability to reconstruct and remove it from light curves ignoring the presence of potential transits (the case of unknown transits search).
-Once transits cleaned from stellar variability, we asses their remaining SNR using \autoref{eq:snr}.
-\bigskip\\
+Once transits cleaned from stellar variability, we asses their remaining SNR using \autoref{eq:snr}.\\
+
 \autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and suggests that this degradation is strongly dependant on the stellar variability characteristics encountered.
@@
     \omega = \frac{\pi}{\tau_v\tau} \quad 
-    \sigma = \delta_v \frac{\delta^2}{4} \quad  \text{and}  \quad  
+    \sigma = \delta_v \frac{\delta}{2} \quad  \text{and}  \quad  
     Q = 10
@@
 \autoref{fig:snr_detrend} shows the remaining SNR values of transits computed this way, after each variability signals with random $(\tau_v, \delta_v)$ have been detrended.
-\bigskip\\
+% \bigskip\\
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
+        \caption{}
+        \label{fig:snr_detrend}
+    \end{centering}
+\end{figure}
 \autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau_v, \delta_v)$ parameter space for which detrending degrades transit SNR to the point of no detection ($SNR < 6$).
 While this region might only represent a fraction of existing exoplanetary systems, we argue that these systems are of great value for the field of exoplanetary science.
-First, as variability is often linked to starspots co-rotation with their star, systems whose  with the stellar photo in is likely to be observed on stars whose equatorial plane lines up with our line of sight, following recent studies on the dynamical evolution of planetary systems \cite{}, hence being most likely to be transited.
+First, as variability is often linked to the presence of starspots, systems whose  with the stellar photo in is likely to be observed on stars whose equatorial plane lines up with our line of sight, following recent studies on the dynamical evolution of planetary systems \cite{}, hence being most likely to be transited.
 As a direct consequence, these same starspots are more likely to be occulted by planetary companions, events whose measurement benefits both the study of stellar atmospheres and their concerning impact on planetary atmospheric retrievals \cite{}.
 Finally, the growing interest of the community for ultra-cool dwarf stars comes with observations done in the infrared, featuring enhanced red noise and variability, leading to lower transit SNR \cite{}.
-Hence, commonly-used detrending technique (such as the biweight filter studied in this section) make transit-search blind to those pristine systems, motivating the need for a more informed transit search algorithm properly accounting for correlated noise.
+Hence, commonly-used detrending technique (such as the biweight filter studied in this section) make transit-search blind to those pristine systems, motivating the need for a more informed transit search algorithm able to deal with correlated noises.
 In the next section we present such an algorithm: \texttt{nuance}.
+
+\newpage
+\section{\texttt{nuance}}\label{nuance}
+
+In this section we describe \texttt{nuance}, an algorithm to search for planetary transits in light curves featuring correlated noises, such as instrumental signals and stellar photometric variability.
+Our approach is based on \cite{foreman2016}, developed to detect planetary transits accounting for a linear model of systematics in K2 light curves.
+\texttt{nuance} differs by modeling the covariance of the light curve with a gaussian process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.\\
+
+\subsection{Preamble}
+
+The flux $f$ of a star measured at time $t$ is sampled and aranged in the $(1\times N)$ column-vector $\bm{f}$.
+This flux contains instrumental signals, stellar variability and a periodic transit signal that we wish to uncover.
+An example of such signal is simulated and shown in \autoref{fig:principle_dataset}.
 
@@
     \begin{centering}
-        \includegraphics[height=10cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
-        \caption{This is a pretty visualization of the Mandelbrot set.}
-        \label{fig:snr_detrend}
+        \includegraphics[width=0.7\linewidth]{../figures/principle_dataset.pdf}
+        \caption{}
+        \label{fig:principle_dataset}
     \end{centering}
@@
 
-\section{\texttt{nuance}}
+Simultaenously to the flux measurement, we obtained a set of $K$ measurements, aranged in the $(K\times N)$ matrix $\bm{X}$, that can be treated as explanatory variables for $\bm{f}$.
+Details about the simulation of these signals are provided in \autoref{app_principle_simulations}.
+\\\\
+Ideally, we would detect such  a periodic transit signal by sampling the posterior likelihood of this data to a full-fledged model including stellar variability (more generally correlated noises), instrumental systematics (modeled with explanatory variables), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$.
+We would then reduce this posterior likelihood to $p(\bm{f}\vert P)$, its marginalized version over all parameters except the period $P$, producing a transit search periodogram.
+However, this approach has two issues: It is highly untractable, and it may lead to multimodal distributions that are hard to interpret.
+\\\\
+Given a period $P$, we instead want to compute the likelihood of a periodic transit signal at the maximum likelihood parameters $T_0$, $D$ and $\Delta$, i.e the periodogram
 
-\section{Injection-recovery on simulated data}
+$$\mathcal{Q}(P) = \max_{T_0, D, \Delta}p(\bm{f} \vert P, T_0 ,D, \Delta)$$
+\\
+We will do that by adopting the strategy of \cite{foreman2016}, and separate the transt search into two components: the \textit{linear search} and the \textit{periodic search}.
+During the \textit{linear search}, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
+Then, the \textit{periodic search} consists in combining these likelihoods to compute the likelihood of a periodic transit signal for a range of periods.
+These combined likelihoods yield a transit-search periodogram on which the periodic transit detection can be based.
+
+\subsection{The linear search}
+
+During the linear search, the goal is to compute the likelihood of a single non-periodic transit signal of epoch $T_0$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths, i.e.
+
+$$\{p(\bm{f} \vert T_i ,D_j, \Delta_k)\}_{i, j, k}$$
+
+To account for correlated noises, we model the light curve $f$ as being drawn from a Gaussian Process such that
+
+$$\bm{f} \sim \mathcal{N}(\bm{w X}, \bm{\Sigma})$$
+
+with mean $\bm{wX}$ (i.e.
+a linear model of the K explanatory variables with coefficients $\bm{w}$) and covariance $\bm{\Sigma}$.
+We include the non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$ to our model by adding it as the last column of the design matrix $\bm{X}$, so that the transit is part of the linear model and the transit depth $\Delta$ can be solved linearly.
+Under this assumption, the log-likelihood of the data to the model, including a single non-periodic transit, can be computed analyticaly
+
+$$\ln p(\bm{f} \vert \bm{X}) = -\frac{1}{2}(\bm{f}-\bm{wX})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{wX}) -  \frac{1}{2}\vert\bm{\Sigma}\vert - \frac{K}{2}\ln 2$$
+
+and the linear search is reduced to the computation of the latter log-likelihood over a grid of epochs and durations only, i.e.\footnote{In what follows, $\Delta_k$ is ommited from the notations, being linearly solved for any $(T_0, D)$}
+
+$$\{\ln p(\bm{f} \vert T_i ,D_j)\}_{i, j}$$
+
+Figure \autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}.
+To compute this grid, we used the same Gaussian Process and design matrix $\bm{X}$ used to generate the data (see annexe).
+
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.75\linewidth]{../figures/principle_linear_search.pdf}
+        \caption{}
+        \label{fig:linear_search}
+    \end{centering}
+\end{figure}
+
+\subsubsection{The periodic search}
+
+We then need to combine the likelihoods computed from the linear search to obtain\footnote{Since we search for the maximum likelihood of $p(\bm{f}\vert...)$, we ommit $\bm{w}$ from the remaining notatitions, being linearly solved and taken at its maximum likelihood value}
+
+$$\max_{T_0, D}p(\bm{f} \vert P, T_0 ,D)$$
+
+If each transits were to be considered truly independant, we would have
+
+$$p(\bm{f} \vert P, T_0 ,D) \sim \prod_k p(\bm{f} \vert T_k ,D_k)$$
+
+where $T_k$ are the epochs matching ($T_0$, $P$) and $D_k$ the durations of the corresponding transits. However, doing so leads to two issues:
+\begin{enumerate}
+
+    \item Transits are not truly independant and, instead of having their own depths $\Delta_k$, should share a common transit depth $\Delta$.
+We show in Annexe that there is an analytical expression for the likelihood of a periodic transit signal of depth $\Delta$ whose individual transit depths $\Delta_k$ with error $\sigma_k$ have been measured independantly:
+
+    $$p(y \vert P, T_0 ,D) =  \prod_k \frac{\sigma_k \bar\sigma \, e^{- \frac{\left(\Delta_k - \Delta\right)^{2}}{2 \sigma_k^{2} + \bar\sigma^{2}}}}{\sigma_k^{2} + \bar\sigma^{2}}$$
+
+    $$\text{with} \quad  
+    \frac{1}{\bar\sigma^2} = \sum_{k}\frac{1}{\sigma_k^2} \quad \text{and} \quad
+    \Delta = \bar\sigma^2 \sum_{k}{\frac{\Delta_k}{\sigma_k^2}}$$
+
+    \item Individual epochs $T_k$ matching ($T_0$, $P$) are not necessarly available in the grid of epochs $\{T_i\}$.
+In \cite{foreman2016} this is solved by using the nearest neighbors in the epochs grid.
+In \texttt{nuance}, we rather interpolate the likelihood grid from $T_i$ to $T_k$ (see \autoref{fig:periodic_search}).
+
+\end{enumerate}
+
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.8
+        \linewidth]{../figures/principle_tiles_combined.pdf}
+        \caption{}
+        \label{fig:periodic_search}
+    \end{centering}
+\end{figure}
+
+For a given period $P$, we can then compute $\max_{(T_0, D)}\{p(\bm{f} \vert P, T_0 ,D)\}$.
+But instead of relying on this maximum likelihood to build the final periodorgram, we compute the SNR of the transit with period $P$ and $(T_0, D)$ maximizing $p(\bm{f} \vert T_0 ,D, P)$, yielding the transit SNR periodogram shown in \autoref{fig:periodic_search}.
+We leave a discussion of this choice for Annexe .
+\\\\
+Once a transit recoverd, we mask the corresponding epochs in the log likelihood linear grid
+
+\subsubsection{An open-source python package}
+
+\section{Injection-recovery on simulated data}\label{simu}
 
@@
         \includegraphics[height=12cm]{../../workflows/synthetic_injection_recovery/figures/final_result.pdf}
-        \caption{This is a pretty visualization of the Mandelbrot set.}
+        \caption{}
         \label{fig:simu}
@@
 
+\section{Comparison with \texttt{Sherlock} on TOI-540}\label{real}
+
+\section{Performances and limitations}\label{perf}
+
+\section{Conclusion}
+
+This simulated dataset constits in a light curve featuring a transit signal of depth $1$\%, duration $0.05$ days and period $1.3$ days using the simple \citealt{protopapas} analytical model.
+It also contains a simulated variability signal drawn from a Gaussian Process with a quasiseparable
+
+\appendix
+\section{Light curve simulations}\label{app_principle_simulations}
 \bibliography{bib}

### 2023-01-17 workflows/synthetic_injection_recovery/scripts/plot_final.py

@@ -30,7 +30,9 @@ im = main.imshow(
         stats.y_edge.min(), 
         stats.y_edge.max()
     ),
-    aspect="auto"
+    aspect="auto",
+    vmin=0.,
+    vmax=1.
 )
 main.set_title("wõtan + TLS", loc="left")
 
@@ -49,11 +51,13 @@ zoomw.imshow(
         stats.y_edge.min(), 
         stats.y_edge.max()
     ),
-    aspect="auto"
+    aspect="auto",
+    vmin=0.,
+    vmax=1.
 )
 
 a = (0.5, 1) # xy of lower left box
-b = (10, 5) # xy of upper right box
+b = (10, 25) # xy of upper right box
 pcolor="0.6"
 p = patches.Polygon([a, (a[0], b[1]), (b[0], b[1]), (b[0], a[1])], fill=False, ec=pcolor)
 main.add_patch(p)
@@ -61,8 +65,8 @@ con = patches.ConnectionPatch(xyA=(a[0], b[1]), xyB=(0.1, b[1]-0.1), coordsA="da
 fig.add_artist(con)
 con = patches.ConnectionPatch(xyA=(b[0], a[1]), xyB=(0.1, 0.1), coordsA="data", coordsB="data", axesA=main, axesB=zoomw, color=pcolor)
 fig.add_artist(con)
-zoomw.set_ylim(0.1, 5)
-zoomw.set_xlim(0.1, 10)
+zoomw.set_ylim(0.1, b[1])
+zoomw.set_xlim(0.1, b[0])
 
 # nuance
 # ------
@@ -81,7 +85,9 @@ main.imshow(
         stats.y_edge.min(), 
         stats.y_edge.max()
     ),
-    aspect="auto"
+    aspect="auto",
+    vmin=0.,
+    vmax=1.
 )
 main.set_title("nuance", loc="left")
 
@@ -98,11 +104,11 @@ zoom.imshow(
         stats.y_edge.min(), 
         stats.y_edge.max()
     ),
-    aspect="auto"
+    aspect="auto",
+    vmin=0.,
+    vmax=1.
 )
 
-a = (0.5, 1) # xy of lower left box
-b = (10, 5) # xy of upper right box
 pcolor="0.6"
 p = patches.Polygon([a, (a[0], b[1]), (b[0], b[1]), (b[0], a[1])], fill=False, ec=pcolor)
[…]

### 2023-01-18 simple/latex/ms.tex
principle and math details

@@
 
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+
 % Begin!
@@
 
-% Authorz list
+% Authors list
 \author{Lionel J. Garcia, Daniel Foreman-Mackey, Fran Pozuelos}
@@
 \label{sec:intro}
-
 Exoplanets, planets outside our solar system, are discovered at an ever-increasing rate.
-Beyond the study of their inner structure and atmosphere, they give a unique glimpse to extrasolar systems formation, dynamics, as well as being probes to understand exoplanets' host stars.
+Beyond the study of their inner structure and atmosphere, they give a unique glimpse to extrasolar systems' formation, dynamics, as well as being probes to understand exoplanets' host stars.
 This is particularly true for systems whose orbital plane is aligned with our line of sight, leading to observable planetary transits.
 Although these transit signals can be seen in the apparent flux received from their stars, they are often mixed with other astrophysical and instrumental signals.
-% together referred to as \textit{correlated noise}. 
 If they can be disentangled from these nuisance signals, transits offer a powerful way to detect exoplanets.
@@
 Due to their nature we will refer to these signal as correlated noises\\
-Widely used transit-search algorithms (Box-Least-Square algorithms \cite{bls}) are capable of detecting transits on light curves only containing transits and white noise. Hence, 
-the simplest way to find periodic transit signals is to first clean a light curve from nuisance signals before performing the search.
+Widely used transit-search algorithms (Box-Least-Square algorithms \cite{bls}) are capable of detecting transits on light curves only containing transits and white noise.
+Hence, the simplest way to find periodic transit signals is to first clean a light curve from nuisance signals before performing the search.
 This strategy is widely adopted by the community, both using physically-motivated systematics models like \cite{everest1, everest2}, or empirical filtering techniques, such as the ones described and implemented in \cite{wotan}.
@@
 In \autoref{nuance} we describe the tractable full-fledged approach of \texttt{nuance}, and its implementation in an open-source Python package.
-In \autoref{simu} we asses the performance of \texttt{nuance} by computing the recovery of planetary transits injected into simulated light curves.
+In \autoref{simu} we assess the performance of \texttt{nuance} by computing the recovery of planetary transits injected into simulated light curves.
 In section 4, we use nuance in a case study, to detect a known planet observed by the T E S S but undetected by the TESS pipeline due to its light curve characteristics.
@@
 \newpage
-\section{The issue with coorelated noise and its detrending}\label{issues}
+\section{The issue with correlated noise and its detrending}\label{issues}
 Two sources of correlated noise particularly justify the need for a detrending step before searching for transits: instrumental noises (such as telescope pointing errors) and stellar variability (induced by  pulsations or starspots).
@@
         \includegraphics[width=8.5cm]{../figures/issue1.pdf}
-        \caption{Illustration of the effect of correlated noise on a single transit singal-to-noise (SNR).
-A 1 hour transit of depth 1\% is simulated on top of white noise during a 24 hours observing window with an exposure time of 1 minute (top).
+        \caption{Illustration of the effect of correlated noise on a single transit signal-to-noise (SNR).
+A 1-hour transit of depth 1\% is simulated on top of white noise as part of a 24-hours observation with an exposure time of 1 minute (top).
 Then, in the bottom plot, correlated noise is added to the transit signal and simulated using a Gaussian Process (GP) with a Matèrn-32 kernel of length-scale 1 hour and amplitude 0.2\%.
@@
 In order to explore the parameter space for which detrending is the most problematic, we simulate light curves including a single transit and a much wider range of correlated noise characteristics defined with two parameters: $\tau_v$ the relative timescale of the variability with respect to the transit duration; and $\delta_v$, the relative amplitude of the variability against transit depth, such that:
-
 \begin{equation}
@@
 \end{equation}
-
 To follow this parametrization, we simulate the photometric variability using a GP with an SHO kernel with hyperparameters \footnote{}:
-
 \begin{equation}\label{eq:params}
@@
 \end{equation}
-
 with $\delta$ and $\tau$ the depth and duration of the simulated transit, similar in all light curves.
@@
 \autoref{fig:snr_detrend} shows the remaining SNR values of transits computed this way, after each variability signals with random $(\tau_v, \delta_v)$ have been detrended.
-% \bigskip\\
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
-        \caption{}
-        \label{fig:snr_detrend}
-    \end{centering}
-\end{figure}
+\\\\
 \autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau_v, \delta_v)$ parameter space for which detrending degrades transit SNR to the point of no detection ($SNR < 6$).
@@
 
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
+        \caption{}
+        \label{fig:snr_detrend}
+    \end{centering}
+\end{figure}
+
 \newpage
@@
 
-In this section we describe \texttt{nuance}, an algorithm to search for planetary transits in light curves featuring correlated noises, such as instrumental signals and stellar photometric variability.
+In this section we describe \texttt{nuance}, an algorithm to search for planetary transits in light curves containing correlated noises, such as instrumental signals and stellar photometric variability.
 Our approach is based on \cite{foreman2016}, developed to detect planetary transits accounting for a linear model of systematics in K2 light curves.
-\texttt{nuance} differs by modeling the covariance of the light curve with a gaussian process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.\\
+\texttt{nuance} differs by modeling the covariance of the light curve with a Gaussian Process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.\\
 
@@
 
-Simultaenously to the flux measurement, we obtained a set of $K$ measurements, aranged in the $(K\times N)$ matrix $\bm{X}$, that can be treated as explanatory variables for $\bm{f}$.
+Simultaenously to the flux measurement, we obtained a set of $M$ measurements, aranged in the $(M\times N)$ matrix $\bm{X}$, that can be treated as explanatory variables for $\bm{f}$.
 Details about the simulation of these signals are provided in \autoref{app_principle_simulations}.
 \\\\
-Ideally, we would detect such  a periodic transit signal by sampling the posterior likelihood of this data to a full-fledged model including stellar variability (more generally correlated noises), instrumental systematics (modeled with explanatory variables), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$.
-We would then reduce this posterior likelihood to $p(\bm{f}\vert P)$, its marginalized version over all parameters except the period $P$, producing a transit search periodogram.
-However, this approach has two issues: It is highly untractable, and it may lead to multimodal distributions that are hard to interpret.
+Ideally, we would detect such  a periodic transit signal by sampling the posterior likelihood of this data to a full-fledge model including stellar variability (more generally correlated noises), instrumental systematics (modeled with explanatory variables), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$.
+We would then reduce the posterior likelihood to $p(\bm{f}\vert P)$, its marginalized version over all parameters except the period $P$, producing a transit search periodogram $\mathcal{Q}(P)$.
+However, this approach has two issues: It is highly intractable, and it may lead to multimodal distributions that are hard to interpret.
 \\\\
-Given a period $P$, we instead want to compute the likelihood of a periodic transit signal at the maximum likelihood parameters $T_0$, $D$ and $\Delta$, i.e the periodogram
-
-$$\mathcal{Q}(P) = \max_{T_0, D, \Delta}p(\bm{f} \vert P, T_0 ,D, \Delta)$$
-\\
-We will do that by adopting the strategy of \cite{foreman2016}, and separate the transt search into two components: the \textit{linear search} and the \textit{periodic search}.
-During the \textit{linear search}, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
-Then, the \textit{periodic search} consists in combining these likelihoods to compute the likelihood of a periodic transit signal for a range of periods.
+Given a period $P$, we instead want to compute the likelihood of a periodic transit signal at the maximum likelihood parameters $\bar T_0$, $\bar D$ and $\bar \Delta$, i.e the periodogram
+\begin{equation}
+    \mathcal{Q}(P) = p(\bm{f} \vert P, \bar T_0 ,\bar D, \bar \Delta)
+\end{equation}
+\begin{equation}
+\text{where} \quad (\bar T_0 ,\bar D, \bar \Delta) = \argmax_{T_0, D, \Delta}p(\bm{f} \vert P, T_0 , D, \Delta)  
+\end{equation}
+We will do that by adopting the strategy of \cite{foreman2016}, and separate the transit search into two components: the \textit{linear search} and the \textit{periodic search}.
+During the \textit{linear search}, the likelihood of a single non-periodic transit is computed for a grid of epochs and durations (but not depths, as we will see).
+Then, the \textit{periodic search} consists in combining these likelihoods to compute the likelihood of the data given a periodic transit signal for a range of periods.
 These combined likelihoods yield a transit-search periodogram on which the periodic transit detection can be based.
 
-\subsection{The linear search}
+\newpage
+\subsection{The linear search}\label{linear_search}
 
-During the linear search, the goal is to compute the likelihood of a single non-periodic transit signal of epoch $T_0$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths, i.e.
-
-$$\{p(\bm{f} \vert T_i ,D_j, \Delta_k)\}_{i, j, k}$$
-
+During the linear search, the goal is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of our data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and (as we don't know yet) depths.
+\\\\
 To account for correlated noises, we model the light curve $f$ as being drawn from a Gaussian Process such that
-
 $$\bm{f} \sim \mathcal{N}(\bm{w X}, \bm{\Sigma})$$
-
 with mean $\bm{wX}$ (i.e.
-a linear model of the K explanatory variables with coefficients $\bm{w}$) and covariance $\bm{\Sigma}$.
-We include the non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$ to our model by adding it as the last column of the design matrix $\bm{X}$, so that the transit is part of the linear model and the transit depth $\Delta$ can be solved linearly.
-Under this assumption, the log-likelihood of the data to the model, including a single non-periodic transit, can be computed analyticaly
-
-$$\ln p(\bm{f} \vert \bm{X}) = -\frac{1}{2}(\bm{f}-\bm{wX})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{wX}) -  \frac{1}{2}\vert\bm{\Sigma}\vert - \frac{K}{2}\ln 2$$
-
-and the linear search is reduced to the computation of the latter log-likelihood over a grid of epochs and durations only, i.e.\footnote{In what follows, $\Delta_k$ is ommited from the notations, being linearly solved for any $(T_0, D)$}
-
-$$\{\ln p(\bm{f} \vert T_i ,D_j)\}_{i, j}$$
-
-Figure \autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}.
+a linear model of the M explanatory variables with coefficients $\bm{w}$) and covariance $\bm{\Sigma}$.
+We append a non-periodic transit signal of epoch $T$, duration $D$ and unit depth (using the simple transit model from \citealt{protopapas}) as the last column of the design matrix $\bm{X}$, so that the transit is part of the linear model and the transit depth $\Delta$ can be solved linearly.
+Under this assumption, the log-likelihood of the data given a single non-periodic transit is analytical
+\begin{equation} \label{eq:linear_search_ll}
+    \ln p(\bm{f} \vert I) = -\frac{1}{2}(\bm{f}-\bm{wX})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{wX}) -  \frac{1}{2}\vert\bm{\Sigma}\vert - \frac{N}{2}\ln 2
+\end{equation}
+and can be computed on a grid of epochs and durations only, since the transit depth is linearly solved for any $(T, D)$.
+This \textit{linear search} lead to the set of  log-likelihoods\footnote{This notation omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{ij}$) as it is linearly solved for any given value of $(T_i, F_j)$ and irrelevant in what follows.}
+$$\ln\mathcal{L} = \{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{ij})\}_{i, j}$$
+were $\Delta_{ij}$ is the depth linearly solved for a given $(D_i, T_j)$.
+\autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}.
 To compute this grid, we used the same Gaussian Process and design matrix $\bm{X}$ used to generate the data (see annexe).
@@
         \includegraphics[width=0.75\linewidth]{../figures/principle_linear_search.pdf}
-        \caption{}
+        \caption{Principle and output of the \textit{linear search}.
+First, a set of durations and depths $\{T_i, D_j\}_{i,j}$ is generated.
+For each pair of index $(i,j)$, the log likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{ij})$ is computed, using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}.
+This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\{\Delta_{ij}, \sigma_{ij}\}_{i, j}$ transit depths and errors.}
         \label{fig:linear_search}
@@
 
+To prepare the next step, the corresponding depths $\Delta_{ij}$ linearly solved for any $(T_i ,D_j)$ are conserved, as well as their associated errors $\sigma_{ij}$ obtained through the least-square solution
+
+\begin{equation} \label{eq:ls}
+    \begin{split}
+        \bm{w} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}\bm{X}^T\bm{\Sigma}^{-1}\bm{f} \\
+        \bm{\sigma} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}
+    \end{split}
+\end{equation}
+
+where $\Delta_{ij} = \bm{w}_M$ and $\sigma_{ij} = \bm{\sigma}_{MM}$
+
 \subsubsection{The periodic search}
 
-We then need to combine the likelihoods computed from the linear search to obtain\footnote{Since we search for the maximum likelihood of $p(\bm{f}\vert...)$, we ommit $\bm{w}$ from the remaining notatitions, being linearly solved and taken at its maximum likelihood value}
+We then need to combine the likelihoods computed from the linear search to obtain
+\begin{equation}\label{eq:goal}
+    p(\bm{f} \vert P, T_0 , D, \Delta)
+\end{equation}
+For a given transit duration $D$, any combination of $(P, T_0)$ leads to K truly independent transits, for which it is tempting to write
+\begin{equation}
+    p(\bm{f} \vert P, T_0 ,D, \Delta) = \prod_k^K p(\bm{f} \vert T_k, D, \Delta_k)
+\end{equation}
+where $\{T_k\}_k$ are the epochs matching $(T_0, P)$ and $\{\Delta_k\}_k$ the corresponding depths. So that
+\begin{equation}
+    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) = \sum_k^K \ln \mathcal{L}_k
+\end{equation}
+This is the joint log-likelihood of transits belonging to the same periodic signal but with varying depths  $\{\Delta_k\}_k$.
+However, individual transits from a periodic signal cannot be considered independent, and should instead share a common transit depth $\Delta$.
+We show in \autoref{combining_transits} that there is an analytical expression for the joint log-likelihood of K individual transits with depths and errors $\{\Delta_k, \sigma_k\}_k$ under a common depth $\Delta$:
+\begin{equation}\label{eq:comb}
+    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
+    \bar\Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right)
+\end{equation}
+\begin{equation}
+    \text{with} \quad  
+    \frac{1}{\sigma^2} = \sum_k^K \frac{1}{\sigma_k^2} \quad \text{and} \quad
+    \Delta = \sigma^2 \sum_k^K {\frac{\Delta_k}{\sigma_k^2}}
+\end{equation}
+While this gives a solution for \autoref{eq:goal}\footnote{Different from the one found in \cite{foreman2016}}, its practical implementation raises two issues.
+One is that individual epochs matching ($T_0$, $P$) are not necessarily available in the grid of epochs $\{T_k\}_k$.
+In \cite{foreman2016} this is solved by using the nearest neighbors in the epochs grid.
+Instead, to allow the efficient computation of \autoref{eq:comb} in a vectorized form, we interpolate the likelihood grid from $\{T_i\}_i$ to a common grid of transit phases $\{\phi_i\}_i$, leading to the periodic search log-likelihoods $\mathcal{P}(P)$ (see \autoref{fig:periodic_search}).
 
-$$\max_{T_0, D}p(\bm{f} \vert P, T_0 ,D)$$
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.8\linewidth]{../figures/principle_periodic.pdf}
+        \caption{On the example dataset, this figure shows how the periodic search work for any given period $P$.
+First, the log-likelihoods $\ln \mathcal{L}$ are phase-folded and interpolated onto a common grid of phases (the white lines mark the edges of each fold).
+Then \autoref{eq:comb} is computed on the full grid $\{\phi_i, D_i\}_i$ allowing the use of efficient matrix optimization.}
+        \label{fig:periodic_search}
+    \end{centering}
+\end{figure}
 
-If each transits were to be considered truly independant, we would have
+The second and final issue is more fundamental, and lies in the fact that for each period we combine a different number of transits.
+Indeed, each likelihood $p(\bm{f} \vert T, D)$ estimated during the linear search is computed using N measurements.
+Hence, combining $K$ transits likelihoods in the periodic search, through $\Delta_k$ and $\sigma_k$ (\autoref{eq:comb}), artificially lead to a likelihood involving $N\times K$ measurements.
+While this problem could be set aside, it involves a normalization issue when trying to compare the joint log-likelihoods $\mathcal{P}(P)$ from one period to another (since for each period the number of observed transits is changing).
+This motivates a final step to produce the transit search periodogram.
 
-$$p(\bm{f} \vert P, T_0 ,D) \sim \prod_k p(\bm{f} \vert T_k ,D_k)$$
+\subsubsection{The transit periodogram}
 
-where $T_k$ are the epochs matching ($T_0$, $P$) and $D_k$ the durations of the corresponding transits. However, doing so leads to two issues:
-\begin{enumerate}
+To overcome the final issue presented in the last subsection, we will base our transit search periodogram not on the periodic transit likelihood computed for each period $P$, but on the real SNR of the periodic signal assesed
 
-    \item Transits are not truly independant and, instead of having their own depths $\Delta_k$, should share a common transit depth $\Delta$.
-We show in Annexe that there is an analytical expression for the likelihood of a periodic transit signal of depth $\Delta$ whose individual transit depths $\Delta_k$ with error $\sigma_k$ have been measured independantly:
-
-    $$p(y \vert P, T_0 ,D) =  \prod_k \frac{\sigma_k \bar\sigma \, e^{- \frac{\left(\Delta_k - \Delta\right)^{2}}{2 \sigma_k^{2} + \bar\sigma^{2}}}}{\sigma_k^{2} + \bar\sigma^{2}}$$
-
-    $$\text{with} \quad  
-    \frac{1}{\bar\sigma^2} = \sum_{k}\frac{1}{\sigma_k^2} \quad \text{and} \quad
-    \Delta = \bar\sigma^2 \sum_{k}{\frac{\Delta_k}{\sigma_k^2}}$$
-
-    \item Individual epochs $T_k$ matching ($T_0$, $P$) are not necessarly available in the grid of epochs $\{T_i\}$.
-In \cite{foreman2016} this is solved by using the nearest neighbors in the epochs grid.
-In \texttt{nuance}, we rather interpolate the likelihood grid from $T_i$ to $T_k$ (see \autoref{fig:periodic_search}).
-
-\end{enumerate}
+normalization
+For a given period $P$, we can then compute $\max_{(T_0, D)}\{p(\bm{f} \vert P, T_0 ,D)\}$.
+But instead of relying on this maximum likelihood to build the final periodorgram, we compute the SNR of the transit with period $P$ and $(T_0, D)$ maximizing $p(\bm{f} \vert T_0 ,D, P)$, yielding the transit SNR periodogram shown in \autoref{fig:periodic_search}.
+We leave a discussion of this choice for Annexe .
 
@@
 
-For a given period $P$, we can then compute $\max_{(T_0, D)}\{p(\bm{f} \vert P, T_0 ,D)\}$.
-But instead of relying on this maximum likelihood to build the final periodorgram, we compute the SNR of the transit with period $P$ and $(T_0, D)$ maximizing $p(\bm{f} \vert T_0 ,D, P)$, yielding the transit SNR periodogram shown in \autoref{fig:periodic_search}.
-We leave a discussion of this choice for Annexe .
-\\\\
-Once a transit recoverd, we mask the corresponding epochs in the log likelihood linear grid
+Once a transit recovered, we mask the corresponding epochs in the log likelihood linear grid
 
@@
 \section{Light curve simulations}\label{app_principle_simulations}
+\section{Combining transits}\label{combining_transits}
+
+\newcommand{\sumTk}{i\neq k}
+From the \textit{linear search}, we retain and index by $k$ the parameters of the $K$ individual transits whose epochs $\{T_k\}_k$ are compatible with a periodic signal of period $P$ and epoch $T_0$.
+From the likelihoods of these transits (computed in \autoref{linear_search}), we want an expression for
+\begin{equation}\label{eq:goal2}
+    p(\bm{f} \vert P, T_0 ,D, \Delta) = \prod_{k\in\mathbb{T}} p(\bm{f} \vert T_k, D, \Delta)
+\end{equation}
+
+i.e., given a depth $D$, the likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$ and a common depth $\Delta$.
+Since only $\{p(\bm{f} \vert T_k, D, \Delta_k)\}_{k}$ is known (i.e.
+transits with different depths), we decompose
+\begin{equation}\label{eq:non_part_of_per}
+    p(\bm{f} \vert T_k, D, \Delta) = \int p(\bm{f} \vert T_k, D, \tilde\Delta)p(\tilde\Delta | \Delta)\, d\tilde\Delta
+\end{equation}
+where $p(\bm{f} \vert T_k, D, \tilde\Delta)$ is the probability of the $k$-th transit to have a depth $\tilde\Delta$ and $p(\tilde\Delta | \Delta)$ the probability to observe the depth $\tilde\Delta$ knowing the existence of a common depth $\Delta$.
+In other words, \autoref{eq:non_part_of_per} involves the likelihood of the non-periodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
+\\\\
+Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$, and with an amplitude $\mathcal{L}_k$, leading to the likelihood:
+$$p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k G_k(\tilde\Delta)$$
+where $G_k$ is the probability density function of $\mathcal{N}(\Delta_k, \sigma_k^2)$. 
+\\\\
+As for the common transit depth $\Delta$, it can be estimated through the joint probability of all other transit depths than $\Delta_k$, such that
+$$\Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2) \quad\text{i.e.}\quad \Delta \sim \mathcal{N}(\Delta, \sigma^2) $$
+a normal distribution defined by
+\begin{equation}\label{eq:params}
+\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad
+\Delta =\sigma^2 \sum_{\sumTk}^K {\frac{\Delta_i}{\sigma_i^2}}
+\end{equation}
+$$p(\tilde\Delta | \Delta) =  \left(\prod_{\sumTk}^K \mathcal{L}_i \right) G(\tilde\Delta)$$
+where $G$ is the probability density function of $\mathcal{N}(\Delta, \sigma^2)$. We can now rewrite \autoref{eq:non_part_of_per} as
+\begin{equation}\label{eq:ice_int}
+    p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i \right)  \int G_k(\tilde\Delta)\, G(\tilde\Delta)\, d\tilde\Delta
+\end{equation}
+The product of two normal distributions is also a normal distribution, so that the integral in \autoref{eq:ice_int} can be obtained analytically. We have
+$$p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i  \right) \frac{1}{\sqrt{2\pi \left(\sigma^{2} + \sigma_{k}^{2}\right)}} e^{- \frac{1}{2}\frac{\left(\Delta -  \Delta_{k}\right)^{2}}{\sigma^{2} + \sigma_{k}^{2}}}$$
+And \autoref{eq:goal} gives 
+\begin{equation}\label{eq:result}
+    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
+    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right)
+\end{equation}
+the log-likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and common depth $\Delta$.
+In order to reduce the number of times \autoref{eq:params} is computed, we adopt the biased estimates
+$$\frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}}$$
+so that $\Delta$ and $\sigma$ are independent of $k$ in the last sum of \autoref{eq:result}.
+
 \bibliography{bib}

### 2023-01-19 simple/latex/ms.tex
finish first draft on nuance theory

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{hyperref}
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\newcommand{\set}[1]{\{\,#1\,\}}
 
@@
 
-\newpage
 \section{\texttt{nuance}}\label{nuance}
@@
 In this section we describe \texttt{nuance}, an algorithm to search for planetary transits in light curves containing correlated noises, such as instrumental signals and stellar photometric variability.
-Our approach is based on \cite{foreman2016}, developed to detect planetary transits accounting for a linear model of systematics in K2 light curves.
-\texttt{nuance} differs by modeling the covariance of the light curve with a Gaussian Process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.\\
-
-\subsection{Preamble}
-
-The flux $f$ of a star measured at time $t$ is sampled and aranged in the $(1\times N)$ column-vector $\bm{f}$.
+\\\\
+Let a flux $f$ of a star measured at time $t$ be sampled and aranged in the $(1\times N)$ column-vector $\bm{f}$.
 This flux contains instrumental signals, stellar variability and a periodic transit signal that we wish to uncover.
@@
 \\\\
-Given a period $P$, we instead want to compute the likelihood of a periodic transit signal at the maximum likelihood parameters $\bar T_0$, $\bar D$ and $\bar \Delta$, i.e the periodogram
-\begin{equation}
-    \mathcal{Q}(P) = p(\bm{f} \vert P, \bar T_0 ,\bar D, \bar \Delta)
-\end{equation}
-\begin{equation}
-\text{where} \quad (\bar T_0 ,\bar D, \bar \Delta) = \argmax_{T_0, D, \Delta}p(\bm{f} \vert P, T_0 , D, \Delta)  
+Given a period $P$, we instead want to compute the likelihood of a periodic transit signal at the maximum likelihood parameters $\hat T_0$, $\hat D$ and $\hat \Delta$, i.e the periodogram
+\begin{equation}\label{eq:periodogram}
+        \begin{gathered}
+        \mathcal{Q}(P) = p(\bm{f} \vert P, \hat T_0 ,\hat D, \hat \Delta) \\
+        \text{where} \quad (\hat T_0 ,\hat D, \hat \Delta) = \argmax_{T_0, D, \Delta}p(\bm{f} \vert P, T_0 , D, \Delta)
+    \end{gathered}
 \end{equation}
@@
 These combined likelihoods yield a transit-search periodogram on which the periodic transit detection can be based.
+\texttt{nuance} differs by modeling the covariance of the light curve with a Gaussian Process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
 
-\newpage
-\subsection{The linear search}\label{linear_search}
+\subsection{The \textit{linear search}}\label{linear_search}
 
-During the linear search, the goal is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of our data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and (as we don't know yet) depths.
+During the \textit{linear search}, the goal is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of our data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths.
 \\\\
@@
 a linear model of the M explanatory variables with coefficients $\bm{w}$) and covariance $\bm{\Sigma}$.
-We append a non-periodic transit signal of epoch $T$, duration $D$ and unit depth (using the simple transit model from \citealt{protopapas}) as the last column of the design matrix $\bm{X}$, so that the transit is part of the linear model and the transit depth $\Delta$ can be solved linearly.
+To account for the presence of a single non-periodic transit of epoch $T$ and duration $D$, we compute and append its signal as the last column of the design matrix $\bm{X}$, using the simple transit model from \citealt{protopapas} with a unitary depth.
+This way, the transit signal is part of the linear model and its depth $\Delta$ can be solved linearly.
 Under this assumption, the log-likelihood of the data given a single non-periodic transit is analytical
@@
 \end{equation}
-and can be computed on a grid of epochs and durations only, since the transit depth is linearly solved for any $(T, D)$.
+and can be computed on a grid of epochs and durations, the transit depth being linearly solved for any $(T, D)$.
 This \textit{linear search} lead to the set of  log-likelihoods\footnote{This notation omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{ij}$) as it is linearly solved for any given value of $(T_i, F_j)$ and irrelevant in what follows.}
@@
 were $\Delta_{ij}$ is the depth linearly solved for a given $(D_i, T_j)$.
-\autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}.
-To compute this grid, we used the same Gaussian Process and design matrix $\bm{X}$ used to generate the data (see annexe).
+\autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}, using the same Gaussian Process and design matrix $\bm{X}$ used to simulate the data (see annexe).
 
@@
         \caption{Principle and output of the \textit{linear search}.
-First, a set of durations and depths $\{T_i, D_j\}_{i,j}$ is generated.
+First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated.
 For each pair of index $(i,j)$, the log likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{ij})$ is computed, using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}.
-This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\{\Delta_{ij}, \sigma_{ij}\}_{i, j}$ transit depths and errors.}
+This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{ij}, \sigma_{ij}}_{i, j}$ transit depths and errors.}
         \label{fig:linear_search}
@@
 To prepare the next step, the corresponding depths $\Delta_{ij}$ linearly solved for any $(T_i ,D_j)$ are conserved, as well as their associated errors $\sigma_{ij}$ obtained through the least-square solution
-
-\begin{equation} \label{eq:ls}
-    \begin{split}
+\begin{equation}\label{eq:ls}
+    \begin{gathered}
         \bm{w} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}\bm{X}^T\bm{\Sigma}^{-1}\bm{f} \\
         \bm{\sigma} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}
-    \end{split}
+    \end{gathered}
 \end{equation}    
-
 where $\Delta_{ij} = \bm{w}_M$ and $\sigma_{ij} = \bm{\sigma}_{MM}$
 
-\subsubsection{The periodic search}
+\subsection{The periodic search}
 
-We then need to combine the likelihoods computed from the linear search to obtain
+We then need to combine the likelihoods computed from the \textit{linear search} to obtain
 \begin{equation}\label{eq:goal}
@@
 For a given transit duration $D$, any combination of $(P, T_0)$ leads to K truly independent transits, for which it is tempting to write
-\begin{equation}
+\begin{equation}\label{eq:attempt}
     p(\bm{f} \vert P, T_0 ,D, \Delta) = \prod_k^K p(\bm{f} \vert T_k, D, \Delta_k)
 \end{equation}
-where $\{T_k\}_k$ are the epochs matching $(T_0, P)$ and $\{\Delta_k\}_k$ the corresponding depths. So that
-\begin{equation}
+where $\set{T_k}_k$ are the epochs matching $(T_0, P)$ and $\set{\Delta_k}_k$ the corresponding depths. So that
+$$
     \ln p(\bm{f} \vert P, T_0 ,D, \Delta) = \sum_k^K \ln \mathcal{L}_k
+$$
+This is the joint log-likelihood of transits belonging to the same periodic signal but with varying depths  $\set{\Delta_k}_k$.
+However, individual transits from a periodic signal cannot be considered independent, and should instead share a common transit depth $\Delta$.
+We show in \autoref{combining_transits} that there is an analytical expression for the joint log-likelihood of K individual transits with depths and errors $\set{\Delta_k, \sigma_k}_k$ assuming a common depth $\Delta$:
+\begin{equation}\label{eq:comb}
+    \begin{gathered}
+        \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
+        \bar\Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right) \\
+        \text{with} \quad  \frac{1}{\sigma^2} = \sum_k^K \frac{1}{\sigma_k^2} \quad \text{and} \quad
+        \Delta = \sigma^2 \sum_k^K {\frac{\Delta_k}{\sigma_k^2}}
+    \end{gathered}
 \end{equation}
-This is the joint log-likelihood of transits belonging to the same periodic signal but with varying depths  $\{\Delta_k\}_k$.
-However, individual transits from a periodic signal cannot be considered independent, and should instead share a common transit depth $\Delta$.
-We show in \autoref{combining_transits} that there is an analytical expression for the joint log-likelihood of K individual transits with depths and errors $\{\Delta_k, \sigma_k\}_k$ under a common depth $\Delta$:
-\begin{equation}\label{eq:comb}
-    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
-    \bar\Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right)
-\end{equation}
-\begin{equation}
-    \text{with} \quad  
-    \frac{1}{\sigma^2} = \sum_k^K \frac{1}{\sigma_k^2} \quad \text{and} \quad
-    \Delta = \sigma^2 \sum_k^K {\frac{\Delta_k}{\sigma_k^2}}
-\end{equation}
-While this gives a solution for \autoref{eq:goal}\footnote{Different from the one found in \cite{foreman2016}}, its practical implementation raises two issues.
-One is that individual epochs matching ($T_0$, $P$) are not necessarily available in the grid of epochs $\{T_k\}_k$.
+While this gives a solution for \autoref{eq:goal}\footnote{Different from the one found in \cite{foreman2016}}, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$.
 In \cite{foreman2016} this is solved by using the nearest neighbors in the epochs grid.
-Instead, to allow the efficient computation of \autoref{eq:comb} in a vectorized form, we interpolate the likelihood grid from $\{T_i\}_i$ to a common grid of transit phases $\{\phi_i\}_i$, leading to the periodic search log-likelihoods $\mathcal{P}(P)$ (see \autoref{fig:periodic_search}).
+Instead, to allow the efficient matrix computation of \autoref{eq:comb}, we interpolate the likelihood grid from $\set{T_i}_i$ to a common grid of transit phases $\set{\phi_i}_i$, leading to the \textit{periodic search} log-likelihood
+$$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi, D)}_{i,j}$$
+shown for few periods in \autoref{fig:periodic_search}.b, a notation where $\Delta_{ij}$ is omitted since being interpolated from the \textit{linear search} using $\phi_i$ and $D_j$, and $T_0 = 0$.
 
@@
     \begin{centering}
-        \includegraphics[width=0.8\linewidth]{../figures/principle_periodic.pdf}
-        \caption{On the example dataset, this figure shows how the periodic search work for any given period $P$.
-First, the log-likelihoods $\ln \mathcal{L}$ are phase-folded and interpolated onto a common grid of phases (the white lines mark the edges of each fold).
-Then \autoref{eq:comb} is computed on the full grid $\{\phi_i, D_i\}_i$ allowing the use of efficient matrix optimization.}
+        \gridline{\fig{../figures/principle_periodic_0.pdf}{0.7\linewidth}{\vspace{-0.5cm}(a) $\ln \mathcal{L}$}}
+        \vspace{-0.6cm}
+        \gridline{\fig{../figures/principle_periodic_1.pdf}{0.6\linewidth}{\vspace{-0.5cm}(b) $\mathcal{P}(P)$}}
+        \caption{On the example dataset, this figure shows how the \textit{periodic search} works at different periods $P$ (including the true period $P = 0.7$ days).
+Given $P$ and $T_0=0$, the log-likelihood $\ln\mathcal{L}$ shown in (a) is phase-folded and interpolated onto a common grid of phases shown in (b).
+As an example, the white lines in (a) mark the edges of each fold for a period of $P=0.5$ days, and the white crosses the epochs $\set{T_k}_{k\in\mathbb{K}}$ matching a particular phase in the grid (reported in (a) and (b)) on which the corresponding $\set{\ln \mathcal{L}_k}_{k\in\mathbb{K}}$ are interpolated and combined using \autoref{eq:comb}.
+To allow the use of efficient matrix computations, this is done for every duration $\set{D_i}_i$ so that $\mathcal{P}$ is computed on the full grid $\set{D_i, \phi_i}_{i, j}$ at once.
+We understand from plots in (a) that a different choice of epoch $T_0$ may only shift the results in phase but do not affect the values of $\mathcal{P}$.
+For this reason, computing $\mathcal{P}$ for $T_0=0$ is sufficient.
+We also notice how the maximum value of $\mathcal{P}(P=0.35)$ is lower than for $P=0.7$ days, a result of combining the log-likelihoods using \autoref{eq:comb} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
         \label{fig:periodic_search}
@@
 
-The second and final issue is more fundamental, and lies in the fact that for each period we combine a different number of transits.
-Indeed, each likelihood $p(\bm{f} \vert T, D)$ estimated during the linear search is computed using N measurements.
-Hence, combining $K$ transits likelihoods in the periodic search, through $\Delta_k$ and $\sigma_k$ (\autoref{eq:comb}), artificially lead to a likelihood involving $N\times K$ measurements.
-While this problem could be set aside, it involves a normalization issue when trying to compare the joint log-likelihoods $\mathcal{P}(P)$ from one period to another (since for each period the number of observed transits is changing).
-This motivates a final step to produce the transit search periodogram.
+\subsection{The transit search periodogram}
 
-\subsubsection{The transit periodogram}
-
-To overcome the final issue presented in the last subsection, we will base our transit search periodogram not on the periodic transit likelihood computed for each period $P$, but on the real SNR of the periodic signal assesed
-
-normalization
-For a given period $P$, we can then compute $\max_{(T_0, D)}\{p(\bm{f} \vert P, T_0 ,D)\}$.
-But instead of relying on this maximum likelihood to build the final periodorgram, we compute the SNR of the transit with period $P$ and $(T_0, D)$ maximizing $p(\bm{f} \vert T_0 ,D, P)$, yielding the transit SNR periodogram shown in \autoref{fig:periodic_search}.
-We leave a discussion of this choice for Annexe .
+Using \autoref{eq:comb}, we can now compute $\ln\mathcal{P}$ for a range of periods (see \autoref{fig:periodic_search}) and build a transit search periodogram using \autoref{eq:periodogram}.
+But a final issue emerges, one that is fundamentally linked to our strategy.
+Each likelihood $p(\bm{f} \vert T, D, \Delta)$ estimated during the \textit{linear search} is computed using N measurements.
+Hence, combining transits in the \textit{periodic search}, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (\autoref{eq:comb}), artificially leads to a likelihood involving $N\times K$ measurements.
+This lead to a normalization issue when trying to compare the joint log-likelihoods $\mathcal{P}(P)$ from one period to another (like the ones in \autoref{fig:periodic_search} that have been normalized for visualization purpose), as the number of observed transits differs from one period to another.
+This motivates a final step to produce the transit search periodogram $Q$.
+\\\\
+For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, we retrieve the maximum likekihood parameters
+\begin{equation}\label{eq:phi0}
+    (\phi_0 ,D) = \argmax_{\phi_i, D_j} \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i, j}
+\end{equation}
+and define $\mathcal{Q}(P)$ as the SNR of the transit of period $P$, epoch $T_0 = \phi_0 P$ duration $D$ and depth $\Delta$, i.e.
+$$
+    \mathcal{Q}(P) = \frac{\Delta}{\sigma}
+$$
+where $\Delta$ and $\sigma$ are obtained using \autoref{eq:comb} with the last column of $X$ containing a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $1$.
+This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.
 
@@
     \begin{centering}
-        \includegraphics[width=0.8
-        \linewidth]{../figures/principle_tiles_combined.pdf}
+        \includegraphics[width=0.62\linewidth]{../figures/principle_Q.pdf}
         \caption{}
-        \label{fig:periodic_search}
+        \label{fig:periodogram}
     \end{centering}
@@
 
-Once a transit recovered, we mask the corresponding epochs in the log likelihood linear grid
+The periodic transit of period $P$ with the maximum SNR, i.e.
+maximizing $\mathcal{Q}$, is adopted as the best candidate, adapting our confidence through the minimum SNR for which we consider a signal to be real.
+The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (\autoref{eq:comb}).
 
+\newpage
 \subsubsection{An open-source python package}
@@
 
+\newpage
 \appendix

### 2023-01-20 simple/latex/ms.tex
simulation appendix started

@@
 \usepackage{bm}
+\usepackage{xcolor}
 
@@
 
-Throughout this paper, we simulate stellar variability (and later model it) thanks to Gaussian Processes (GP), employing the physically-motivated stochastically-driven damped simple harmonic oscillator kernel (SHO) presented in \citealt{celerite1} (see Annexe ...).
-We generate stellar variability thanks to the \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}} Python package, providing an implententation of the scalable GP method from \citealt{celerite2} powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
+Throughout this paper, we simulate stellar variability (and later model it) thanks to Gaussian Processes (GP), employing the physically-motivated stochastically-driven damped simple harmonic oscillator kernel (SHO) presented in \citealt{celerite1}.
+Extensive details about the following simulations can be found in \autoref{app_principle_simulations}.
 
@@
 and can be computed on a grid of epochs and durations, the transit depth being linearly solved for any $(T, D)$.
-This \textit{linear search} lead to the set of  log-likelihoods\footnote{This notation omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{ij}$) as it is linearly solved for any given value of $(T_i, F_j)$ and irrelevant in what follows.}
-$$\ln\mathcal{L} = \{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{ij})\}_{i, j}$$
-were $\Delta_{ij}$ is the depth linearly solved for a given $(D_i, T_j)$.
+This \textit{linear search} lead to the set of  log-likelihoods\footnote{This notation omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{i,j}$) as it is linearly solved for any given value of $(T_i, F_j)$ and irrelevant in what follows.}
+$$\ln\mathcal{L} = \{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})\}_{i, j}$$
+were $\Delta_{i,j}$ is the depth linearly solved for a given $(D_i, T_j)$.
 \autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}, using the same Gaussian Process and design matrix $\bm{X}$ used to simulate the data (see annexe).
@@
 First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated.
-For each pair of index $(i,j)$, the log likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{ij})$ is computed, using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}.
-This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{ij}, \sigma_{ij}}_{i, j}$ transit depths and errors.}
+For each pair of index $(i,j)$, the log likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed, using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}.
+This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors.}
         \label{fig:linear_search}
@@
 
-To prepare the next step, the corresponding depths $\Delta_{ij}$ linearly solved for any $(T_i ,D_j)$ are conserved, as well as their associated errors $\sigma_{ij}$ obtained through the least-square solution
+To prepare the next step, the corresponding depths $\Delta_{i,j}$ linearly solved for any $(T_i ,D_j)$ are conserved, as well as their associated errors $\sigma_{i,j}$ obtained through the least-square solution
 \begin{equation}\label{eq:ls}
@@
 \end{equation}    
-where $\Delta_{ij} = \bm{w}_M$ and $\sigma_{ij} = \bm{\sigma}_{MM}$
+where $\Delta_{i,j} = \bm{w}_M$ and $\sigma_{i,j} = \bm{\sigma}_{MM}$
 
@@
 $$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi, D)}_{i,j}$$
-shown for few periods in \autoref{fig:periodic_search}.b, a notation where $\Delta_{ij}$ is omitted since being interpolated from the \textit{linear search} using $\phi_i$ and $D_j$, and $T_0 = 0$.
+shown for few periods in \autoref{fig:periodic_search}.b, a notation where $\Delta_{i,j}$ is omitted since being interpolated from the \textit{linear search} using $\phi_i$ and $D_j$, and $T_0 = 0$.
 
@@
     \begin{centering}
-        \gridline{\fig{../figures/principle_periodic_0.pdf}{0.7\linewidth}{\vspace{-0.5cm}(a) $\ln \mathcal{L}$}}
+        \gridline{\fig{../figures/principle_periodic_0.pdf}{0.72\linewidth}{\vspace{-0.5cm}(a) $\ln \mathcal{L}$}}
         \vspace{-0.6cm}
@@
     \begin{centering}
-        \includegraphics[width=0.62\linewidth]{../figures/principle_Q.pdf}
+        \includegraphics[width=0.65\linewidth]{../figures/principle_Q.pdf}
         \caption{}
@@
 \section{Light curve simulations}\label{app_principle_simulations}
+
+This paper relies on simulated light curves featuring transit signals, instrumental signals and correlated noises such as stellar variability.
+Let such a flux $f$ of a star simulated at time $t$ be sampled and aranged in the $(1\times N)$ column-vector $\bm{f}$ associated to the column-vector of times $\bm{t}$.
+An example of such signal is simulated and shown in \autoref{fig:principle_dataset}.
+
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.8\linewidth]{../figures/principle_dataset_decomposed.pdf}
+        \caption{}
+        \label{fig:principle_dataset}
+    \end{centering}
+\end{figure}
+In what follows, we describe the models used to simulate the components of this signal, i.e.
+the periodic transit $\mathcal{T}$, the instrumental systematics $\mathcal{S}$ and the correlated noise $\mathcal{V}$, such that
+$$
+    f(t) = \textcolor{gray}{\mathcal{T}}(t) %+ \textcolor{Gray}{\mathcal{S}}(t) + \textcolor{Gray}{\mathcal{V}}(t)
+$$
+computed for $N=2880$ times $t$ corresponding to an observation of 4 days with an exposure time of 2 minutes. 
+\subsection{Transit signal}
+We simulate the periodic transit signal $\mathcal{T}$ using the simple model described in \cite{protopapas} where a transit of period $P$, epoch $T_0$, duration $D$ and unitary depth observed at time $t$ is given by
+\begin{equation}
+    \begin{gathered}
+        \mathcal{T}_c(t, P, T_0, D) = \frac{1}{2}\tanh\left(c\left(\theta - \frac{1}{2}\right)\right) - \frac{1}{2}\tanh\left(c\left(\theta + \frac{1}{2}\right)\right) \\
+        \text{with}\quad\theta = \frac{P}{\pi  D}\sin\left(\frac{\pi(t-T_0)}{P}\right)
+    \end{gathered}
+\end{equation}
+where the dimensionless parameter c controls the roundness of the transit depth ($c\gg1$ corresponding to a box-shaped transit as shown in \autoref{fig:protopapas}).
+This analytical model is fully empirical but easily differentiable.
+From this expression, we get the signal of the single non-periodic transit of unitary depth
+\begin{equation}\label{eq:protopapas_single}
+    \tau_c(t, T_0, D) = \lim_{P\to \infty} \mathcal{T}(t, P, T_0, D) = -\frac{1}{2}\tan\left(\frac{c}{2}(1-T_0+t)\right) -\frac{1}{2}\tan\left(\frac{c}{2}(1+T_0-t)\right)
+\end{equation}
+with epoch $T_0$ and duration $D$.
+
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.9\linewidth]{../figures/protopapas.pdf}
+        \caption{Simulations of the single non-periodic transit signal $\tau_c(t, T_0, D)$ (\autoref{eq:protopapas_single}) for different values of $c$}
+        \label{fig:protopapas}
+    \end{centering}
+\end{figure}
+In our study, we simulate all transit signals with $c=12$.
+The periodic transit signal $\mathcal{T}$ seen in \autoref{fig:simu} corresponds to $\mathcal{T} = \mathcal{T}_{c=12}(t, P=0.7, T_0=0.2, D=0.05)$, all parameters in unit of \textit{days}.
+
+\subsection{Instrumental signals}
+We simulate instrumental signals as a linear model of $M$ explanatory variables arranged in the $(M\times N)$ design matrix $\bm{X}$. Hence,
+$$
+    \mathcal{S} \sim \bm{\mathcal{S}} = \bm{w}\bm{X}
+$$
+where the vector $\bm{w}$ are linear coefficients of the model.
+The simulated flux shown in \autoref{fig:simu} contains a linear model where the $M=4$ columns of the design matrix $\bm{X}$ are given by $\bm{X}_{i} = \bm{t}^i$ (i.e.
+$\bm{X}$ is the Vandermonde matrix order $3$ of time $t$) and $\bm{w} = [1.0\quad0.0005\quad\min0.0002\quad-0.0005].$
+
+\subsection{Correlated noise}
+We generate stellar variability thanks to the \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}} Python package, providing an implententation of the scalable GP method from \citealt{celerite2} powered by
+\texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
 \section{Combining transits}\label{combining_transits}

### 2023-01-23 simple/latex/ms.tex
simulated principle dataset description

@@
 \usepackage{bm}
-\usepackage{xcolor}
+\usepackage[dvipsnames]{xcolor}
 
@@
     \begin{centering}
-        \includegraphics[width=0.75\linewidth]{../figures/principle_linear_search.pdf}
+        \includegraphics[width=0.9\linewidth]{../figures/principle_linear_search.pdf}
         \caption{Principle and output of the \textit{linear search}.
@@
     \begin{centering}
-        \gridline{\fig{../figures/principle_periodic_0.pdf}{0.72\linewidth}{\vspace{-0.5cm}(a) $\ln \mathcal{L}$}}
+        \gridline{\fig{../figures/principle_periodic_0.pdf}{0.9\linewidth}{\vspace{-0.5cm}(a) $\ln \mathcal{L}$}}
         \vspace{-0.6cm}
-        \gridline{\fig{../figures/principle_periodic_1.pdf}{0.6\linewidth}{\vspace{-0.5cm}(b) $\mathcal{P}(P)$}}
+        \gridline{\fig{../figures/principle_periodic_1.pdf}{0.7\linewidth}{\vspace{-0.5cm}(b) $\mathcal{P}(P)$}}
         \caption{On the example dataset, this figure shows how the \textit{periodic search} works at different periods $P$ (including the true period $P = 0.7$ days).
@@
     \begin{centering}
-        \includegraphics[width=0.65\linewidth]{../figures/principle_Q.pdf}
+        \includegraphics[width=0.8\linewidth]{../figures/principle_Q.pdf}
         \caption{}
@@
 
-This paper relies on simulated light curves featuring transit signals, instrumental signals and correlated noises such as stellar variability.
-Let such a flux $f$ of a star simulated at time $t$ be sampled and aranged in the $(1\times N)$ column-vector $\bm{f}$ associated to the column-vector of times $\bm{t}$.
-An example of such signal is simulated and shown in \autoref{fig:principle_dataset}.
+\definecolor{blue_S}{rgb}{0.12156862745098039, 0.4666666666666667, 0.7058823529411765}
+\definecolor{red_V}{rgb}{0.8392156862745098, 0.15294117647058825, 0.1568627450980392}
+
+% This paper relies on simulated light curves featuring periodic transit signals, instrumental signals and correlated noises such as stellar variability.
+To do so, we draw $\bm{f}$ from a Gaussian Process of mean $\bm{\mu}$ with $\mu_i = {\color{Gray}{\mathcal{T}}}(t_i) + {\color{blue_S}{\mathcal{S}}}(t_i)$, where and covariance matrix $\bm{C}$ with $C_{i, j} = {\color{red_V}\nu}(t_i, t_j)$
+
+Let $f$ be the simulated flux of a star sampled and aranged in the $(1\times N)$ column-vector $\bm{f}$ associated to the column-vector of times $\bm{t}$. We construct $\bm{f}$ such that
+$$
+    \bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{C})
+$$
+The mean $\bm{\mu}$ is built such that $\mu_i = \mathcal{T}(t_i) + \mathcal{S}(t_i)$ where $\mathcal{T}$ is a periodic transit signal and $\mathcal{S}$ an instrumental signal.
+The covariance matrix $\bm{C}$ is built such that $C_{i, j} = \nu(t_i, t_j)$ where $\nu$ is a covariance function accounting for correlated noise in the form of stellar variabilty with added white noise.
+An example of such signal is simulated and shown in \autoref{fig:app_principle_dataset}.
 
@@
         \includegraphics[width=0.8\linewidth]{../figures/principle_dataset_decomposed.pdf}
-        \caption{}
-        \label{fig:principle_dataset}
+        \caption{Example dataset sampled at $N=2880$ times corresponding to an observation of 4 days with an exposure time of 2 minutes.
+The mean of this signal consists in a periodic transit signal plus instrumental signals.
+Stellar variability and white noise are simulated by modeling the covariance matrix of the signal with Gaussian Processes .}
+        \label{fig:app_principle_dataset}
     \end{centering}
 \end{figure}
-In what follows, we describe the models used to simulate the components of this signal, i.e.
-the periodic transit $\mathcal{T}$, the instrumental systematics $\mathcal{S}$ and the correlated noise $\mathcal{V}$, such that
-$$
-    f(t) = \textcolor{gray}{\mathcal{T}}(t) %+ \textcolor{Gray}{\mathcal{S}}(t) + \textcolor{Gray}{\mathcal{V}}(t)
-$$
-computed for $N=2880$ times $t$ corresponding to an observation of 4 days with an exposure time of 2 minutes. 
-\subsection{Transit signal}
+In what follows, we describe the models used to simulate the components of the signal $f$, i.e.
+the periodic transit $\mathcal{T}$, the instrumental systematics $\mathcal{S}$ and the covariance function $\nu$ describing stellar variability and white noise.
+
+\subsection{Transit signal $\mathcal{T}$}
 We simulate the periodic transit signal $\mathcal{T}$ using the simple model described in \cite{protopapas} where a transit of period $P$, epoch $T_0$, duration $D$ and unitary depth observed at time $t$ is given by
@@
 In our study, we simulate all transit signals with $c=12$.
-The periodic transit signal $\mathcal{T}$ seen in \autoref{fig:simu} corresponds to $\mathcal{T} = \mathcal{T}_{c=12}(t, P=0.7, T_0=0.2, D=0.05)$, all parameters in unit of \textit{days}.
+The periodic transit signal $\mathcal{T}$ seen in \autoref{fig:app_principle_dataset} corresponds to $\mathcal{T} = \mathcal{T}_{c=12}(\bm{t}, P=0.7, T_0=0.2, D=0.05)$, all parameters in unit of \textit{days}.
 
-\subsection{Instrumental signals}
+\subsection{Instrumental signals $\mathcal{S}$}
 We simulate instrumental signals as a linear model of $M$ explanatory variables arranged in the $(M\times N)$ design matrix $\bm{X}$. Hence,
 $$
-    \mathcal{S} \sim \bm{\mathcal{S}} = \bm{w}\bm{X}
+    \mathcal{S} = \bm{w}\bm{X}
 $$
 where the vector $\bm{w}$ are linear coefficients of the model.
-The simulated flux shown in \autoref{fig:simu} contains a linear model where the $M=4$ columns of the design matrix $\bm{X}$ are given by $\bm{X}_{i} = \bm{t}^i$ (i.e.
-$\bm{X}$ is the Vandermonde matrix order $3$ of time $t$) and $\bm{w} = [1.0\quad0.0005\quad\min0.0002\quad-0.0005].$
+The simulated flux shown in \autoref{fig:app_principle_dataset} contains a linear model where the $M=4$ columns of the design matrix $\bm{X}$ are given by $\bm{X}_{i} = \bm{t}^i$ (i.e.
+$\bm{X}$ is the Vandermonde matrix order $3$ of time $t$) and $\bm{w} = [1.0\quad0.0005\quad\text{-}0.0002\quad\text{-}0.0005].$
 
-\subsection{Correlated noise}
-We generate stellar variability thanks to the \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}} Python package, providing an implententation of the scalable GP method from \citealt{celerite2} powered by
-\texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
+\subsection{Stellar variability $\nu$} 
+We simulate correlated noise thanks to Gaussian Processes.
+In our study, we are particularly interested in stellar variability and its effect on transit detection.
+For this reason, we use a physically-motivated kernel, describing the covariance of a stochastically-driven damped harmonic oscillator (SHO, \cite{, celerite, celerite2}) taking the form
+
+\begin{equation}
+    \begin{gathered}
+        k(\tau) = \sigma^2\,\exp\left(-\frac{\omega\,\tau}{2\,Q}\right)
+        \left\{\begin{array}{ll}
+            1 + \omega\,\tau & \mbox{for } Q = 1/2 \\
+            \cosh(f\,\omega\,\tau/2\,Q) + \sinh(f\,\omega\,\tau/2\,Q)/f
+                & \mbox{for } Q < 1/2 \\
+            \cos(g\,\omega\,\tau/2\,Q) + \sin(g\,\omega\,\tau/2\,Q)/g
+                & \mbox{for } Q > 1/2
+        \end{array}\right. \\
+        \text{where}\quad \tau = |t_i - t_j|\text{,}\quad f = \sqrt{1 - 4\,Q^2} \quad \text{and}\quad g = \sqrt{4\,Q^2 - 1}
+    \end{gathered}
+\end{equation}
+We use this kernel through its implementation in the \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}} Python package, providing an implententation of the quasi-separable kernel from \citealt{celerite2} powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
+The stellar variability signal in \autoref{fig:app_principle_dataset} has been sampled from a Gaussian Process with an SHO kernel of parameters $\omega = \pi/6D$ (i.e.
+a period equal to 12 times the duration $D$ of the simulated transit), $Q=45$ and $\sigma=\Delta$, the depth of the simulated transit.
+An extra term $\sigma_f^2=0.001^2$ is added to the covariance matrix, corresponding to the variance of the simulated measurement $f$ and leading to the white noise observed in \autoref{fig:app_principle_dataset}.
+
+\newpage
 \section{Combining transits}\label{combining_transits}

### 2023-01-23 simple/latex/ms.tex
larger figures + refactor synthetic IR
typos and syntax reading

@@
 \newcommand{\set}[1]{\{\,#1\,\}}
+\newcommand{\nuance}{\texttt{nuance}}
+\newcommand{\TODO}{\textcolor{red}{TODO}}
 
@@
 \begin{abstract}
-    We present \texttt{nuance}, an algorithm to search for planetary transits in light curves featuring correlated noises, such as instrumental signals and stellar photometric variability.
-Simpler approaches consist in searching for transits in light curves cleaned from correlated noise, where only transit signals would remain.
-However, we show that commonly used detrending techniques strongly affect transits signal-to-noise, up to the point of no detection when the noise characteristics is close to that of the transits.
-Focusing on stellar variability, we explore the parameter space for which this degradation occur, and quantify its effect on variety of cases.
+    We present \nuance{}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
+To deal with these nuisance signals, a common approach consist in cleaning a light curve from correlated noise before searching for transits.
+However, we show that this approach, based on an incomplete light curve modelling, strongly degrades transit signals, up to the point of no detection.
+As this degradation depends on the noise characteristics, we explore the parameter space for which it occurs, and quantify its effect on a wide variety of cases.
+We show that \nuance{} outperforms commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth.
+We perform our tests by injecting transits on synthetic light curves, and on known TESS candidates light curves to assess the performance of our algorithm on realistic multi-planetary systems datasets.
+Beyond its detection efficiency, we make \nuance{} tractable, \TODO{} times faster than current alternatives based on brute force detrending, and available as a well documented open-source Python package.
 \end{abstract}
@@
 However, disentangling these signals comes with many challenges.
-Due to their nature we will refer to these signal as correlated noises\\
-Widely used transit-search algorithms (Box-Least-Square algorithms \cite{bls}) are capable of detecting transits on light curves only containing transits and white noise.
+Due to their nature we will refer to these nuisance signals as correlated noise.
+\\\\
+Widely used transit-search algorithms (\cite{bls}) are capable of detecting transits on light curves containing only transits and white noise.
 Hence, the simplest way to find periodic transit signals is to first clean a light curve from nuisance signals before performing the search.
 This strategy is widely adopted by the community, both using physically-motivated systematics models like \cite{everest1, everest2}, or empirical filtering techniques, such as the ones described and implemented in \cite{wotan}.
-However, when correlated noises start resembling transits, this cleaning step (often referred to as \textit{detrending}) strongly degrades their detectability.
-In such cases, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noises, and asses the likelihood of the model to the data on a wide parameter space, an approach largely avoided due to its untractable nature.
-Nonetheless, \citealt{kovacs2016} ask the question: \textit{Periodic transit and variability search with simultaneous systematics filtering: Is it worth it?}.
-The latter study discards the general benefit of using a full-fledged approach, but it fails at exploring the light-curves characteristics for which it becomes necessary.\\
+However, when correlated noise start resembling transits, this cleaning step (often referred to as \textit{detrending}) is believed to degrade their detectability.
+In this case, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noise, and asses the likelihood of the data to the transit model on a wide parameter space, an approach largely avoided due to its intractable nature.
+Nonetheless, \citealt{kovacs2016} ask: \textit{Periodic transit and variability search with simultaneous systematics filtering: Is it worth it?}, and answer \textit{no}.
+Although the latter study discards the benefit of using a full-fledged approach in the general case, it fails at exploring the light-curves characteristics for which it becomes necessary.\\
 
@@
 Hence, these hidden system are pristine.$
-We present \texttt{nuance}, an algorithm using linear models and Gaussian processes (GP) to simultaneously search for transits while modeling correlated noise in a tractable way, such as instrumental signals and stellar photometric variability.
-In \autoref{issues}, we describe the issues inherent to the two-step approach described earlier, and study the parameter space for which commonly used detrending techniques degrade transit signals to the point of no detection.
-In \autoref{nuance} we describe the tractable full-fledged approach of \texttt{nuance}, and its implementation in an open-source Python package.
-In \autoref{simu} we assess the performance of \texttt{nuance} by computing the recovery of planetary transits injected into simulated light curves.
+We present \nuance, an algorithm using linear models and Gaussian processes to simultaneously search for transits while modeling correlated noise in a tractable way, such as instrumental signals and stellar photometric variability.
+In \autoref{issues}, we describe the issues inherent to the detrending approach described earlier, and study the parameter space for which commonly used detrending techniques degrade transit signals to the point of no detection.
+In \autoref{nuance} we describe the tractable full-fledged approach of  \nuance, and its implementation in an open-source Python package.
+In \autoref{simu} we assess the performances of \nuance by computing the recovery of planetary transits injected into simulated light curves.
 In section 4, we use nuance in a case study, to detect a known planet observed by the T E S S but undetected by the TESS pipeline due to its light curve characteristics.
@@
 
-\newpage
 \section{The issue with correlated noise and its detrending}\label{issues}
 Two sources of correlated noise particularly justify the need for a detrending step before searching for transits: instrumental noises (such as telescope pointing errors) and stellar variability (induced by pulsations or starspots).
-In this section, we discuss the impact of such correlated noises and their detrending on transits detectability.
-For simplicity, we will model transit signals using an analytic empirical model described in \citealt{protopapas} (see Annexe).\\
+In this section, we discuss the impact of such correlated noise and its detrending on transits detectability.
 
@@
 To study transits detectability, we will focus on the signal-to-noise (SNR) of a unique event, reduced to the simplified expression (\citealt{pont2006}, Equation 12):
-
 \begin{equation}\label{eq:snr}
@@
 \end{equation}
-
-where $df$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits (unity here since we consider a single transit), and $\sigma_w$ and $\sigma_c$ are the white and correlated noise standard deviations.
-To show the impact of correlated noise on transit detectability, we simulate a unique transit signal and compute its SNR using \autoref{eq:snr}, both in the absence and presence of correlated noise (\autoref{fig:issue1}).
+where $df$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits ($N_{tr}=1$ here since we consider a single transit), and $\sigma_w^2$ and $\sigma_c^2$ are the white and correlated noise variances.
+To show the impact of correlated noise on transit detectability, we simulate a unique transit signal (using \autoref{eq:protopapas_single}) and compute its SNR using \autoref{eq:snr}, both in the absence and presence of correlated noise (\autoref{fig:issue1}).
 
@@
 
-Instrumental signals have the benefit to be shared among light curves of stars observed with the same instrument, strongly correlated with measurements from the experimental setup (like detector's temperature, pointing error, sky background or airmass), hence we will make the strong assumption that their detrending based on an incomplete model of the light curve, one that ignore transit signals (because unknowns), do not affect the latter.
+Instrumental signals have the benefit to be shared among light curves of stars observed with the same instrument, strongly correlated with measurements from the experimental setup (like detector's temperature, pointing error, sky background or airmass), hence we will make the strong assumption that their detrending based on an incomplete model of the light curve, one that ignore transit signals (because unknowns), do not affect the searched signals.
 In opposition, stellar variability is generally unknown and harder to correlate with simultaneous measurements.
 This gave rise to two types of treatments in order to reconstruct and detrend stellar variability, and perform transit search on \textit{flattened} light curves.
-One is physically-motivated and make use of Gaussian Processes (e.g.
+One is physically-motivated and makes use of Gaussian Processes (e.g.
 \cite{k2sc}).
-The other is empirical and make use of filtering algorithms (a wide variety being described in \cite{wotan}).
-In this section, we will study in details the effect of both approaches on transit detectability, by studying the degradation of a unique transit SNR for a wide variety of stellar variability characteristics.\\
-
+The other is empirical and makes use of filtering algorithms (a wide variety being described in \cite{wotan}).
+In this section, we will study in details the effect of both approaches on transit detectability, by studying the degradation of a unique transit SNR for a wide variety of stellar variability characteristics.
+\\\\
 Throughout this paper, we simulate stellar variability (and later model it) thanks to Gaussian Processes (GP), employing the physically-motivated stochastically-driven damped simple harmonic oscillator kernel (SHO) presented in \citealt{celerite1}.
@@
         \includegraphics[width=\linewidth]{../figures/issue2.pdf}
-        \caption{}
+        \caption{\TODO}
         \label{fig:issue2}
@@
 
-In \autoref{fig:issue2}, we simulate a light curve featuring a single transit (again using the model from \citealt{protopapas}), in addition with stellar variability of different timescales and amplitudes.
+In \autoref{fig:issue2}, we simulate a light curve featuring a single transit in addition with stellar variability of different timescales and amplitudes.
 In case \textit{a} (purple in \autoref{fig:issue2}), we model and detrend these signals using the widely-adopted Tukey's biweight filtering method presented in \citealt{tukey} and its implementatation from \texttt{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}.
 In case \textit{b} (red in \autoref{fig:issue2}) we employ the same GP model used to simulate stellar variability to reconstruct and remove it from light curves ignoring the presence of potential transits (the case of unknown transits search).
-Once transits cleaned from stellar variability, we asses their remaining SNR using \autoref{eq:snr}.\\
-
+Once transits cleaned from stellar variability, we asses their remaining SNR using \autoref{eq:snr}.
+\\\\
 \autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and suggests that this degradation is strongly dependant on the stellar variability characteristics encountered.
-In order to explore the parameter space for which detrending is the most problematic, we simulate light curves including a single transit and a much wider range of correlated noise characteristics defined with two parameters: $\tau_v$ the relative timescale of the variability with respect to the transit duration; and $\delta_v$, the relative amplitude of the variability against transit depth, such that:
+In order to explore the parameter space for which detrending is the most problematic, we simulate light curves including a single transit and a much wider range of correlated noise characteristics defined with two parameters: $\tau$ the relative timescale of the variability with respect to the transit duration; and $\delta$, the relative amplitude of the variability against transit depth, such that:
 \begin{equation}
-    \tau_v \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}} \quad \text{and} \quad 
-    \delta_v \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}} 
+    \tau \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}} \quad \text{and} \quad 
+    \delta \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}} 
 \end{equation}
 To follow this parametrization, we simulate the photometric variability using a GP with an SHO kernel with hyperparameters \footnote{}:
-\begin{equation}\label{eq:params}
-    \omega = \frac{\pi}{\tau_v\tau} \quad 
-    \sigma = \delta_v \frac{\delta}{2} \quad  \text{and}  \quad  
+\begin{equation}\label{eq:relative_params}
+    \omega = \frac{\pi}{\tau D} \quad 
+    \sigma = \delta \frac{\Delta}{2} \quad  \text{and}  \quad  
     Q = 10
 \end{equation}
-with $\delta$ and $\tau$ the depth and duration of the simulated transit, similar in all light curves.
-We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability signals.
-For $\tau_v=1$ and $\delta_v=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:params} correspond to a periodic signal with a period half that of the transit duration, and a peak to peak amplitude two times that of the transit depth, i.e.
+with $\Delta$ and $D$ the depth and duration of the simulated transit, similar in all light curves. 
+% We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability signals. 
+For $\tau=1$ and $\delta=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a periodic signal with a period half that of the transit duration, and a peak to peak amplitude two times that of the transit depth, i.e.
 strongly resembling the simulated transit signal.
-As in \autoref{fig:issue2}, we reconstruct the variability using a biweight filter with a window length of $3\tau$ (three times the transit duration, an optimal value according to \citealt{wotan}).
+As in \autoref{fig:issue2}, we reconstruct the variability using a biweight filter with a window length of $3D$ (three times the transit duration, an optimal value according to \citealt{wotan}).
 We then subtract the variability from each light-curve, estimate the resulting transit depth (using the in-transit minimum flux) and compute its SNR.
-\autoref{fig:snr_detrend} shows the remaining SNR values of transits computed this way, after each variability signals with random $(\tau_v, \delta_v)$ have been detrended.
+\autoref{fig:snr_detrend} shows the remaining SNR values of transits computed this way, after each variability signals with random $(\tau, \delta)$ have been detrended.
 \\\\
-\autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau_v, \delta_v)$ parameter space for which detrending degrades transit SNR to the point of no detection ($SNR < 6$).
-While this region might only represent a fraction of existing exoplanetary systems, we argue that these systems are of great value for the field of exoplanetary science.
-First, as variability is often linked to the presence of starspots, systems whose  with the stellar photo in is likely to be observed on stars whose equatorial plane lines up with our line of sight, following recent studies on the dynamical evolution of planetary systems \cite{}, hence being most likely to be transited.
+\autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which detrending degrades transit SNR to the point of no detection ($SNR < 6$).
+While this region might only represent a fraction of existing exoplanetary systems, we argue that these systems are of great value for the field of Exoplanetary Science.
+Indeed, as variability is often linked to the presence of starspots, systems whose  with the stellar photo
+
+% in is likely to be observed on stars whose equatorial plane lines up with our line of sight, following recent studies on the dynamical evolution of planetary systems \cite{}, hence being most likely to be transited.
+\TODO
+
 As a direct consequence, these same starspots are more likely to be occulted by planetary companions, events whose measurement benefits both the study of stellar atmospheres and their concerning impact on planetary atmospheric retrievals \cite{}.
 Finally, the growing interest of the community for ultra-cool dwarf stars comes with observations done in the infrared, featuring enhanced red noise and variability, leading to lower transit SNR \cite{}.
-Hence, commonly-used detrending technique (such as the biweight filter studied in this section) make transit-search blind to those pristine systems, motivating the need for a more informed transit search algorithm able to deal with correlated noises.
+Hence, commonly-used detrending technique (such as the biweight filter studied in this section) make transit-search blind to those pristine systems, motivating the need for a more informed transit search algorithm able to deal with correlated noise.
 In the next section we present such an algorithm: \texttt{nuance}.
@@
         \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
-        \caption{}
+        \caption{\TODO}
         \label{fig:snr_detrend}
@@
 
-In this section we describe \texttt{nuance}, an algorithm to search for planetary transits in light curves containing correlated noises, such as instrumental signals and stellar photometric variability.
+\texttt{nuance} is an algorithm capable of searching for planetary transits in light curves containing correlated noise, such as instrumental signals and stellar photometric variability.
 \\\\
@@
         \includegraphics[width=0.7\linewidth]{../figures/principle_dataset.pdf}
-        \caption{}
+        \caption{\TODO}
         \label{fig:principle_dataset}
@@
 \\\\
-Ideally, we would detect such  a periodic transit signal by sampling the posterior likelihood of this data to a full-fledge model including stellar variability (more generally correlated noises), instrumental systematics (modeled with explanatory variables), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$.
+Ideally, we would detect such  a periodic transit signal by sampling the posterior likelihood of this data to a full-fledge model including stellar variability (more generally correlated noise), instrumental systematics (modeled with explanatory variables), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$.
 We would then reduce the posterior likelihood to $p(\bm{f}\vert P)$, its marginalized version over all parameters except the period $P$, producing a transit search periodogram $\mathcal{Q}(P)$.
@@
 We will do that by adopting the strategy of \cite{foreman2016}, and separate the transit search into two components: the \textit{linear search} and the \textit{periodic search}.
-During the \textit{linear search}, the likelihood of a single non-periodic transit is computed for a grid of epochs and durations (but not depths, as we will see).
+During the \textit{linear search}, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
 Then, the \textit{periodic search} consists in combining these likelihoods to compute the likelihood of the data given a periodic transit signal for a range of periods.
-These combined likelihoods yield a transit-search periodogram on which the periodic transit detection can be based.
+These combined likelihoods yield a transit-search periodogram on which the periodic transit detection is based.
 \texttt{nuance} differs by modeling the covariance of the light curve with a Gaussian Process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
@@
 \\\\
-To account for correlated noises, we model the light curve $f$ as being drawn from a Gaussian Process such that
+To account for correlated noise, we model the light curve $f$ as being drawn from a Gaussian Process such that
 $$\bm{f} \sim \mathcal{N}(\bm{w X}, \bm{\Sigma})$$
@@
 a linear model of the M explanatory variables with coefficients $\bm{w}$) and covariance $\bm{\Sigma}$.
-To account for the presence of a single non-periodic transit of epoch $T$ and duration $D$, we compute and append its signal as the last column of the design matrix $\bm{X}$, using the simple transit model from \citealt{protopapas} with a unitary depth.
+To account for the presence of a single non-periodic transit of epoch $T$ and duration $D$, we compute and append its signal as the last column of the design matrix $\bm{X}$, using the simple transit model from \citealt{protopapas} with a unitary depth (see \autoref{app_proto}).
 This way, the transit signal is part of the linear model and its depth $\Delta$ can be solved linearly.
-Under this assumption, the log-likelihood of the data given a single non-periodic transit is analytical
+Under this assumption, the likelihood of the data given a single non-periodic transit is analytical
 \begin{equation} \label{eq:linear_search_ll}
@@
 and can be computed on a grid of epochs and durations, the transit depth being linearly solved for any $(T, D)$.
-This \textit{linear search} lead to the set of  log-likelihoods\footnote{This notation omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{i,j}$) as it is linearly solved for any given value of $(T_i, F_j)$ and irrelevant in what follows.}
+This \textit{linear search} lead to the set of likelihoods\footnote{This notation omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{i,j}$) as it is linearly solved for any given value of $(T_i, F_j)$ and irrelevant in what follows.}
 $$\ln\mathcal{L} = \{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})\}_{i, j}$$
-were $\Delta_{i,j}$ is the depth linearly solved for a given $(D_i, T_j)$.
-\autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}, using the same Gaussian Process and design matrix $\bm{X}$ used to simulate the data (see annexe).
+were $\Delta_{i,j}$ is the depth linearly solved for a given $(T_i, D_j)$.
+\autoref{fig:linear_search} shows this likelihood grid computed for the simulated dataset shown in \autoref{fig:principle_dataset}, using the same Gaussian Process and design matrix $\bm{X}$ used to simulate the data (see \autoref{app_gp}).
 
@@
     \begin{centering}
-        \includegraphics[width=0.9\linewidth]{../figures/principle_linear_search.pdf}
+        \includegraphics[width=0.95\linewidth]{../figures/principle_linear_search.pdf}
         \caption{Principle and output of the \textit{linear search}.
 First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated.
-For each pair of index $(i,j)$, the log likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed, using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}.
-This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors.}
+For each pair of index $(i,j)$, the likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed, using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}.
+This process yields the grid of likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly.}
         \label{fig:linear_search}
@@
 $$
-This is the joint log-likelihood of transits belonging to the same periodic signal but with varying depths  $\set{\Delta_k}_k$.
+This is the joint likelihood of transits belonging to the same periodic signal but with varying depths  $\set{\Delta_k}_k$.
 However, individual transits from a periodic signal cannot be considered independent, and should instead share a common transit depth $\Delta$.
-We show in \autoref{combining_transits} that there is an analytical expression for the joint log-likelihood of K individual transits with depths and errors $\set{\Delta_k, \sigma_k}_k$ assuming a common depth $\Delta$:
+We show in \autoref{combining_transits} that there is an analytical expression for the joint likelihood of K individual transits with depths and errors $\set{\Delta_k, \sigma_k}_k$ assuming a common depth $\Delta$:
 \begin{equation}\label{eq:comb}
@@
 \end{equation}
-While this gives a solution for \autoref{eq:goal}\footnote{Different from the one found in \cite{foreman2016}}, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$.
-In \cite{foreman2016} this is solved by using the nearest neighbors in the epochs grid.
+While this gives a solution for \autoref{eq:goal}, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$.
+In \cite{foreman2016} a similar issue is solved by using the nearest neighbors in the epochs grid.
 Instead, to allow the efficient matrix computation of \autoref{eq:comb}, we interpolate the likelihood grid from $\set{T_i}_i$ to a common grid of transit phases $\set{\phi_i}_i$, leading to the \textit{periodic search} log-likelihood
 $$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi, D)}_{i,j}$$
-shown for few periods in \autoref{fig:periodic_search}.b, a notation where $\Delta_{i,j}$ is omitted since being interpolated from the \textit{linear search} using $\phi_i$ and $D_j$, and $T_0 = 0$.
+shown for few periods in \autoref{fig:periodic_search}.b.
+In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the \textit{linear search} using $\phi_i$, $D_j$ and $T_0 = 0$.
 
@@
     \begin{centering}
-        \gridline{\fig{../figures/principle_periodic_0.pdf}{0.9\linewidth}{\vspace{-0.5cm}(a) $\ln \mathcal{L}$}}
+        \gridline{\fig{../figures/principle_periodic_0.pdf}{0.95\linewidth}{\vspace{-0.5cm}(a) $\ln \mathcal{L}$}}
         \vspace{-0.6cm}
-        \gridline{\fig{../figures/principle_periodic_1.pdf}{0.7\linewidth}{\vspace{-0.5cm}(b) $\mathcal{P}(P)$}}
+        \gridline{\fig{../figures/principle_periodic_1.pdf}{0.72\linewidth}{\vspace{-0.5cm}(b) $\mathcal{P}(P)$}}
         \caption{On the example dataset, this figure shows how the \textit{periodic search} works at different periods $P$ (including the true period $P = 0.7$ days).
-Given $P$ and $T_0=0$, the log-likelihood $\ln\mathcal{L}$ shown in (a) is phase-folded and interpolated onto a common grid of phases shown in (b).
-As an example, the white lines in (a) mark the edges of each fold for a period of $P=0.5$ days, and the white crosses the epochs $\set{T_k}_{k\in\mathbb{K}}$ matching a particular phase in the grid (reported in (a) and (b)) on which the corresponding $\set{\ln \mathcal{L}_k}_{k\in\mathbb{K}}$ are interpolated and combined using \autoref{eq:comb}.
-To allow the use of efficient matrix computations, this is done for every duration $\set{D_i}_i$ so that $\mathcal{P}$ is computed on the full grid $\set{D_i, \phi_i}_{i, j}$ at once.
-We understand from plots in (a) that a different choice of epoch $T_0$ may only shift the results in phase but do not affect the values of $\mathcal{P}$.
+Given $P$ and $T_0=0$, the likelihood $\ln\mathcal{L}$ shown in (a) is phase-folded and interpolated onto a common grid of phases shown in (b).
+As an example, the white lines in (a) mark the edges of each fold for a period of $P=0.5$ days, and the white crosses show the epochs $\set{T_k}_{k\in\mathbb{K}}$ matching a particular phase in the grid (reported in (a) and (b)) on which the corresponding $\set{\ln \mathcal{L}_k}_{k\in\mathbb{K}}$ are interpolated and combined using \autoref{eq:comb}.
+To allow the use of efficient matrix computations, this is done for all durations $\set{D_i}_i$ so that $\mathcal{P}$ is computed on the full grid $\set{D_i, \phi_i}_{i, j}$ at once.
+We understand from the folded likelihoods plots in (b) that a different choice of epoch $T_0$ may only shift the results in phase but do not affect the values of $\mathcal{P}$.
 For this reason, computing $\mathcal{P}$ for $T_0=0$ is sufficient.
-We also notice how the maximum value of $\mathcal{P}(P=0.35)$ is lower than for $P=0.7$ days, a result of combining the log-likelihoods using \autoref{eq:comb} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
+We also notice how the maximum value of $\mathcal{P}(P=0.7/2)$ is lower than for $P=0.7$ days, a result of combining the log-likelihoods using \autoref{eq:comb} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
         \label{fig:periodic_search}
@@
 
-Using \autoref{eq:comb}, we can now compute $\ln\mathcal{P}$ for a range of periods (see \autoref{fig:periodic_search}) and build a transit search periodogram using \autoref{eq:periodogram}.
+Using \autoref{eq:comb}, we can now compute $\ln\mathcal{P}$ for a range of periods and build a transit search periodogram using \autoref{eq:periodogram}.
 But a final issue emerges, one that is fundamentally linked to our strategy.
@@
 This motivates a final step to produce the transit search periodogram $Q$.
-\\\\
 For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, we retrieve the maximum likekihood parameters
@@
 \end{equation}
-and define $\mathcal{Q}(P)$ as the SNR of the transit of period $P$, epoch $T_0 = \phi_0 P$ duration $D$ and depth $\Delta$, i.e.
+and define $\mathcal{Q}(P)$ as the SNR of the transit of period $P$, epoch $T_0 = \phi_0 P$, duration $D$ and depth $\Delta$, i.e.
 $$
@@
         \includegraphics[width=0.8\linewidth]{../figures/principle_Q.pdf}
-        \caption{}
+        \caption{\TODO}
         \label{fig:periodogram}
@@
 The periodic transit of period $P$ with the maximum SNR, i.e.
-maximizing $\mathcal{Q}$, is adopted as the best candidate, adapting our confidence through the minimum SNR for which we consider a signal to be real.
+maximizing $\mathcal{Q}$, is adopted as the best candidate, basing our confidence in this signal through its SNR.
 The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (\autoref{eq:comb}).
@@
 \newpage
-\subsubsection{An open-source python package}
+\subsection{An open-source python package}
+\TODO
 
 \section{Injection-recovery on simulated data}\label{simu}
-
+\TODO
 \begin{figure}[H]
@@
         \includegraphics[height=12cm]{../../workflows/synthetic_injection_recovery/figures/final_result.pdf}
-        \caption{}
+        \caption{\TODO}
         \label{fig:simu}
@@
 
-\subsection{Transit signal $\mathcal{T}$}
+\subsection{Transit signal $\mathcal{T}$}\label{app_proto}
 We simulate the periodic transit signal $\mathcal{T}$ using the simple model described in \cite{protopapas} where a transit of period $P$, epoch $T_0$, duration $D$ and unitary depth observed at time $t$ is given by
@@
 
-\subsection{Stellar variability $\nu$} 
+\subsection{Stellar variability $\nu$}\label{app_gp}
 We simulate correlated noise thanks to Gaussian Processes.

### 2023-01-23 simple/latex/ms.tex
add proof and test

@@
 \documentclass{aastex631}
+% \usepackage{showyourwork}
 \usepackage{float}
@@
 \usepackage{bm}
-\usepackage[dvipsnames]{xcolor}
 
@@
 \newcommand{\TODO}{\textcolor{red}{TODO}}
+
+% Cortex style
+\input{\cortex}
 
@@
         \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
-        \caption{\TODO}
+        \caption{\TODO{}, and redo plot with top arrows and fix $tau_v$ $delta_v$}
         \label{fig:snr_detrend}
@@
 The product of two normal distributions is also a normal distribution, so that the integral in \autoref{eq:ice_int} can be obtained analytically. We have
-$$p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i  \right) \frac{1}{\sqrt{2\pi \left(\sigma^{2} + \sigma_{k}^{2}\right)}} e^{- \frac{1}{2}\frac{\left(\Delta -  \Delta_{k}\right)^{2}}{\sigma^{2} + \sigma_{k}^{2}}}$$
+\begin{proof}{integral_sympy}
+    p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i  \right) \frac{1}{\sqrt{2\pi \left(\sigma^{2} + \sigma_{k}^{2}\right)}} e^{- \frac{1}{2}\frac{\left(\Delta -  \Delta_{k}\right)^{2}}{\sigma^{2} + \sigma_{k}^{2}}}
+\end{proof}
 And \autoref{eq:goal} gives 

### 2023-01-26 simple/latex/ms.tex
paper: package and simulated ir

@@
 \usepackage{bm}
+\usepackage{listings}
 
@@
 \newcommand{\TODO}{\textcolor{red}{TODO}}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{gray!5},
+    commentstyle=\color{gray!70},
+    keywordstyle=\color{Bittersweet},
+    stringstyle=\color{RoyalBlue},
+    basicstyle=\fontsize{7.5}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont,
+    breakatwhitespace=false,         
+    breaklines=true,
+    rulecolor=\color{black!15},
+    numbers=left,
+    numberstyle=\fontsize{7}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont\color{gray!50},
+    framerule=0pt,
+    breakindent=0pt,
+    resetmargins=true,
+    numbersep=5pt,
+    frame=single,
+    aboveskip=1em,
+    belowskip=1em,
+}
+\lstset{style=mystyle}
 
@@
 To deal with these nuisance signals, a common approach consist in cleaning a light curve from correlated noise before searching for transits.
-However, we show that this approach, based on an incomplete light curve modelling, strongly degrades transit signals, up to the point of no detection.
+However, we show that this approach, based on an incomplete modelling of light curves, strongly degrades transit signals, up to the point of no detection.
 As this degradation depends on the noise characteristics, we explore the parameter space for which it occurs, and quantify its effect on a wide variety of cases.
-We show that \nuance{} outperforms commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth.
+We show that \nuance{} outperforms the detection capabilities of commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth.
 We perform our tests by injecting transits on synthetic light curves, and on known TESS candidates light curves to assess the performance of our algorithm on realistic multi-planetary systems datasets.
@@
 
-\subsection{The effect of detrending on transits detectability}
+\subsection{The effect of detrending on transits detectability}\label{detrending_effect}
 
@@
 This gave rise to two types of treatments in order to reconstruct and detrend stellar variability, and perform transit search on \textit{flattened} light curves.
-One is physically-motivated and makes use of Gaussian Processes (e.g.
-\cite{k2sc}).
+One is physically-motivated and makes use of Gaussian Processes \citep[e.g.]{k2sc}.
 The other is empirical and makes use of filtering algorithms (a wide variety being described in \cite{wotan}).
@@
         \includegraphics[width=\linewidth]{../figures/issue2.pdf}
-        \caption{\TODO}
+        \caption{Simulation of a single non-periodic transit plus various variability signals using the formalism described in \autoref{app_principle_simulations}.
+For each simulation, the variability is fitted using a biweight filter \citep[in violet]{wotan} and a Gaussian Process (in red, descibed in \autoref{app_gp}, in each case with the same kernel used to generate the data).
+The fitted variability is then subtracted and the transit SNR computed using \autoref{eq:snr}, with the transit depth estimated as the minimum in-transit flux.}
         \label{fig:issue2}
@@
 In order to explore the parameter space for which detrending is the most problematic, we simulate light curves including a single transit and a much wider range of correlated noise characteristics defined with two parameters: $\tau$ the relative timescale of the variability with respect to the transit duration; and $\delta$, the relative amplitude of the variability against transit depth, such that:
-\begin{equation}
+\begin{equation}\label{eq:relative_param_space}
     \tau \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}} \quad \text{and} \quad 
@@
         \includegraphics[width=0.7\linewidth]{../figures/principle_dataset.pdf}
-        \caption{\TODO}
+        \caption{Simulated light curve containing a periodic transit signal, instrumental systematics, and stellar variability, all described in \autoref{app_principle_simulations}.}
         \label{fig:principle_dataset}
@@
         \includegraphics[width=0.8\linewidth]{../figures/principle_Q.pdf}
-        \caption{\TODO}
+        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ of matching transits is computed using \autoref{eq:comb}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
         \label{fig:periodogram}
@@
 \subsection{An open-source python package}
-\TODO
+We implement the methods described in this paper in the open-source Python package \nuance{}, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} with an extensive user documentation\footnote{\href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}}.
+\\\\
+To instantiate a search, a user can start by creating a \texttt{Nuance} object with
+\begin{lstlisting}[language=Python]
+    from nuance import Nuance
 
-\section{Injection-recovery on simulated data}\label{simu}
+    nu = Nuance(time, flux, error, kernel, X)
+\end{lstlisting}
+where \texttt{kernel} is the kernel of the Gaussian Process used to model corelated noise, and \texttt{X} the design matrix of the linear model.
+In \nuance{}, Gaussian Processes are created through the \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}} Python package powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}, allowing for custom kernels to be built and highly tractable computations.
+We can then define a set of epochs \texttt{t0} and durations \texttt{Ds} and run the \textit{linear search} with
+\begin{lstlisting}[language=Python]
+    import numpy as np
+
+    t0s = time.copy()
+    # a range of 20 durations
+    Ds = np.linspace(0.01, 0.2, 10)
+    nu.linear_search(t0s, Ds)
+\end{lstlisting}
+Finally, the \textit{periodic search} is run with
+\begin{lstlisting}[language=Python]
+    # a wide range of periods
+    periods = np.linspace(0.1, 5, 2000)
+    search = nu.periodic_search(periods)
+\end{lstlisting}
+From this \texttt{search} object can be computed the best transiting candidate parameters (\texttt{search.best}), the $\mathcal{Q}$ periodogram (\texttt{search.Q\_snr}) as well as valuable information about the transit search.
+The \texttt{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \texttt{nuance} being that the \textit{linear search} only needs to be performed ones, and reused for the search of several transiting candidates.
+An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.
+
+\newpage
+\section{Validation using transit injection-recovery}\label{simu}
+In order to validate our algorithm, we assess the transit search capabilities of \texttt{nuance} against commonly used techniques.
+As demonstrated in \autoref{detrending_effect}, the full-fledged modeling offered by \texttt{nuance} is not always necessary, and becomes beneficial for particular noise characteristics that are relative to the searched transit parameters.
+Hence, our comparison will be done on the relative parameter space $(\tau, \delta)$ described in \autoref{eq:relative_param_space}, focusing on its formulation by assuming correlated noise in the form of stellar variability (\autoref{eq:relative_params}) simulated using a Gaussian Process with an SHO kernel (\autoref{app_gp}).
+Ultimately, we demonstrate in this section how \texttt{nuance} compares with commonly used techniques, depending on the similarity between stellar variability and the searched transit signals.
+\\\\
+In what follows, we inject transits in simulated or real datasets, and employ algorithms to recover these transits with the right parameters (hence \textit{injection-recovery}).
+We consider the transit to be found if the absolute difference between the periods $dP$ and epochs $dT_0$ of the recovered and injected planet follow
+$$dP<0.1\;\text{days} \hspace{0.5cm} \text{and} \hspace{0.5cm} dT_0<0.01\;\text{days}$$
+
+\newcommand{\wtls}{\texttt{wõtan+TLS}}
+
+\subsection{Comparison with \wtls{} on a simulated light curves}
+
+We start by comparing \texttt{naunce} to a relatively simple method on simulated light curves.
+As in \autoref{detrending_effect}, this simple method consist in searching for planetary transits by first modeling stellar variability using an optimal biweight filter (implemented in \texttt{wõtan}, \citealt{wotan}), subtracting it from the light-curve, and perform the search on the detrended light curve using an optimal box-least-square algorithm (\texttt{TLS}, \citealt{tls}).
+We name this method \wtls{}.
+\\\\
+The dataset consists in 10 000 light curves simulated using the model described in \autoref{app_principle_simulations}.
+We simulate a common periodic transit among all light curves, of period $P=1.1$ days, epoch $T_0=0.2$ days, duration $D=0.04$ days and depth $\Delta=2\%$.
+Each light curve consist in a 3 days observation, with an exposure time of 2 minutes, leading to $N=2160$ data points with a normal error of $0.1\%$.
+\\\\
+For a given pair of $(\tau, \delta)$, we simulate stellar variability using a Gaussian Process with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params}, computed with respect to the injected transit parameters $D$ and $\Delta$.
+The same kernel is used for the search with \texttt{nuance}, an optimal choice on equal footing with the optimal biweight filter employed in the \wtls{} search.
+We generate 10 000 pairs of $(\tau, \delta)$ such that
+$$\tau \sim \mathcal{U}(0.1, 5)\hspace{0.5cm}\text{and}\hspace{0.5cm} \delta \sim \mathcal{U}(0.1, 10)$$
+with $\mathcal{U}(a, b)$ a uniform distribution of lower bound $a$ and upper bound $b$.
+
+\vfill
 \TODO
@@
 
-\section{Comparison with \texttt{Sherlock} on TOI-540}\label{real}
+\subsection{Comparison with \texttt{Sherlock} on TOI-540 real light curves}\label{real}
 
@@
 From this expression, we get the signal of the single non-periodic transit of unitary depth
-\begin{equation}\label{eq:protopapas_single}
+\begin{proof}{single_transits}\label{eq:protopapas_single}
     \tau_c(t, T_0, D) = \lim_{P\to \infty} \mathcal{T}(t, P, T_0, D) = -\frac{1}{2}\tan\left(\frac{c}{2}(1-T_0+t)\right) -\frac{1}{2}\tan\left(\frac{c}{2}(1+T_0-t)\right)
-\end{equation}
+\end{proof}
 with epoch $T_0$ and duration $D$.
@@
 The product of two normal distributions is also a normal distribution, so that the integral in \autoref{eq:ice_int} can be obtained analytically. We have
-\begin{proof}{integral_sympy}
+\begin{proof}{gaussians_product_integral}
     p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i  \right) \frac{1}{\sqrt{2\pi \left(\sigma^{2} + \sigma_{k}^{2}\right)}} e^{- \frac{1}{2}\frac{\left(\Delta -  \Delta_{k}\right)^{2}}{\sigma^{2} + \sigma_{k}^{2}}}

### 2023-01-26 simple/latex/ms.tex
paper: simulated injection recovery results

@@
 where \texttt{kernel} is the kernel of the Gaussian Process used to model corelated noise, and \texttt{X} the design matrix of the linear model.
-In \nuance{}, Gaussian Processes are created through the \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}} Python package powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}, allowing for custom kernels to be built and highly tractable computations.
+\nuance{} exploits the Gaussian Process implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}} allowing for custom kernels to be built and highly tractable computations.
 We can then define a set of epochs \texttt{t0} and durations \texttt{Ds} and run the \textit{linear search} with
@@
 \end{lstlisting}
-From this \texttt{search} object can be computed the best transiting candidate parameters (\texttt{search.best}), the $\mathcal{Q}$ periodogram (\texttt{search.Q\_snr}) as well as valuable information about the transit search.
+From this \texttt{search} object, the best transiting candidate parameters can be computed (\texttt{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\texttt{search.Q\_snr}), together with valuable information about the transit search.
 The \texttt{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \texttt{nuance} being that the \textit{linear search} only needs to be performed ones, and reused for the search of several transiting candidates.
@@
 \newpage
-\section{Validation using transit injection-recovery}\label{simu}
+\section{Validation with injection-recovery}\label{simu}
 In order to validate our algorithm, we assess the transit search capabilities of \texttt{nuance} against commonly used techniques.
@@
 In what follows, we inject transits in simulated or real datasets, and employ algorithms to recover these transits with the right parameters (hence \textit{injection-recovery}).
-We consider the transit to be found if the absolute difference between the periods $dP$ and epochs $dT_0$ of the recovered and injected planet follow
+We consider the transit to be found if the absolute difference between the periods $dP$ and epochs $dT_0$ of the recovered and injected planets follow
 $$dP<0.1\;\text{days} \hspace{0.5cm} \text{and} \hspace{0.5cm} dT_0<0.01\;\text{days}$$
@@
 We start by comparing \texttt{naunce} to a relatively simple method on simulated light curves.
-As in \autoref{detrending_effect}, this simple method consist in searching for planetary transits by first modeling stellar variability using an optimal biweight filter (implemented in \texttt{wõtan}, \citealt{wotan}), subtracting it from the light-curve, and perform the search on the detrended light curve using an optimal box-least-square algorithm (\texttt{TLS}, \citealt{tls}).
+As in \autoref{detrending_effect}, this simple method consist in searching for planetary transits by first modeling stellar variability using an optimal biweight filter (implemented in \texttt{wõtan} \citep{wotan} using a window size three times that of the injected transit duration), subtracting it from the light-curve, and then perform the search on the detrended light curve using an optimal box-least-square algorithm (\texttt{TLS}, \citealt{tls}).
 We name this method \wtls{}.
@@
 We generate 10 000 pairs of $(\tau, \delta)$ such that
-$$\tau \sim \mathcal{U}(0.1, 5)\hspace{0.5cm}\text{and}\hspace{0.5cm} \delta \sim \mathcal{U}(0.1, 10)$$
-with $\mathcal{U}(a, b)$ a uniform distribution of lower bound $a$ and upper bound $b$.
+$$\tau \sim \mathcal{U}(0.1, 40)\hspace{0.5cm}\text{and}\hspace{0.5cm} \delta \sim \mathcal{U}(0.1, 100)$$
+where $\mathcal{U}(a, b)$ denotes a uniform distribution of lower bound $a$ and upper bound $b$.
+\\\\
+The results of these injection-recovery are shown in \autoref{fig:simu} and highlight particularly well the benefit of \nuance{} against \wtls{} on transits corresponding to $\delta > 3$ and $\tau > 2$, i.e.
+with relatively low transit depth compared to the stellar variability amplitude, and a relatively small duration compared to the stellar variability period.
+On average, searching a single light curve with \nuance{} took 9 seconds, compared to 8 seconds with \wtls{}.
+Beyond its increased sensitivity, this metric demonstrates the tractability of \nuance{}, studied in more details in \autoref{}.
+\\\\
+About these results, we note these transit searches were done in a particularly optimal setup, on light curves not all physically realistic, hence demonstrating the performance of \nuance{} on a theoretical basis only.
+In the next subsection, we perform transits injection-recovery on a real dataset, using more realistic tools and on light curves featuring known planetary transits, hence validating \nuance{} on a realistic multi-planetary dataset.
 
-\vfill
-\TODO
 \begin{figure}[H]

### 2023-01-26 workflows/synthetic_injection_recovery/scripts/plot_final.py

@@ -7,8 +7,8 @@ from matplotlib import patches
 
 nuance_main, wotan_main, nuance_zoom, wotan_zoom = snakemake.input
 
-ylabel = r"$\delta_v \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$"
-xlabel = r"$\tau_v \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$"
+ylabel = r"$\delta$"
+xlabel = r"$\tau$"
 
 fig = plt.figure(None, (7, 6.5))
 
@@ -34,10 +34,10 @@ im = main.imshow(
     vmin=0.,
     vmax=1.
 )
-main.set_title("wõtan + TLS", loc="left")
+main.set_title("wõtan + TLS\n\n")
 
 
-zoomw = plt.subplot(2, 2, 2)
+zoomw = plt.subplot(2, 2, 3)
 bins = (30,30)
 df = pd.read_csv(wotan_zoom)
 detected, _var, _amp = df.values.T.astype(float)
@@ -56,22 +56,23 @@ zoomw.imshow(
     vmax=1.
 )
 
-a = (0.5, 1) # xy of lower left box
+a = (0.01,0.01) # xy of lower left box
 b = (10, 25) # xy of upper right box
 pcolor="0.6"
 p = patches.Polygon([a, (a[0], b[1]), (b[0], b[1]), (b[0], a[1])], fill=False, ec=pcolor)
 main.add_patch(p)
 con = patches.ConnectionPatch(xyA=(a[0], b[1]), xyB=(0.1, b[1]-0.1), coordsA="data", coordsB="data", axesA=main, axesB=zoomw, color=pcolor)
 fig.add_artist(con)
-con = patches.ConnectionPatch(xyA=(b[0], a[1]), xyB=(0.1, 0.1), coordsA="data", coordsB="data", axesA=main, axesB=zoomw, color=pcolor)
+con = patches.ConnectionPatch(xyA=(b[0], a[1]), xyB=(b[0], b[1]-0.1), coordsA="data", coordsB="data", axesA=main, axesB=zoomw, color=pcolor)
 fig.add_artist(con)
 zoomw.set_ylim(0.1, b[1])
 zoomw.set_xlim(0.1, b[0])
+zoomw.set_xlabel(xlabel, fontsize=14)
+zoomw.set_ylabel(ylabel, fontsize=14)
 
 # nuance
[…]

### 2023-01-27 latex/ms.tex
Refactoring all + paper: thanks

--- 2023-01-26 simple/latex/ms.tex
@@
 % Authors list
-\author{Lionel J. Garcia, Daniel Foreman-Mackey, Fran Pozuelos}
+\author{Lionel J. Garcia}
+\affiliation{Astrobiology Research Unit, Universit\'{e} de Li\`{e}ge, All\'{e}e du 6 Ao\^{u}t 19C, B-4000 Li\`{e}ge, Belgium}
+\author{Daniel Foreman-Mackey}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
+\author{Francisco J. Pozuelos}
+
+\correspondingauthor{Lionel J. Garcia}
+\email{lionel\_garcia@live.fr}
 
@@
 To deal with these nuisance signals, a common approach consist in cleaning a light curve from correlated noise before searching for transits.
-However, we show that this approach, based on an incomplete modelling of light curves, strongly degrades transit signals, up to the point of no detection.
-As this degradation depends on the noise characteristics, we explore the parameter space for which it occurs, and quantify its effect on a wide variety of cases.
+However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signals, up to the point of no detection.
+As this degradation depends on the correlated noise characteristics, we explore the parameter space for which transits are altered, and quantify this effect on a wide variety of cases.
 We show that \nuance{} outperforms the detection capabilities of commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth.
@@
     \begin{centering}
-        \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
+        \includegraphics[height=9cm]{../workflows/cleaning_snr/figures/simu1/result.pdf}
         \caption{\TODO{}, and redo plot with top arrows and fix $tau_v$ $delta_v$}
@@
     \begin{centering}
-        \includegraphics[height=12cm]{../../workflows/synthetic_injection_recovery/figures/final_result.pdf}
+        \includegraphics[height=12cm]{../workflows/synthetic_injection_recovery/figures/final_result.pdf}
         \caption{\TODO}
@@
 \section{Performances and limitations}\label{perf}
+\TODO
 
 \section{Conclusion}
-
-This simulated dataset constits in a light curve featuring a transit signal of depth $1$\%, duration $0.05$ days and period $1.3$ days using the simple \citealt{protopapas} analytical model.
-It also contains a simulated variability signal drawn from a Gaussian Process with a quasiseparable
+\TODO
+\\\\
+We would like to thank Julien De Wit and Prajwal Niraula for meaningful discussions at the begining of the project, Germain Garcia for his useful insights about numerical optimization, and Michaël Gillon for his overall support.
+Co-author Daniel Foreman-Mackey helped in shaping the mathematical formalism of our method and introduced the use of \texttt{JAX} for its tractable implementation.
+Co-author Francisco J.
+Pozuelos was key in validating \texttt{nuance} on real datasets.
 

### 2023-01-27 src/scripts/plot_issue_1.py

@@ -1,132 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import jax, jaxopt
-jax.config.update("jax_enable_x64", True)
-import jax.numpy as jnp
-from tinygp import kernels, GaussianProcess
-import matplotlib.pyplot as plt
-import tinygp
-import paths
-
-def white_red(time, flux, error, plot=False, mean=0.):
-    init = {
-        "log_diag": 2*np.log(np.mean(error)),
-        "log_scale": np.log(1), # 5 mins
-        "log_amp": np.log(np.std(flux))
-    }
-
-    def build_gp(params):
-        kernel = jnp.exp(params["log_amp"]) * kernels.ExpSquared(jnp.exp(params["log_scale"]))
-        return GaussianProcess(kernel, time, diag=jnp.exp(params["log_diag"]), mean=mean)
-
-    @jax.jit
-    def loss(params):
-        gp = build_gp(params)
-        return -gp.log_probability(flux)
-    
-    opt = jaxopt.ScipyMinimize(fun=loss)
-    params = opt.run(init).params  
-    w = np.sqrt(np.exp(params['log_diag']))
-    
-    if plot:
-        gp = build_gp(params)
-        _, cond = gp.condition(y=flux, X_test=time)
-        mu = cond.mean
-        plt.plot(time, flux, ".", c="0.8")
-        plt.plot(time, mu, c="k")
-        noise = w
-        plt.fill_between(time, mu + noise, mu - noise, color="C0", alpha=0.3)
-
-    return w, np.sqrt(np.exp(params['log_amp']))
-
-# Simulation
-# ------------
-from nuance.utils import transit
-np.random.seed(43)
-
-fluxes = []
-snrs = []
-
-# transit
-time = np.linspace(0, 3, 1000)
-dur = 0.1
-depth = 0.8e-2
-n = dur/np.median(np.diff(time))
-signal = transit(time, time.mean(), dur, depth, c=10) + 1.
-
-# white noise
-wsigma = 1.5e-3
-white_noise = np.random.normal(0, wsigma, len(time))
-flux = white_noise + signal
-
-# No systematics
-# --------------
-
[…]

### 2023-02-06 latex/ms.tex
readme title + paper typo
add nuance github link

@@
     We present \nuance{}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
-To deal with these nuisance signals, a common approach consist in cleaning a light curve from correlated noise before searching for transits.
+To deal with these nuisance signals, a common approach consists in cleaning a light curve from correlated noise before searching for transits.
 However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signals, up to the point of no detection.
@@
 We perform our tests by injecting transits on synthetic light curves, and on known TESS candidates light curves to assess the performance of our algorithm on realistic multi-planetary systems datasets.
-Beyond its detection efficiency, we make \nuance{} tractable, \TODO{} times faster than current alternatives based on brute force detrending, and available as a well documented open-source Python package.
+Beyond its detection efficiency, we make \nuance{} tractable, \TODO{} times faster than current alternatives based on brute force detrending, and available as a well documented open-source Python package\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}}.
 \end{abstract}

### 2023-02-11 latex/ms.tex
commit dev
fix proof of periodic likelihood

@@
 This \textit{linear search} lead to the set of likelihoods\footnote{This notation omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{i,j}$) as it is linearly solved for any given value of $(T_i, F_j)$ and irrelevant in what follows.}
-$$\ln\mathcal{L} = \{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})\}_{i, j}$$
+$$\set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})}_{i, j}$$
 were $\Delta_{i,j}$ is the depth linearly solved for a given $(T_i, D_j)$.
@@
 \\\\
-Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$, and with an amplitude $\mathcal{L}_k$, leading to the likelihood:
-$$p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k G_k(\tilde\Delta)$$
-where $G_k$ is the probability density function of $\mathcal{N}(\Delta_k, \sigma_k^2)$. 
+Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$, and with an amplitude $\mathcal{L}_k$, leading to the likelihood function:
+$$p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k\exp \left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right)$$
 \\\\
 As for the common transit depth $\Delta$, it can be estimated through the joint probability of all other transit depths than $\Delta_k$, such that
-$$\Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2) \quad\text{i.e.}\quad \Delta \sim \mathcal{N}(\Delta, \sigma^2) $$
-a normal distribution defined by
+$$\Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2)$$
+with 
 \begin{equation}\label{eq:params}
-\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad
+\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \hspace{0.5cm} \text{and} \hspace{0.5cm}
 \Delta =\sigma^2 \sum_{\sumTk}^K {\frac{\Delta_i}{\sigma_i^2}}
 \end{equation}
-$$p(\tilde\Delta | \Delta) =  \left(\prod_{\sumTk}^K \mathcal{L}_i \right) G(\tilde\Delta)$$
-where $G$ is the probability density function of $\mathcal{N}(\Delta, \sigma^2)$. We can now rewrite \autoref{eq:non_part_of_per} as
+Hence
+$$p(\tilde\Delta | \Delta) = \frac{1}{\sqrt{2\pi\sigma^2}}x\exp \left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)$$
+We can now rewrite \autoref{eq:non_part_of_per} as
 \begin{equation}\label{eq:ice_int}
-    p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i \right)  \int G_k(\tilde\Delta)\, G(\tilde\Delta)\, d\tilde\Delta
+    p(\bm{f} \vert T_k, D, \Delta) =  \frac{\mathcal{L}_k}{\sqrt{2\pi\sigma^2}} \int \exp\left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right)\, \exp\left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)\, d\tilde\Delta
 \end{equation}
-The product of two normal distributions is also a normal distribution, so that the integral in \autoref{eq:ice_int} can be obtained analytically. We have
+The integral in \autoref{eq:ice_int} is a product of gaussian integrals that can be obtained analytically and we have
 \begin{proof}{gaussians_product_integral}
-    p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i  \right) \frac{1}{\sqrt{2\pi \left(\sigma^{2} + \sigma_{k}^{2}\right)}} e^{- \frac{1}{2}\frac{\left(\Delta -  \Delta_{k}\right)^{2}}{\sigma^{2} + \sigma_{k}^{2}}}
+    p(\bm{f} \vert T_k, D, \Delta) = \mathcal{L}_k  \sqrt{\frac{2\pi\sigma_{k}^2}{\sigma^{2} + \sigma_{k}^{2}}} \exp\left(-\frac{1}{2}\frac{(\Delta_k-\Delta)^2}{\sigma_k^2 + \sigma^2}\right)
 \end{proof}
@@
 \begin{equation}\label{eq:result}
-    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
-    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right)
+    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(2\pi\sigma_{k}^2) + \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
+    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}}\right)
 \end{equation}
@@
 In order to reduce the number of times \autoref{eq:params} is computed, we adopt the biased estimates
-$$\frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}}$$
+$$\frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \hspace{0.5cm}\text{and}\hspace{0.5cm} \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}}$$
 so that $\Delta$ and $\sigma$ are independent of $k$ in the last sum of \autoref{eq:result}.

### 2023-02-11 latex/ms.tex
fix again proof of period. likelihood

@@
 Hence
-$$p(\tilde\Delta | \Delta) = \frac{1}{\sqrt{2\pi\sigma^2}}x\exp \left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)$$
+$$p(\tilde\Delta | \Delta) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp \left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)$$
 We can now rewrite \autoref{eq:non_part_of_per} as
@@
 \begin{proof}{gaussians_product_integral}
-    p(\bm{f} \vert T_k, D, \Delta) = \mathcal{L}_k  \sqrt{\frac{2\pi\sigma_{k}^2}{\sigma^{2} + \sigma_{k}^{2}}} \exp\left(-\frac{1}{2}\frac{(\Delta_k-\Delta)^2}{\sigma_k^2 + \sigma^2}\right)
+    p(\bm{f} \vert T_k, D, \Delta) = \mathcal{L}_k  \sqrt{\frac{\sigma_{k}^2}{\sigma^{2} + \sigma_{k}^{2}}} \exp\left(-\frac{1}{2}\frac{(\Delta_k-\Delta)^2}{\sigma_k^2 + \sigma^2}\right)
 \end{proof}
@@
 \begin{equation}\label{eq:result}
-    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(2\pi\sigma_{k}^2) + \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
+    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(\sigma_{k}^2) - \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
     \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}}\right)

### 2023-02-11 latex/ms.tex
update period. likelihood demo in main

@@
     \begin{gathered}
-        \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
-        \bar\Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right) \\
+        \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(\sigma_{k}^2) - \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
+        \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}}\right) \\
         \text{with} \quad  \frac{1}{\sigma^2} = \sum_k^K \frac{1}{\sigma_k^2} \quad \text{and} \quad
@@
 \end{equation}
+
 While this gives a solution for \autoref{eq:goal}, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$.

### 2023-05-01 latex/ms.tex
add better real injections

@@
 \begin{equation}\label{eq:relative_param_space}
-    \tau \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}} \quad \text{and} \quad 
-    \delta \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}} 
+    \tau \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}} \hspace{0.5cm} \text{and} \hspace{0.5cm} 
+    \delta \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}.
 \end{equation}
@@
 \begin{equation}\label{eq:relative_params}
-    \omega = \frac{\pi}{\tau D} \quad 
-    \sigma = \delta \frac{\Delta}{2} \quad  \text{and}  \quad  
-    Q = 10
+    \omega = \frac{\pi}{\tau D}, \hspace{0.5cm} 
+    \sigma = \delta \frac{\Delta}{2} \hspace{0.5cm}  \text{and}  \hspace{0.5cm}  
+    Q = 10,
 \end{equation}
@@
 \newpage
-\section{Validation with injection-recovery}\label{simu}
+\section{Validation on simulated light curves}\label{simu}
 In order to validate our algorithm, we assess the transit search capabilities of \texttt{nuance} against commonly used techniques.
@@
 \newcommand{\wtls}{\texttt{wõtan+TLS}}
-
-\subsection{Comparison with \wtls{} on a simulated light curves}
 
@@
 About these results, we note these transit searches were done in a particularly optimal setup, on light curves not all physically realistic, hence demonstrating the performance of \nuance{} on a theoretical basis only.
-In the next subsection, we perform transits injection-recovery on a real dataset, using more realistic tools and on light curves featuring known planetary transits, hence validating \nuance{} on a realistic multi-planetary dataset.
+In the next section, we perform transits injection-recovery on a real dataset, using more realistic tools and on light curves featuring known planetary transits, hence validating \nuance{} on a realistic multi-planetary dataset.
 
@@
 \end{figure}
+
+\section{Validation on real TESS light curves}\label{real}
+
+The parameter space ($\tau$, $\delta$) presented in is convenient from a signal analysis standpoint, but does not easily translate into physical cases.
+In order to assess the performance of \texttt{nuance} on real light curves, we rather evaluate its transit recovery capabilities in the planetary (period, radius) parameter space, more often encountered in the literature (e.g.
+).
 

### 2023-05-01 workflows/real_injection_recovery/lib/tls.py

@@ -1,0 +1,407 @@
+from transitleastsquares.main import *
+
+
+class tls(transitleastsquares):
+    def power(self, periods, durations=None, **kwargs):
+        """Compute the periodogram for a set of user-defined parameters"""
+        self, kwargs = validate_args(self, kwargs)
+
+        if self.verbose:
+            print(tls_constants.TLS_VERSION)
+
+        if durations is None:
+            durations = duration_grid(
+                periods, shortest=1 / len(self.t), log_step=self.duration_grid_step
+            )
+
+        maxwidth_in_samples = int(numpy.max(durations) * numpy.size(self.y))
+        if maxwidth_in_samples % 2 != 0:
+            maxwidth_in_samples = maxwidth_in_samples + 1
+        lc_cache_overview, lc_arr = get_cache(
+            durations=durations,
+            maxwidth_in_samples=maxwidth_in_samples,
+            per=self.per,
+            rp=self.rp,
+            a=self.a,
+            inc=self.inc,
+            ecc=self.ecc,
+            w=self.w,
+            u=self.u,
+            limb_dark=self.limb_dark,
+            verbose=self.verbose,
+        )
+
+        if self.verbose:
+            print(
+                "Searching "
+                + str(len(self.y))
+                + " data points, "
+                + str(len(periods))
+                + " periods from "
+                + str(round(min(periods), 3))
+                + " to "
+                + str(round(max(periods), 3))
+                + " days"
+            )
+
+        # Python 2 multiprocessing with "partial" doesn't work
+        # For now, only single-threading in Python 2 is supported
+        if sys.version_info[0] < 3:
+            self.use_threads = 1
+            warnings.warn("This TLS version supports no multithreading on Python 2")
+
+        if self.verbose:
+            if self.use_threads == multiprocessing.cpu_count():
+                print("Using all " + str(self.use_threads) + " CPU threads")
+            else:
+                print(
+                    "Using "
+                    + str(self.use_threads)
+                    + " of "
+                    + str(multiprocessing.cpu_count())
+                    + " CPU threads"
+                )
+
+        if self.show_progress_bar:
+            bar_format = "{desc}{percentage:3.0f}%|{bar}| {n_fmt}/{total_fmt} periods | {elapsed}<{remaining}"
+            pbar = tqdm(total=numpy.size(periods), smoothing=0.3, bar_format=bar_format)
+
+        if tls_constants.PERIODS_SEARCH_ORDER == "ascending":
[…]

### 2023-05-01 workflows/real_injection_recovery/notebooks/lib.py

@@ -1,0 +1,119 @@
+from tinygp import kernels
+import jax
+import jaxopt
+import numpy as np
+import jax.numpy as jnp
+from tinygp import kernels, GaussianProcess
+import matplotlib.pyplot as plt
+
+jax.config.update("jax_enable_x64", True)
+
+def plot_kernel(kernel, lag=None, **kwargs):
+    if lag is None:
+        lag = np.linspace(0, 30, 1000)
+    k = kernel(lag, lag[:1])
+    plt.plot(lag, k/k[0], **kwargs)
+    plt.xlabel("dx")
+    plt.ylabel("k(dx)")
+    
+def print_params(params):
+    return {n.split("log_")[-1] if "log" in n else n: f"{np.exp(p):.2e}" if "log" in n else p for n, p in params.items()}
+
+# GP
+def build_gp(params, x):
+    
+    short = kernels.quasisep.Matern32(sigma=jnp.exp(params["log_sigma_short"]), scale=jnp.exp(params["log_scale_short"]))
+    long = kernels.quasisep.Matern32(sigma=jnp.exp(params["log_sigma_long"]), scale=jnp.exp(params["log_scale_long"]))
+    periodic = kernels.quasisep.SHO(omega=jnp.exp(params["log_omega"]), quality=jnp.exp(params["log_quality"]), sigma=jnp.exp(params["log_sigma"]))
+    
+    kernel = periodic + short + long
+    
+    return GaussianProcess(kernel, x, diag=jnp.exp(2 * params["log_jitter"]), mean=0.)
+
+@jax.jit
+def solve(x, y, X, solver):
+    Liy = solver.solve_triangular(y)
+    LiX = solver.solve_triangular(X.T)
+    LiXT = LiX.T
+    w = jax.numpy.linalg.lstsq(LiXT@LiX, LiXT@Liy)[0]
+    v = jax.numpy.linalg.lstsq(LiXT@LiX, jnp.ones_like(w))[0]
+    return w, v
+
+@jax.jit
+def neg_log_likelihood(params, x, y, X):
+    gp = build_gp(params, x)
+    w = solve(x, y, X, gp.solver)[0]
+    return - gp.log_probability(y - w@X)
+
+@jax.jit
+def _eval(params, x, y, X):
+    gp = build_gp(params, x)
+    w, v = solve(x, y, X, gp.solver)
+    return - gp.log_probability(y - w@X), w, v
+    
+def model(x, y, X, mask=None):
[…]

### 2023-05-01 workflows/real_injection_recovery/scripts/clean_flares.py

@@ -1,0 +1,58 @@
+import pickle
+
+import jax
+import matplotlib.pyplot as plt
+import numpy as np
+import yaml
+from nuance import Nuance
+from nuance.kernels import rotation
+
+i = 0
+data_path, gp_path = snakemake.input
+
+data = pickle.load(open(data_path, "rb"))
+gp_params = yaml.full_load(open(gp_path, "r"))
+
+build_gp, init = rotation(data["star_period"], np.mean(data["error"]), long_scale=0.5)
+gp = build_gp(gp_params, data["time"])
+nu = Nuance(data["time"], data["flux"], gp=gp)
+
+
+def mask_flares(time, flux, iterations=3, sigma=4, gp=None):
+    _, mu, _ = nu.gp_optimization(build_gp)
+
+    mask = np.ones_like(time).astype(bool)
+    window = 30
+
+    for _ in range(iterations):
+        m = mu(gp_params)
+        r = flux - m
+        mask_up = r < np.std(r[mask]) * sigma
+
+        # mask around flares
+        ups = np.flatnonzero(~mask_up)
+        if len(ups) > 0:
+            mask[
+                np.hstack(
+                    [
+                        np.arange(max(u - window, 0), min(u + window, len(time)))
+                        for u in ups
+                    ]
+                )
+            ] = False
+        _, mu, _ = nu.gp_optimization(build_gp, mask)
+
+    return ~mask
+
+
+mask = ~mask_flares(data["time"], data["flux"], gp=gp)
+
+# we remove the first 1000 points
+mask[0:1000] = False
+
+cleaned_data = data.copy()
+cleaned_data["flux"] = data["flux"][mask]
+cleaned_data["time"] = data["time"][mask]
+cleaned_data["error"] = data["error"][mask]
+
+pickle.dump(cleaned_data, open(snakemake.output[0], "wb"))

### 2023-05-01 workflows/real_injection_recovery/scripts/concatenate.py

@@ -1,0 +1,52 @@
+import numpy as np
+import pandas as pd
+import yaml
+
+
+def right_candidate(t0, period, true_t0, true_period, verbose=False):
+    t0_check = (
+        np.abs((t0 - true_t0 + 0.5 * true_period) % true_period - 0.5 * true_period)
+        % period
+        < 0.01
+    )
+    period_check = np.abs(period - true_period) < 0.01
+    period_check |= np.abs(2 * period - true_period) < 0.01
+    period_check |= np.abs(period / 2 - true_period) < 0.01
+    same = period_check  # np.logical_and(t0_check, period_check)
+    if verbose:
+        if not same:
+            if not t0_check:
+                output = f"t0 differ: {t0:.2e} {true_t0:.2e}"
+            if not period_check:
+                output = f"period differ: {period:.2e} {true_period:.2e}"
+        else:
+            output = "match"
+        return same, output
+    else:
+        return same
+
+
+def get_result(f, tool):
+    d = yaml.safe_load(open(f, "r"))
+    return {
+        f"{tool}_found_t0": d["found_t0"],
+        f"{tool}_found_period": d["found_period"],
+        f"{tool}_found": right_candidate(
+            d["found_t0"], d["found_period"], d["transit_t0"], d["planet_period"]
+        ),
+    }, d
+
+
+nuance_files = snakemake.input.nuance
+tls_files = snakemake.input.tls
+
+results = []
+
+for i, (nuance_file, tls_file) in enumerate(zip(nuance_files, tls_files)):
+    nuance_results, injected = get_result(nuance_file, "nuance")
+    tls_results, _ = get_result(tls_file, "tls")
+    results.append({**nuance_results, **tls_results, **injected})
+
+results = pd.DataFrame(results)
+
+results.to_csv(snakemake.output[0], index=False)

### 2023-05-01 workflows/real_injection_recovery/scripts/download.py

@@ -1,0 +1,57 @@
+import pickle
+
+import astropy.units as u
+import lightkurve as lk
+import numpy as np
+import pandas as pd
+
+tic = int(snakemake.wildcards["target"])
+filepath = snakemake.output[0]
+
+targets = pd.read_csv("data/tess-ultra-fast-rotators-brighter-mag14.csv")[
+    ["Name", "LS_Period", "Amplitude", "teff_val"]
+]
+
+targets = targets.sort_values("LS_Period")
+target = targets[targets.Name == tic]
+
+name = f"TIC {tic}"
+search_result = lk.search_lightcurve(name, author="SPOC", exptime=120)
+
+# data
+klc = search_result[0].download()
+times = klc.time.to_value("mjd")
+fluxes = klc.pdcsap_flux.to_value().filled(np.nan)
+errors = klc.flux_err.to_value().filled(np.nan)
+masks = [
+    np.isnan(f) | np.isnan(e) | np.isnan(t) for f, e, t in zip(fluxes, errors, times)
+]
+times = [t[~m] for t, m in zip(times, masks)]
+
+# masking
+time = np.hstack(times)
+time -= np.min(time)
+original_flux = np.hstack([f[~m] for f, m in zip(fluxes, masks)])
+original_error = np.hstack([e[~m] for e, m in zip(errors, masks)])
+
+flux_median = np.median(original_flux)
+flux = original_flux / flux_median
+error = original_error / flux_median
+
+# stellar parameters
+url = f"https://exofop.ipac.caltech.edu/tess/download_stellar.php?id={tic}"
+star = pd.read_csv(url, delimiter="|", index_col=1).iloc[0]
+
+pickle.dump(
+    {
+        "flux": flux,
+        "time": time,
+        "error": error,
+        "star_period": float(target.LS_Period),
+        "star_amplitude": float(target.Amplitude),
+        "star_radius": star["Radius (R_Sun)"],
+        "star_mass": star["Mass (M_Sun)"],
+    },
+    open(filepath, "wb"),
+)
+print(f"Downloaded TIC {tic} data (first sector)")

### 2023-05-01 workflows/real_injection_recovery/scripts/inject.py

@@ -1,0 +1,38 @@
+import pickle
+
+import numpy as np
+from nuance.star import Star
+from nuance.utils import periodic_transit
+
+i = int(snakemake.wildcards.lc)
+data_path, params_path = snakemake.input
+data = pickle.load(open(data_path, "rb"))
+periods, radii = pickle.load(open(params_path, "rb"))
+
+star = Star(
+    data["star_radius"], data["star_mass"], data["star_amplitude"], data["star_period"]
+)
+
+np.random.seed(i)
+period = periods[i]
+radius = radii[i]
+tau = star.period2tau(period)
+delta = star.radius2delta(radius)
+duration = star.transit_duration(period)
+depth = star.transit_depth(radius)
+
+t0 = np.random.rand() * period
+
+transit = depth * periodic_transit(data["time"], t0, duration, P=period)
+injected_flux = data["flux"] + transit
+
+data["flux"] = injected_flux
+data["transit_t0"] = t0
+data["transit_duration"] = duration
+data["transit_depth"] = depth
+data["delta"] = delta
+data["tau"] = tau
+data["planet_period"] = period
+data["planet_radius"] = radius
+output = snakemake.output[0]
+pickle.dump(data, open(output, "wb"))

### 2023-05-01 workflows/real_injection_recovery/scripts/optimize_gp.py

@@ -1,0 +1,36 @@
+import pickle
+
+import matplotlib.pyplot as plt
+import numpy as np
+import yaml
+from nuance import Nuance
+from nuance.kernels import rotation
+
+input = snakemake.input[0]
+data = pickle.load(open(input, "rb"))
+
+# only one continuous segment
+dt = np.diff(data["time"])
+mask = data["time"] < data["time"][np.flatnonzero(dt > 0.5)[0]]
+data["flux"] = data["flux"][mask]
+data["time"] = data["time"][mask]
+data["error"] = data["error"][mask]
+# sigma clip
+mask = np.abs(data["flux"] - np.median(data["flux"])) < 3 * np.std(data["flux"])
+data["flux"] = data["flux"][mask]
+data["time"] = data["time"][mask]
+data["error"] = data["error"][mask]
+
+build_gp, init = rotation(data["star_period"], np.mean(data["error"]), long_scale=0.5)
+nu = Nuance(data["time"], data["flux"], data["error"])
+optimize, mu, nll = nu.gp_optimization(build_gp)
+
+new = optimize(
+    init, ["log_sigma", "log_short_scale", "log_short_sigma", "log_long_sigma"]
+)
+new = optimize(new)
+
+gp_output = snakemake.output[0]
+yaml.safe_dump(
+    {name: float(value) for name, value in new.items()}, open(gp_output, "w")
+)

### 2023-05-01 workflows/real_injection_recovery/scripts/params.py

@@ -1,0 +1,35 @@
+import pickle
+
+import numpy as np
+from nuance.star import Star
+
+input = snakemake.input[0]
+data = pickle.load(open(input, "rb"))
+n = snakemake.params["n"]
+
+star = Star(
+    data["star_radius"], data["star_mass"], data["star_amplitude"], data["star_period"]
+)
+
+min_period = 0.2
+max_period = 10.0
+max_snr = 30.0
+min_snr = 4.0
+# -----
+
+dt = np.median(np.diff(data["time"]))
+sigma = np.mean(data["error"])
+radii = np.linspace(0.1, 50, 100000)
+N = len(data["time"])
+
+max_radius = star.min_radius(max_period, max_snr, N, sigma)
+min_radius = star.min_radius(max_period, min_snr, N, sigma)
+
+period_radius = np.array(
+    np.meshgrid(
+        np.linspace(min_period, max_period, n), np.linspace(min_radius, max_radius, n)
+    )
+)
+period_radius = period_radius.reshape(2, n * n)
+output = snakemake.output[0]
+pickle.dump(period_radius, open(output, "wb"))

### 2023-05-01 workflows/real_injection_recovery/scripts/plots/plot_comparison.py

@@ -1,0 +1,92 @@
+import pickle
+
+import pandas as pd
+import yaml
+
+df = pd.read_csv(snakemake.input[0])
+gp_params = yaml.full_load(open(snakemake.input[1], "r"))
+data = pickle.load(open(snakemake.input[2], "rb"))
+tic = snakemake.wildcards.target
+
+
+tls, nuance, radii, periods, tau, delta = (
+    df["tls_found"].values,
+    df["nuance_found"].values,
+    df["planet_radius"].values,
+    df["planet_period"].values,
+    df["tau"].values,
+    df["delta"].values,
+)
+
+import numpy as np
+from nuance.star import Star
+
+jitter = np.exp(gp_params["log_jitter"])
+star = Star(
+    data["star_radius"], data["star_mass"], data["star_amplitude"], data["star_period"]
+)
+
+unique_periods = np.unique(periods)
+min_radii = star.min_radius(unique_periods, 6, len(data["time"]), jitter)
+
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+# setting up ranges
+taus_range = np.min(tau), np.max(tau)
+deltas_range = np.min(delta), np.max(delta)
+periods_range = np.min(periods), np.max(periods)
+radii_range = np.min(radii), np.max(radii)
+extent = (*periods_range, *radii_range)
+
+# figure
+fig = plt.figure(None, (8.5, 3.5))
+sec_color = "0.6"
+
+ax = plt.subplot(122)
+plt.imshow(
+    nuance.reshape(20, 20), extent=extent, origin="lower", aspect="auto", cmap="Greys_r"
+)
+plt.plot(unique_periods, min_radii, label="SNR = 6")
+plt.xlim(periods_range)
+plt.ylim(radii_range)
+plt.xlabel("Period (days)")
+ax.set_title("nuance")
+# secondary axes
+secax = ax.secondary_yaxis("right")
+radii_ticks = ax.get_yticks()
+delta_ticks = star.radius2delta(radii_ticks)
+secax.set_yticks(radii_ticks, [f"{t:.1f}" for t in delta_ticks])
+secax.set_ylabel(r"$\delta$", color=sec_color)
+secax.tick_params(axis="y", colors=sec_color)
+
+secax = ax.secondary_xaxis("top")
+period_ticks = ax.get_xticks()
[…]

### 2023-05-01 workflows/real_injection_recovery/scripts/plots/plot_lc.py

@@ -1,0 +1,42 @@
+import pickle
+
+import jax
+import matplotlib.pyplot as plt
+import numpy as np
+import yaml
+from nuance import Nuance
+from nuance.kernels import rotation
+
+data_path = snakemake.input[0]
+gp_path = snakemake.input[1]
+info_path = snakemake.input[2]
+
+data = pickle.load(open(data_path, "rb"))
+gp_params = yaml.full_load(open(gp_path, "r"))
+info = yaml.full_load(open(info_path, "r"))
+
+build_gp, init = rotation(data["star_period"], np.mean(data["error"]), long_scale=0.5)
+gp = build_gp(gp_params, data["time"])
+
+nu = Nuance(data["time"], data["flux"], gp=gp)
+optimize, mu, nll = nu.gp_optimization(build_gp)
+
+
+plt.figure(None, (8.5, 3))
+plt.xlabel(
+    f"time - {info['first_exposure_time']:.2f} ${info['first_exposure_time_format']}$"
+)
+plt.ylabel("differential flux")
+plt.title(f"TIC {info['tic']} PDCSAP light curve (sector {int(info['sector'])})")
+plt.plot(data["time"], data["flux"], ".", c="0.8", label="cleaned data")
+gp_mean = mu(gp_params)
+idxs = [0, *np.flatnonzero(np.diff(data["time"]) > 10 / 60 / 24)]
+for i in range(len(idxs) - 1):
+    x = data["time"][idxs[i] + 1 : idxs[i + 1]]
+    y = gp_mean[idxs[i] + 1 : idxs[i + 1]]
+    plt.plot(x, y, "k", label="GP mean" if i == 0 else None)
+plt.legend(loc="lower right")
+plt.xlim(0, 3)
+
+plt.tight_layout()
+plt.savefig(snakemake.output[0])

### 2023-05-01 workflows/real_injection_recovery/scripts/tls_search.py

@@ -1,0 +1,59 @@
+import sys
+
+import numpy as np
+from nuance.utils import clean_periods
+
+sys.path.append("lib")
+import pickle
+
+import numpy as np
+import yaml
+from tls import tls
+from wotan import flatten
+
+lc_file = snakemake.input[0]
+periods_file = snakemake.input[1]
+periods = np.load(periods_file)
+data = pickle.load(open(lc_file, "rb"))
+
+
+def trend(time, flux, window_length):
+    return (
+        flatten(time, flux + 1000.0, window_length=window_length, return_trend=True)[1]
+        - 1000.0
+    )
+
+
+def search(time, flux, window_length, periods, verbose=False):
+    flatten_trend = trend(time, flux, window_length)
+    flatten_flux = flux - flatten_trend
+    flatten_flux -= np.mean(flatten_flux)
+    flatten_flux += 1.0
+
+    model = tls(time, flatten_flux, verbose=verbose)
+    results = model.power(
+        periods,
+        verbose=verbose,
+        use_threads=1,
+        show_progress_bar=verbose,
+        durations=[0.01, data["transit_duration"]],
+    )
+
+    return results["T0"], results["period"]
+
+
+t0, period = search(
+    data["time"], data["flux"], 3 * data["transit_duration"], periods, verbose=True
+)
+
+output = snakemake.output[0]
+result = data.copy()
+del result["flux"]
+del result["error"]
+del result["time"]
+result.update(
+    dict(zip(["found_t0", "found_duration", "found_period"], [t0, -1, period]))
+)
+yaml.safe_dump(
+    {name: float(value) for name, value in result.items()}, open(output, "w")
+)

### 2023-05-01 workflows/real_injection_recovery_2/scripts/bspline_tls.py

@@ -1,0 +1,49 @@
+import sys
+
+import numpy as np
+
+sys.path.append("lib")
+import pickle
+
+import numpy as np
+import yaml
+from scipy.interpolate import BSpline, splrep
+from tls import tls
+
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+periods = np.load(snakemake.input.periods)
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+verbose = True
+mask = np.ones_like(data["time"], dtype=bool)
+time, flux, error = data["time"], data["flux"], data["error"]
+
+for i in range(2):
+    tck = splrep(
+        time[mask],
+        flux[mask],
+        w=1 / error[mask],
+    )
+    trend = BSpline(*tck)(time)
+    mask &= np.abs(flux - trend) < 3 * np.std(flux - trend)
+
+trend = BSpline(*tck)(time)
+
+flatten_trend = BSpline(*tck)(time)
+flatten_flux = flux - flatten_trend
+flatten_flux -= np.mean(flatten_flux)
+flatten_flux += 1.0
+
+model = tls(time, flatten_flux, verbose=verbose)
+results = model.power(
+    periods,
+    verbose=verbose,
+    use_threads=1,
+    show_progress_bar=verbose,
+    durations=[0.01, data["transit_duration"]],
+)
+
+t0, period, power = results["T0"], results["period"], results["power"]
+pickle.dump(
+    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
+    open(snakemake.output[0], "wb"),
+)

### 2023-05-01 workflows/real_injection_recovery_2/scripts/inject.py

@@ -1,0 +1,39 @@
+import pickle
+
+import numpy as np
+import yaml
+from nuance.star import Star
+from nuance.utils import periodic_transit
+
+i = int(snakemake.wildcards.lc)
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+periods, radii = pickle.load(open(snakemake.input.params, "rb"))
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+
+star = Star(
+    info["star_radius"], info["star_mass"], info["star_amplitude"], info["star_period"]
+)
+
+np.random.seed(i)
+period = periods[i]
+radius = radii[i]
+tau = star.period2tau(period)
+delta = star.radius2delta(radius)
+duration = star.transit_duration(period)
+depth = star.transit_depth(radius)
+
+t0 = np.random.rand() * period
+
+transit = depth * periodic_transit(data["time"], t0, duration, P=period)
+injected_flux = data["flux"] + transit
+
+data["flux"] = injected_flux
+data["transit_t0"] = t0
+data["transit_duration"] = duration
+data["transit_depth"] = depth
+data["delta"] = delta
+data["tau"] = tau
+data["planet_period"] = period
+data["planet_radius"] = radius
+output = snakemake.output[0]
+pickle.dump(data, open(output, "wb"))

### 2023-05-01 workflows/real_injection_recovery_2/scripts/periods.py

@@ -1,0 +1,11 @@
+import numpy as np
+import yaml
+from nuance.utils import clean_periods
+
+info = yaml.safe_load(open(snakemake.input[0], "r"))
+
+periods = np.linspace(0.3, 11.0, 8000)
+periods = clean_periods(periods, info["star_period"])
+
+output = snakemake.output[0]
+np.save(output, periods)

### 2023-05-01 workflows/real_injection_recovery_2/scripts/plots/plot_comparison.py

@@ -1,0 +1,85 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+from nuance.star import Star
+
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+trend = pickle.load(open(snakemake.input.trend, "rb"))["trend"]
+
+
+methods = {
+    "bspline": r"bspline + TLS",
+    "wotan3D": r"wotan$_{3D}$ + TLS",
+    "harmonics": r"harmonics + TLS",
+    "nuance": r"nuance",
+}
+star = Star(
+    info["star_radius"], info["star_mass"], info["star_amplitude"], info["star_period"]
+)
+sec_color = "0.7"
+
+# grid of plot with ratio 1:3
+fig, axes = plt.subplots(
+    nrows=len(methods), ncols=2, gridspec_kw={"width_ratios": [3, 1]}, figsize=(8.5, 10)
+)
+for i, (name, title) in enumerate(methods.items()):
+    ax = axes[i, 1]
+    df = pd.read_csv(f"../data/140212114/recovered/{name}/results.csv")
+    trend = pickle.load(open(f"../data/140212114/recovered/{name}/0.params", "rb"))[
+        "trend"
+    ]
+    df["found"] = df.apply(
+        lambda row: right_candidate(
+            row["t0"], row["period"], row["true_t0"], row["true_period"]
+        ),
+        axis=1,
+    )
+    radius, period, found, tau, delta = df[
+        ["radius", "true_period", "found", "tau", "delta"]
+    ].values.T
+
+    # setting up ranges
+    taus_range = np.min(tau), np.max(tau)
+    deltas_range = np.min(delta), np.max(delta)
+    periods_range = np.min(periods), np.max(periods)
+    radii_range = np.min(radius), np.max(radius)
+    extent = (*periods_range, *radii_range)
+
+    ax.imshow(
+        found.reshape((10, 10)).astype(bool),
+        extent=[*periods_range, *radii_range],
+        aspect="auto",
+        cmap="Greys_r",
+        origin="lower",
+    )
+
+    secax = ax.secondary_yaxis("right")
+    radii_ticks = ax.get_yticks()
+    delta_ticks = star.radius2delta(radii_ticks)
[…]

### 2023-05-01 workflows/real_injection_recovery_2/scripts/tls_search.py

@@ -1,0 +1,86 @@
+import sys
+
+import numpy as np
+
+sys.path.append("lib")
+import pickle
+
+import numpy as np
+import yaml
+from tls import tls
+from wotan import flatten
+
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+periods = np.load(snakemake.input.periods)
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+
+time, flux = data["time"], data["flux"]
+verbose = True
+
+from functools import partial
+
+# wotan 3D
+# --------
+
+flatten_trend = flatten(
+    time, flux, window_length=data["transit_duration"] * 3, return_trend=True
+)[1]
+flatten_flux = flux - flatten_trend
+flatten_flux -= np.mean(flatten_flux)
+flatten_flux += 1.0
+
+model = tls(time, flatten_flux, verbose=verbose)
+results = model.power(
+    periods,
+    verbose=verbose,
+    use_threads=1,
+    show_progress_bar=verbose,
+    durations=[0.01, data["transit_duration"]],
+)
+
+t0, period, power = results["T0"], results["period"], results["power"]
+pickle.dump(
+    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
+    open(snakemake.output.wotan3D, "wb"),
+)
+
+
+# harmonics
+# ---------
+
+
+# makes an harmonics design matrix of time
+def make_harmonics(time, period, nharmonics=4):
+    # make design matrix
+    X = np.ones((len(time), 2 * nharmonics + 1))
+    X[:, 1] = time
+    for i in range(1, nharmonics):
+        X[:, 2 * i + 1] = np.sin(2 * np.pi * i * time / period)
+        X[:, 2 * i + 2] = np.cos(2 * np.pi * i * time / period)
+
+    return X
+
+
+X = make_harmonics(time, info["star_period"], nharmonics=2)
+# solve for the coefficients
+coeffs = np.linalg.solve(X.T @ X, X.T @ flux)
+# make the model
+flatten_trend = X @ coeffs
+flatten_flux = flux - flatten_trend
+flatten_flux -= np.mean(flatten_flux)
+flatten_flux += 1.0
+
+model = tls(time, flatten_flux, verbose=verbose)
+results = model.power(
+    periods,
+    verbose=verbose,
+    use_threads=1,
+    show_progress_bar=verbose,
+    durations=[0.01, data["transit_duration"]],
+)
+
[…]

### 2023-05-01 workflows/toi540_injection_recovery/scripts/gp.py

@@ -55,7 +55,7 @@ for i in range(3):
     optimize, mu, plot_kernel = nu.gp_optimization(utils.build_gp, mask=mask)
     new_params = optimize(new_params, ["log_sigma", "log_jitter"])
     new_params = optimize(new_params, ["log_quality", "log_sigma", "log_jitter"])
-    new_params = optimize(new_params)
+    # we do not optimize for the period as it is found independently
 
 # Saving
 

### 2023-05-01 workflows/toi540_injection_recovery/scripts/search.py

@@ -13,41 +13,45 @@ params = yaml.full_load(open(snakemake.input[0], "r"))
 df = pd.read_csv(snakemake.input[1])
 
 time, flux, error = df.values.T
-t_0 = time[0]
-time -= t_0
+if len(time) > 0:
+    t_0 = time[0]
+    time -= t_0
 
-X = utils.poly_X(time, 3)
-kernel = utils.build_gp(params, time).kernel
-nu = Nuance(time, flux, kernel=kernel, error=error, X=X)
+    X = utils.poly_X(time, 3)
+    kernel = utils.build_gp(params, time).kernel
+    nu = Nuance(time, flux, kernel=kernel, error=error, X=X)
 
-# Mask flares
-window = 30
-r = flux - np.mean(flux)
-mask = np.array(r < np.std(r) * 4)
-ups = np.flatnonzero(~mask)
-if len(ups) > 0:
-    mask[
-        np.hstack(
-            [np.arange(max(u - window, 0), min(u + window, len(nu.time))) for u in ups]
-        )
-    ] = False
+    # Mask flares
+    # In what Fran Pozuelos sent me, the falres are sigma clipped. To remove their
+    # full extended signal, I locate and mask 30 points on each side of "holes" in
+    # time
+    window = 30
+    mask = np.pad(np.array(np.diff(time) < 3 * np.median(np.diff(time))), (0, 1))
+    ups = np.flatnonzero(~mask)
+    if len(ups) > 0:
+        mask[
+            np.hstack(
+                [
+                    np.arange(max(u - window, 0), min(u + window, len(nu.time)))
+                    for u in ups
+                ]
+            )
+        ] = False
 
-nu = Nuance(time[mask], flux[mask], error[mask], kernel=kernel, X=X[:, mask])
+    nu = Nuance(time[mask], flux[mask], error[mask], kernel=kernel, X=X[:, mask])
 
-# linear search
-t0s = time.copy()
-Ds = np.linspace(0.01, 0.1, 10)
-nu.linear_search(t0s, Ds)
+    # linear search
+    t0s = time.copy()
+    Ds = np.linspace(0.01, 0.07, 6)
+    nu.linear_search(t0s, Ds)
 
-# First search
-periods = np.linspace(0.8, 12.0, 15000)
-search = nu.periodic_search(periods)
+    # We mask directly the first planet signal
+    nu = nu.mask(0.9236153776066658, 0.02, 1.2390692712847524)
 
[…]

### 2023-09-19 latex/paper.tex
feat: major update

\documentclass{aastex631}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{float}

% commands
\newcommand{\nuance}{\texttt{nuance}}
\newcommand{\TODO}{\texttt{todo}}
\newcommand{\set}[1]{\{\,#1\,\}}
\newcommand{\footlink}[1]{\footnote{\url{#1}}}
\newcommand{\wtls}{\texttt{biweight+BLS}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% no indent
\setlength\parindent{0pt}

% code style
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{white},
    commentstyle=\color{gray!70},
    keywordstyle=\color{Bittersweet},
    stringstyle=\color{RoyalBlue},
    basicstyle=\fontsize{7.5}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont,
    breakatwhitespace=false,         
    breaklines=true,
    rulecolor=\color{black!15},
    numbers=none,
    numberstyle=\fontsize{7}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont\color{gray!50},
    framerule=0pt,
    breakindent=5pt,
    resetmargins=true,
    numbersep=10pt,
    frame=single,
    aboveskip=1em,
    belowskip=1em,
    xleftmargin=6pt,
    framexleftmargin=4pt
}
\lstset{style=mystyle}

% figure options
\graphicspath{{../figures}}

\begin{document}

\title{\texttt{nuance}: Detection of planetary transits in the presence of correlated noise}

% author info
\author{Lionel J. Garcia}
\author{Daniel Foreman-Mackey}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
\correspondingauthor{Lionel J. Garcia}
\email{lionel\_garcia@live.fr}

\begin{abstract}
    We present \texttt{nuance}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability. To deal with these nuisance signals, a common approach consists in cleaning a light curve from correlated noise before searching for transits. However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signals, up to the point of no detection. As this degradation depends on the correlated noise characteristics, we explore the parameter space for which transits are altered, and quantify this effect on a wide variety of cases. We show that \nuance{} outperforms the detection capabilities of commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth. We perform our tests by injecting transits on synthetic light curves, and on known TESS candidates light curves to assess the performance of our algorithm on realistic multi-planetary systems datasets. Beyond its detection efficiency, we make \nuance{} tractable, \TODO{} times faster than current alternatives based on brute force detrending, and available as a well documented open-source Python package\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}}.
\end{abstract}

\section*{Introduction}
Transiting exoplanets are keystone objects for the field of exoplanetary science, but detecting transits in light curves featuring stellar variability and instrumental signals remains a challenge \citep{Pont2006,Howell2016}. For this reason, known transiting exoplanets tend to be found around quieter stars, or belong to the population of close-in giants whose transit signals dominate over stellar rotational variability \citep{Simpson2023}. However, discovering transiting exoplanets around active stars holds few promises. First, as younger stars are more active\footnote{approximately down to spectral type MV ()} \citep{Skumanich1972}, being able to detect planets transiting active stars will favor the discovery of young planetary systems \citep[e.g.][]{Newton2022}. Second, as stellar variability may originate from surface active regions (such as starpots), the increased discovery of occulting companions will increase our chances to map the photosphere of active stars \citep[e.g.][]{Morris2017}, benefiting both the study of stellar atmospheres and the concerning impact of their non-uniformity on planetary atmosphere retrievals \citep{rackham2018}. Overall, enabling the detection of transits in light curves with high levels of correlated noises will greatly benefit the study of terrestrial exoplanets around late M-dwarfs, usually observed at lower SNR and more likely to display photometric variability (e.g. \citealt{Murray2020}).
\\\\
Commonly used transit-search algorithms, such as the Box-Least-Square algorithm \citep[BLS,][]{bls} are capable of detecting transits in light curves containing only transit signals and white noise. Using this method, the simplest way to detect transits in a light curve featuring correlated noise (either astrophysical or instrumental), is to first clean it from nuisance signals before performing the search. This strategy is widely adopted by the community, both using physically-motivated systematic models like \cite{everest1, everest2}, or filtering techniques (\citealt{Jenkins2010}, \citealt{wotan}). However, when correlated noise starts resembling transits, this cleaning step (often referred to as \textit{detrending}) is believed to degrade their detectability \cite[see subsection 4.3 of][]{wotan}. In this case, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noise, and to compute the likelihood of the data to the transit model on a wide parameter space (an approach largely avoided due to its intractable nature). Nonetheless, \cite{kovacs2016} ask: \textit{Periodic transit and variability search with simultaneous systematics filtering: Is it worth it?}. The latter study explores a handful of cases and generally discards the benefit of using a full-fledged approach. However, it does not explore the light-curves characteristics for which such a complete modeling becomes necessary.
\\\\
In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, an we present \nuance{}, a method to search for transit signals while modeling correlated noises in a tractable way.

\section{The degradation of transit detectability}\label{issues}

A strong assumption when using BLS-type algorithms is that the searched dataset only contains transit signals and white noise. Hence, two sources of correlated noise particularly justify the need for a detrending step before searching for transits: instrumental signals (such as those induced by telescope pointing errors) and astrophysical signals (such as stellar variability induced by pulsations or starspots). In this section, we qualitatively study the effect of correlated noise and its detrending on transits detectability.

\subsection{Transit detectability}

One way to study the detectability of a unique transit signal is to compute its signal-to-noise-ratio (SNR, \citealt{Pont2006}, Equation 12):
\begin{equation}\label{eq:snr}
  SNR= \frac{\Delta}{\sqrt{\frac{\sigma_w^2}{n} + \frac{\sigma_c^2}{N_{tr}}}}
\end{equation}
where $\Delta$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits ($N_{tr}=1$ here), and $\sigma_w^2$ and $\sigma_c^2$ are the white and correlated noise variances. \autoref{fig:issue1} shows this metric computed for a unique transit observed in the absence (grey) and presence (red) of correlated noise.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=8.5cm]{issue1.pdf}
        \caption{Illustration of the effect of correlated noise on a single transit SNR. A 1-hour transit signal of depth 1\% is generated on top of white noise (variance of $0.0015^2$) as part of a 24-hours observation with an exposure time of 1 minute (top). Then, in the bottom plot, correlated noise is added, generated using a Gaussian Process with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%. The SNR on the right of each light curve is computed using \autoref{eq:snr}. Models used to simulate these data are provided in \autoref{signals_simulations}.}
        \label{fig:issue1}
    \end{centering}
\end{figure}

As illustrated in \autoref{fig:issue1}, the presence of correlated noise strongly decreases the transit signal SNR, which would ultimately limit its detectability.

\subsection{Detrending methods and their effects}\label{detrending_effect}
The presence of instrumental correlated noise motivated the development of systematics detrending algorithms, such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}) or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}). Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal and can be modeled using contemporaneous measurements (like detector's temperature, pointing error, sky background or airmass time series). In opposition, stellar variability and other astrophysical signals cannot be correlated with simultaneous measurements. This gave rise to several treatments in order to reconstruct and detrend stellar variability. One of them is physically-motivated and makes use of Gaussian processes (e.g. \citealt{k2sc}), another is empirical and makes use of filtering algorithms (\citealt{Jenkins2010}, \citealt{wotan}).

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.85\linewidth]{issue2.pdf}
        \caption{(Top left) Simulated observation spanning 3 days, including a transit signal with a depth of 0.8\%, duration of 0.1 day, and measurement error of 1.5\% (using the model from \cite{protopapas} described in \autoref{app_proto}). (Left) Correlated noise in the form of stellar variability with different timescales and amplitudes added to the unique transit signal. This signal is then reconstructed using (in purple) Tukey's bi-weight filter \citep{wotan} with an optimal window size of three times the transit duration, and (in red) a Gaussian Process with the same kernel used to simulate the data. In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr} with the transit depth estimated as the minimum in-transit flux.}
        \label{fig:issue2}
    \end{centering}
\end{figure}

In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a Gaussian Process with an SHO kernel described in \autoref{app_gp}. For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same Gaussian Process from which the data has been sampled. We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.

\noindent \autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that the degradation is strongly dependant on the correlated noise characteristics encountered.\\\\
In order to explore the parameter space for which detrending is the most problematic, we employ the method previously described to simulate 10 000 light curves containing a transit signal and correlated noise in the form of stellar variability. For each light curve, we reconstruct the variability signal and detrend it using an optimal bi-weight filter before estimating the remaining transit SNR. For convenience, the simulated light curves share a common transit added on top of white noise, and a variability signal only defined by two parameters: $\tau$, the relative timescale of the variability with respect to the transit duration; and $\delta$, the relative amplitude of the variability against transit depth, corresponding to an SHO kernel with hyperparameters chosen as: 
\begin{equation}\label{eq:relative_params}
    \omega = \frac{\pi}{\tau D}, \hspace{0.5cm} 
    \sigma = \delta \frac{\Delta}{2} \hspace{0.5cm}  \text{and}  \hspace{0.5cm}  
    Q = 10,
\end{equation}
with $\Delta=1\%$ and $D=0.04$ days, the depth and duration of the simulated transit. 
% We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability signals.

For $\tau=1$ and $\delta=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a variability signal with a period half that of the transit duration, and a standard deviation two times that of the transit depth, i.e. strongly resembling the simulated transit signal.

\autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 6$). Hence, bi-weight filter detrending makes transit-search blind to many systems. While this study should be extended to other detrending techniques, it highlights the need for a more informed transit search algorithm able to deal with correlated noise.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.7\linewidth]{../workflows/cleaning_snr/figures/result.pdf}
        \caption{SNR of a unique transit after detrending light curves using an optimal bi-weight filter. All light curves correspond to a 2.8 days observation with a cadence of 2 minutes, and contain a unique transit of duration 1 hour and a depth of 1\%, added on top of white noise with a standard deviation of 0.5\% plus stellar variability. Light curves at the top and right side of the central plot are shown with their corresponding $\tau$ and $\delta$ values. The amplitude of the variability increases with $\delta$ and the timescale of the variability increases with $\tau$.}
        \label{fig:snr_detrend}
    \end{centering}
\end{figure}

\newpage
\section{\textsf{nuance}}\label{nuance}

\textsf{nuance} is an algorithm capable of searching for planetary transits in light curves containing correlated noise, such as instrumental signals and photometric stellar variability.
\\\\
Let assume that the flux $f$ of a star is observed and arranged in the vector $\bm{f}$ of size $N$, associated to the vector of times $\bm{t}$. This flux, shown in \autoref{fig:linear_search}, contains instrumental signals, stellar variability and a periodic transit signal that we wish to uncover. We assume that a set of $M$ observed measurements (such as the position of the star on the detector or the sky background) taken at the same time as the flux can be treated as explanatory variables for $\bm{f}$. These measurements constitute the columns of the $(N\times M)$ \textit{design matrix} $\bm{X}$.\\\\

Ideally, we would detect the periodic transit signal in this flux by sampling the posterior likelihood of this data to a full-fledge model including stellar variability (more generally correlated noise), instrumental systematic signals (modeled with explanatory variables or treated as correlate noise), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$. We would then reduce the posterior likelihood to $p(\bm{f}\vert P)$, its marginalized version over all parameters except the period $P$, producing a transit search periodogram $\mathcal{Q}(P)$. However, this approach has two issues: It is highly intractable, and it may lead to multimodal distributions that are hard to interpret.
\\\\
Given a period $P$, we instead want to compute the likelihood of a periodic transit signal at the maximum likelihood parameters $\hat T_0$, $\hat D$ and $\hat \Delta$, i.e the periodogram
\begin{equation}\label{eq:periodogram}
        \mathcal{Q}(P) = p(\bm{f} \vert P, \hat T_0 ,\hat D, \hat \Delta)
\end{equation}
This is done by adopting the strategy of \cite{foreman2016}, and separate the transit search into two components: the linear search and the periodic search. During the linear search, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths. Then, the periodic search consists in combining these likelihoods to compute the likelihood of the data given a periodic transit signal for a range of periods. These combined likelihoods yield a transit-search periodogram on which the periodic transit detection is based. \textsf{nuance} differs from \cite{foreman2016} and other existing transit search algorithms as it models the covariance of the light curve with a Gaussian Process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable. This way, \textsf{nuance} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR.\\\\
We note that the approach employed by \textsf{nuance} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.

\subsection{The linear search}\label{linear_search}

During the linear search, the goal is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of the data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths.
\\\\
To account for correlated noise, the light curve $f$ is modeled as being drawn from a Gaussian Process, like the simulated light curves used in \autoref{issues}, such that
\begin{equation*}
    \bm{f} \sim \mathcal{N}(\bm{X w}, \bm{\Sigma}),
\end{equation*}
with mean $\bm{Xw}$ (i.e. a linear model of the M explanatory variables with coefficients $\bm{w}$) and covariance $\bm{\Sigma}$. To account for the presence of a single non-periodic transit of epoch $T$ and duration $D$, this signal is computed and appended as the last column of the design matrix $\bm{X}$, using the simple transit model from \cite{protopapas} with a unitary depth (\autoref{eq:protopapas}). This way, the transit signal is part of the linear model and its depth $\Delta$ can be solved linearly. Under this assumption, the log-likelihood of the data given a single non-periodic transit is \citep{Rasmussen2005}
\begin{equation} \label{eq:linear_search_ll}
    \ln p(\bm{f} \vert I) = -\frac{1}{2}(\bm{f}-\bm{Xw})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{Xw}) -  \frac{1}{2}\ln\vert\bm{\Sigma}\vert - \frac{N}{2}\ln 2\pi,
\end{equation}
where the parameters vector $\bm{w}$ and their errors $\bm{\sigma}$ are computed using the generalized least-square solution
\begin{equation}\label{eq:ls}
    \bm{w} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}\bm{X}^T\bm{\Sigma}^{-1}\bm{f} \hspace{0.5cm} \text{and} \hspace{0.5cm} \bm{\sigma} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1},
\end{equation} 
with $\bm{\Sigma}$ the covariance matrix modeled using the Gaussian Process. Hence, $\ln p(\bm{f} \vert I)$ can be computed on a grid of epochs and durations, the transit depth being linearly solved for any $(T, D)$. In this equation. This linear search leads to the set of likelihoods
\begin{equation*}
    \set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})}_{i, j},
\end{equation*}
were $\Delta_{i,j}$ is the depth linearly solved for a given $(T_i, D_j)$\footnote{The expression of $\set{\ln\mathcal{L}_{i,j}}_{i, j}$ omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{i,j}$) as it is also linearly solved for any given value of $(T_i, D_j)$ and irrelevant in what follows.}. \autoref{fig:linear_search} shows this likelihood grid computed for a simulated dataset, using the same Gaussian Process  and design matrix $\bm{X}$ used to simulate the data.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{principle_linear_search.pdf}
        \caption{Principle and output of the linear search. The simulated dataset (top) corresponds to the one shown and described in \autoref{fig:app_principle_dataset}. First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated. For each pair of indices $(i,j)$, the likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}. This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly using \autoref{eq:ls}.}
        \label{fig:linear_search}
    \end{centering}
\end{figure}

\noindent To prepare for the next step, the corresponding depths $\Delta_{i,j}$ linearly solved for any $(T_i ,D_j)$ are stored, as well as their associated uncertainties $\sigma_{i,j}$, corresponding to
\begin{equation*}
    \Delta_{i,j} = \bm{w}_M \hspace{0.5cm}\text{and}\hspace{0.5cm} \sigma_{i,j} = \bm{\sigma}_{MM},
\end{equation*}
M denoting the index of the last column of the design matrix $\bm{X}$, where the transit signal is contained.

\subsection{The periodic search}

We then need to combine the likelihoods computed from the linear search to obtain
\begin{equation*}
    p(\bm{f} \vert P, T_0 , D, \Delta),
\end{equation*}
i.e.\,the probability of a periodic transit of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$ given the data $\bm{f}$. For a given transit duration $D$, any combination of $(P, T_0)$ leads to K transits, for which it is tempting to write
\begin{equation}\label{eq:attempt}
    p(\bm{f} \vert P, T_0 ,D, \Delta) = \prod_k^K p(\bm{f} \vert T_k, D, \Delta_k),
\end{equation}
where $\set{T_k}_k$ are the epochs matching $(T_0, P)$ and $\set{\Delta_k}_k$ the corresponding depths. So that
\begin{equation*}
    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) = \sum_k^K \ln \mathcal{L}_k.
\end{equation*}
This is the joint likelihood of transits belonging to the same periodic signal but with varying depths  $\set{\Delta_k}_k$. However, individual transits from a periodic signal cannot be considered independent, and should instead be found periodically and share a common transit depth $\Delta$. To this end, it can be shown (see \autoref{proof}) that there is an analytical expression for the joint likelihood of K individual transits with depths and errors $\set{\Delta_k, \sigma_k}_k$ assuming a common depth $\Delta$, corresponding to 
\begin{equation}\label{eq:result}
    \begin{gathered}
        \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(\sigma_{k}^2) - \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
        \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}}\right) \\
        \text{with} \quad  \frac{1}{\sigma^2} = \sum_k^K \frac{1}{\sigma_k^2} \quad \text{and} \quad
        \Delta = \sigma^2 \sum_k^K {\frac{\Delta_k}{\sigma_k^2}}.
    \end{gathered}
\end{equation}

While \autoref{eq:result} takes a closed form, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$. In \cite{foreman2016}, a similar issue is solved by using the nearest neighbors in the epochs grid. Instead, to allow the efficient matrix computation of \autoref{eq:result}, the likelihood grid is linearly interpolated from $\set{T_i}_i$ to a common grid of transit phases $\set{\phi_i}_i$, leading to the periodic search log-likelihood
$$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi, D)}_{i,j}$$
shown for few periods in \autoref{fig:periodic_search} (b). In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the linear search using $\phi_i$, $D_j$ and $T_0 = 0$.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{principle_periodic_0.pdf}\\
    {(a) $\ln \mathcal{L}$}

    \centering
    \includegraphics[width=0.8\linewidth]{principle_periodic_1.pdf}\\
    {(b) $\mathcal{P}(P)$}
\end{figure}

\begin{figure}[H]
    \caption{Applied to the dataset shown \autoref{fig:linear_search}, this figure shows how the periodic search works at different periods $P$ (including the true period $P = 0.7$ days). Given different periods $P$ and $T_0=0$, the likelihood $\ln\mathcal{L}$ shown in (a) is phase-folded and interpolated onto a common grid of phases shown in (b). As an example, the white lines in (a) mark the edges of each fold for a period of $P=0.5$ days, and the white crosses show the epochs $\set{T_k}_{k\in\mathbb{K}}$ matching a particular phase in the grid (reported in (a) and (b)) on which the corresponding $\set{\ln \mathcal{L}_k}_{k\in\mathbb{K}}$ are interpolated and combined using \autoref{eq:result}. To allow the use of efficient matrix computations, this is done for all durations $\set{D_i}_i$ so that $\mathcal{P}$ is computed on the full grid $\set{D_i, \phi_i}_{i, j}$ at once. We understand from the folded likelihoods plots in (b) that a different choice of epoch $T_0$ may only shift the results in phase but do not affect the values of $\mathcal{P}$. For this reason, computing $\mathcal{P}$ for $T_0=0$ is sufficient. We also notice how the maximum value of $\mathcal{P}(P=0.7/2)$ (left plot of (b)) is lower than for $P=0.7$ days, a result of combining the log-likelihoods using \autoref{eq:result} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
    \label{fig:periodic_search}
\end{figure}

\subsection{The transit search periodogram}

Using \autoref{eq:result}, we can now compute $\ln\mathcal{P}$ for a range of periods and build a transit search periodogram using \autoref{eq:periodogram}. But a final issue emerges, one that is fundamentally linked to our strategy. Each likelihood $p(\bm{f} \vert T, D, \Delta)$ estimated during the linear search is computed using N measurements. Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving $N\times K$ measurements. This lead to a normalization issue when trying to compare the joint log-likelihoods $\mathcal{P}(P)$ from one period to another, as the number of observed transits differs from one period to another. This motivates a final step to produce the transit search periodogram $\mathcal{Q}$. For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, the maximum likelihood parameters
\begin{equation}\label{eq:phi0}
    (\phi_0 ,D) = \argmax_{\phi_i, D_j} \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i, j}
\end{equation}
are retrieved, and $\mathcal{Q}(P)$ is defined as the SNR of the transit of period $P$, epoch $T_0 = \phi_0 P$, duration $D$ and depth $\Delta$, i.e.
\begin{equation*}
    \mathcal{Q}(P) = \frac{\Delta}{\sigma},
\end{equation*}
where $\Delta$ and $\sigma$ are obtained using \autoref{eq:sigma_delta} with the last column of $X$ containing a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $1$. This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.\\\\
The periodic transit of period $P$ with the maximum SNR, i.e. maximizing $\mathcal{Q}$, is adopted as the best candidate, basing the confidence in this signal through its SNR. The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (given by \autoref{eq:sigma_delta}).

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{principle_Q.pdf}
        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
        \label{fig:periodogram}
    \end{centering}
\end{figure}

\subsection{An open-source python package}
Methods presented in this paper are made accessible through the \textsf{nuance} open-source Python package, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} and released on the Python Package Index (PyPI)\footlink{https://pypi.org/project/nuance/}. 
\\\\
To instantiate a search, a user can start by creating a \textsf{Nuance} object with
\begin{lstlisting}[language=Python]
from nuance import Nuance

nu = Nuance(time, flux, gp=gp, X=X)
\end{lstlisting}
where \textsf{gp} is a \textsf{tinygp} Gaussian Process instance and \textsf{X} the design matrix of the linear model. \textsf{nuance} exploits the use of \textsf{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package powered by \textsf{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}, allowing for custom kernels to be built and highly tractable computations. We can then define a set of epochs \textsf{t0s} and durations \textsf{Ds} and run the linear search with
\begin{lstlisting}[language=Python,linewidth=\linewidth]
import numpy as np

t0s = time.copy()
# a range of 10 durations
Ds = np.linspace(0.01, 0.2, 10)
nu.linear_search(t0s, Ds)
\end{lstlisting}
Finally, the periodic search is run with
\begin{lstlisting}[language=Python]
# a wide range of periods
periods = np.linspace(0.1, 5, 2000)
search = nu.periodic_search(periods)
\end{lstlisting}
From this \texttt{search} object, the best transiting candidate parameters can be computed (\textsf{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\textsf{search.Q\_snr}), together with valuable information about the transit search. The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \textsf{nuance} being that the linear search only needs to be performed ones, and reused for the search of several transiting candidates. An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.

\section{Control test against BLS}\label{control}

To start testing \textsf{nuance} against existing methods, a simple light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ (relative flux), observed for a duration of 6 days with a cadence of 2 minutes.
In this signal, box-shaped transits are injected with periods randomly sampled from 0.3 to 2.5 days, a unique transit duration of 50 minutes, and depths randomly sampled from $2.6 \times 10^{-4}$ to $6.8 \times 10^{-4}$ (using the simple model from \cite{protopapas} presented in \autoref{app_proto} with $c=500$). By design, these injected signals have an SNR ranging from 5 to 30 (with $\sigma_r = 0$ in \autoref{eq:snr}). For each light curve, transits are searched using two different methods: \textsf{nuance}, using its implementation from the Python package described in the previous section; and \texttt{BLS}, the box-least-square algorithm \citep{bls}, using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}. For \nuance{}\\\\
For both methods, 3000 trial periods from 0.2 to 2.6 days are searched, with a single trial duration fixed to the unique known duration of 50 minutes. A transit signal is considered detected if the absolute difference between the injected and the recovered period is less than 0.01 day. To ease the detection criteria, transit periods recovered at half or twice the injected period (aka \textit{aliases}) are considered as being detected. For this reason, detected transit epochs are not considered. Results from this \textit{injection-recovery} procedure are shown in \autoref{fig:control}.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.85\linewidth]{../workflows/sanity/figures/control_test.pdf}
        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise (whose characteristics are described in the main text), using \texttt{BLS} and \textsf{nuance}. The color scale indicates the recovery of transits in the corresponding (period, depth) parameters space, white for a full recovery and black for no detection.}
        \label{fig:control}
    \end{centering}
\end{figure}

These results demonstrate the qualitative match between the detection capabilities of \textsf{nuance} and \texttt{BLS} on light curves with no correlated noise. Explaining the subtle differences observed between the two methods when only white noise is present is beyond the scope of this paper, and we will assume that any differences observed in the following sections are due to the different treatments of correlated noise.

\section{\textsf{nuance} \textit{vs.} \texttt{biweight+BLS} on simulated light curves}\label{simu}
\autoref{fig:snr_detrend} shows that \textsf{nuance}'s full-fledged modeling capabilities may not always be necessary and may only be beneficial for certain noise characteristics, relative to the searched transit parameters. Here, we evaluate the performance of \textsf{nuance} in the relative parameter space $(\tau, \delta)$ described in \autoref{eq:relative_params} and see when its specific treatment of correlated noise in the transit search becomes necessary.\\\\
We compare \textsf{nuance} to the approach that involves removing stellar variability from the light curves before performing the search on a detrended dataset. Similar to \autoref{detrending_effect}, an optimal bi-weight filter implemented in the \textsf{wõtan} Python package\footlink{https://github.com/hippke/wotan} is used with a window size three times that of the injected transit duration. A transit search on the detrended light curve is then performed using the \texttt{BLS} algorithm, like in \autoref{control} using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}. This strategy is widely used in the community and is denoted \texttt{biweight+BLS}. In what follows, the transit detection criteria are the same as the ones used in \autoref{control}, i.e. that a transit is considered recovered if the absolute difference between the injected and recovered period is less than 0.01 days, including periods found at half or twice the injected ones (aliases) and ignoring the value of the recovered transit epoch.

%% STOPPED HERE

\subsection{Dataset}\label{sim_dataset}
The dataset consists in 4000 light curves simulated using the model fully described in \autoref{signals_simulations}. We simulate a common periodic transit added to all light curves, of period $P=1.1$ days, epoch $T_0=0.2$ days, duration $D=0.04$ days and depth $\Delta=1\%$. Each light curve consists in a 4 days observation with an exposure time of 2 minutes, leading to $N=2880$ data points with a normal error of $0.1\%$. As in \autoref{control}, only box-shaped transits are injected, using the model from \cite{protopapas} with $c=500$.
\\\\
For a given pair of $(\tau, \delta)$, we simulate stellar variability using a Gaussian Process with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params} (except for $Q$, described below), computed with respect to the injected transit parameters $D$ and $\Delta$. The same kernel is used for the search with \textsf{nuance}, an optimal choice on equal footing with the optimal $3\times D$ window size of the bi-weight filter employed in the \wtls{} search. While this kernel is optimal, a comprehensive discussion on the use of non-optimal kernels is discussed in . 4000 pairs of $(\tau, \delta)$ are generated such that
\begin{equation*}
    \tau \sim \mathcal{U}(0.1, 10), \hspace{0.5cm} \delta \sim \mathcal{U}(0.1, 25) \hspace{0.5cm}\text{and}\hspace{0.5cm} Q \sim \mathcal{U}(10, 100)
\end{equation*}
where $\mathcal{U}(a, b)$ denotes a uniform distribution of lower bound $a$ and upper bound $b$.

\subsection{Results}
The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against the \wtls{} strategy on transits with relatively low depth compared to the stellar variability amplitude, and a relatively small duration compared to the stellar variability period. This empirical statement only concerns light curves with a given amount of white noise, and may vary depending on the length of the observing window or the number of transits. For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperforms \wtls{} would only apply to this specific example. While more examples are considered in the following \autoref{real}, proper guidelines on when to use \textsf{nuance} instead of more classic detrending techniques are discussed in \autoref{when}.
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.85\linewidth]{../workflows/synthetic-injection-recovery/figures/synthetic_ir.pdf}
        \caption{Injection-recovery results on 4000 simulated light curves described in \autoref{sim_dataset}. The color scale represents the fraction of recovered transits, white if all injected transits are recovered in a given portion of the $(\tau, \delta)$ parameter space, black if none are recovered.}
        \label{fig:simu}
    \end{centering}
\end{figure}

\noindent This injection-recovery is done in a particularly optimal setup, on light curves not all physically realistic and using an optimal Gaussian Process kernel, hence demonstrating the performance of \textsf{nuance} only on a purely experimental basis. In the next section, we perform transits injection-recovery on real space-based light curves.

\section{Validation on TESS light curves}\label{real}

The parameter space ($\tau$, $\delta$) introduced in \autoref{detrending_effect} is very convenient from a signal processing point of view, but lacks physical reality. Very low values of $\tau$ and $\delta$ often translate into non-physical exoplanets properties as well as unrealistic stellar variability, with a strong dependence on the host star properties. \\\\
In order to assess the performance of \textsf{nuance} on real light curves, its transit recovery capabilities are rather evaluated in the planetary (period, radius) parameter space, more often encountered in the literature (e.g. Figure 12 of \citealt{Delrez2022}). In what follows, transits described by the simple model of \cite{protopapas} (see \autoref{app_proto}) are injected into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}). We focus this proof of concept on light curves from a list of $\sim$600 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{} (see \autoref{when}). For each of the studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) on a single sector (the first being observed for each target) spanning an average of 27 days.\\\\
In the following sections, \textsf{nuance} is compared to other transit search strategies employing the \texttt{BLS} algorithm, each time after detrending the light curves with three different techniques:
\begin{itemize}
    \item \texttt{bspline+BLS} employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the \texttt{BLS} algorithm.
    \item \texttt{biweight+BLS} employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package \citep{wotan} with an optimal window size of $3\times D$ (i.e.\, three times the transit duration), followed by a search with the \texttt{BLS} algorithm.
    \item \texttt{harmonics+BLS} employs a linear harmonic detrending, where the light curve is modeled as a Fourier series including four harmonics of the stellar rotation period found by \cite{Ramsay2020}, with coefficients found through ordinary least square. This detrending is followed by a search with the \texttt{BLS} algorithm.
\end{itemize}

\subsection{Stellar variability kernel}
A Gaussian Process kernel with a single SHO kernel is not representative of realistic stellar variability. Instead, a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), being representative of a wide range of stochastic variability in stellar time series from rotation to pulsations\footlink{https://celerite2.readthedocs.io/en/latest/api/python}, is adopted, with the addition of one short scale and one long term exponential term. This full rotation kernel is expressed as
\begin{equation*}
    k = k_1 + k_2 + k_3 + k_4
\end{equation*}
with
\begin{itemize}
    \item $k_1$ a SHO kernel with hyperparameters \begin{equation*}
        Q_1 = 1/2 + Q_0 + \delta Q\,, \hspace{0.5cm}
        \omega_1 = \frac{4\,\pi\,Q_1}{P\,\sqrt{4\,Q_1^2 - 1}} \hspace{0.5cm} \text{and} \hspace{0.5cm}
        S_1 = \frac{\sigma^2}{(1 + f)\,\omega_1\,Q_1}.
    \end{equation*}
    \item $k_2$ a SHO kernel with hyperparameters \begin{equation*}\begin{gathered}
        Q_2 = 1/2 + Q_0\,, \hspace{0.5cm}
        \omega_2 = 2 \omega_1 \hspace{0.5cm} \text{and} \hspace{0.5cm}
        S_2 = \frac{f\,\sigma^2}{(1 + f)\,\omega_2\,Q_2},
    \end{gathered}\end{equation*}
\end{itemize}
where $Q_0$ is the quality factor for the secondary oscillation, $\delta Q$ is the difference between the quality factors of the first and the second modes, $f$ is the fractional amplitude of the secondary mode compared to the primary and $\sigma$ is the standard deviation of the process. The kernels $k_3$ and $k_4$ are expressed as 
\begin{equation*}
        k(t, t')=\sigma^2\,\exp\left(-\frac{\vert t - t' \vert}{\ell}\right),
\end{equation*}
with $\ell$ and $\sigma$ the scale and standard deviation of the process. These are meant to model short and long timescale non-periodic correlated noise In total, the rotation kernel $k$ has 8 hyperparameters.

\subsection{Light curve cleaning}
As the described light curve model does not account for stellar flares, the light curve of each target is cleaned using an iterative sigma clipping approach. For each iteration, points 4 times above the standard deviation of the full light curve (previously subtracted by its median) are identified. Then, the 30 adjacent points on each side of the found outliers are masked. This way, large flare signals and their expected ingress and egress are masked, using a total of 3 iterations. As PDCSAP light curves often start with a ramp-like signal, the first 300 points (as well as the last 300 points) of each continuous observation are masked. Finally, each light curve is normalized by its median value. Such a cleaned light curve is shown in \autoref{fig:cleaned}.
\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{153193474_clean.png}
    \caption{Cleaned single-sector light curve of the target TIC 153193474. The light curve is normalized by its median value.}
    \label{fig:cleaned}
\end{figure}

\subsection{Injected transits}
For each target light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injections and recovery per target. Orbital periods $P$ are sampled on a regular grid between 0.4 and 10 days, and planetary radii $R_p$ are sampled on a regular grid designed to lead a minimum transit SNR of 4 and a maximum of 30. Using \autoref{eq:snr} with $\sigma_r = 0$, the planetary radius leading to a transit with a desired SNR $s$ is given by
\begin{equation*}
    R_p = R_{\star}\,n^{-\frac{1}{4}} \sqrt{\sigma s}\ %\sqrt[\leftroot{0}\uproot{2}4]{n}}. 
\end{equation*}
with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.

\subsection{Gaussian Process hyperparameters optimization}
The hyperparameters of the employed rotation kernel are optimized on cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize}\footlink{https://jax.readthedocs.io/en/latest/_autosummary/jax.scipy.optimize.minimize.html} wrapper provided by the \textsf{jaxopt} Python package, taking advantage of the \textsf{JAX} implementation of \textsf{tinygp}. As correlated noise is expected to affect the light curve uncertainty estimates performed by SPOC, the diagonal of the full covariance matrix of the data (i.e.\,their uncertainty, assuming homoscedasticity) is held free, increasing the number of optimized parameters to 9. The optimization is performed using the \textsf{BFGS} algorithm \citep{Fletcher1987}, minimizing the negative log-likelihood of the data as expressed in \autoref{eq:linear_search_ll} (without transit), i.e.\,accounting for a linear systematic model of the data in addition of stellar variability. For simplicity, and to adopt a uniform treatment for all target light curves, a design matrix $\bm{X}$ with a single constant column is adopted, such that the systematic model only consists in a single parameter corresponding to the mean value of the differential flux (expected to be close to 1) solved linearly.

\subsection{Detection criteria}
Because we also consider planets recovered with half or twice the real period, we ease the detection criteria by ignoring the match of the recovered transit epoch with the real epoch. Note that a visual vetting ensured that the found epochs were consistent with the ones injected for cases where the planet was considered detected.

\subsection{Injection-recovery studies}
\subsubsection{TIC 140212114: fast sinusoidal variability}\label{140212114}
TIC 140212114 has an effective temperature of 3595 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=140212114} (M1.5-type star, \citealt{Rajpurohit2013}) with a rotation period of 0.14 days \citep{Ramsay2020}. For each technique, the left plot of \autoref{fig:140212114} shows a portion of TIC 140212114 cleaned light curve with an injected transit signal (hardly distinguishable due to correlated noise of the light curve compared to the transit depth). On these plots the black lines correspond to the trends computed for this injected light curve subtracted from it before being searched for transits. This detrending only applies to techniques other than \textsf{nuance}, where instead the black line corresponds to the mean of the Gaussian Process model. For each technique, the right plot of \autoref{fig:140212114} shows the results of the transit search for each of the 10 $\times$ 10 injected light curves, reported in the planetary (period, radius) parameter space. A black square denotes that the transit signal has not been detected, a gray square denotes a signal detected at an alias period ($P/2$ or $2P$), and a white square denotes a transit signal detected with the correct period. On the right plots of \autoref{fig:140212114}, secondary axes show the values of the injected transits in the $(\tau, \delta)$ relative parameter space.\\\\
Results from \autoref{fig:140212114} are discussed in \autoref{tess_results}.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{140212114.pdf}
        \caption{Results of the transits injection-recovery on TIC 140212114 single-sector light curve. Left: cleaned light curve with computed trend overplotted in black (except for \textsf{nuance} where it corresponds to the mean of the Gaussian Process model). Right: Results of the transit search where a black square denotes a transit signal not detected, gray a signal detected at an alias period ($P/2$ or $2P$), and  white a signal detected with the correct period. On the right plots, secondary axes show the $(\tau, \delta)$ relative parameter space. For each method, the upper right legend on the left plot indicates its ranking based on the percent of recovered transit signals (where a transit with an aliased period counts as being detected). The \texttt{harmonics+TLS} method is the best after \textsf{nuance}, but with 35\% less transits detected in total. 
        }
        \label{fig:140212114}
    \end{centering}
\end{figure}

\subsubsection*{TIC 452793374: high-amplitude red noise}
TIC 452793374 has an effective temperature of 3377 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=452793374} (M2.5-type star, \citealt{Rajpurohit2013}) with a rotation period of 0.33 days \citep{Ramsay2020}. Injection-recovery results are shown in \autoref{fig:452793374} and discussed in \autoref{tess_results}.

\subsubsection*{TIC 153193474: non-sinusoidal variability}
TIC 153193474 has an effective temperature of 3351 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=27734884} (M2.5-type star, \citealt{Rajpurohit2013}) with a rotation period of 0.11 days \citep{Ramsay2020}. Injection-recovery results are shown in \autoref{fig:153193474} and discussed in \autoref{tess_results}.

\subsubsection*{TIC 306331621: slow sinusoidal variability}
TIC 306331621 has an effective temperature of 3067 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=306331621} (M5-type star, \citealt{Rajpurohit2013}) with a rotation period of 1.00 days \citep{Ramsay2020}. It was chosen for its relatively slow rotation (compared to the three other targets) and higher white noise compared to the amplitude of its variability. Injection-recovery results are shown in \autoref{fig:306331621} and discussed in \autoref{tess_results}.

\subsection{Results}\label{tess_results}

From the injection-recovery conducted on TIC 140212114, TIC 452793374, TIC 153193474 and TIC 306331621, it can be concluded that searching for transits directly with \textsf{nuance} offers a clear advantage over any of the presented detrending methods combined with \texttt{BLS}. While not fully described in this dissertation, injection-recovery on additional targets lead to a similar conclusion.\\\\
Targets with light curves featuring stable sinusoidal stellar variability, i.e.\,with a high quality factor $Q$ (see the SHO kernel), are the ones that benefit less from \textsf{nuance}, as a set of harmonic signals can be combined linearly to robustly model and detrend the periodic signal. In the cases of TIC 452793374 and TIC 306331621, an additional $\sim$6\% of planets are detected by \textsf{nuance} over the probed planetary (period, radius) parameter space. From these two examples, it cannot be concluded if this difference is solely due to the difference observed between \textsf{nuance} and \texttt{BLS} in \autoref{control} . While red noise was properly detrended by a bi-weight filter in the case of TIC 452793374, additional tests must be done to show that \textsf{nuance} clearly outperforms this method when any red noise is present in the light curve (as partially observed in \autoref{simu}).
\\\\
Finally, targets with non-sinusoidal or short scale variability are the ones that benefit the most from \textsf{nuance}, as no other method is able to robustly detrend their light curves  without removing transit signals. This is well illustrated by the study of TIC 140212114 (\autoref{fig:140212114}) and TIC 153193474 (\autoref{fig:153193474}).
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{452793374.pdf}
        \caption{Results of the transits injection-recovery on TIC 452793374 single-sector light curve. See \autoref{140212114} and legend of \autoref{fig:140212114} for a detailed description of this figure. The \texttt{biweight+TLS} method is the best after \textsf{nuance}, with 6\% less transits detected in total that could be due to the sole difference between \textsf{nuance} and \texttt{BLS} discussed in \autoref{control}. It indicates that a more flexible filter is necessary for this light curve with significantly higher red noise, making \textsf{nuance} not necessary.
        }
        \label{fig:452793374}
    \end{centering}
\end{figure}

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{153193474.pdf}
        \caption{Results of the transits injection-recovery on TIC 153193474 single-sector light curve. See \autoref{140212114} and legend of \autoref{fig:140212114} for a detailed description of this figure. The \texttt{harmonics+TLS} method is the best after \textsf{nuance}, with 62\% less transits detected in total, indicating that all presented methods are unable to detrend without removing many of the injected transit signals (which is not surprising when considering the trend being built, except in the case of \texttt{bspline+TLS}, but being apparently too flexible).}
        \label{fig:153193474}
    \end{centering}
\end{figure}

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{306331621.pdf}
        \caption{Results of the transits injection-recovery on TIC 306331621 single-sector light curve. See \autoref{140212114} and legend of \autoref{fig:140212114} for a detailed description of this figure. The \texttt{harmonics+TLS} method is the best after \textsf{nuance}, with 7\% less transits detected in total that could, again, be due to the sole difference between \textsf{nuance} and \texttt{BLS} discussed in \autoref{control}.}
        \label{fig:306331621}
    \end{centering}
\end{figure}

\newpage
\section{Discussion}\label{perf}

The results from \autoref{real} show that \textsf{nuance} always outperforms other methods when searching for transit signals in light curves containing correlated noise. While performances with other methods can be very comparable, \textsf{nuance} alleviates the need to choose for a specific detrending technique and rather exploit the complete modeling of the light curve using a physically interpretable Gaussian Process.\\\\
In few cases, \textsf{nuance} offers a clear advantage in comparison with other methods, in particular for low SNR transit signals in light curves containing fast or non-sinusoidal stellar variability. In these cases, \textsf{nuance} is able to robustly detect transits, and provide a unique option in comparison with techniques that likely degrade the searched signals through detrending.

\subsection*{Processing time}
As of June 2023, \textsf{nuance} is very much a work in progress and a full assessment of its performances is still ongoing. \autoref{fig:blsvsnuance} is a preliminary study of the computational cost of \textsf{nuance} against \texttt{BLS}. For reasons described in the next section, the linear search is first considered separately from the periodic search.

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{nuance_vs_bls_time.pdf}
        \caption{Preliminary study of \textsf{nuance} processing time against \texttt{BLS}. The lighter blue curve shows the performance of \textsf{nuance} when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together (the only option when using \texttt{BLS}).}
        \label{fig:blsvsnuance}
    \end{centering}
\end{figure}

\noindent In \autoref{fig:blsvsnuance}, the processing time of the linear search from \textsf{nuance} is recorded against the number of points in the light curve. One advantage of \textsf{nuance} is that the linear search can be performed separately on different continuous observations, and then combined in the periodic search. Hence, if searching for transits in separate observations with similar durations (such as different TESS sectors or different ground-based nightly observations), the computational cost of \textsf{nuance} grows linearly with the number of observations. Such a parallel cannot be made with \texttt{BLS}, as the algorithm considers all observations as a single one. This way, after assuming that 10 000 points observations can be combined into the periodic search (roughly corresponding to the number of data points in continuous TESS observations for example), \textsf{nuance} will constantly perform a single order of magnitude slower than \texttt{BLS}, only considering the linear search. For this reason, \textsf{nuance} must not be employed in the general case, but rather used when light curves contain correlated noise with characteristics described in this chapter. Proper guidelines and metrics to assess the presence of such noise, which would indicate if \textsf{nuance} is required, are currently being developed.\\\\
For reference, the complete search (including the periodic search) of a single TESS sector for TIC 140212114 (15992 points partly shown in  \autoref{fig:140212114}) took a total of 30 seconds, executed on a 10 core Apple M1 Pro chip. The latter processing time would grow linearly with the number of sectors being processed.

\subsection*{Periodic search}

The periodic search of \textsf{nuance} can be done separately from the linear search. For a given light curve, a single linear search can be performed then masked and combined iteratively to search for multiple transiting exoplanets. Hence, \autoref{fig:blsvsnuance} only shows the processing time of the linear search, assuming that a single planet is searched for. As of June 2023, the periodic search in the current implementation of \textsf{nuance} is the least optimized part. But recent developments indicate that the periodic search can be performed in a time equal to the complete processing time of the \texttt{BLS} algorithm. Hence, as the \texttt{BLS} algorithm grows linearly with the number of planets being searched (because reran every time on masked light curves), the increased computing time for \textsf{nuance} would be equal, bridging the gap between the computing time of  \texttt{BLS} and \textsf{nuance} when searching for multiplanetary transiting systems.

\subsection*{Design matrix}

In this chapter, starting from \autoref{detrending_effect}, it was assumed that systematic signals (as opposed to astrophysical signals) were completely corrected for without affecting the searched transit signals. For this reason, all analyses in this chapter were done with a single column design matrix $\bm{X}$, corresponding to the mean, constant, differential flux (ideally 1) of a light curve. In practice, \textsf{nuance} has been developed to linearly model systematic signals through more complex design matrices, in addition with its capability to model correlated noise while searching for transits. This aspect is completely unexploited in the previous comparisons, and would require to search for transits in completely raw light curves, which comes with few challenges. Given a light curve, this approach involves choosing a design matrix containing specific explanatory measurements, hence requiring a proper model comparison. In addition, large design matrices are expected to increase \textsf{nuance} processing time, something that needs to be accounted for in its performance assessment.\\\\
Used in the context of the SPECULOOS survey, \textsf{nuance} is currently being tested on light curves containing systematic signals that can be modeled linearly using polynomials of common explanatory variables (such as airmass, background flux, pointing error or PSF full-width at half-maximum). On light curves spanning only few hours, with less than a thousand photometric measurements, the use of a linear model to detrend the light curve ignoring the presence of transits is highly susceptible to remove the signals of interest, giving a clear advantage to a full-fledge method like \textsf{nuance}. The same applies for filter-based detrending algorithms. Proving the later empirical statement is out of the scope of this chapter and will be treated in a following publication. However, \textsf{nuance} might be particularly well suited to search for transits in ground-based observations, where the detrending of sparse observations is believed to remove the lowest SNR transit signals.\\\\
Overall, \textsf{nuance} is a powerful tool to search for transits in light curves containing systematic signals and correlated noise, characteristic of late M-dwarfs light curves. Due to several advances in computational methods, as well as its analytical formalism, \textsf{nuance}  remains highly tractable, and may already be applied to space-based continuous observations. \textsf{nuance} is currently being tested on sparse observations, offering a unique tool to search for transiting exoplanet in the light curves of SPECULOOS and other ground-based surveys.

\subsection{When to use it?}\label{when}

% APPENDIX

\newpage
\appendix
\section{Light curves simulations}\label{signals_simulations}

In order to study the effect of correlated noise on transit search, this paper relies on transit light curve simulations including realistic effects of stellar variability and instrumental signals. The following describes how such signals are modeled.
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.75\linewidth]{principle_dataset_decomposed.pdf}
        \caption{Example dataset sampled at $N=2880$ times corresponding to an observation of 4 days with an exposure time of 2 minutes. The mean of this signal consists in a periodic transit signal of period $P=0.7$ days, duration $D=0.05$ days and depth of 2\% ($\mathcal{T}$ in gray) plus instrumental signals ($\mathcal{S}$ in blue). Correlated noise in the form of stellar variability is simulated by modeling the covariance matrix of the signal with a Gaussian process ($\mathcal{V}$ in red) including a diagonal variance of $0.001^2$ corresponding to white noise ($\epsilon$ in light grey). This simulated signal is not intended to be physically realistic.}
        \label{fig:app_principle_dataset}
    \end{centering}
\end{figure}
\noindent Let $f$ be the simulated flux of a star sampled and arranged in the vector $\bm{f}$ associated to the vector of times $\bm{t}$, such that
\begin{equation*}
    \bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{C}),
\end{equation*}
i.e. that $\bm{f}$ is drawn from a Gaussian process of mean $\bm{\mu}$ and covariance matrix $\bm{C}$. In this equation, $\bm{\mu}$ is such that its $i$-th element is defined by  $\mu_i = \mathcal{T}(t_i) + \mathcal{S}(t_i)$ where $t_i$ is the $i$-th time of observation, $\mathcal{T}$ is a periodic transit function and $\mathcal{S}$ a function describing the instrumental part of the signal (both described below). The covariance matrix $\bm{C}$ is built such that $C_{i, j} = k(t_i, t_j)$ where $k$ is a covariance function (or \textit{kernel}) accounting for correlated noise in the form of stellar variability (with added white noise). An example of such signal is simulated and shown in \autoref{fig:app_principle_dataset}.

\subsection{Transit signal $\mathcal{T}$}\label{app_proto}
The periodic transit signal $\mathcal{T}$ is simulated using the simple model described in \cite{protopapas}, where a transit of period $P$, epoch $T_0$, duration $D$ and unitary depth observed at time $t$ is given by
\begin{equation}\label{eq:protopapas}
    \begin{gathered}
        \mathcal{T}_c(t, P, T_0, D) = \frac{1}{2}\tanh\left(c\left[\theta - \frac{1}{2}\right]\right) - \frac{1}{2}\tanh\left(c\left[\theta + \frac{1}{2}\right]\right), \\
        \text{with}\quad\theta = \frac{P}{\pi  D}\sin\left(\frac{\pi(t-T_0)}{P}\right),
    \end{gathered}
\end{equation}
where the dimensionless parameter $c$ controls the roundness of the transit depth ($c\gg1$ corresponding to a box-shaped transit as shown in \autoref{fig:protopapas}). This analytical model is fully empirical but easily differentiable.
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{protopapas.pdf}
        \caption{Simulations of a single transit signal (\autoref{eq:protopapas}) shown for different values of $c$.}
        \label{fig:protopapas}
    \end{centering}
\end{figure}
\noindent In this paper, unless specified, all transits are simulated with $c=12$, a value arbitrarily chosen that can be fined-tuned in real applications using the limb-darkening coefficients of a given star. The periodic transit signal $\mathcal{T}$ seen in \autoref{fig:app_principle_dataset} corresponds to $\mathcal{T} = 0.02 \times \mathcal{T}_{c=12}(\bm{t}, P=0.7, T_0=0.2, D=0.05)$, all parameters in unit of days.

\subsection{Instrumental signals $\mathcal{S}$}
Instrumental signals are simulated as a linear model of $M$ explanatory variables arranged in the $(N\times M)$ design matrix $\bm{X}$. Hence,
$$
    \mathcal{S} = \bm{X w},
$$
where the vector $\bm{w}$ contains the linear coefficients of the model. The simulated flux shown in \autoref{fig:app_principle_dataset} contains a linear model where the $M=4$ columns of the design matrix $\bm{X}$ are given by $\bm{X}_{i} = \bm{t}^i$ (i.e. $\bm{X}$ is the Vandermonde matrix order $3$ of time $t$) and $\bm{w} = [1.0\quad0.0005\quad\text{-}0.0002\quad\text{-}0.0005].$

\subsection{Stellar variability $\nu$}\label{app_gp}
As this chapter focuses on stellar variability and its effect on transit detection, We employ a simple physically-motivated Gaussian Process kernel, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) taking the form 
\begin{equation}
    \begin{gathered}
        k(\tau) = \sigma^2\,\exp\left(-\frac{\omega\,\tau}{2\,Q}\right)
        \left\{\begin{array}{ll}
            1 + \omega\,\tau & \mbox{for } Q = 1/2 \\
            \cosh(f\,\omega\,\tau/2\,Q) + \sinh(f\,\omega\,\tau/2\,Q)/f
                & \mbox{for } Q < 1/2 \\
            \cos(g\,\omega\,\tau/2\,Q) + \sin(g\,\omega\,\tau/2\,Q)/g
                & \mbox{for } Q > 1/2
        \end{array}\right. \\
        \text{where}\quad \tau = |t_i - t_j|\text{,}\quad f = \sqrt{1 - 4\,Q^2} \quad \text{and}\quad g = \sqrt{4\,Q^2 - 1}
    \end{gathered}
\end{equation}
where $Q$ is the quality factor of the oscillator, $\omega$ its pulsation and $\sigma$ the amplitude of the kernel function. Gaussian Process computations in this paper use the implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package exposing the quasi-separable kernels from \cite{celerite2} and powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}. The stellar variability signal in \autoref{fig:app_principle_dataset} has been sampled from a Gaussian Process with an SHO kernel of parameters $\omega = \pi/6D$ (i.e. a period equal to 12 times the duration $D$ of the simulated transit), $Q=45$ and $\sigma=0.02$, the depth of the simulated transit. An extra term $\sigma_f^2=0.001^2$ is added to the diagonal of the covariance matrix, corresponding to the variance of the simulated measurement $f$ and leading to the white noise observed in \autoref{fig:app_principle_dataset}.

\section{Proof for the periodic search expression}\label{proof}

\newcommand{\sumTk}{i\neq k}
From the linear search presented in \autoref{linear_search}, we retain and index by $k$ the parameters of the $K$ individual transits whose epochs $\{T_k\}_k$ are compatible with a periodic signal of period $P$ and epoch $T_0$. From the likelihoods of these transits (computed in \autoref{linear_search}), we want an expression for
\begin{equation*}
    p(\bm{f} \vert P, T_0 ,D, \Delta) = \prod_{k\in\mathbb{T}} p(\bm{f} \vert T_k, D, \Delta),
\end{equation*}
i.e., given a depth $D$, the likelihood of the data given a periodic transit signal of period $P$, epoch $T_0$ and a common depth $\Delta$. Since only $\{p(\bm{f} \vert T_k, D, \Delta_k)\}_{k}$ is known (i.e. transits with different depths), we decompose
\begin{equation}\label{eq:non_part_of_per}
    p(\bm{f} \vert T_k, D, \Delta) = \int p(\bm{f} \vert T_k, D, \tilde\Delta)p(\tilde\Delta | \Delta)\, d\tilde\Delta,
\end{equation}
where $p(\bm{f} \vert T_k, D, \tilde\Delta)$ is the probability of the $k$-th transit to have a depth $\tilde\Delta$ and $p(\tilde\Delta | \Delta)$ the probability to observe the depth $\tilde\Delta$ knowing the existence of a common depth $\Delta$. In other words, \autoref{eq:non_part_of_per} involves the likelihood of the non-periodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
\\\\
Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$ and an amplitude $\mathcal{L}_k$, leading to the likelihood function
\begin{equation*}
    p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k\exp \left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right).
\end{equation*}
As for the common transit depth $\Delta$, it can be estimated through the joint probability of all other transit depths than $\Delta_k$, such that
\begin{equation*}
    \Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2),
\end{equation*}
with 
\begin{equation}\label{eq:params}
\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \hspace{0.5cm} \text{and} \hspace{0.5cm}
\Delta =\sigma^2 \sum_{\sumTk}^K {\frac{\Delta_i}{\sigma_i^2}}.
\end{equation}
Hence
\begin{equation*}
    p(\tilde\Delta | \Delta) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp \left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right).
\end{equation*}
We can now rewrite \autoref{eq:non_part_of_per} as
\begin{equation*}
    p(\bm{f} \vert T_k, D, \Delta) =  \frac{\mathcal{L}_k}{\sqrt{2\pi\sigma^2}} \int \exp\left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right)\, \exp\left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)\, d\tilde\Delta.
\end{equation*}
The integral in this equation is a product of gaussian integrals that can be obtained analytically, leading to
\begin{equation*}
    p(\bm{f} \vert T_k, D, \Delta) = \mathcal{L}_k  \sqrt{\frac{\sigma_{k}^2}{\sigma^{2} + \sigma_{k}^{2}}} \exp\left(-\frac{1}{2}\frac{(\Delta_k-\Delta)^2}{\sigma_k^2 + \sigma^2}\right).
\end{equation*}
Finally,
\begin{equation}
    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(\sigma_{k}^2) - \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}}\right),
\end{equation}
the log-likelihood of the data given a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and common depth $\Delta$. In order to reduce the number of times \autoref{eq:params} is computed, we adopt the biased estimates
\begin{equation}\label{eq:sigma_delta}
    \frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \hspace{0.5cm}\text{and}\hspace{0.5cm} \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}},
\end{equation}
so that $\Delta$ and $\sigma$ are independent of $k$ in the last sum of \autoref{eq:result}.

\bibliography{ref}

\end{document}

### 2023-09-19 lib/forben.py

@@ -1,0 +1,41 @@
+import numpy as np
+from astropy.stats import LombScargle
+
+
+def bens_detrend(time, flux, n=5):
+    """iterative simple harmonics detrending"""
+
+    def rotation_period(time, flux):
+        """rotation period based on LS periodogram"""
+
+        ls = LombScargle(time, flux)
+        frequency, power = ls.autopower(
+            minimum_frequency=1 / 10, maximum_frequency=1 / 0.1
+        )
+        period = 1 / frequency[np.argmax(power)]
+        return period
+
+    def subtract_sinusoid(time, flux, period):
+        """subtract a sinusoid with period from flux"""
+        X = np.vstack(
+            [
+                np.ones(len(time)),
+                np.cos(2 * np.pi / period * time),
+                np.sin(2 * np.pi / period * time),
+            ]
+        ).T
+
+        w = np.linalg.lstsq(X, flux, rcond=None)[0]
+        model = np.dot(X, w)
+
+        return flux - model, model
+
+    _flux = flux.copy()
+    model = np.zeros_like(flux)
+
+    for _ in range(n):
+        period = rotation_period(time, _flux)
+        _flux, _model = subtract_sinusoid(time, _flux, period)
+        model += _model
+
+    return _flux + 1.0, model

### 2023-09-19 workflows/cleaning_snr/scripts/plot_snr_cleaning.py

@@ -1,146 +1,0 @@
-import sys
-
-import numpy as np
-import pandas as pd
-import yaml
-from scipy.stats import binned_statistic_2d
-
-sys.path.append("./lib")
-import matplotlib.pyplot as plt
-from make_lightcurve import make_lc
-from make_params import make
-from matplotlib.patches import ConnectionPatch
-from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-from utils import depth as true_depth
-from utils import error
-
-df = pd.read_csv(snakemake.input[0])
-
-n = 3  # number of small plots
-scale = 1.2
-W = n + 1
-H = n + 1
-
-fig = plt.figure(None, (8.5, 7))
-
-
-def I(i, j):
-    return int(j * W + i) + 1
-
-
-for i in range(H * W):
-    ax = plt.subplot(H, W, i + 1)
-
-## Main plot
-main = plt.subplot(H, W, (I(0, 1), I(W - 2, H - 1)))
-
-bins = (30, 30)
-snr, var, amp = df.values.T.astype(float)
-stats = binned_statistic_2d(var, amp, snr, bins=bins)
-im = main.imshow(
-    stats.statistic.T,
-    origin="lower",
-    extent=(
-        stats.x_edge.min(),
-        stats.x_edge.max(),
-        stats.y_edge.min(),
-        stats.y_edge.max(),
-    ),
-    aspect="auto",
-    vmax=5,
-)
-
-main.set_ylabel(
-    r"$\delta \propto \frac{\mathrm{variability\; amplitude}}{\mathrm{transit\;depth}}$",
-    fontsize=14,
-)
-main.set_xlabel(
-    r"$\tau \propto \frac{\mathrm{variability\; timescale}}{\mathrm{transit\;duration}}$",
-    fontsize=14,
-)
-params = true_depth
-original_snr = true_depth / np.sqrt(np.mean(error) ** 2 / n)
-
-main.set_title(f"Detrending effect on SNR", loc="left")
-main.text(
-    0.1,
-    stats.y_edge.max() + 0.26,
-    f"original SNR: {original_snr:.2f}",
-    color="0.3",
-    va="center",
-)
-
-# removing some axes
-# ------------------
-ax = plt.subplot(H, W, I(W - 1, 0))
-plt.axis("off")
-
[…]

### 2023-09-19 workflows/real_injection_recovery/scripts/bspline_tls.py

@@ -1,49 +1,0 @@
-import sys
-
-import numpy as np
-
-sys.path.append("lib")
-import pickle
-
-import numpy as np
-import yaml
-from scipy.interpolate import BSpline, splrep
-from tls import tls
-
-data = pickle.load(open(snakemake.input.fluxes, "rb"))
-periods = np.load(snakemake.input.periods)
-info = yaml.safe_load(open(snakemake.input.info, "r"))
-verbose = True
-mask = np.ones_like(data["time"], dtype=bool)
-time, flux, error = data["time"], data["flux"], data["error"]
-
-for i in range(2):
-    tck = splrep(
-        time[mask],
-        flux[mask],
-        w=1 / error[mask],
-    )
-    trend = BSpline(*tck)(time)
-    mask &= np.abs(flux - trend) < 3 * np.std(flux - trend)
-
-trend = BSpline(*tck)(time)
-
-flatten_trend = BSpline(*tck)(time)
-flatten_flux = flux - flatten_trend
-flatten_flux -= np.mean(flatten_flux)
-flatten_flux += 1.0
-
-model = tls(time, flatten_flux, verbose=verbose)
-results = model.power(
-    periods,
-    verbose=verbose,
-    use_threads=1,
-    show_progress_bar=verbose,
-    durations=[0.01, data["transit_duration"]],
-)
-
-t0, period, power = results["T0"], results["period"], results["power"]
-pickle.dump(
-    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
-    open(snakemake.output[0], "wb"),
-)

### 2023-09-19 workflows/real_injection_recovery/scripts/concatenate.py

@@ -1,55 +1,0 @@
-import pickle
-
-import numpy as np
-import pandas as pd
-
-
-def right_candidate(t0, period, true_t0, true_period, verbose=False):
-    t0_check = (
-        np.abs((t0 - true_t0 + 0.5 * true_period) % true_period - 0.5 * true_period)
-        % period
-        < 0.01
-    )
-    period_check = np.abs(period - true_period) < 0.01
-    period_check |= np.abs(2 * period - true_period) < 0.01
-    period_check |= np.abs(period / 2 - true_period) < 0.01
-    same = period_check  # np.logical_and(t0_check, period_check)
-    if verbose:
-        if not same:
-            if not t0_check:
-                output = f"t0 differ: {t0:.2e} {true_t0:.2e}"
-            if not period_check:
-                output = f"period differ: {period:.2e} {true_period:.2e}"
-        else:
-            output = "match"
-        return same, output
-    else:
-        return same
-
-
-def get_result(injected_file, recovered_file):
-    injected = pickle.load(open(injected_file, "rb"))
-    recovered = pickle.load(open(recovered_file, "rb"))
-    t0, period = recovered["t0"], recovered["period"]
-    true_t0, true_period = injected["transit_t0"], injected["planet_period"]
-    return {
-        "t0": t0,
-        "period": period,
-        "true_t0": true_t0,
-        "true_period": true_period,
-        "found": right_candidate(t0, period, true_t0, true_period),
-        "radius": injected["planet_radius"],
-        "tau": injected["tau"],
-        "delta": injected["delta"],
-    }
-
-
-recovered = snakemake.input.recovered
-injected = snakemake.input.injected
-
-results = []
-
-for i, (a, b) in enumerate(zip(injected, recovered)):
-    results.append(get_result(a, b))
-results = pd.DataFrame(results)
-results.to_csv(snakemake.output[0], index=False)

### 2023-09-19 workflows/real_injection_recovery/scripts/download.py

@@ -1,78 +1,0 @@
-import pickle
-
-import astropy.units as u
-import lightkurve as lk
-import numpy as np
-import pandas as pd
-import yaml
-
-tic = int(snakemake.wildcards["target"])
-filepath = snakemake.output[0]
-
-targets = pd.read_csv("data/tess-ultra-fast-rotators-brighter-mag14.csv")[
-    ["Name", "LS_Period", "Amplitude", "teff_val"]
-]
-
-targets = targets.sort_values("LS_Period")
-target = targets[targets.Name == tic]
-
-name = f"TIC {tic}"
-search_result = lk.search_lightcurve(name, author="SPOC", exptime=120)
-
-# data
-klc = search_result[0].download()
-times = klc.time.to_value("btjd")
-fluxes = klc.pdcsap_flux.to_value().filled(np.nan)
-errors = klc.flux_err.to_value().filled(np.nan)
-mask = np.isnan(fluxes) | np.isnan(errors) | np.isnan(times)
-original_time = times[~mask]
-original_flux = fluxes[~mask]
-original_error = errors[~mask]
-
-splits = np.array_split(
-    np.arange(len(original_time)), np.flatnonzero(np.diff(original_time) > 0.1) + 1
-)
-
-n = 300
-time = np.hstack([original_time[split[n:-n]] for split in splits])
-flux = np.hstack([original_flux[split[n:-n]] for split in splits])
-error = np.hstack([original_error[split[n:-n]] for split in splits])
-
-# first exposure
-first_exposure = klc.time[0]
-time_format = rf"{first_exposure.format.upper()}_{{{first_exposure.scale.upper()}}}"
-first_exposure_time = first_exposure.value
-
-flux_median = np.median(original_flux)
-time = time - first_exposure_time
-flux /= flux_median
-error /= flux_median
-
-# stellar parameters
-url = f"https://exofop.ipac.caltech.edu/tess/download_stellar.php?id={tic}"
-star = pd.read_csv(url, delimiter="|", index_col=1).iloc[0]
-
-pickle.dump(
-    {
-        "flux": flux,
-        "time": time,
-        "error": error,
-    },
-    open(snakemake.output.fluxes, "wb"),
-)
-
-yaml.safe_dump(
-    {
-        "tic": int(tic),
-        "time0": float(first_exposure_time),
-        "time0_format": str(time_format),
-        "star_period": float(target.LS_Period),
[…]

### 2023-09-19 workflows/real_injection_recovery/scripts/plot_comparison.py

@@ -1,171 +1,0 @@
-import pickle
-
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import yaml
-from matplotlib import gridspec
-from nuance.star import Star
-
-info = yaml.safe_load(open(snakemake.input.info, "r"))
-data = pickle.load(open(snakemake.input.first_injected, "rb"))
-tic = snakemake.wildcards.target
-results = {
-    method: snakemake.input.results[i]
-    for i, method in enumerate(snakemake.params.methods)
-}
-first_recovered = {
-    method: snakemake.input.first_recovered[i]
-    for i, method in enumerate(snakemake.params.methods)
-}
-
-from nuance import Nuance
-from nuance.kernels import rotation
-
-split_idxs = [
-    0,
-    *np.flatnonzero(np.diff(data["time"]) > 10 / 60 / 24),
-    len(data["time"]),
-]
-
-# nuance trend
-gp_params = yaml.safe_load(open(snakemake.input.gp, "r"))
-build_gp, _ = rotation(info["star_period"])
-nu = Nuance(data["time"], data["flux"], gp=build_gp(gp_params, data["time"]))
-nuance_trend = np.sum(nu.models(0, 0, 1), 0)
-
-
-def right_candidate(t0, period, true_t0, true_period, verbose=False):
-    t0_check = (
-        np.abs((t0 - true_t0 + 0.5 * true_period) % true_period - 0.5 * true_period)
-        % period
-        < 0.01
-    )
-    period_check = np.abs(period - true_period) <= 0.01
-    alias_check = np.abs(2 * period - true_period) <= 0.01
-    alias_check |= np.abs(period / 2 - true_period) <= 0.01
-    if period_check:
-        return 1
-    elif alias_check:
-        return 0.5
-    else:
-        return 0
-
-
-methods = snakemake.params.methods
-
-total_found = {}
-for method in methods.keys():
-    df = pd.read_csv(results[method])
-    df["found"] = df.apply(
-        lambda row: right_candidate(
-            row["t0"], row["period"], row["true_t0"], row["true_period"]
-        ),
-        axis=1,
-    )
-    total_found[method] = np.count_nonzero(df.found)
-
-# sorted name of methods total_found
-sorted_methods = sorted(total_found, key=total_found.get, reverse=True)
[…]

### 2023-09-19 workflows/real_injection_recovery_gp_opt/lib/tls.py

@@ -1,407 +1,0 @@
-from transitleastsquares.main import *
-
-
-class tls(transitleastsquares):
-    def power(self, periods, durations=None, **kwargs):
-        """Compute the periodogram for a set of user-defined parameters"""
-        self, kwargs = validate_args(self, kwargs)
-
-        if self.verbose:
-            print(tls_constants.TLS_VERSION)
-
-        if durations is None:
-            durations = duration_grid(
-                periods, shortest=1 / len(self.t), log_step=self.duration_grid_step
-            )
-
-        maxwidth_in_samples = int(numpy.max(durations) * numpy.size(self.y))
-        if maxwidth_in_samples % 2 != 0:
-            maxwidth_in_samples = maxwidth_in_samples + 1
-        lc_cache_overview, lc_arr = get_cache(
-            durations=durations,
-            maxwidth_in_samples=maxwidth_in_samples,
-            per=self.per,
-            rp=self.rp,
-            a=self.a,
-            inc=self.inc,
-            ecc=self.ecc,
-            w=self.w,
-            u=self.u,
-            limb_dark=self.limb_dark,
-            verbose=self.verbose,
-        )
-
-        if self.verbose:
-            print(
-                "Searching "
-                + str(len(self.y))
-                + " data points, "
-                + str(len(periods))
-                + " periods from "
-                + str(round(min(periods), 3))
-                + " to "
-                + str(round(max(periods), 3))
-                + " days"
-            )
-
-        # Python 2 multiprocessing with "partial" doesn't work
-        # For now, only single-threading in Python 2 is supported
-        if sys.version_info[0] < 3:
-            self.use_threads = 1
-            warnings.warn("This TLS version supports no multithreading on Python 2")
-
-        if self.verbose:
-            if self.use_threads == multiprocessing.cpu_count():
-                print("Using all " + str(self.use_threads) + " CPU threads")
-            else:
-                print(
-                    "Using "
-                    + str(self.use_threads)
-                    + " of "
-                    + str(multiprocessing.cpu_count())
-                    + " CPU threads"
-                )
-
-        if self.show_progress_bar:
-            bar_format = "{desc}{percentage:3.0f}%|{bar}| {n_fmt}/{total_fmt} periods | {elapsed}<{remaining}"
-            pbar = tqdm(total=numpy.size(periods), smoothing=0.3, bar_format=bar_format)
-
-        if tls_constants.PERIODS_SEARCH_ORDER == "ascending":
[…]

### 2023-09-19 workflows/real_injection_recovery_gp_opt/scripts/bls.py

@@ -1,110 +1,0 @@
-import sys
-
-import numpy as np
-
-sys.path.append("lib")
-import pickle
-
-import numpy as np
-import yaml
-from astropy.timeseries import BoxLeastSquares
-from scipy.interpolate import BSpline, splrep
-from wotan import flatten
-
-data = pickle.load(open(snakemake.input.fluxes, "rb"))
-periods = np.load(snakemake.input.periods)
-info = yaml.safe_load(open(snakemake.input.info, "r"))
-
-time, flux, error = data["time"], data["flux"], data["error"]
-verbose = True
-
-# biweight
-# --------
-
-flatten_trend = flatten(
-    time, flux, window_length=data["transit_duration"] * 3, return_trend=True
-)[1]
-flatten_flux = flux - flatten_trend
-flatten_flux -= np.mean(flatten_flux)
-flatten_flux += 1.0
-
-model = BoxLeastSquares(time, flatten_flux, dy=error)
-results = model.power(periods, 0.03)
-
-t0 = 0.0
-period = results.period[np.argmax(results.power)]
-power = results.power
-
-pickle.dump(
-    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
-    open(snakemake.output.wotan3D, "wb"),
-)
-
-
-# harmonics
-# ---------
-
-
-# makes an harmonics design matrix of time
-def make_harmonics(time, period, nharmonics=4):
-    # make design matrix
-    X = np.ones((len(time), 2 * nharmonics + 1))
-    X[:, 1] = time
-    for i in range(1, nharmonics):
-        X[:, 2 * i + 1] = np.sin(2 * np.pi * i * time / period)
-        X[:, 2 * i + 2] = np.cos(2 * np.pi * i * time / period)
-
-    return X
-
-
-X = make_harmonics(time, info["star_period"], nharmonics=2)
-# solve for the coefficients
-coeffs = np.linalg.solve(X.T @ X, X.T @ flux)
-# make the model
-flatten_trend = X @ coeffs
-flatten_flux = flux - flatten_trend
-flatten_flux -= np.mean(flatten_flux)
-flatten_flux += 1.0
-
-model = BoxLeastSquares(time, flatten_flux, dy=error)
-results = model.power(periods, 0.03)
-
-t0 = 0.0
-period = results.period[np.argmax(results.power)]
-power = results.power
-
-pickle.dump(
[…]

### 2023-09-19 workflows/real_injection_recovery_gp_opt/scripts/nuance_search.py

@@ -1,27 +1,0 @@
-import pickle
-
-import numpy as np
-import yaml
-from nuance import Nuance
-from nuance.kernels import rotation
-
-gp_params = yaml.full_load(open(snakemake.input.gp, "r"))
-data = pickle.load(open(snakemake.input.fluxes, "rb"))
-periods = np.load(snakemake.input.periods)
-info = yaml.safe_load(open(snakemake.input.info, "r"))
-
-build_gp, _ = rotation(info["star_period"])
-gp = build_gp(gp_params, data["time"])
-
-nu = Nuance(data["time"], data["flux"], gp=gp)
-
-nu.linear_search(data["time"], np.array([0.01, data["transit_duration"]]))
-
-search = nu.periodic_search(periods)
-
-output = snakemake.output[0]
-t0, _, period = search.best
-pickle.dump(
-    {"t0": t0, "period": period, "power": search.Q_snr, "trend": None},
-    open(snakemake.output[0], "wb"),
-)

### 2023-09-19 workflows/real_injection_recovery_gp_opt/scripts/periods.py

@@ -1,11 +1,0 @@
-import numpy as np
-import yaml
-from nuance.utils import clean_periods
-
-info = yaml.safe_load(open(snakemake.input[0], "r"))
-
-periods = np.linspace(0.3, 11.0, 8000)
-periods = clean_periods(periods, info["star_period"])
-
-output = snakemake.output[0]
-np.save(output, periods)

### 2023-09-19 workflows/real_injection_recovery_gp_opt/scripts/plots/info.py

@@ -1,36 +1,0 @@
-import lightkurve as lk
-import pandas as pd
-import yaml
-
-tic = int(snakemake.wildcards.target)
-info_output = snakemake.output[0]
-
-targets = pd.read_csv("data/tess-ultra-fast-rotators-brighter-mag14.csv")[
-    ["Name", "LS_Period", "Amplitude", "teff_val"]
-]
-
-targets = targets.sort_values("LS_Period")
-target = targets[targets.Name == tic]
-
-name = f"TIC {tic}"
-search_result = lk.search_lightcurve(name, author="SPOC", exptime=120)
-
-# sector
-sector = search_result[0].mission[0].split(" ")[-1]
-sector = int(sector.lstrip("0"))
-
-# first exposure
-first_exposure = search_result[0].download().time[0]
-time_format = rf"{first_exposure.format.upper()}_{{{first_exposure.scale.upper()}}}"
-first_exposure_time = first_exposure.value
-
-info = {
-    "tic": int(tic),
-    "sector": int(sector),
-    "ramsay_teff": float(target.teff_val.values[0]),
-    "ramsay_period": float(target.LS_Period.values[0]),
-    "first_exposure_time": float(first_exposure_time),
-    "first_exposure_time_format": time_format,
-}
-
-yaml.safe_dump(info, open(info_output, "w"))

### 2023-09-19 workflows/real_injection_recovery_gp_opt/scripts/tls_search.py

@@ -1,86 +1,0 @@
-import sys
-
-import numpy as np
-
-sys.path.append("lib")
-import pickle
-
-import numpy as np
-import yaml
-from tls import tls
-from wotan import flatten
-
-data = pickle.load(open(snakemake.input.fluxes, "rb"))
-periods = np.load(snakemake.input.periods)
-info = yaml.safe_load(open(snakemake.input.info, "r"))
-
-time, flux = data["time"], data["flux"]
-verbose = True
-
-from functools import partial
-
-# wotan 3D
-# --------
-
-flatten_trend = flatten(
-    time, flux, window_length=data["transit_duration"] * 3, return_trend=True
-)[1]
-flatten_flux = flux - flatten_trend
-flatten_flux -= np.mean(flatten_flux)
-flatten_flux += 1.0
-
-model = tls(time, flatten_flux, verbose=verbose)
-results = model.power(
-    periods,
-    verbose=verbose,
-    use_threads=1,
-    show_progress_bar=verbose,
-    durations=[0.01, data["transit_duration"]],
-)
-
-t0, period, power = results["T0"], results["period"], results["power"]
-pickle.dump(
-    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
-    open(snakemake.output.wotan3D, "wb"),
-)
-
-
-# harmonics
-# ---------
-
-
-# makes an harmonics design matrix of time
-def make_harmonics(time, period, nharmonics=4):
-    # make design matrix
-    X = np.ones((len(time), 2 * nharmonics + 1))
-    X[:, 1] = time
-    for i in range(1, nharmonics):
-        X[:, 2 * i + 1] = np.sin(2 * np.pi * i * time / period)
-        X[:, 2 * i + 2] = np.cos(2 * np.pi * i * time / period)
-
-    return X
-
-
-X = make_harmonics(time, info["star_period"], nharmonics=2)
-# solve for the coefficients
-coeffs = np.linalg.solve(X.T @ X, X.T @ flux)
-# make the model
-flatten_trend = X @ coeffs
-flatten_flux = flux - flatten_trend
-flatten_flux -= np.mean(flatten_flux)
-flatten_flux += 1.0
-
-model = tls(time, flatten_flux, verbose=verbose)
-results = model.power(
-    periods,
-    verbose=verbose,
-    use_threads=1,
-    show_progress_bar=verbose,
-    durations=[0.01, data["transit_duration"]],
-)
-
[…]

### 2023-09-19 workflows/sanity/scripts/bls_search.py

@@ -4,26 +4,21 @@ from astropy.timeseries import BoxLeastSquares
 
 time, _, error = np.load(snakemake.input.lc)
 flux = np.load(snakemake.input.injected)
-period_range = snakemake.config["period_range"]
-n_periods = snakemake.config["n_periods"]
 duration = snakemake.config["duration"]
-params = np.load(snakemake.input.params)
-i = int(snakemake.wildcards.i)
-periods = np.linspace(-0.1 + period_range[0], 0.1 + period_range[1], n_periods)
+params = yaml.full_load(open(snakemake.input.params))
+period_range = snakemake.config["period_range"]
+periods = np.linspace(
+    -0.1 + period_range[0], 0.1 + period_range[1], snakemake.config["n_periods"]
+)
 
 # search
 # ------
-periods = np.linspace(0.5, 6, 3000)
 model = BoxLeastSquares(time, flux, dy=error)
-results = model.power(periods, 0.03)
+results = model.power(periods, duration)
 
 yaml.safe_dump(
     {
-        "t0": float(0.0),
         "period": float(results.period[np.argmax(results.power)]),
-        "true_period": float(params[i, 0]),
-        "true_t0": float(0),
-        "true_depth": float(params[i, 1]),
     },
     open(snakemake.output[0], "w"),
 )

### 2023-09-19 workflows/sanity/scripts/compute_depth.py

@@ -1,0 +1,18 @@
+import yaml
+import numpy as np
+
+time, flux, error = np.load(snakemake.input[0])
+period_range = snakemake.config["period_range"]
+snr_range = [4.0, 30.0]
+duration = float(snakemake.config["duration"])
+
+# calculate the minimum and maximum depth
+minimum_number_of_transit_points = len(time) * duration / period_range[1]
+min_depth = snr_range[0] * np.median(error) / np.sqrt(minimum_number_of_transit_points)
+
+maximum_number_of_transit_points = len(time) * duration / period_range[0]
+max_depth = snr_range[1] * np.median(error) / np.sqrt(maximum_number_of_transit_points)
+
+# save the results
+with open(snakemake.output[0], "w") as f:
+    yaml.safe_dump({"min_depth": float(min_depth), "max_depth": float(max_depth)}, f)

### 2023-09-19 workflows/sanity/scripts/concat.py

@@ -1,0 +1,24 @@
+import pandas as pd
+import yaml
+
+results = []
+
+for p, b, n in zip(
+    snakemake.input.params,
+    snakemake.input.bls,
+    snakemake.input.nuance,
+):
+    params = yaml.full_load(open(p, "r"))
+    bls = yaml.full_load(open(b, "r"))
+    nuance = yaml.full_load(open(n, "r"))
+    results.append(
+        {
+            "period": params["period"],
+            "depth": params["depth"],
+            "bls_period": bls["period"],
+            "nuance_period": nuance["period"],
+        }
+    )
+
+df = pd.DataFrame(results)
+df.to_csv(snakemake.output[0], index=False)

### 2023-09-19 workflows/sanity/scripts/inject.py

@@ -1,13 +1,17 @@
 import numpy as np
+import yaml
 from nuance.utils import periodic_transit
 
 time, flux, error = np.load(snakemake.input.lc)
-params = np.load(snakemake.input.params)
-i = int(snakemake.wildcards.i)
+params = yaml.full_load(open(snakemake.input.params))
 duration = snakemake.config["duration"]
+i = int(snakemake.wildcards.i)
 
 np.random.seed(i)
-period, depth = params[i]
-injected_flux = flux + depth * periodic_transit(time, 0, duration, period)
+period = params["period"]
+t0 = params["t0"]
+depth = params["depth"]
+# c = 500 to have a box shaped transit
+injected_flux = flux + depth * periodic_transit(time, t0, duration, period, c=500)
 
 np.save(snakemake.output[0], injected_flux)

### 2023-09-19 workflows/sanity/scripts/nuance_search.py

@@ -4,12 +4,12 @@ from nuance import Nuance
 
 time, _, error = np.load(snakemake.input.lc)
 flux = np.load(snakemake.input.injected)
-period_range = snakemake.config["period_range"]
-n_periods = snakemake.config["n_periods"]
 duration = snakemake.config["duration"]
-params = np.load(snakemake.input.params)
-i = int(snakemake.wildcards.i)
-periods = np.linspace(-0.1 + period_range[0], 0.1 + period_range[1], n_periods)
+params = yaml.full_load(open(snakemake.input.params))
+period_range = snakemake.config["period_range"]
+periods = np.linspace(
+    -0.1 + period_range[0], 0.1 + period_range[1], snakemake.config["n_periods"]
+)
 
 # search
 # ------
@@ -22,9 +22,6 @@ yaml.safe_dump(
     {
         "t0": float(t0),
         "period": float(period),
-        "true_period": float(params[i, 0]),
-        "true_t0": float(0),
-        "true_depth": float(params[i, 1]),
     },
     open(snakemake.output[0], "w"),
 )

### 2023-09-19 workflows/sanity/scripts/params.py

@@ -3,26 +3,13 @@ import yaml
 
 time, flux, error = np.load(snakemake.input[0])
 period_range = snakemake.config["period_range"]
-snr_range = [4.0, 30.0]
 duration = float(snakemake.config["duration"])
-n = int(snakemake.config["n"])
+depths = yaml.full_load(open(snakemake.input[1]))
 
-# calculate the minimum and maximum depth
-minimum_number_of_transit_points = len(time) * duration / period_range[1]
-min_depth = snr_range[0] * np.median(error) / np.sqrt(minimum_number_of_transit_points)
-
-maximum_number_of_transit_points = len(time) * duration / period_range[0]
-max_depth = snr_range[1] * np.median(error) / np.sqrt(maximum_number_of_transit_points)
+period = np.random.uniform(period_range[0], period_range[1])
+depth = np.random.uniform(depths["min_depth"], depths["max_depth"])
+t0 = np.random.uniform(0, period)
 
 
-periods = np.linspace(period_range[0], period_range[1], n)
-depths = np.linspace(min_depth, max_depth, n)
-
-# grid
-periods, depths = np.meshgrid(periods, depths)
-# as pairs
-periods = periods.flatten()
-depths = depths.flatten()
-period_depth = np.array([periods, depths]).T
-
-np.save(snakemake.output[0], period_depth)
+with open(snakemake.output[0], "w") as f:
+    yaml.safe_dump({"period": float(period), "depth": float(depth), "t0": float(t0)}, f)

### 2023-09-19 workflows/sanity/scripts/plot.py

@@ -1,8 +1,9 @@
-import matplotlib.patches as mpatches
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
-import yaml
+from scipy.stats import binned_statistic_2d
+
+scatter = False
 
 
 def is_found(period, true_period):
@@ -12,61 +13,87 @@ def is_found(period, true_period):
     if period_check:
         return 1
     elif alias_check:
-        return 0.7
+        return 0.5
     else:
         return 0
 
 
-results_files = {
-    "TLS": snakemake.input.tls,
-    "BLS": snakemake.input.bls,
-    "nuance": snakemake.input.nuance,
-}
+df = pd.read_csv(snakemake.input[0])
 
+df["bls_found"] = df.apply(
+    lambda row: is_found(row["bls_period"], row["period"]), axis=1
+)
+df["nuance_found"] = df.apply(
+    lambda row: is_found(row["nuance_period"], row["period"]), axis=1
+)
 
-methods = ["TLS", "BLS", "nuance"]
-plt.figure(None, (7.5, 2.7))
+bins = (23, 23)
+fig = plt.figure(None, (7.5, 3.5))
+cmap = "Greys_r"
 
-for i, method in enumerate(methods):
-    ax = plt.subplot(1, len(methods), i + 1)
-    files = results_files[method]
-    results = pd.DataFrame([yaml.safe_load(open(f, "r")) for f in files])
-    results["found"] = results.apply(
-        lambda row: is_found(row["period"], row["true_period"]), axis=1
-    )
-
-    period, depth, found = results[["true_period", "true_depth", "found"]].values.T
-    period_range = [period.min(), period.max()]
-    depth_range = np.array([depth.min(), depth.max()]) * 1e4
-    # plt.scatter(period, depth, c=found, s=10, cmap="Greys_r")
-
-    ax.imshow(
-        found.reshape(20, 20),
-        extent=[*period_range, *depth_range],
-        aspect="auto",
-        cmap="Greys_r",
+ax = plt.subplot(121)
+period, depth, found = df[["period", "depth", "bls_found"]].values.T
+depth *= 1e4
+stats = binned_statistic_2d(period, depth, found, bins=bins)
+if scatter:
+    plt.scatter(period, depth, c=found, s=10, cmap=cmap)
+else:
+    im = plt.imshow(
+        stats.statistic.T,
         origin="lower",
[…]

### 2023-09-19 workflows/sanity/scripts/tls_search.py

@@ -1,37 +1,0 @@
-import sys
-
-import numpy as np
-
-sys.path.append("../lib")
-import yaml
-from tls import tls
-
-time, _, error = np.load(snakemake.input.lc)
-flux = np.load(snakemake.input.injected)
-period_range = snakemake.config["period_range"]
-n_periods = snakemake.config["n_periods"]
-duration = snakemake.config["duration"]
-params = np.load(snakemake.input.params)
-i = int(snakemake.wildcards.i)
-periods = np.linspace(-0.1 + period_range[0], 0.1 + period_range[1], n_periods)
-
-model = tls(time, flux, verbose=False)
-results = model.power(
-    periods,
-    verbose=False,
-    use_threads=1,
-    show_progress_bar=False,
-    durations=[0.01, duration],
-)
-
-t0, period, power = results["T0"], results["period"], results["power"]
-yaml.safe_dump(
-    {
-        "t0": float(t0),
-        "period": float(period),
-        "true_period": float(params[i, 0]),
-        "true_t0": float(0),
-        "true_depth": float(params[i, 1]),
-    },
-    open(snakemake.output[0], "w"),
-)

### 2023-09-19 workflows/synthetic/scripts/plot.py

@@ -1,77 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-from scipy.stats import binned_statistic_2d
-
-
-def is_found(period, true_period):
-    period_check = np.abs(period - true_period) < 0.01
-    alias_check = np.abs(2 * period - true_period) < 0.01
-    alias_check |= np.abs(period / 2 - true_period) < 0.01
-    if period_check:
-        return 1
-    elif alias_check:
-        return 1
-    else:
-        return 0
-
-
-df = pd.read_csv(snakemake.input[0])
-
-df["bls_found"] = df.apply(
-    lambda row: is_found(row["bls_period"], row["period"]), axis=1
-)
-df["nuance_found"] = df.apply(
-    lambda row: is_found(row["nuance_period"], row["period"]), axis=1
-)
-
-bins = (22, 22)
-fig = plt.figure(None, (7.5, 3.5))
-cmap = "Greys_r"
-
-ax = plt.subplot(121)
-bins = (22, 22)
-tau, delta, found = df[["tau", "delta", "bls_found"]].values.T
-stats = binned_statistic_2d(tau, delta, found, bins=bins)
-im = plt.imshow(
-    stats.statistic.T,
-    origin="lower",
-    extent=(
-        stats.x_edge.min(),
-        stats.x_edge.max(),
-        stats.y_edge.min(),
-        stats.y_edge.max(),
-    ),
-    aspect="auto",
-    cmap=cmap,
-)
-ax.set_xlabel(r"$\tau$")
-ax.set_ylabel(r"$\delta$")
-ax.set_title("biweight+BLS")
-
-ax = plt.subplot(122)
-bins = (22, 22)
-tau, delta, found = df[["tau", "delta", "nuance_found"]].values.T
-stats = binned_statistic_2d(tau, delta, found, bins=bins)
-im = plt.imshow(
-    stats.statistic.T,
-    origin="lower",
-    extent=(
-        stats.x_edge.min(),
-        stats.x_edge.max(),
-        stats.y_edge.min(),
-        stats.y_edge.max(),
-    ),
-    aspect="auto",
-    cmap=cmap,
-)
-ax.set_xlabel(r"$\tau$")
-ax.set_title("nuance")
-axins = ax.inset_axes((1.07, 0.05, 0.05, 0.4))
[…]

### 2023-09-19 workflows/tess_injection_recovery/scripts/bls.py

@@ -1,0 +1,166 @@
+import sys
+
+import numpy as np
+
+sys.path.append("lib")
+import pickle
+
+import numpy as np
+import yaml
+from astropy.timeseries import BoxLeastSquares
+from scipy.interpolate import BSpline, splrep
+from wotan import flatten
+
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+periods = np.load(snakemake.input.periods)
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+
+time, flux, error = data["time"], data["flux"], data["error"]
+verbose = True
+
+# biweight
+# --------
+
+flatten_trend = flatten(
+    time, flux, window_length=data["transit_duration"] * 3, return_trend=True
+)[1]
+flatten_flux = flux - flatten_trend
+flatten_flux -= np.mean(flatten_flux)
+flatten_flux += 1.0
+
+model = BoxLeastSquares(time, flatten_flux, dy=error)
+results = model.power(periods, 0.03)
+
+t0 = 0.0
+period = results.period[np.argmax(results.power)]
+power = results.power
+
+pickle.dump(
+    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
+    open(snakemake.output.wotan3D, "wb"),
+)
+
+
+# harmonics
+# ---------
+
+
+# makes an harmonics design matrix of time
+def make_harmonics(time, period, nharmonics=4):
+    # make design matrix
+    X = np.ones((len(time), 2 * nharmonics + 1))
+    X[:, 1] = time
+    for i in range(1, nharmonics):
+        X[:, 2 * i + 1] = np.sin(2 * np.pi * i * time / period)
+        X[:, 2 * i + 2] = np.cos(2 * np.pi * i * time / period)
+
+    return X
+
+
+X = make_harmonics(time, info["star_period"], nharmonics=2)
+# solve for the coefficients
+coeffs = np.linalg.solve(X.T @ X, X.T @ flux)
+# make the model
+flatten_trend = X @ coeffs
+flatten_flux = flux - flatten_trend
+flatten_flux -= np.mean(flatten_flux)
+flatten_flux += 1.0
+
+model = BoxLeastSquares(time, flatten_flux, dy=error)
+results = model.power(periods, 0.03)
+
+t0 = 0.0
+period = results.period[np.argmax(results.power)]
+power = results.power
+
+pickle.dump(
+    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
[…]

### 2023-09-19 workflows/tess_injection_recovery/scripts/concatenate.py

@@ -1,0 +1,59 @@
+import pickle
+
+import numpy as np
+import pandas as pd
+import yaml
+
+
+def right_candidate(t0, period, true_t0, true_period, verbose=False):
+    t0_check = (
+        np.abs((t0 - true_t0 + 0.5 * true_period) % true_period - 0.5 * true_period)
+        % period
+        < 0.01
+    )
+    period_check = np.abs(period - true_period) < 0.01
+    period_check |= np.abs(2 * period - true_period) < 0.01
+    period_check |= np.abs(period / 2 - true_period) < 0.01
+    same = period_check  # np.logical_and(t0_check, period_check)
+    if verbose:
+        if not same:
+            if not t0_check:
+                output = f"t0 differ: {t0:.2e} {true_t0:.2e}"
+            if not period_check:
+                output = f"period differ: {period:.2e} {true_period:.2e}"
+        else:
+            output = "match"
+        return same, output
+    else:
+        return same
+
+
+def get_result(injected_file, recovered_file, fresh_file):
+    injected = pickle.load(open(injected_file, "rb"))
+    recovered = pickle.load(open(recovered_file, "rb"))
+    fresh = yaml.safe_load(open(fresh_file, "r"))
+    t0, period = recovered["t0"], recovered["period"]
+    true_t0, true_period = injected["transit_t0"], injected["planet_period"]
+    return {
+        "t0": t0,
+        "period": period,
+        "true_t0": true_t0,
+        "true_period": true_period,
+        "found": right_candidate(t0, period, true_t0, true_period),
+        "radius": injected["planet_radius"],
+        "tau": injected["tau"],
+        "delta": injected["delta"],
+        "fresh_snr": fresh["snr"],
+    }
+
+
+recovered = snakemake.input.recovered
+injected = snakemake.input.injected
+fresh = snakemake.input.fresh
+
+results = []
+
+for i, (a, b) in enumerate(zip(injected, recovered)):
+    results.append(get_result(a, b, fresh))
+results = pd.DataFrame(results)
+results.to_csv(snakemake.output[0], index=False)

### 2023-09-19 workflows/tess_injection_recovery/scripts/download.py

@@ -1,0 +1,92 @@
+import pickle
+
+import astropy.units as u
+import lightkurve as lk
+import numpy as np
+import pandas as pd
+import yaml
+
+tic = int(snakemake.wildcards["target"])
+filepath = snakemake.output[0]
+
+targets = pd.read_csv("static/tess-ultra-fast-rotators-brighter-mag14.csv")[
+    ["Name", "LS_Period", "Amplitude", "teff_val"]
+]
+
+targets = targets.sort_values("LS_Period")
+target = targets[targets.Name == tic]
+
+name = f"TIC {tic}"
+search_result = lk.search_lightcurve(name, author="SPOC", exptime=120)
+
+# data
+klc = search_result[0].download()
+times = klc.time.to_value("btjd")
+fluxes = klc.pdcsap_flux.to_value().filled(np.nan)
+errors = klc.flux_err.to_value().filled(np.nan)
+mask = np.isnan(fluxes) | np.isnan(errors) | np.isnan(times)
+original_time = times[~mask]
+original_flux = fluxes[~mask]
+original_error = errors[~mask]
+
+splits = np.array_split(
+    np.arange(len(original_time)), np.flatnonzero(np.diff(original_time) > 0.1) + 1
+)
+
+n = 300
+time = np.hstack([original_time[split[n:-n]] for split in splits])
+flux = np.hstack([original_flux[split[n:-n]] for split in splits])
+error = np.hstack([original_error[split[n:-n]] for split in splits])
+
+# first exposure
+first_exposure = klc.time[0]
+time_format = rf"{first_exposure.format.upper()}_{{{first_exposure.scale.upper()}}}"
+first_exposure_time = first_exposure.value
+
+flux_median = np.median(original_flux)
+time = time - first_exposure_time
+flux /= flux_median
+error /= flux_median
+
+# stellar parameters
+url = f"https://exofop.ipac.caltech.edu/tess/download_stellar.php?id={tic}"
+star = pd.read_csv(url, delimiter="|", index_col=1).iloc[0]
+
+pickle.dump(
+    {
+        "flux": flux,
+        "time": time,
+        "error": error,
+    },
+    open(snakemake.output.fluxes, "wb"),
+)
+
+
+star_dict = {
+    "tic": int(tic),
+    "time0": float(first_exposure_time),
+    "time0_format": str(time_format),
[…]

### 2023-09-19 workflows/tess_injection_recovery/scripts/nuance_search.py

@@ -1,0 +1,27 @@
+import pickle
+
+import numpy as np
+import yaml
+from nuance import Nuance
+from nuance.kernels import rotation
+
+gp_params = yaml.full_load(open(snakemake.input.gp, "r"))
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+periods = np.load(snakemake.input.periods)
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+
+build_gp, _ = rotation(info["star_period"])
+gp = build_gp(gp_params, data["time"])
+
+nu = Nuance(data["time"], data["flux"], gp=gp, c=500)
+
+nu.linear_search(data["time"], np.array([0.01, data["transit_duration"]]))
+
+search = nu.periodic_search(periods)
+
+output = snakemake.output[0]
+t0, _, period = search.best
+pickle.dump(
+    {"t0": t0, "period": period, "power": search.Q_snr, "trend": None},
+    open(snakemake.output[0], "wb"),
+)

### 2023-09-19 workflows/tess_injection_recovery/scripts/plot_comparison.py

@@ -1,0 +1,180 @@
+import pickle
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import yaml
+from matplotlib import gridspec
+from nuance.star import Star
+
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+data = pickle.load(open(snakemake.input.first_injected, "rb"))
+tic = snakemake.wildcards.target
+results = {
+    method: snakemake.input.results[i]
+    for i, method in enumerate(snakemake.params.methods)
+}
+first_recovered = {
+    method: snakemake.input.first_recovered[i]
+    for i, method in enumerate(snakemake.params.methods)
+}
+
+n_methods = len(snakemake.params.methods)
+
+from nuance import Nuance
+from nuance.kernels import rotation
+
+n = snakemake.config["n"]
+
+split_idxs = [
+    0,
+    *np.flatnonzero(np.diff(data["time"]) > 10 / 60 / 24),
+    len(data["time"]),
+]
+
+# nuance trend
+gp_params = yaml.safe_load(open(snakemake.input.gp, "r"))
+build_gp, _ = rotation(info["star_period"])
+nu = Nuance(data["time"], data["flux"], gp=build_gp(gp_params, data["time"]))
+nuance_trend = np.sum(nu.models(0, 0, 1), 0)
+
+
+def right_candidate(t0, period, true_t0, true_period, verbose=False):
+    t0_check = (
+        np.abs((t0 - true_t0 + 0.5 * true_period) % true_period - 0.5 * true_period)
+        % period
+        < 0.01
+    )
+    period_check = np.abs(period - true_period) <= 0.01
+    alias_check = np.abs(2 * period - true_period) <= 0.01
+    alias_check |= np.abs(period / 2 - true_period) <= 0.01
+    if period_check:
+        return 1
+    elif alias_check:
+        return 1.0
+    else:
+        return 0
+
+
+methods = snakemake.params.methods
+
+total_found = {}
+for method in methods.keys():
+    df = pd.read_csv(results[method])
+    df["found"] = df.apply(
+        lambda row: right_candidate(
+            row["t0"], row["period"], row["true_t0"], row["true_period"]
+        ),
+        axis=1,
+    )
+    total_found[method] = np.count_nonzero(df.found)
+
+# sorted name of methods total_found
[…]

### 2023-09-19 workflows/tess_injection_recovery/scripts/plot_nude_search copy.py

@@ -1,0 +1,28 @@
+import pickle
+
+import numpy as np
+import yaml
+from nuance import Nuance
+from nuance.kernels import rotation
+import matplotlib.pyplot as plt
+from nuance.utils import plot_search
+
+gp_params = yaml.full_load(open(snakemake.input.gp, "r"))
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+periods = np.load(snakemake.input.periods)
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+
+build_gp, _ = rotation()
+
+gp = build_gp(gp_params, data["time"])
+nu = Nuance(data["time"], data["flux"], gp=gp)
+
+Ds = np.linspace(0.01, 0.03, 5)
+# Ds = np.array([0.01, data["transit_duration"]])
+nu.linear_search(data["time"], Ds)
+
+search = nu.periodic_search(periods)
+
+plt.figure(figsize=(15, 5))
+plot_search(nu, search)
+plt.savefig(snakemake.output[0])

### 2023-09-19 workflows/tess_injection_recovery/scripts/plot_results.py

@@ -1,0 +1,97 @@
+# could not finish this one, but the idea is to plot the difference in detection rate
+
+import pandas as pd
+import numpy as np
+from glob import glob
+from pathlib import Path
+
+methods = {
+    "bls_bspline": "bspline + BLS",
+    "bls_wotan3D": "biweight + BLS",
+    "bls_harmonics": "harmonics + BLS",
+    "bens": "sinusoids + BLS (Ben's)",
+    "nuance": "nuance",
+}
+
+
+targets = [int(Path(f).stem) for f in glob("../figures/searched/*")]
+
+result_df = pd.DataFrame(columns=targets)
+
+
+def right_candidate(t0, period, true_t0, true_period, verbose=False):
+    t0_check = (
+        np.abs((t0 - true_t0 + 0.5 * true_period) % true_period - 0.5 * true_period)
+        % period
+        < 0.01
+    )
+    period_check = np.abs(period - true_period) <= 0.01
+    alias_check = np.abs(2 * period - true_period) <= 0.01
+    alias_check |= np.abs(period / 2 - true_period) <= 0.01
+    if period_check:
+        return 1
+    elif alias_check:
+        return 1
+    else:
+        return 0
+
+
+for target in targets:
+    for method in methods:
+        df = pd.read_csv(
+            f"../data/{target}/recovered/{method}/results.csv", index_col=0
+        )
+        df["found"] = df.apply(
+            lambda row: right_candidate(0, row["period"], 0, row["true_period"]),
+            axis=1,
+        )
+        total_found = np.count_nonzero(df.found) / len(df)
+
+        result_df.loc[method, target] = total_found * 100
+
+    result_df.loc["fresh_snr", target] = df.fresh_snr.values[0]
+
+
+gp_characteristics = {}
+
+import yaml
+
+for target in targets:
+    gp_characteristics[target] = yaml.safe_load(open(f"../data/{target}/gp.yaml"))
+
+gp_characteristics = pd.DataFrame(gp_characteristics)
+
+
+import matplotlib.pyplot as plt
+
+snr = 6
+
+# exclude nuance
+good_ones = result_df.loc[:, (result_df.loc["fresh_snr"] < snr).values]
+difference = good_ones.iloc[-2] - good_ones.drop("nuance").max(axis=0)
+gp_char = gp_characteristics[good_ones.columns]
+
+# grid of plots
+
[…]

### 2023-10-16 latex/paper.tex
update

@@
 
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.8\linewidth]{principle_Q.pdf}
+        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
+        \label{fig:periodogram}
+    \end{centering}
+\end{figure}
+
 Using \autoref{eq:result}, we can now compute $\ln\mathcal{P}$ for a range of periods and build a transit search periodogram using \autoref{eq:periodogram}.
@@
 The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (given by \autoref{eq:sigma_delta}).
-
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=0.8\linewidth]{principle_Q.pdf}
-        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
-        \label{fig:periodogram}
-    \end{centering}
-\end{figure}
 
@@
 
+\newpage
 \section{Validation on TESS light curves}\label{real}
 
-The parameter space ($\tau$, $\delta$) introduced in \autoref{detrending_effect} is very convenient from a signal processing point of view, but lacks physical reality.
+% In order to validate
+
+% The parameter space ($\tau$, $\delta$) introduced in \autoref{detrending_effect} is very convenient from a signal processing point of view, but lacks physical reality.
 Very low values of $\tau$ and $\delta$ often translate into non-physical exoplanets properties as well as unrealistic stellar variability, with a strong dependence on the host star properties.
 \\\\
-In order to assess the performance of \textsf{nuance} on real light curves, its transit recovery capabilities are rather evaluated in the planetary (period, radius) parameter space, more often encountered in the literature (e.g.
+% In order to assess the performance of \textsf{nuance} on real light curves, its transit recovery capabilities are rather evaluated in the planetary (period, radius) parameter space, more often encountered in the literature (e.g.
 Figure 12 of \citealt{Delrez2022}).
-In what follows, transits described by the simple model of \cite{protopapas} (see \autoref{app_proto}) are injected into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}).
-We focus this proof of concept on light curves from a list of $\sim$600 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{} (see \autoref{when}).
-For each of the studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) on a single sector (the first being observed for each target) spanning an average of 27 days.\\\\
-In the following sections, \textsf{nuance} is compared to other transit search strategies employing the \texttt{BLS} algorithm, each time after detrending the light curves with three different techniques:
+In what follows, transits described by the simple model of \cite{protopapas} (see \autoref{app_proto}) are
+
+In order to assess the performance of \textsf{nuance} on real datasets, we inject and recover transits into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}).
+The goal of this transit injection-recovery is to compare \textsf{nuance} to other transit search strategies, and to understand the light curves characteristics for which it becomes beneficial.\\\\
+We focus this proof of concept on light curves from a list of $\sim$438 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{}.
+For each of the studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) of a single sector (the first being observed for each target) spanning on average 10 days.\\\\
+In this injection-recovery tests, \textsf{nuance} is compared to other transit search strategies employing the \texttt{BLS} algorithm, for each of them after detrending the light curves with a different technique:
 \begin{itemize}
-    \item \texttt{bspline+BLS} employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the \texttt{BLS} algorithm.
+    \item \texttt{bspline+BLS} employs a B-spline model \footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the \texttt{BLS} algorithm.
     \item \texttt{biweight+BLS} employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package \citep{wotan} with an optimal window size of $3\times D$ (i.e.\, three times the transit duration), followed by a search with the \texttt{BLS} algorithm.
@@
 This detrending is followed by a search with the \texttt{BLS} algorithm.
+    \item \texttt{iterative+BLS} iteratively detrends the light curve with a sinusoidal signal fitted to the data, each time using the dominant period of the residuals found using a lomb-scargle periodogram (5 iterations).
+This detrending is followed by a search with the \texttt{BLS} algorithm.
 \end{itemize}
 
+\subsection{Light curve cleaning and transits injection}
+
+As some of these techniques can be affected by gaps in the data, we only use continuous measurements from half a TESS sector.
+We assume that all methods (including nuance) are based on an incomplete model of the data that do not account for stellar flares.
+For this reason, the light curve of each target is cleaned using an iterative sigma clipping approach.
+For each iteration, points 3 times above the standard deviation of the full light curve (previously subtracted by its median) are identified.
+Then, the 30 adjacent points on each side of the found outliers are masked.
+This way, large flare signals and their expected ingress and egress are masked, using a total of 3 iterations.
+As PDCSAP light curves often start with a ramp-like signal, the first 300 points (as well as the last 300 points) of each continuous observation are masked.
+Finally, each light curve is normalized by its median value.
+Such a cleaned light curve is shown in \autoref{fig:cleaned} (bottom top).
+We note that the gaps left after sigma clipping may be problematic for some of the detrending techniques (such as \texttt{bspline+BLS}).
+However, adopting this flare cleaning step and analyzing light curves with few small gaps is a practice commonly found in the literature.\\\\
+
+For each light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injections and recovery per target.
+Orbital periods $P$ are sampled on a regular grid between 0.4 and 5 days, and planetary radii $R_p$ are sampled on a regular grid designed to lead a minimum transit SNR of 4 and a maximum of 30.
+Using \autoref{eq:snr} with $\sigma_r = 0$, the planetary radius leading to a transit with a desired SNR $s$ is given by
+\begin{equation*}
+    R_p = R_{\star}\,n^{-\frac{1}{4}} \sqrt{\sigma s}\ %\sqrt[\leftroot{0}\uproot{2}4]{n}}. 
+\end{equation*}
+with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
+    \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692.
+Top plot shows how much of the data is truncated and sigma clipped, as to result in an almost continuous light curve shown in the bottom plot.
+On this bottom plot, the black line corresponds to the mean of one of the Gaussian Process model (whith hyperparameters optimized after each transit signal is injected).}
+    \label{fig:cleaned}
+\end{figure}
+
 \subsection{Stellar variability kernel}
-A Gaussian Process kernel with a single SHO kernel is not representative of realistic stellar variability.
-Instead, a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), being representative of a wide range of stochastic variability in stellar time series from rotation to pulsations\footlink{https://celerite2.readthedocs.io/en/latest/api/python}, is adopted, with the addition of one short scale and one long term exponential term.
-This full rotation kernel is expressed as
+One of the particularity of nuance is to model a light curve as a Gaussian Process.
+To model these TESS light curves, we employ a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
+In order to account for additional corelated noises, we complement this kernel with a short and a long-timescale exponential term, so that the full kernel can be expressed as
 \begin{equation*}
@@
 with $\ell$ and $\sigma$ the scale and standard deviation of the process.
-These are meant to model short and long timescale non-periodic correlated noise In total, the rotation kernel $k$ has 8 hyperparameters.
+These are meant to model short and long-timescale non-periodic correlated noise.
+In total, the rotation kernel $k$ has 8 hyperparameters.\\\\
 
-\subsection{Light curve cleaning}
-As the described light curve model does not account for stellar flares, the light curve of each target is cleaned using an iterative sigma clipping approach.
-For each iteration, points 4 times above the standard deviation of the full light curve (previously subtracted by its median) are identified.
-Then, the 30 adjacent points on each side of the found outliers are masked.
-This way, large flare signals and their expected ingress and egress are masked, using a total of 3 iterations.
-As PDCSAP light curves often start with a ramp-like signal, the first 300 points (as well as the last 300 points) of each continuous observation are masked.
-Finally, each light curve is normalized by its median value.
-Such a cleaned light curve is shown in \autoref{fig:cleaned}.
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\linewidth]{153193474_clean.png}
-    \caption{Cleaned single-sector light curve of the target TIC 153193474. The light curve is normalized by its median value.}
-    \label{fig:cleaned}
-\end{figure}
-
-\subsection{Injected transits}
-For each target light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injections and recovery per target.
-Orbital periods $P$ are sampled on a regular grid between 0.4 and 10 days, and planetary radii $R_p$ are sampled on a regular grid designed to lead a minimum transit SNR of 4 and a maximum of 30.
-Using \autoref{eq:snr} with $\sigma_r = 0$, the planetary radius leading to a transit with a desired SNR $s$ is given by
-\begin{equation*}
-    R_p = R_{\star}\,n^{-\frac{1}{4}} \sqrt{\sigma s}\ %\sqrt[\leftroot{0}\uproot{2}4]{n}}. 
-\end{equation*}
-with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.
-
-\subsection{Gaussian Process hyperparameters optimization}
-The hyperparameters of the employed rotation kernel are optimized on cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize}\footlink{https://jax.readthedocs.io/en/latest/_autosummary/jax.scipy.optimize.minimize.html} wrapper provided by the \textsf{jaxopt} Python package, taking advantage of the \textsf{JAX} implementation of \textsf{tinygp}.
+The hyperparameters of this kernel are optimized on trimmed and cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize}\footlink{https://jax.readthedocs.io/en/latest/_autosummary/jax.scipy.optimize.minimize.html} wrapper provided by the \textsf{jaxopt} Python package, and taking advantage of the \textsf{JAX} implementation of \textsf{tinygp}.
 As correlated noise is expected to affect the light curve uncertainty estimates performed by SPOC, the diagonal of the full covariance matrix of the data (i.e.\,their uncertainty, assuming homoscedasticity) is held free, increasing the number of optimized parameters to 9.
@@
 
-\subsection{Detection criteria}
-Because we also consider planets recovered with half or twice the real period, we ease the detection criteria by ignoring the match of the recovered transit epoch with the real epoch.
-Note that a visual vetting ensured that the found epochs were consistent with the ones injected for cases where the planet was considered detected.
+\subsection{Search parameters and transit detection criteria}
+For all techniques, we only search for transits with a duration fixed to the duration of the injected transits.
+This is mainly done for computational efficiency and allows for a narrower comparison between \texttt{nuance} and all \texttt{BLS}-based techniques.
+Finally, the search is done over 4000 trial periods linearly sampled from 0.3 to 6 days.
+Injection-recoveries using a wider parameter space are discussed in \autoref{}.\\\\
+Like in previous sections, we consider the detection of period aliases and ignore the match between the injected and recovered transit epochs (although we visually vet that the found epochs were consistent with the ones injected for cases where the planet was considered detected).
+We define true positives has transit signals with an SNR greater than 5 recovered with the correct period (or an alias) and an SNR greater than 5.
+False positives are signals detected with an SNR greater than 5 whereas the injected signal was lower than 5.
 
-\subsection{Injection-recovery studies}
-\subsubsection{TIC 140212114: fast sinusoidal variability}\label{140212114}
-TIC 140212114 has an effective temperature of 3595 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=140212114} (M1.5-type star, \citealt{Rajpurohit2013}) with a rotation period of 0.14 days \citep{Ramsay2020}.
-For each technique, the left plot of \autoref{fig:140212114} shows a portion of TIC 140212114 cleaned light curve with an injected transit signal (hardly distinguishable due to correlated noise of the light curve compared to the transit depth).
-On these plots the black lines correspond to the trends computed for this injected light curve subtracted from it before being searched for transits.
-This detrending only applies to techniques other than \textsf{nuance}, where instead the black line corresponds to the mean of the Gaussian Process model.
-For each technique, the right plot of \autoref{fig:140212114} shows the results of the transit search for each of the 10 $\times$ 10 injected light curves, reported in the planetary (period, radius) parameter space.
-A black square denotes that the transit signal has not been detected, a gray square denotes a signal detected at an alias period ($P/2$ or $2P$), and a white square denotes a transit signal detected with the correct period.
-On the right plots of \autoref{fig:140212114}, secondary axes show the values of the injected transits in the $(\tau, \delta)$ relative parameter space.\\\\
-Results from \autoref{fig:140212114} are discussed in \autoref{tess_results}.
+\subsection{Results}
+\autoref{fig:onesearch} shows the injection-recovery process and result for TIC 1019692.
+As with this target, we find that \texttt{nuance} leads to the highest number of true positives for 89\% of all targets, and the lowest rate of false positives for 69\% of them.
+When considering both true positives and false negatives, \texttt{nuance} is the best option in 65\% of all cases.\\\\
+\autoref{}
 
-\begin{figure}[H]
+\begin{figure}
     \begin{centering}
-        \includegraphics[width=0.8\linewidth]{140212114.pdf}
-        \caption{Results of the transits injection-recovery on TIC 140212114 single-sector light curve.
+        \includegraphics[width=0.8\linewidth]{../workflows/tess_injection_recovery/figures/searched/1019692.pdf}
+        \caption{Results of the transits injection-recovery on TIC 1019692 half-sector light curve.
 Left: cleaned light curve with computed trend overplotted in black (except for \textsf{nuance} where it corresponds to the mean of the Gaussian Process model).
@@
 For each method, the upper right legend on the left plot indicates its ranking based on the percent of recovered transit signals (where a transit with an aliased period counts as being detected).
-The \texttt{harmonics+TLS} method is the best after \textsf{nuance}, but with 35\% less transits detected in total.
         }
-        \label{fig:140212114}
+        \label{fig:onesearch}
     \end{centering}
 \end{figure}
-
-\subsubsection*{TIC 452793374: high-amplitude red noise}
-TIC 452793374 has an effective temperature of 3377 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=452793374} (M2.5-type star, \citealt{Rajpurohit2013}) with a rotation period of 0.33 days \citep{Ramsay2020}.
-Injection-recovery results are shown in \autoref{fig:452793374} and discussed in \autoref{tess_results}.
-
-\subsubsection*{TIC 153193474: non-sinusoidal variability}
-TIC 153193474 has an effective temperature of 3351 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=27734884} (M2.5-type star, \citealt{Rajpurohit2013}) with a rotation period of 0.11 days \citep{Ramsay2020}.
-Injection-recovery results are shown in \autoref{fig:153193474} and discussed in \autoref{tess_results}.
-
-\subsubsection*{TIC 306331621: slow sinusoidal variability}
-TIC 306331621 has an effective temperature of 3067 $\pm$ 157 K\footlink{https://exofop.ipac.caltech.edu/tess/target.php?id=306331621} (M5-type star, \citealt{Rajpurohit2013}) with a rotation period of 1.00 days \citep{Ramsay2020}.
-It was chosen for its relatively slow rotation (compared to the three other targets) and higher white noise compared to the amplitude of its variability.
-Injection-recovery results are shown in \autoref{fig:306331621} and discussed in \autoref{tess_results}.
 
@@
 \end{figure}
+
+\newpage
+\section{Extensive comparison on TIC ...}
 

### 2023-10-16 workflows/different_kernels/scripts/download.py

@@ -1,0 +1,75 @@
+import pickle
+
+import astropy.units as u
+import lightkurve as lk
+import numpy as np
+import pandas as pd
+import yaml
+
+tic = int(snakemake.wildcards["target"])
+filepath = snakemake.output[0]
+
+name = f"TIC {tic}"
+search_result = lk.search_lightcurve(name, author="SPOC", exptime=120)
+
+# data
+klc = search_result[0].download()
+times = klc.time.to_value("btjd")
+fluxes = klc.pdcsap_flux.to_value().filled(np.nan)
+errors = klc.flux_err.to_value().filled(np.nan)
+mask = np.isnan(fluxes) | np.isnan(errors) | np.isnan(times)
+original_time = times[~mask]
+original_flux = fluxes[~mask]
+original_error = errors[~mask]
+
+splits = np.array_split(
+    np.arange(len(original_time)), np.flatnonzero(np.diff(original_time) > 0.1) + 1
+)
+
+n = 300
+time = np.hstack([original_time[split[n:-n]] for split in splits])
+flux = np.hstack([original_flux[split[n:-n]] for split in splits])
+error = np.hstack([original_error[split[n:-n]] for split in splits])
+
+# first exposure
+first_exposure = klc.time[0]
+time_format = rf"{first_exposure.format.upper()}_{{{first_exposure.scale.upper()}}}"
+first_exposure_time = first_exposure.value
+
+flux_median = np.median(original_flux)
+time = time - first_exposure_time
+flux /= flux_median
+error /= flux_median
+
+# stellar parameters
+url = f"https://exofop.ipac.caltech.edu/tess/download_stellar.php?id={tic}"
+star = pd.read_csv(url, delimiter="|", index_col=1).iloc[0]
+
+pickle.dump(
+    {
+        "flux": flux,
+        "time": time,
+        "error": error,
+    },
+    open(snakemake.output.fluxes, "wb"),
+)
+
+
+star_dict = {
+    "tic": int(tic),
+    "time0": float(first_exposure_time),
+    "time0_format": str(time_format),
+    "star_radius": float(star["Radius (R_Sun)"]),
+    "star_mass": float(star["Mass (M_Sun)"]),
+    "star_teff": float(star["Teff (K)"]),
+    "star_logg": float(star["log(g)"]),
+    "sector": int(klc.sector),
+    "logg": float(star["log(g)"]),
+}
+
+yaml.safe_dump(
+    star_dict,
[…]

### 2023-10-16 workflows/different_kernels/scripts/inject.py

@@ -1,0 +1,39 @@
+import pickle
+
+import numpy as np
+import yaml
+from nuance.star import Star
+from nuance.utils import periodic_transit
+
+i = int(snakemake.wildcards.lc)
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+periods, radii = pickle.load(open(snakemake.input.params, "rb"))
+info = yaml.safe_load(open(snakemake.input.info, "r"))
+
+star = Star(
+    info["star_radius"], info["star_mass"], info["star_amplitude"], info["star_period"]
+)
+
+np.random.seed(i)
+period = periods[i]
+radius = radii[i]
+tau = star.period2tau(period)
+delta = star.radius2delta(radius)
+duration = star.transit_duration(period)
+depth = star.transit_depth(radius)
+
+t0 = np.random.rand() * period
+
+transit = depth * periodic_transit(data["time"], t0, duration, P=period, c=500)
+injected_flux = data["flux"] + transit
+
+data["flux"] = injected_flux
+data["transit_t0"] = t0
+data["transit_duration"] = duration
+data["transit_depth"] = depth
+data["delta"] = delta
+data["tau"] = tau
+data["planet_period"] = period
+data["planet_radius"] = radius
+output = snakemake.output[0]
+pickle.dump(data, open(output, "wb"))

### 2023-10-16 workflows/different_kernels/scripts/optimize_gp.py

@@ -1,0 +1,36 @@
+import pickle
+
+import matplotlib.pyplot as plt
+import numpy as np
+import yaml
+from nuance import Nuance
+from nuance.kernels import rotation
+
+data = pickle.load(open(snakemake.input.fluxes, "rb"))
+period = yaml.safe_load(open(snakemake.input.star_period, "r"))["period"]
+
+# only one continuous segment
+dt = np.diff(data["time"])
+mask = data["time"] < data["time"][np.flatnonzero(dt > 0.5)[0]]
+data["flux"] = data["flux"][mask]
+data["time"] = data["time"][mask]
+data["error"] = data["error"][mask]
+# sigma clip
+mask = np.abs(data["flux"] - np.median(data["flux"])) < 3 * np.std(data["flux"])
+data["flux"] = data["flux"][mask]
+data["time"] = data["time"][mask]
+data["error"] = data["error"][mask]
+
+build_gp, init = rotation(period, np.mean(data["error"]), long_scale=0.5)
+nu = Nuance(data["time"], data["flux"], data["error"])
+optimize, mu, nll = nu.gp_optimization(build_gp)
+
+new = optimize(
+    init, ["log_sigma", "log_short_scale", "log_short_sigma", "log_long_sigma"]
+)
+new = optimize(new)
+
+gp_output = snakemake.output[0]
+yaml.safe_dump(
+    {name: float(value) for name, value in new.items()}, open(gp_output, "w")
+)

### 2023-10-16 workflows/sanity/scripts/plot.py

@@ -29,7 +29,7 @@ df["nuance_found"] = df.apply(
 
 bins = (23, 23)
 fig = plt.figure(None, (7.5, 3.5))
-cmap = "Greys_r"
+cmap = "viridis"
 
 ax = plt.subplot(121)
 period, depth, found = df[["period", "depth", "bls_found"]].values.T

### 2023-10-16 workflows/tess_injection_recovery/scripts/bls.py

@@ -11,6 +11,30 @@ from astropy.timeseries import BoxLeastSquares
 from scipy.interpolate import BSpline, splrep
 from wotan import flatten
 
+
+def bls(time, flux, error):
+    model = BoxLeastSquares(time, flux, dy=error)
+    return model.power(periods, 0.03, objective="snr")
+
+
+def save_result(trend, results, filename):
+    t0 = 0.0
+    i = np.argmax(results.power)
+    period = results.period[i]
+    power = results.power
+
+    pickle.dump(
+        {
+            "t0": t0,
+            "period": period,
+            "power": power,
+            "trend": trend,
+            "snr": power[i],
+        },
+        open(filename, "wb"),
+    )
+
+
 data = pickle.load(open(snakemake.input.fluxes, "rb"))
 periods = np.load(snakemake.input.periods)
 info = yaml.safe_load(open(snakemake.input.info, "r"))
@@ -22,23 +46,20 @@ verbose = True
 # --------
 
 flatten_trend = flatten(
-    time, flux, window_length=data["transit_duration"] * 3, return_trend=True
+    time,
+    flux,
+    window_length=data["transit_duration"] * 3,
+    return_trend=True,
+    robust=True,
 )[1]
+
 flatten_flux = flux - flatten_trend
 flatten_flux -= np.mean(flatten_flux)
 flatten_flux += 1.0
 
-model = BoxLeastSquares(time, flatten_flux, dy=error)
-results = model.power(periods, 0.03)
+results = bls(time, flatten_flux, error)
 
-t0 = 0.0
-period = results.period[np.argmax(results.power)]
-power = results.power
-
-pickle.dump(
-    {"t0": t0, "period": period, "power": power, "trend": flatten_trend},
-    open(snakemake.output.wotan3D, "wb"),
-)
+save_result(flatten_trend, results, snakemake.output.wotan3D)
 
 
 # harmonics
@@ -57,7 +78,7 @@ def make_harmonics(time, period, nharmonics=4):
     return X
 
 
-X = make_harmonics(time, info["star_period"], nharmonics=2)
+X = make_harmonics(time, info["star_period"], nharmonics=5)
 # solve for the coefficients
 coeffs = np.linalg.solve(X.T @ X, X.T @ flux)
 # make the model
@@ -66,17 +87,9 @@ flatten_flux = flux - flatten_trend
 flatten_flux -= np.mean(flatten_flux)
[…]

### 2023-10-16 workflows/tess_injection_recovery/scripts/concatenate.py

@@ -32,7 +32,7 @@ def get_result(injected_file, recovered_file, fresh_file):
     injected = pickle.load(open(injected_file, "rb"))
     recovered = pickle.load(open(recovered_file, "rb"))
     fresh = yaml.safe_load(open(fresh_file, "r"))
-    t0, period = recovered["t0"], recovered["period"]
+    t0, period, snr = recovered["t0"], recovered["period"], recovered["snr"]
     true_t0, true_period = injected["transit_t0"], injected["planet_period"]
     return {
         "t0": t0,
@@ -44,6 +44,7 @@ def get_result(injected_file, recovered_file, fresh_file):
         "tau": injected["tau"],
         "delta": injected["delta"],
         "fresh_snr": fresh["snr"],
+        "snr": snr,
     }
 
 

### 2023-10-16 workflows/tess_injection_recovery/scripts/nuance_search.py

@@ -13,15 +13,16 @@ info = yaml.safe_load(open(snakemake.input.info, "r"))
 build_gp, _ = rotation(info["star_period"])
 gp = build_gp(gp_params, data["time"])
 
-nu = Nuance(data["time"], data["flux"], gp=gp, c=500)
+nu = Nuance(data["time"], data["flux"], gp=gp, c=100)
 
 nu.linear_search(data["time"], np.array([0.01, data["transit_duration"]]))
 
 search = nu.periodic_search(periods)
 
 output = snakemake.output[0]
-t0, _, period = search.best
+t0, D, period = search.best
+snr = nu.snr(t0, D, period)
 pickle.dump(
-    {"t0": t0, "period": period, "power": search.Q_snr, "trend": None},
+    {"t0": t0, "period": period, "power": search.Q_snr, "trend": None, "snr": snr},
     open(snakemake.output[0], "wb"),
 )

### 2023-10-16 workflows/tess_injection_recovery/scripts/optimize_gp.py

@@ -9,9 +9,11 @@ from nuance.kernels import rotation
 data = pickle.load(open(snakemake.input.fluxes, "rb"))
 info = yaml.safe_load(open(snakemake.input.info, "r"))
 
-# only one continuous segment
+# only one continuous segment (removed later so commented)
 dt = np.diff(data["time"])
-mask = data["time"] < data["time"][np.flatnonzero(dt > 0.5)[0]]
+mask = np.ones_like(
+    data["time"], dtype=bool
+)  # < data["time"][np.flatnonzero(dt > 0.5)[0]]
 data["flux"] = data["flux"][mask]
 data["time"] = data["time"][mask]
 data["error"] = data["error"][mask]

### 2023-10-16 workflows/tess_injection_recovery/scripts/params.py

@@ -13,16 +13,16 @@ star = Star(
 )
 
 min_period = 0.4
-max_period = 10.0
+max_period = 5.0
 max_snr = 30.0
-min_snr = 4.0
+min_snr = 2.0
 # -----
 
 dt = np.median(np.diff(data["time"]))
 sigma = np.mean(data["error"])
 N = len(data["time"])
 
-max_radius = star.min_radius(max_period, max_snr, N, sigma)
+max_radius = star.min_radius(min_period, max_snr, N, sigma)
 min_radius = star.min_radius(max_period, min_snr, N, sigma)
 
 period_radius = np.array(

### 2023-10-16 workflows/tess_injection_recovery/scripts/periods.py

@@ -4,7 +4,7 @@ from nuance.utils import clean_periods
 
 info = yaml.safe_load(open(snakemake.input[0], "r"))
 
-periods = np.linspace(0.3, 11.0, 8000)
+periods = np.linspace(0.3, 6.0, 4000)
 periods = clean_periods(periods, info["star_period"])
 
 output = snakemake.output[0]

### 2023-10-16 workflows/tess_injection_recovery/scripts/plot_comparison.py

@@ -51,7 +51,7 @@ def right_candidate(t0, period, true_t0, true_period, verbose=False):
     if period_check:
         return 1
     elif alias_check:
-        return 1.0
+        return 0.5
     else:
         return 0
 

### 2023-10-16 workflows/tess_injection_recovery/scripts/plot_lc.py

@@ -1,0 +1,73 @@
+import pickle
+
+import jax
+import matplotlib.pyplot as plt
+import numpy as np
+import yaml
+from nuance import Nuance
+from nuance.kernels import rotation
+
+data_path = snakemake.input.fluxes
+gp_path = snakemake.input.gp
+info_path = snakemake.input.info
+raw_path = snakemake.input.raw
+
+original_data = pickle.load(open(raw_path, "rb"))
+data = pickle.load(open(data_path, "rb"))
+gp_params = yaml.full_load(open(gp_path, "r"))
+info = yaml.full_load(open(info_path, "r"))
+
+build_gp, init = rotation(info["star_period"], np.mean(data["error"]), long_scale=0.5)
+gp = build_gp(gp_params, data["time"])
+
+nu = Nuance(data["time"], data["flux"], gp=gp)
+optimize, mu, nll = nu.gp_optimization(build_gp)
+gp_mean = mu(gp_params)
+
+
+plt.figure(None, (8.5, 5))
+ax = plt.subplot(211)
+# ax.xaxis.set_ticklabels([])
+ax.set_title(
+    f"TIC {info['tic']} PDCSAP cleaned light curve (sector {int(info['sector'])})"
+)
+plt.plot(
+    original_data["time"],
+    original_data["flux"],
+    ".",
+    c="k",
+    label="original data",
+    ms=3,
+)
+plt.plot(data["time"], data["flux"], ".", c="0.8", ms=3.5)
+plt.ylabel("diff. flux")
+plt.xlim(original_data["time"][0], original_data["time"][-1])
+ylim = ax.get_ylim()
+plt.legend()
+
+plt.subplot(212)
+plt.plot(
+    data["time"], data["flux"], ".", c="0.8", label="trimmed and cleaned data", ms=3.5
+)
+split_idxs = [
+    0,
+    *np.flatnonzero(np.diff(data["time"]) > 10 / 60 / 24),
+    len(data["time"]),
+]
+
+_ = True
+for i in range(len(split_idxs) - 1):
+    x = data["time"][split_idxs[i] + 1 : split_idxs[i + 1]]
+    y = gp_mean[split_idxs[i] + 1 : split_idxs[i + 1]]
+    plt.plot(x, y, "k", label="GP mean" if _ else None)
+    _ = False
+plt.xlim(data["time"][0], data["time"][-1])
+plt.legend()
[…]

### 2023-12-02 latex/paper.tex
dev

@@
 % commands
-\newcommand{\nuance}{\texttt{nuance}}
+\newcommand{\nuancemethod}{\textit{nuance}}
+\newcommand{\nuance}{\nuancemethod{}}
+\newcommand{\nuancecode}{\textsf{nuance}}
+
 \newcommand{\TODO}{\texttt{todo}}
@@
 \affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
+\author{Dax L. Feliz}
+\author{Catriona A. Murray}
+\author{Francisco J. Pozuelos}
 \correspondingauthor{Lionel J. Garcia}
-\email{lionel\_garcia@live.fr}
+\email{lgarcia@flatironinstitute.org}
 
@@
 The latter study explores a handful of cases and generally discards the benefit of using a full-fledged approach.
-However, it does not explore the light-curves characteristics for which such a complete modeling becomes necessary.
-\\\\
-In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, an we present \nuance{}, a method to search for transit signals while modeling correlated noises in a tractable way.
+However, it does not explore the light-curves characteristics for which such a complete modeling becomes necessary.\\\\
+In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, when referring to the \nuancemethod{} method, we typeset \nuancemethod{} in italics and, in reference to the implementation of this model, we typeset \nuancecode{} in sans-serif.}, a method to search for transit signals while modeling correlated noises in a tractable way.
+In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on planetary signals detectability.
+In \autoref{nuance}, we present \nuance{}.
+In \autoref{results}, we test the performance of \nuance{} on a wide variety of cases, and compare it to commonly used transit search algorithms.
+Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, and we conclude in \autoref{conclusion}.
 
-\section{The degradation of transit detectability}\label{issues}
+\newpage
+\section{The issue}\label{issues}
 
@@
 One of them is physically-motivated and makes use of Gaussian processes (e.g.
-\citealt{k2sc}), another is empirical and makes use of filtering algorithms (\citealt{Jenkins2010}, \citealt{wotan}).
+\citealt{k2sc}), another is empirical and makes use of filtering algorithms (\citealt{Jenkins2010}, \citealt{wotan}).\\\\
+
+In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a Gaussian Process with an SHO kernel described in \autoref{app_gp}.
+For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same Gaussian Process from which the data has been sampled.
+We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
+\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that the degradation is strongly dependant on the correlated noise characteristics encountered.
 
@@
 \end{figure}
-
-In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a Gaussian Process with an SHO kernel described in \autoref{app_gp}.
-For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same Gaussian Process from which the data has been sampled.
-We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
-
-\noindent \autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that the degradation is strongly dependant on the correlated noise characteristics encountered.\\\\
 In order to explore the parameter space for which detrending is the most problematic, we employ the method previously described to simulate 10 000 light curves containing a transit signal and correlated noise in the form of stellar variability.
@@
 For $\tau=1$ and $\delta=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a variability signal with a period half that of the transit duration, and a standard deviation two times that of the transit depth, i.e.
-strongly resembling the simulated transit signal.
-
+strongly resembling the simulated transit signal.\\\\
 \autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 6$).
@@
 
-\textsf{nuance} is an algorithm capable of searching for planetary transits in light curves containing correlated noise, such as instrumental signals and photometric stellar variability.
+\textit{nuance} is an algorithm capable of searching for planetary transits in light curves containing correlated noise, such as instrumental signals and photometric stellar variability.
 \\\\
@@
 \end{equation}
-This is done by adopting the strategy of \cite{foreman2016}, and separate the transit search into two components: the linear search and the periodic search.
-During the linear search, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
+This is done by adopting the strategy of \cite{foreman2016}, and separate the transit search into two components: the \textit{linear search} and the periodic search.
+During the \textit{linear search}, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
 Then, the periodic search consists in combining these likelihoods to compute the likelihood of the data given a periodic transit signal for a range of periods.
@@
 
-During the linear search, the goal is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of the data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths.
+During the \textit{linear search}, the goal is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of the data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths.
 \\\\
@@
 In this equation.
-This linear search leads to the set of likelihoods
+This \textit{linear search} leads to the set of likelihoods
 \begin{equation*}
@@
         \includegraphics[width=0.8\linewidth]{principle_linear_search.pdf}
-        \caption{Principle and output of the linear search.
+        \caption{Principle and output of the \textit{linear search}.
 The simulated dataset (top) corresponds to the one shown and described in \autoref{fig:app_principle_dataset}.
@@
 
-We then need to combine the likelihoods computed from the linear search to obtain
+We then need to combine the likelihoods computed from the \textit{linear search} to obtain
 \begin{equation*}
@@
 $$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi, D)}_{i,j}$$
-shown for few periods in \autoref{fig:periodic_search} (b). In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the linear search using $\phi_i$, $D_j$ and $T_0 = 0$.
+shown for few periods in \autoref{fig:periodic_search} (b).
+In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the \textit{linear search} using $\phi_i$, $D_j$ and $T_0 = 0$.
 
@@
 But a final issue emerges, one that is fundamentally linked to our strategy.
-Each likelihood $p(\bm{f} \vert T, D, \Delta)$ estimated during the linear search is computed using N measurements.
+Each likelihood $p(\bm{f} \vert T, D, \Delta)$ estimated during the \textit{linear search} is computed using N measurements.
 Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving $N\times K$ measurements.
@@
 \textsf{nuance} exploits the use of \textsf{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package powered by \textsf{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}, allowing for custom kernels to be built and highly tractable computations.
-We can then define a set of epochs \textsf{t0s} and durations \textsf{Ds} and run the linear search with
+We can then define a set of epochs \textsf{t0s} and durations \textsf{Ds} and run the \textit{linear search} with
 \begin{lstlisting}[language=Python,linewidth=\linewidth]
@@
 From this \texttt{search} object, the best transiting candidate parameters can be computed (\textsf{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\textsf{search.Q\_snr}), together with valuable information about the transit search.
-The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \textsf{nuance} being that the linear search only needs to be performed ones, and reused for the search of several transiting candidates.
+The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \textsf{nuance} being that the \textit{linear search} only needs to be performed ones, and reused for the search of several transiting candidates.
 An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.
 
-\section{Control test against BLS}\label{control}
+\subsection{Comparison with BLS}\label{control}
 
@@
     \begin{centering}
-        \includegraphics[width=0.85\linewidth]{../workflows/sanity/figures/control_test.pdf}
+        \includegraphics[width=0.85\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
         \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise (whose characteristics are described in the main text), using \texttt{BLS} and \textsf{nuance}.
@@
 
-\section{\textsf{nuance} \textit{vs.} \texttt{biweight+BLS} on simulated light curves}\label{simu}
+\section{Results}\label{results}
+\subsection{Comparison with \texttt{biweight+BLS} on simulated light curves}\label{simu}
 \autoref{fig:snr_detrend} shows that \textsf{nuance}'s full-fledged modeling capabilities may not always be necessary and may only be beneficial for certain noise characteristics, relative to the searched transit parameters.
@@
 
-\subsection{Dataset}\label{sim_dataset}
+\subsubsection*{Dataset}\label{sim_dataset}
 The dataset consists in 4000 light curves simulated using the model fully described in \autoref{signals_simulations}.
@@
 
-\subsection{Results}
+\subsubsection*{Results}
 The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against the \wtls{} strategy on transits with relatively low depth compared to the stellar variability amplitude, and a relatively small duration compared to the stellar variability period.
@@
 \newpage
-\section{Validation on TESS light curves}\label{real}
+\subsection{Comparison with various techniques on limited TESS light curves}\label{real}
 
@@
 The goal of this transit injection-recovery is to compare \textsf{nuance} to other transit search strategies, and to understand the light curves characteristics for which it becomes beneficial.\\\\
-We focus this proof of concept on light curves from a list of $\sim$438 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{}.
+We focus this proof of concept on light curves from a list of 438 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{}.
 For each of the studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) of a single sector (the first being observed for each target) spanning on average 10 days.\\\\
@@
 \begin{itemize}
-    \item \texttt{bspline+BLS} employs a B-spline model \footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the \texttt{BLS} algorithm.
+    \item \texttt{bspline+BLS} employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the \texttt{BLS} algorithm.
     \item \texttt{biweight+BLS} employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package \citep{wotan} with an optimal window size of $3\times D$ (i.e.\, three times the transit duration), followed by a search with the \texttt{BLS} algorithm.
@@
 
-\subsection{Light curve cleaning and transits injection}
+\subsubsection*{Light curve cleaning and transits injection}
 
@@
 Finally, each light curve is normalized by its median value.
-Such a cleaned light curve is shown in \autoref{fig:cleaned} (bottom top).
+Such a cleaned light curve is shown in \autoref{fig:cleaned}.
 We note that the gaps left after sigma clipping may be problematic for some of the detrending techniques (such as \texttt{bspline+BLS}).
 However, adopting this flare cleaning step and analyzing light curves with few small gaps is a practice commonly found in the literature.\\\\
-
 For each light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injections and recovery per target.
@@
     \centering
-    \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
+    \includegraphics[width=0.9\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
     \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692.
-Top plot shows how much of the data is truncated and sigma clipped, as to result in an almost continuous light curve shown in the bottom plot.
-On this bottom plot, the black line corresponds to the mean of one of the Gaussian Process model (whith hyperparameters optimized after each transit signal is injected).}
+Top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
+On this bottom plot, the black line corresponds to the mean of one of the Gaussian Process model (with hyperparameters optimized here on the cleaned light curve).}
     \label{fig:cleaned}
@@
 
-\subsection{Stellar variability kernel}
+\subsubsection*{Stellar variability kernel}\label{rotation_kernel}
 One of the particularity of nuance is to model a light curve as a Gaussian Process.
@@
 For simplicity, and to adopt a uniform treatment for all target light curves, a design matrix $\bm{X}$ with a single constant column is adopted, such that the systematic model only consists in a single parameter corresponding to the mean value of the differential flux (expected to be close to 1) solved linearly.
+The initial value for the rotation period of the star is found using a lomb-scargle periodogram \citep{Lomb1976, Scargle1982} implemented in \textsf{astropy}\footlink{https://docs.astropy.org/en/stable/timeseries/lombscargle.html}.\\\\
 
-\subsection{Search parameters and transit detection criteria}
+\subsubsection*{Search parameters and transit detection criteria}
 For all techniques, we only search for transits with a duration fixed to the duration of the injected transits.
@@
 Finally, the search is done over 4000 trial periods linearly sampled from 0.3 to 6 days.
-Injection-recoveries using a wider parameter space are discussed in \autoref{}.\\\\
-Like in previous sections, we consider the detection of period aliases and ignore the match between the injected and recovered transit epochs (although we visually vet that the found epochs were consistent with the ones injected for cases where the planet was considered detected).
-We define true positives has transit signals with an SNR greater than 5 recovered with the correct period (or an alias) and an SNR greater than 5.
-False positives are signals detected with an SNR greater than 5 whereas the injected signal was lower than 5.
+Realistic transit searches on a wider parameter space (e.g.
+multiple trial durations) are discussed in \autoref{}.\\\\
+Like in previous sections, we consider the detection of period aliases and ignore the match between the injected and recovered transit epochs (although we visually vet that the found epochs were consistent with the ones injected for cases where the planet was detected).
+We consider a transit detectable if its original SNR is greater than 6.
+Hence we define \textit{true positives} has detectable transits recovered with the correct period (or an alias) and a measured SNR greater or equal to 6, and \textit{false positives} as non-detectable transits recovered with a measured SNR greater or equal to 6.
 
-\subsection{Results}
-\autoref{fig:onesearch} shows the injection-recovery process and result for TIC 1019692.
-As with this target, we find that \texttt{nuance} leads to the highest number of true positives for 89\% of all targets, and the lowest rate of false positives for 69\% of them.
-When considering both true positives and false negatives, \texttt{nuance} is the best option in 65\% of all cases.\\\\
-\autoref{}
-
-\begin{figure}
+\newpage
+\subsubsection*{Results}
+An example of the transit injection-recovery and its result is shown in \autoref{1019692} for the target TIC 1019692.
+\autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in).
+Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (10950 with the non-dectable ones), and the lowest number of false negatives (\autoref{fig:allsearch}).
+\begin{figure}[H]
     \begin{centering}
-        \includegraphics[width=0.8\linewidth]{../workflows/tess_injection_recovery/figures/searched/1019692.pdf}
-        \caption{Results of the transits injection-recovery on TIC 1019692 half-sector light curve.
-Left: cleaned light curve with computed trend overplotted in black (except for \textsf{nuance} where it corresponds to the mean of the Gaussian Process model).
-Right: Results of the transit search where a black square denotes a transit signal not detected, gray a signal detected at an alias period ($P/2$ or $2P$), and  white a signal detected with the correct period.
-On the right plots, secondary axes show the $(\tau, \delta)$ relative parameter space.
-For each method, the upper right legend on the left plot indicates its ranking based on the percent of recovered transit signals (where a transit with an aliased period counts as being detected).
-        }
-        \label{fig:onesearch}
+        \includegraphics[width=0.9\linewidth]{../workflows/tess_injection_recovery/figures/true_and_false_positives_bars.pdf}
+        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals.}
+        \label{fig:allsearch}
     \end{centering}
 \end{figure}
-
-\subsection{Results}\label{tess_results}
-
-From the injection-recovery conducted on TIC 140212114, TIC 452793374, TIC 153193474 and TIC 306331621, it can be concluded that searching for transits directly with \textsf{nuance} offers a clear advantage over any of the presented detrending methods combined with \texttt{BLS}.
-While not fully described in this dissertation, injection-recovery on additional targets lead to a similar conclusion.\\\\
-Targets with light curves featuring stable sinusoidal stellar variability, i.e.\,with a high quality factor $Q$ (see the SHO kernel), are the ones that benefit less from \textsf{nuance}, as a set of harmonic signals can be combined linearly to robustly model and detrend the periodic signal.
-In the cases of TIC 452793374 and TIC 306331621, an additional $\sim$6\% of planets are detected by \textsf{nuance} over the probed planetary (period, radius) parameter space.
-From these two examples, it cannot be concluded if this difference is solely due to the difference observed between \textsf{nuance} and \texttt{BLS} in \autoref{control} .
-While red noise was properly detrended by a bi-weight filter in the case of TIC 452793374, additional tests must be done to show that \textsf{nuance} clearly outperforms this method when any red noise is present in the light curve (as partially observed in \autoref{simu}).
-\\\\
-Finally, targets with non-sinusoidal or short scale variability are the ones that benefit the most from \textsf{nuance}, as no other method is able to robustly detrend their light curves  without removing transit signals.
-This is well illustrated by the study of TIC 140212114 (\autoref{fig:140212114}) and TIC 153193474 (\autoref{fig:153193474}).
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{452793374.pdf}
-        \caption{Results of the transits injection-recovery on TIC 452793374 single-sector light curve.
-See \autoref{140212114} and legend of \autoref{fig:140212114} for a detailed description of this figure.
-The \texttt{biweight+TLS} method is the best after \textsf{nuance}, with 6\% less transits detected in total that could be due to the sole difference between \textsf{nuance} and \texttt{BLS} discussed in \autoref{control}.
-It indicates that a more flexible filter is necessary for this light curve with significantly higher red noise, making \textsf{nuance} not necessary.
-        }
-        \label{fig:452793374}
-    \end{centering}
-\end{figure}
+While the performances of other methods highly depend on the characteristics of the variability, \nuance{} leads to the highest number of true positives in 79\% of cases, and the lowest number of false positives in 72\% of cases (\autoref{fig:allsearch_relative}).
+When considering both true and false positives simultaneously, \nuance{} can be considered the best technique in only 55\% of cases.
+However, when considering transits with a relative timescale $\tau < 5$ (transits that justified the development of this algorithm), \nuance{} is the best technique in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives.
 
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{153193474.pdf}
-        \caption{Results of the transits injection-recovery on TIC 153193474 single-sector light curve.
-See \autoref{140212114} and legend of \autoref{fig:140212114} for a detailed description of this figure.
-The \texttt{harmonics+TLS} method is the best after \textsf{nuance}, with 62\% less transits detected in total, indicating that all presented methods are unable to detrend without removing many of the injected transit signals (which is not surprising when considering the trend being built, except in the case of \texttt{bspline+TLS}, but being apparently too flexible).}
-        \label{fig:153193474}
-    \end{centering}
-\end{figure}
-
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{306331621.pdf}
-        \caption{Results of the transits injection-recovery on TIC 306331621 single-sector light curve.
-See \autoref{140212114} and legend of \autoref{fig:140212114} for a detailed description of this figure.
-The \texttt{harmonics+TLS} method is the best after \textsf{nuance}, with 7\% less transits detected in total that could, again, be due to the sole difference between \textsf{nuance} and \texttt{BLS} discussed in \autoref{control}.}
-        \label{fig:306331621}
+        \includegraphics[width=0.9\linewidth]{../workflows/tess_injection_recovery/figures/true_false_positives.pdf}
+        \caption{Rate of true and false positives of \nuance{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
+The solid red line corresponds to the maximum of true positives among all methods on the top panel, and the minimum of false positives on the bottom panel.}
+        \label{fig:allsearch_relative}
     \end{centering}
@@
 \newpage
-\section{Extensive comparison on TIC ...}
+\subsection{Comparison for a multi-sector TESS candidate: TOI-540}\label{toi540}
 
-\newpage
-\section{Discussion}\label{perf}
-
-The results from \autoref{real} show that \textsf{nuance} always outperforms other methods when searching for transit signals in light curves containing correlated noise.
-While performances with other methods can be very comparable, \textsf{nuance} alleviates the need to choose for a specific detrending technique and rather exploit the complete modeling of the light curve using a physically interpretable Gaussian Process.\\\\
-In few cases, \textsf{nuance} offers a clear advantage in comparison with other methods, in particular for low SNR transit signals in light curves containing fast or non-sinusoidal stellar variability.
-In these cases, \textsf{nuance} is able to robustly detect transits, and provide a unique option in comparison with techniques that likely degrade the searched signals through detrending.
-
-\subsection*{Processing time}
-As of June 2023, \textsf{nuance} is very much a work in progress and a full assessment of its performances is still ongoing.
-\autoref{fig:blsvsnuance} is a preliminary study of the computational cost of \textsf{nuance} against \texttt{BLS}.
-For reasons described in the next section, the linear search is first considered separately from the periodic search.
+In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b.
+We download the 2 min cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
+Like in the previous section, we use a Lomb-Scargle periodogram and identify the 0.72 days rotation period of the star, that we use as an initial value to optimize the kernel described in \autoref{rotation_kernel} on each sector independently.
+Here again, we employ an upper-sigma-clipping to mask flares out of the data.
+The resulting light curve for sector 4 and its mean model is shown in \autoref{fig:toi540_clean}.
 
@@
     \begin{centering}
-        \includegraphics[width=0.8\linewidth]{nuance_vs_bls_time.pdf}
-        \caption{Preliminary study of \textsf{nuance} processing time against \texttt{BLS}.
-The lighter blue curve shows the performance of \textsf{nuance} when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together (the only option when using \texttt{BLS}).}
+        \includegraphics[width=0.9\linewidth]{../workflows/comparison_toi/figures/TOI 540/4.pdf}
+        \caption{Sector 4 light curve of TOI 540.
+The clean signal (gray points) has been mask masked for flares (black points), and the black line corresponds to the mean of the Gaussian Process model.}
+        \label{fig:toi540_clean}
+    \end{centering}
+\end{figure}
+
+For each sector, we perform the \textit{linear search} of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
+We then perform the periodic search on all sectors combined, using a concatenation of all linear searches.
+By adopting this by-sector GP modeling of the light-curve, the \textit{linear search} of \nuance{} scales linearly with the number of sectors being processed.
+This approach is adopted for efficiency but also to encapsulate the changing properties of stellar variability from one sector to another, often separated by year-long gaps.
+The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days.
+This search, using \nuance{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the \texttt{BLS} algorithm (denoted \texttt{biweight+BLS} and described in \autoref{real}) on all sectors combined.
+Since we don't know the transit duration a-priori, we employ a sliding window duration of 5 hours, small enough to encompass the stellar variability of TOI-540 but large enough not to detrend existing transits with durations lower tham 1.5 hours.
+The results of this comparison are shown in \autoref{fig:toi540_periodograms}.
+
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.9\linewidth]{../workflows/comparison_toi/figures/TOI 540/periodograms.pdf} 
+        \caption{Transit search SNR periodograms of TOI-540 using \texttt{biweight+BLS} and \nuance{}.
+After a first periodic search (left panel), the epochs corresponding to the maximum-SNR transit are masked before the second search is performed (right panel).}
+        \label{fig:toi540_periodograms}
+    \end{centering}
+\end{figure}
+
+After a first periodic search, the times falling in the detected periodic transit signal (the one with an SNR higher than the detection limit set to 7) are masked.
+In practice, this is done by masking the linear search products \\\\
+As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{biweight+BLS} and \texttt{nuance} are very similar.
+This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b (using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ min of TOI-540 b \citep{TOI540}).
+Hence, as $\tau < 7$, the use of \nuance{} on TOI-540 is not expected to be very beneficial (see \autoref{fig:snr_detrend}).
+Nonetheless, the \nuance{} periodogram of TOI 540 features less spurious SNR peaks, largely due to the penalty naturally occurring when single transits with different depths are periodically combined.
+In the second search (right panel) of \autoref{fig:toi540_periodograms}, we also notice a higher number of peaks that would lead to false positive detections of transits in the \texttt{biweight+BLS} case.
+The proper treatment of correlated noise in \nuance{}, as observed in \autoref{real}, makes these peaks non-significant, avoiding a large number of false detections.\\\\
+We note that finding a TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques.
+We reserve the search for such new transit signals to a follow-up publication.
+
+\section{Discussion}\label{discussion}
+
+In the previous section we demonstrated the capability of \nuance{} to search for synthetic or known transit signals, in simulated or real datasets.
+Here, we discuss the caveats of this algorithm, the advantages and limitations of the \nuancecode{} implementation, and future prospects for its extension.
+
+\subsection{Processing time}
+
+\noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple non-optimized GP with a squared exponential kernel.
+These are compared to the processing time of \texttt{biweight+BLS} (cf.\;\autoref{simu}) separated into the \textit{biweight} filtering step and the \textsf{BLS} search.
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{../workflows/benchmark/figures/nuance_vs_bls.pdf}
+        \caption{Study of \textsf{nuance} (black line) processing time against \texttt{biweight+BLS} (cf.\;\autoref{simu}, blue line).
+The gray curve shows the performance of \textsf{nuance} \textit{linear search} when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together.
+This study is performed on both a single core of an Apple M2 Max chip (red line) and all of its 12 cores (black line).
+While not being shown, we verify that both the \textsf{BLS} algorithm and \nuancemethod{}processing times scale linearly with the number of probed transit durations and orbital periods.}
         \label{fig:blsvsnuance}
@@
 
-\noindent In \autoref{fig:blsvsnuance}, the processing time of the linear search from \textsf{nuance} is recorded against the number of points in the light curve.
-One advantage of \textsf{nuance} is that the linear search can be performed separately on different continuous observations, and then combined in the periodic search.
-Hence, if searching for transits in separate observations with similar durations (such as different TESS sectors or different ground-based nightly observations), the computational cost of \textsf{nuance} grows linearly with the number of observations.
-Such a parallel cannot be made with \texttt{BLS}, as the algorithm considers all observations as a single one.
-This way, after assuming that 10 000 points observations can be combined into the periodic search (roughly corresponding to the number of data points in continuous TESS observations for example), \textsf{nuance} will constantly perform a single order of magnitude slower than \texttt{BLS}, only considering the linear search.
-For this reason, \textsf{nuance} must not be employed in the general case, but rather used when light curves contain correlated noise with characteristics described in this chapter.
-Proper guidelines and metrics to assess the presence of such noise, which would indicate if \textsf{nuance} is required, are currently being developed.\\\\
-For reference, the complete search (including the periodic search) of a single TESS sector for TIC 140212114 (15992 points partly shown in  \autoref{fig:140212114}) took a total of 30 seconds, executed on a 10 core Apple M1 Pro chip.
-The latter processing time would grow linearly with the number of sectors being processed.
+As see in \autoref{fig:blsvsnuance}, most of the computational cost of \nuancecode{} and the \texttt{biweight+BLS} method comes from the \textit{linear search} and the \textit{biweight} filtering step.
+One advantage of \nuancemethod{} is that the \textit{linear search} can be performed separately on different continuous observations, and then combined in the periodic search.
+Hence, if searching for transits in separate observations with approximately similar durations, such as different TESS sectors or different ground-based nightly observations, the computational cost of \nuancecode{} grows linearly with the number of observations (see grey line in \autoref{fig:blsvsnuance}).
+Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform about one order of magnitude slower than more conventional techniques associated with \textsf{BLS}}.
+The second advantage of \nuancecode{} is its \textsf{JAX} implementation, allowing the linear search to be run in parallel.
+As a reference, the 12 cores of an Apple M2 Max chip, searching for TOI-540 b transits with nuance (\autoref{toi540}) took 3 minutes in total, 1 minute for the 5 sectors linear searches, and 2 minutes for the combined periodic search.
+In comparison, the search with \texttt{biweight+BLS} took about 4 seconds.\\\\
+% As the periodic search is expected to be improved by lower lever optimizations (using a compiled language), we reserve its benchmark for a future work.
+Thanks to its \textsf{JAX} implementation, we also note the possibility of running \nuancecode{} on GPUs, which is out of the scope of the present comparison.\\\\
+Because of its computational cost, \nuancecode{} must not be employed in the general case, but rather adopted when light curves contain correlated noise with specific characteristics.
+If employing a \textit{biweight} filter for detrending, these characteristics correspond to the ones discussed in \autoref{issues}.
+But as these strongly depend on the type of detrending technique employed, we do not provide general guidelines as when \nuancemethod{} should be preferred over a specific detrending technique.
+Nevertheless, extensive benchmarks and guidelines are reserved to future developments and will be progressively shared on \nuancecode{}'s online documentation\footlink{https://nuance.readthedocs.io/en/latest/}.
 
-\subsection*{Periodic search}
+\subsection{Systematics modeling}
+Throughout the paper, a single column design matrix $\bm{X}$, corresponding to the mean differential flux (ideally unitary), was employed, hence assuming that the instrumental systematics signals were non-existent.
+In practice, \nuancemethod{} has been developed to linearly model systematic signals through more complex design matrices (as in \citealt{foreman2016}), in addition with its capability to model correlated noise while searching for transits.
+This feature is intentionally unexploited in the comparisons presented in \autoref{results}, as detrending light curves assuming a linear systematics model, such as PLD co-trending vectors \citep{pld}, is highly incomplete if applied on data while ignoring the presence of other astrophysical signals.
+Comparisons involving more complex design matrices would also be sensitive to the choice of linear components, and would have unwanted repercussions on their results.\\\\
+As an illustration, the NEMESIS pipeline \citep{nemesis} starts processing the differential light curves by employing a linear systematics detrending using a least square fit of the data with a reduced PLD basis, before smoothing the signal from stellar variability using an approach similar to the one employed in the \texttt{biweight+BLS} approach (cf.\;\autoref{real}), hence detrending the systematics with an incomplete model that does not account for stellar variability.
+To account for stellar variability while fitting the linear systematics model to the data, a step further would be to use a GP, such as done in the EVEREST \citep{everest2} pipeline.
+However this would also involve some potential degradation of the transit signals (see e.g.\;\autoref{fig:issue2}), with a hardly distinguishable origin.
+For these reasons, and to keep our comparisons as targeted as possible, we do not compare commonly used systematics detrending approaches and decided to focus our study on the comparison with stellar variability detrending techniques (although these two aspects often overlap in the literature).\\\\
+Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsly is extremely promising for the search of transiting exoplanets in ground-based data, that usually suffer from daily interruptions.
+In this respect, we note the similarity of our \textit{linear search} (cf\;\autoref{linear_search}) to the one presented in \citealt{Berta2012}, that focused on the detection of single eclipses in the MEarth light curves \citep{Irwin2009}.
+Similarly, \nuancemethod{} would highly benefit the search for transiting exoplanets around M-dwarf type stars, such as the ones observed by the SPECULOOS survey \citep{speculoos} whose monitoring suffers from both increased red noise (due to atmospheric and instrumental thermal effects discussed e.g.\;in \citealt{Berta2012} and  \citealt{Pedersen2023}) and enhanced stellar variability \citep{Murray2020}.
+We reserve this promising application to a future study.
 
-The periodic search of \textsf{nuance} can be done separately from the linear search.
-For a given light curve, a single linear search can be performed then masked and combined iteratively to search for multiple transiting exoplanets.
-Hence, \autoref{fig:blsvsnuance} only shows the processing time of the linear search, assuming that a single planet is searched for.
-As of June 2023, the periodic search in the current implementation of \textsf{nuance} is the least optimized part.
-But recent developments indicate that the periodic search can be performed in a time equal to the complete processing time of the \texttt{BLS} algorithm.
-Hence, as the \texttt{BLS} algorithm grows linearly with the number of planets being searched (because reran every time on masked light curves), the increased computing time for \textsf{nuance} would be equal, bridging the gap between the computing time of  \texttt{BLS} and \textsf{nuance} when searching for multiplanetary transiting systems.
+\subsection{The choice of kernel}
+While not being discussed in our study, the efficiency of \nuancemethod{} to detect transits in correlated noise is highly dependant on the design of its GP kernel.
+In the ensemble comparison of \autoref{real}, the goal was to choose a kernel and an optimization strategy suited to most of the studied light curves, leading to few outliers in the results that were indicative of a badly designed and/or optimized kernel.
+An alternative, recommended for more realistic blind searches, is to perform model comparison on well selected kernels, and to adapt the optimization strategy to each dataset.\\\\
+When using \nuancecode{} on TESS light curves for example, it must be noted that the observed light curve variability might originate from a contaminated photometric aperture, so that a physically-interpretable GP kernel representing a single star activity is not necessarily optimal.
 
-\subsection*{Design matrix}
-
-In this chapter, starting from \autoref{detrending_effect}, it was assumed that systematic signals (as opposed to astrophysical signals) were completely corrected for without affecting the searched transit signals.
-For this reason, all analyses in this chapter were done with a single column design matrix $\bm{X}$, corresponding to the mean, constant, differential flux (ideally 1) of a light curve.
-In practice, \textsf{nuance} has been developed to linearly model systematic signals through more complex design matrices, in addition with its capability to model correlated noise while searching for transits.
-This aspect is completely unexploited in the previous comparisons, and would require to search for transits in completely raw light curves, which comes with few challenges.
-Given a light curve, this approach involves choosing a design matrix containing specific explanatory measurements, hence requiring a proper model comparison.
-In addition, large design matrices are expected to increase \textsf{nuance} processing time, something that needs to be accounted for in its performance assessment.\\\\
-Used in the context of the SPECULOOS survey, \textsf{nuance} is currently being tested on light curves containing systematic signals that can be modeled linearly using polynomials of common explanatory variables (such as airmass, background flux, pointing error or PSF full-width at half-maximum).
-On light curves spanning only few hours, with less than a thousand photometric measurements, the use of a linear model to detrend the light curve ignoring the presence of transits is highly susceptible to remove the signals of interest, giving a clear advantage to a full-fledge method like \textsf{nuance}.
-The same applies for filter-based detrending algorithms.
-Proving the later empirical statement is out of the scope of this chapter and will be treated in a following publication.
-However, \textsf{nuance} might be particularly well suited to search for transits in ground-based observations, where the detrending of sparse observations is believed to remove the lowest SNR transit signals.\\\\
-Overall, \textsf{nuance} is a powerful tool to search for transits in light curves containing systematic signals and correlated noise, characteristic of late M-dwarfs light curves.
-Due to several advances in computational methods, as well as its analytical formalism, \textsf{nuance}  remains highly tractable, and may already be applied to space-based continuous observations.
-\textsf{nuance} is currently being tested on sparse observations, offering a unique tool to search for transiting exoplanet in the light curves of SPECULOOS and other ground-based surveys.
-
-\subsection{When to use it?}\label{when}
+\section{Conclusion}
 
@@
 \newcommand{\sumTk}{i\neq k}
-From the linear search presented in \autoref{linear_search}, we retain and index by $k$ the parameters of the $K$ individual transits whose epochs $\{T_k\}_k$ are compatible with a periodic signal of period $P$ and epoch $T_0$.
+From the \textit{linear search} presented in \autoref{linear_search}, we retain and index by $k$ the parameters of the $K$ individual transits whose epochs $\{T_k\}_k$ are compatible with a periodic signal of period $P$ and epoch $T_0$.
 From the likelihoods of these transits (computed in \autoref{linear_search}), we want an expression for
@@
 
+\section{Injection-recovery on TIC 1019692}\label{1019692}
+
+\begin{figure}
+    \begin{centering}
+        \includegraphics[width=0.8\linewidth]{../workflows/tess_injection_recovery/figures/searched/1019692.pdf}
+        \caption{Results of the transits injection-recovery on TIC 1019692 half-sector light curve.
+Left: cleaned light curve with computed trend overplotted in black (except for \textsf{nuance} where it corresponds to the mean of the Gaussian Process model).
+Right: Results of the transit search where a black square denotes a transit signal not detected, gray a signal detected at an alias period ($P/2$ or $2P$), and  white a signal detected with the correct period.
+On the right plots, secondary axes show the $(\tau, \delta)$ relative parameter space.
+For each method, the upper right legend on the left plot indicates its ranking based on the percent of recovered transit signals (where a transit with an aliased period counts as being detected).
+        }
+        \label{fig:onesearch}
+    \end{centering}
+\end{figure}
+
 \bibliography{ref}

### 2023-12-02 workflows/benchmark/scripts/bls.py

@@ -1,0 +1,31 @@
+import yaml
+import numpy as np
+import yaml
+from astropy.timeseries import BoxLeastSquares
+from time import time as _time
+from wotan import flatten
+
+times = {}
+
+n_durations = int(eval(snakemake.wildcards.durations.split("_")[-1]))
+periods = np.linspace(*[eval(i) for i in snakemake.wildcards.periods.split("_")[1:]])
+durations = np.linspace(0.01, 0.05, n_durations)
+time, flux, error = np.load(snakemake.input[0])
+n_points = int(snakemake.wildcards.n_points)
+
+t0 = _time()
+
+flatten_trend = flatten(
+    time, flux, window_length=0.2, return_trend=True
+)[1]
+flatten_flux = flux - flatten_trend
+flatten_flux -= np.mean(flatten_flux)
+flatten_flux += 1.0
+time = _time() - t0
+times["biweight"] = float(time)
+
+model = BoxLeastSquares(time, flatten_flux, dy=error)
+results = model.power(periods, durations)#, objective="snr")
+time = _time() - t0
+times["bls"] = float(time)
+yaml.safe_dump(times, open(snakemake.output[0], "w"))

### 2023-12-02 workflows/benchmark/scripts/nuance.py

@@ -1,0 +1,32 @@
+import yaml
+import numpy as np
+import os
+from multiprocessing import cpu_count
+
+cores = int(eval(snakemake.wildcards.cores))
+os.environ["XLA_FLAGS"] = f"--xla_force_host_platform_device_count={cores}"
+
+import jax
+
+from astropy.timeseries import BoxLeastSquares
+from nuance import Nuance
+from time import time as _time
+
+
+
+n_durations = int(eval(snakemake.wildcards.durations))
+periods = np.linspace(*[eval(i) for i in snakemake.wildcards.periods.split("_")])
+durations = np.linspace(0.01, 0.05, n_durations)
+time, flux, error = np.load(snakemake.input[0])
+n_points = int(snakemake.wildcards.n_points)
+
+times = {}
+
+nu = Nuance(time, flux, error)
+t0 = _time()
+nu.linear_search(time, durations)
+times["linear"] = float(_time() - t0)
+search = nu.periodic_search(periods)
+times["all"] = float(_time() - t0)
+times["cpu_counts"] = jax.device_count(), cores
+yaml.safe_dump(times, open(snakemake.output[0], "w"))

### 2023-12-02 workflows/comparison_toi/scripts/biweight_bls_search_1.py

@@ -1,0 +1,58 @@
+import pickle
+from nuance.combined import CombinedNuance
+
+periods, durations = snakemake.params
+nus = [pickle.load(open(f, "rb")) for f in snakemake.input]
+nu_s = CombinedNuance(nus)
+len(nu_s.time)
+
+import numpy as np
+from astropy.timeseries import BoxLeastSquares
+from wotan import flatten
+from tqdm import tqdm
+
+
+def bls(time, flux, error):
+    model = BoxLeastSquares(time, flux, dy=error)
+    return model.power(periods, durations, objective="snr")
+
+
+verbose = True
+
+# biweight
+# --------
+
+windows = np.linspace(30 / 60 / 24, 5 / 24, 15)
+results_ = []
+flatten_fluxes = []
+
+for window in tqdm(windows):
+    flatten_trend = np.hstack(
+        [
+            flatten(
+                n.time,
+                n.flux,
+                window_length=window,
+                return_trend=True,
+                robust=True,
+            )[1]
+        ]
+        for n in nus
+    )[0]
+
+    flatten_flux = nu_s.flux - flatten_trend
+    flatten_flux -= np.mean(flatten_flux)
+    flatten_flux += 1.0
+
+    error = np.median([np.median(np.sqrt(n.gp.variance)) for n in nus]) * np.ones_like(
+        flatten_flux
+    )
+
+    results = bls(nu_s.time, flatten_flux, error)
+    results_.append(results)
+    flatten_fluxes.append(flatten_flux)
+
+i = np.argmax([np.max(r.power) for r in results_])
+
+pickle.dump(results_[i], open(snakemake.output[0], "wb"))
+np.save(snakemake.output[1], [nu_s.time, flatten_fluxes[i], error])

### 2023-12-02 workflows/comparison_toi/scripts/biweight_bls_search_2.py

@@ -1,0 +1,19 @@
+import numpy as np
+from astropy.timeseries import BoxLeastSquares
+from nuance.utils import phase
+
+periods, durations = snakemake.params[0:2]
+
+
+def bls(time, flux, error):
+    model = BoxLeastSquares(time, flux, dy=error)
+    return model.power(periods, durations, objective="snr")
+
+
+time, flux, error = np.load(snakemake.input[0])
+t0, D, P = np.load(snakemake.input[1])
+
+mask = phase(time, t0, P) >= 2 * D
+
+results = bls(time[mask], flux[mask], error[mask])
+pickle.dump(results, open(snakemake.output[0], "wb"))

### 2023-12-02 workflows/comparison_toi/scripts/periodic_search_1.py

@@ -1,0 +1,11 @@
+import pickle
+import numpy as np
+from nuance.combined import CombinedNuance
+
+
+nu_s = CombinedNuance([pickle.load(open(f, "rb")) for f in snakemake.input])
+len(nu_s.time)
+
+combined_search = nu_s.periodic_search(snakemake.params[0])
+np.save(snakemake.output[1], np.array(combined_search.best))
+pickle.dump(combined_search, open(snakemake.output[0], "wb"))

### 2023-12-02 workflows/comparison_toi/scripts/plot_periodograms.py

@@ -1,0 +1,28 @@
+import pickle
+import matplotlib.pyplot as plt
+
+plt.figure(figsize=(9, 3))
+plt.subplot(121)
+results = pickle.load(open(snakemake.input[0], "rb"))
+search = pickle.load(open(snakemake.input[1], "rb"))
+
+plt.plot(search.periods, search.Q_snr, c="C0", label="nuance")
+plt.plot(results.period, results.power, c="0.7", label="biweight+BLS", alpha=0.8)
+plt.title("First transit search", loc="left")
+plt.ylabel("SNR")
+plt.xlabel("Period (days)")
+y = plt.ylim()
+
+plt.subplot(122)
+results = pickle.load(open(snakemake.input[2], "rb"))
+search = pickle.load(open(snakemake.input[3], "rb"))
+
+plt.plot(search.periods, search.Q_snr, c="C0", label="nuance")
+plt.plot(results.period, results.power, c="0.7", label="biweight+BLS", alpha=0.8)
+plt.legend(loc="lower right")
+plt.title("Second transit search", loc="left")
+plt.xlabel("Period (days)")
+plt.ylim(y)
+
+plt.tight_layout()
+plt.savefig(snakemake.output[0])

### 2023-12-02 workflows/control_test_bls/scripts/plot.py

@@ -1,0 +1,99 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+from scipy.stats import binned_statistic_2d
+
+scatter = False
+
+
+def is_found(period, true_period):
+    period_check = np.abs(period - true_period) < 0.01
+    alias_check = np.abs(2 * period - true_period) < 0.01
+    alias_check |= np.abs(period / 2 - true_period) < 0.01
+    if period_check:
+        return 1
+    elif alias_check:
+        return 0.5
+    else:
+        return 0
+
+
+df = pd.read_csv(snakemake.input[0])
+
+df["bls_found"] = df.apply(
+    lambda row: is_found(row["bls_period"], row["period"]), axis=1
+)
+df["nuance_found"] = df.apply(
+    lambda row: is_found(row["nuance_period"], row["period"]), axis=1
+)
+
+bins = (23, 23)
+fig = plt.figure(None, (7.5, 3.5))
+cmap = "Greys_r"
+
+ax = plt.subplot(121)
+period, depth, found = df[["period", "depth", "bls_found"]].values.T
+depth *= 1e4
+stats = binned_statistic_2d(period, depth, found, bins=bins)
+if scatter:
+    plt.scatter(period, depth, c=found, s=10, cmap=cmap)
+else:
+    im = plt.imshow(
+        stats.statistic.T,
+        origin="lower",
+        extent=(
+            stats.x_edge.min(),
+            stats.x_edge.max(),
+            stats.y_edge.min(),
+            stats.y_edge.max(),
+        ),
+        aspect="auto",
+        cmap=cmap,
+    )
+
+ax.text(
+    0.05,
+    0.92,
+    f"{len(period)} samples (binned)",
+    va="center",
+    ha="left",
+    transform=ax.transAxes,
+)
+
+xlabel = r"period (days)"
+ax.set_xlabel(xlabel)
+ax.set_ylabel(r"depth ($10^{-4}$)")
+ax.set_title("BLS")
+
+ax = plt.subplot(122)
+period, depth, found = df[["period", "depth", "nuance_found"]].values.T
+depth *= 1e4
+
+if scatter:
+    plt.scatter(period, depth, c=found, s=10, cmap=cmap)
+else:
+    stats = binned_statistic_2d(period, depth, found, bins=bins)
+    im = plt.imshow(
+        stats.statistic.T,
[…]

### 2023-12-02 workflows/performances/scripts/make_lightcurve.py

@@ -1,14 +1,0 @@
-import numpy as np
-
-n_points = int(snakemake.wildcards.n_points)
-n_X = int(snakemake.wildcards.n_X)
-
-np.random.seed(n_points)
-time = np.linspace(0, 3, n_points)
-flux = np.random.normal(1.0, 0.01, size=len(time))
-error = np.ones_like(time) * 0.0001
-
-np.save(
-    snakemake.output[0],
-    np.array([time, flux, error]),
-)

### 2023-12-02 workflows/performances/scripts/tls_search.py

@@ -1,34 +1,0 @@
-import sys
-
-import numpy as np
-
-sys.path.append("lib")
-import pickle
-
-import numpy as np
-import yaml
-from tls import tls
-from wotan import flatten
-
-period_range = snakemake.params.period_range
-time, flux, _ = np.load(snakemake.input[0])
-n_points = int(snakemake.wildcards.n_points)
-n_X = int(snakemake.wildcards.n_X)
-n_periods = int(snakemake.wildcards.n_periods)
-periods = np.linspace(period_range[0], period_range[1], n_periods)
-
-verbose = True
-
-flatten_trend = flatten(time, flux, window_length=0.01, return_trend=True)[1]
-flatten_flux = flux - flatten_trend
-flatten_flux -= np.mean(flatten_flux)
-flatten_flux += 1.0
-
-model = tls(time, flatten_flux, verbose=verbose)
-results = model.power(
-    periods,
-    verbose=verbose,
-    use_threads=1,
-    show_progress_bar=verbose,
-    durations=[0.01, 0.02],
-)

### 2023-12-02 workflows/toi540_injection_recovery/lib/utils.py

@@ -1,22 +1,0 @@
-from tinygp import kernels, GaussianProcess
-import jax.numpy as jnp
-import numpy as np
-
-
-def poly_X(x, order):
-    X = x ** np.arange(0, order)[:, None]
-    X[1::] -= X[1::].mean(1)[:, None]
-    X[1::] /= X[1::].std(1)[:, None]
-
-    return X
-
-
-def build_gp(params, x):
-    periodic = kernels.quasisep.SHO(
-        omega=jnp.exp(params["log_omega"]),
-        quality=jnp.exp(params["log_quality"]),
-        sigma=jnp.exp(params["log_sigma"]),
-    )
-    return GaussianProcess(
-        periodic, x, diag=jnp.exp(2 * params["log_jitter"]), mean=0.0
-    )

### 2023-12-02 workflows/toi540_injection_recovery/scripts/gp.py

@@ -1,69 +1,0 @@
-import pandas as pd
-from nuance import Nuance
-import sys
-import jax.numpy as jnp
-import yaml
-
-sys.path.append("lib")
-import utils
-import numpy as np
-
-df = pd.read_csv(snakemake.input[0])
-time, flux, error = df.values.T
-time -= time[0]
-
-X = utils.poly_X(time, 3)
-
-nu = Nuance(time, flux, np.mean(error), X=X, compute=False)
-
-period = 17.0 / 24.0
-
-init_params = {
-    "log_omega": jnp.log(2 * jnp.pi / period),
-    "log_quality": jnp.log(50.0),
-    "log_sigma": jnp.log(np.std(flux)),
-    "log_jitter": jnp.log(np.mean(error)),
-}
-
-# Masked optimization (for flares and transits)
-mask = np.ones_like(nu.time).astype(bool)
-new_params = init_params.copy()
-window = 30
-
-for i in range(3):
-    if i == 0:
-        m = np.mean(flux)
-    else:
-        m = np.array(mu(new_params))
-    r = flux - m
-    mask_up = r < np.std(r[mask]) * 3
-    mask_down = r > -np.std(r[mask]) * 2
-
-    # mask around flares
-    ups = np.flatnonzero(~mask_up)
-    if len(ups) > 0:
-        mask_up[
-            np.hstack(
-                [
-                    np.arange(max(u - window, 0), min(u + window, len(nu.time)))
-                    for u in ups
-                ]
-            )
-        ] = False
-    mask = np.logical_and(mask_up, mask_down)
-
-    optimize, mu, plot_kernel = nu.gp_optimization(utils.build_gp, mask=mask)
-    new_params = optimize(new_params, ["log_sigma", "log_jitter"])
-    new_params = optimize(new_params, ["log_quality", "log_sigma", "log_jitter"])
-    # we do not optimize for the period as it is found independently
-
-# Saving
-
-final_params = {
-    "log_omega": float(new_params["log_omega"]),
-    "log_quality": float(new_params["log_quality"]),
-    "log_jitter": float(new_params["log_jitter"]),
-    "log_sigma": float(new_params["log_sigma"]),
-}
-
-yaml.safe_dump(final_params, open(snakemake.output[0], "w"))

### 2023-12-02 workflows/toi540_injection_recovery/scripts/search.py

@@ -1,57 +1,0 @@
-import pandas as pd
-import numpy as np
-import yaml
-import sys
-import jax
-
-sys.path.append("lib")
-import utils
-from nuance import Nuance
-
-# Load data
-params = yaml.full_load(open(snakemake.input[0], "r"))
-df = pd.read_csv(snakemake.input[1])
-
-time, flux, error = df.values.T
-if len(time) > 0:
-    t_0 = time[0]
-    time -= t_0
-
-    X = utils.poly_X(time, 3)
-    kernel = utils.build_gp(params, time).kernel
-    nu = Nuance(time, flux, kernel=kernel, error=error, X=X)
-
-    # Mask flares
-    # In what Fran Pozuelos sent me, the falres are sigma clipped. To remove their
-    # full extended signal, I locate and mask 30 points on each side of "holes" in
-    # time
-    window = 30
-    mask = np.pad(np.array(np.diff(time) < 3 * np.median(np.diff(time))), (0, 1))
-    ups = np.flatnonzero(~mask)
-    if len(ups) > 0:
-        mask[
-            np.hstack(
-                [
-                    np.arange(max(u - window, 0), min(u + window, len(nu.time)))
-                    for u in ups
-                ]
-            )
-        ] = False
-
-    nu = Nuance(time[mask], flux[mask], error[mask], kernel=kernel, X=X[:, mask])
-
-    # linear search
-    t0s = time.copy()
-    Ds = np.linspace(0.01, 0.07, 6)
-    nu.linear_search(t0s, Ds)
-
-    # We mask directly the first planet signal
-    nu = nu.mask(0.9236153776066658, 0.02, 1.2390692712847524)
-
-    # Search of the second
-    periods = np.linspace(0.9, 10.1, 8000)
-    search = nu.periodic_search(periods)
-
-    yaml.safe_dump([float(p) for p in search.best], open(snakemake.output[0], "w"))
-else:
-    yaml.safe_dump([0], open(snakemake.output[0], "w"))

### 2023-12-02 latex/paper.tex
aal figures go in workflows

@@
     \begin{centering}
-        \includegraphics[width=8.5cm]{issue1.pdf}
+        \includegraphics[width=8.5cm]{../workflows/plot_issues/figures/issue1.pdf}
         \caption{Illustration of the effect of correlated noise on a single transit SNR.
@@
     \begin{centering}
-        \includegraphics[width=0.85\linewidth]{issue2.pdf}
+        \includegraphics[width=1\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
         \caption{(Top left) Simulated observation spanning 3 days, including a transit signal with a depth of 0.8\%, duration of 0.1 day, and measurement error of 1.5\% (using the model from \cite{protopapas} described in \autoref{app_proto}).
@@
     \begin{centering}
-        \includegraphics[width=0.8\linewidth]{principle_linear_search.pdf}
+        \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
         \caption{Principle and output of the \textit{linear search}.
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{principle_periodic_0.pdf}\\
+    \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_periodic_0.pdf}\\
     {(a) $\ln \mathcal{L}$}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{principle_periodic_1.pdf}\\
+    \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
     {(b) $\mathcal{P}(P)$}
@@
     \begin{centering}
-        \includegraphics[width=0.8\linewidth]{principle_Q.pdf}
+        \includegraphics[width=0.85\linewidth]{principle_Q.pdf}
         \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}

### 2023-12-02 workflows/plot_issues/scripts/plot_issue_2.py

@@ -5,10 +5,43 @@ import tinygp
 import jax
 import matplotlib.pyplot as plt
 jax.config.update("jax_enable_x64", True)
-import sys
-sys.path.append("./lib")
-sys.path.append("../lib")
-from utils import white_red
+
+import jax.numpy as jnp
+from tinygp import kernels, GaussianProcess
+import matplotlib.pyplot as plt
+
+
+def white_red(time, flux, error, plot=False, mean=0.):
+    init = {
+        "log_diag": 2*np.log(np.mean(error)),
+        "log_scale": np.log(1), # 5 mins
+        "log_amp": np.log(np.std(flux))
+    }
+
+    def build_gp(params):
+        kernel = jnp.exp(params["log_amp"]) * kernels.ExpSquared(jnp.exp(params["log_scale"]))
+        return GaussianProcess(kernel, time, diag=jnp.exp(params["log_diag"]), mean=mean)
+
+    @jax.jit
+    def loss(params):
+        gp = build_gp(params)
+        return -gp.log_probability(flux)
+    
+    opt = jaxopt.ScipyMinimize(fun=loss)
+    params = opt.run(init).params  
+    w = np.sqrt(np.exp(params['log_diag']))
+    
+    if plot:
+        gp = build_gp(params)
+        _, cond = gp.condition(y=flux, X_test=time)
+        mu = cond.mean
+        plt.plot(time, flux, ".", c="0.8")
+        plt.plot(time, mu, c="k")
+        noise = w
+        plt.fill_between(time, mu + noise, mu - noise, color="C0", alpha=0.3)
+
+    return w, np.sqrt(np.exp(params['log_amp']))
+
 
 seed = 42
 
@@ -141,4 +174,4 @@ plt.text(6.5, 1.005, f"{original_snr:.2f}", ha="center", fontsize=14, color=txt_
 
 plt.axis("off")
 plt.tight_layout()
-plt.savefig("figures/issue2.pdf")
+plt.savefig(snakemake.output[0])

### 2023-12-02 workflows/different_kernels/scripts/clean.py
chore: cleaning a bit

@@ -1,55 +1,0 @@
-import pickle
-
-import jax
-import matplotlib.pyplot as plt
-import numpy as np
-import yaml
-from nuance import Nuance
-from nuance.kernels import rotation
-
-i = 0
-
-data = pickle.load(open(snakemake.input.fluxes, "rb"))
-info = yaml.safe_load(open(snakemake.input.info, "r"))
-gp_params = yaml.full_load(open(snakemake.input.gp, "r"))
-
-build_gp, init = rotation(info["star_period"], np.mean(data["error"]), long_scale=0.5)
-gp = build_gp(gp_params, data["time"])
-nu = Nuance(data["time"], data["flux"], gp=gp)
-
-
-def mask_flares(time, flux, iterations=3, sigma=4, gp=None):
-    _, mu, _ = nu.gp_optimization(build_gp)
-
-    mask = np.ones_like(time).astype(bool)
-    window = 30
-
-    for _ in range(iterations):
-        m = mu(gp_params)
-        r = flux - m
-        mask_up = r < np.std(r[mask]) * sigma
-
-        # mask around flares
-        ups = np.flatnonzero(~mask_up)
-        if len(ups) > 0:
-            mask[
-                np.hstack(
-                    [
-                        np.arange(max(u - window, 0), min(u + window, len(time)))
-                        for u in ups
-                    ]
-                )
-            ] = False
-        _, mu, _ = nu.gp_optimization(build_gp, mask)
-
-    return ~mask
-
-
-mask = ~mask_flares(data["time"], data["flux"], gp=gp)
-
-cleaned_data = data.copy()
-cleaned_data["flux"] = data["flux"][mask]
-cleaned_data["time"] = data["time"][mask]
-cleaned_data["error"] = data["error"][mask]
-
-pickle.dump(cleaned_data, open(snakemake.output[0], "wb"))

### 2023-12-02 workflows/different_kernels/scripts/download.py

@@ -1,75 +1,0 @@
-import pickle
-
-import astropy.units as u
-import lightkurve as lk
-import numpy as np
-import pandas as pd
-import yaml
-
-tic = int(snakemake.wildcards["target"])
-filepath = snakemake.output[0]
-
-name = f"TIC {tic}"
-search_result = lk.search_lightcurve(name, author="SPOC", exptime=120)
-
-# data
-klc = search_result[0].download()
-times = klc.time.to_value("btjd")
-fluxes = klc.pdcsap_flux.to_value().filled(np.nan)
-errors = klc.flux_err.to_value().filled(np.nan)
-mask = np.isnan(fluxes) | np.isnan(errors) | np.isnan(times)
-original_time = times[~mask]
-original_flux = fluxes[~mask]
-original_error = errors[~mask]
-
-splits = np.array_split(
-    np.arange(len(original_time)), np.flatnonzero(np.diff(original_time) > 0.1) + 1
-)
-
-n = 300
-time = np.hstack([original_time[split[n:-n]] for split in splits])
-flux = np.hstack([original_flux[split[n:-n]] for split in splits])
-error = np.hstack([original_error[split[n:-n]] for split in splits])
-
-# first exposure
-first_exposure = klc.time[0]
-time_format = rf"{first_exposure.format.upper()}_{{{first_exposure.scale.upper()}}}"
-first_exposure_time = first_exposure.value
-
-flux_median = np.median(original_flux)
-time = time - first_exposure_time
-flux /= flux_median
-error /= flux_median
-
-# stellar parameters
-url = f"https://exofop.ipac.caltech.edu/tess/download_stellar.php?id={tic}"
-star = pd.read_csv(url, delimiter="|", index_col=1).iloc[0]
-
-pickle.dump(
-    {
-        "flux": flux,
-        "time": time,
-        "error": error,
-    },
-    open(snakemake.output.fluxes, "wb"),
-)
-
-
-star_dict = {
-    "tic": int(tic),
-    "time0": float(first_exposure_time),
-    "time0_format": str(time_format),
-    "star_radius": float(star["Radius (R_Sun)"]),
-    "star_mass": float(star["Mass (M_Sun)"]),
-    "star_teff": float(star["Teff (K)"]),
-    "star_logg": float(star["log(g)"]),
-    "sector": int(klc.sector),
-    "logg": float(star["log(g)"]),
-}
-
-yaml.safe_dump(
-    star_dict,
[…]

### 2023-12-02 workflows/different_kernels/scripts/inject.py

@@ -1,39 +1,0 @@
-import pickle
-
-import numpy as np
-import yaml
-from nuance.star import Star
-from nuance.utils import periodic_transit
-
-i = int(snakemake.wildcards.lc)
-data = pickle.load(open(snakemake.input.fluxes, "rb"))
-periods, radii = pickle.load(open(snakemake.input.params, "rb"))
-info = yaml.safe_load(open(snakemake.input.info, "r"))
-
-star = Star(
-    info["star_radius"], info["star_mass"], info["star_amplitude"], info["star_period"]
-)
-
-np.random.seed(i)
-period = periods[i]
-radius = radii[i]
-tau = star.period2tau(period)
-delta = star.radius2delta(radius)
-duration = star.transit_duration(period)
-depth = star.transit_depth(radius)
-
-t0 = np.random.rand() * period
-
-transit = depth * periodic_transit(data["time"], t0, duration, P=period, c=500)
-injected_flux = data["flux"] + transit
-
-data["flux"] = injected_flux
-data["transit_t0"] = t0
-data["transit_duration"] = duration
-data["transit_depth"] = depth
-data["delta"] = delta
-data["tau"] = tau
-data["planet_period"] = period
-data["planet_radius"] = radius
-output = snakemake.output[0]
-pickle.dump(data, open(output, "wb"))

### 2023-12-02 workflows/different_kernels/scripts/optimize_gp.py

@@ -1,36 +1,0 @@
-import pickle
-
-import matplotlib.pyplot as plt
-import numpy as np
-import yaml
-from nuance import Nuance
-from nuance.kernels import rotation
-
-data = pickle.load(open(snakemake.input.fluxes, "rb"))
-period = yaml.safe_load(open(snakemake.input.star_period, "r"))["period"]
-
-# only one continuous segment
-dt = np.diff(data["time"])
-mask = data["time"] < data["time"][np.flatnonzero(dt > 0.5)[0]]
-data["flux"] = data["flux"][mask]
-data["time"] = data["time"][mask]
-data["error"] = data["error"][mask]
-# sigma clip
-mask = np.abs(data["flux"] - np.median(data["flux"])) < 3 * np.std(data["flux"])
-data["flux"] = data["flux"][mask]
-data["time"] = data["time"][mask]
-data["error"] = data["error"][mask]
-
-build_gp, init = rotation(period, np.mean(data["error"]), long_scale=0.5)
-nu = Nuance(data["time"], data["flux"], data["error"])
-optimize, mu, nll = nu.gp_optimization(build_gp)
-
-new = optimize(
-    init, ["log_sigma", "log_short_scale", "log_short_sigma", "log_long_sigma"]
-)
-new = optimize(new)
-
-gp_output = snakemake.output[0]
-yaml.safe_dump(
-    {name: float(value) for name, value in new.items()}, open(gp_output, "w")
-)

### 2023-12-02 workflows/different_kernels/scripts/star_period.py

@@ -1,19 +1,0 @@
-import numpy as np
-from astropy.stats import LombScargle
-import pickle as pkl
-import yaml
-
-data = pkl.load(open(snakemake.input[0], "rb"))
-
-
-def rotation_period(time, flux):
-    """rotation period based on LS periodogram"""
-
-    ls = LombScargle(time, flux)
-    frequency, power = ls.autopower(minimum_frequency=1 / 5, maximum_frequency=1 / 0.1)
-    period = 1 / frequency[np.argmax(power)]
-    return period
-
-
-period = rotation_period(data["time"], data["flux"])
-yaml.safe_dump({"period": float(period)}, open(snakemake.output[0], "w"))

### 2023-12-02 latex/paper.tex
chore: cleaning a bit
chore: remove unused

@@
-\documentclass{aastex631}
+\documentclass[modern]{aastex631}
 \usepackage{amsmath}
@@
     stringstyle=\color{RoyalBlue},
-    basicstyle=\fontsize{7.5}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont,
+    basicstyle=\fontsize{8.5}{13}\fontfamily{DejaVuSansMono-TLF}\selectfont,
     breakatwhitespace=false,         
@@
     \begin{centering}
-        \includegraphics[width=8.5cm]{../workflows/plot_issues/figures/issue1.pdf}
+        \includegraphics[width=0.6\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
         \caption{Illustration of the effect of correlated noise on a single transit SNR.
@@
     \begin{centering}
-        \includegraphics[width=0.7\linewidth]{../workflows/cleaning_snr/figures/result.pdf}
+        \includegraphics[width=0.9\linewidth]{../workflows/cleaning_snr/figures/result.pdf}
         \caption{SNR of a unique transit after detrending light curves using an optimal bi-weight filter.
@@
     \begin{centering}
-        \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
+        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
         \caption{Principle and output of the \textit{linear search}.
@@
     \centering
-    \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_periodic_0.pdf}\\
+    \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_periodic_0.pdf}\\
     {(a) $\ln \mathcal{L}$}
-
-    \centering
-    \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
-    {(b) $\mathcal{P}(P)$}
 \end{figure}
@@
 \begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
+    {(b) $\mathcal{P}(P)$}
+
     \caption{Applied to the dataset shown \autoref{fig:linear_search}, this figure shows how the periodic search works at different periods $P$ (including the true period $P = 0.7$ days).
@@
 \subsection{The transit search periodogram}
-
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=0.85\linewidth]{principle_Q.pdf}
-        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
-        \label{fig:periodogram}
-    \end{centering}
-\end{figure}
 
@@
 where $\Delta$ and $\sigma$ are obtained using \autoref{eq:sigma_delta} with the last column of $X$ containing a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $1$.
-This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.\\\\
+This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_Q.pdf}
+        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
+        \label{fig:periodogram}
+    \end{centering}
+\end{figure}
 The periodic transit of period $P$ with the maximum SNR, i.e.
@@
     \begin{centering}
-        \includegraphics[width=0.85\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
+        \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
         \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise (whose characteristics are described in the main text), using \texttt{BLS} and \textsf{nuance}.
@@
     \begin{centering}
-        \includegraphics[width=0.85\linewidth]{../workflows/synthetic-injection-recovery/figures/synthetic_ir.pdf}
+        \includegraphics[width=\linewidth]{../workflows/synthetic-injection-recovery/figures/synthetic_ir.pdf}
         \caption{Injection-recovery results on 4000 simulated light curves described in \autoref{sim_dataset}.
@@
 We note that the gaps left after sigma clipping may be problematic for some of the detrending techniques (such as \texttt{bspline+BLS}).
-However, adopting this flare cleaning step and analyzing light curves with few small gaps is a practice commonly found in the literature.\\\\
+However, adopting this flare cleaning step and analyzing light curves with few small gaps is a practice commonly found in the literature.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
+    \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692.
+Top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
+On this bottom plot, the black line corresponds to the mean of one of the Gaussian Process model (with hyperparameters optimized here on the cleaned light curve).}
+    \label{fig:cleaned}
+\end{figure}
 For each light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injections and recovery per target.
@@
 with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.9\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
-    \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692.
-Top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
-On this bottom plot, the black line corresponds to the mean of one of the Gaussian Process model (with hyperparameters optimized here on the cleaned light curve).}
-    \label{fig:cleaned}
-\end{figure}
 
@@
 \autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in).
-Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (10950 with the non-dectable ones), and the lowest number of false negatives (\autoref{fig:allsearch}).
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=0.9\linewidth]{../workflows/tess_injection_recovery/figures/true_and_false_positives_bars.pdf}
-        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals.}
-        \label{fig:allsearch}
-    \end{centering}
-\end{figure}
+Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (10950 with the non-dectable ones), and the lowest number of false negatives (\autoref{fig:allsearch}).\\\\
 While the performances of other methods highly depend on the characteristics of the variability, \nuance{} leads to the highest number of true positives in 79\% of cases, and the lowest number of false positives in 72\% of cases (\autoref{fig:allsearch_relative}).
@@
     \begin{centering}
-        \includegraphics[width=0.9\linewidth]{../workflows/tess_injection_recovery/figures/true_false_positives.pdf}
+        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/true_false_positives.pdf}
         \caption{Rate of true and false positives of \nuance{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
@@
         \label{fig:allsearch_relative}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/true_and_false_positives_bars.pdf}
+        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals.}
+        \label{fig:allsearch}
     \end{centering}
@@
     \begin{centering}
-        \includegraphics[width=0.9\linewidth]{../workflows/comparison_toi/figures/TOI 540/4.pdf}
+        \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/4.pdf}
         \caption{Sector 4 light curve of TOI 540.
@@
     \begin{centering}
-        \includegraphics[width=0.9\linewidth]{../workflows/comparison_toi/figures/TOI 540/periodograms.pdf} 
+        \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/periodograms.pdf} 
         \caption{Transit search SNR periodograms of TOI-540 using \texttt{biweight+BLS} and \nuance{}.
@@
 
-% APPENDIX
+We would like to thank Julien De Wit, Prajwal Niraula and Francisco J.
+Pozuelos for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
 
@@
     \begin{centering}
-        \includegraphics[width=0.75\linewidth]{principle_dataset_decomposed.pdf}
+        \includegraphics[width=0.75\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
         \caption{Example dataset sampled at $N=2880$ times corresponding to an observation of 4 days with an exposure time of 2 minutes.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{protopapas.pdf}
+        \includegraphics[width=\linewidth]{../workflows/principle/figures/protopapas.pdf}
         \caption{Simulations of a single transit signal (\autoref{eq:protopapas}) shown for different values of $c$.}

### 2023-12-02 lib/forben.py

@@ -1,41 +1,0 @@
-import numpy as np
-from astropy.stats import LombScargle
-
-
-def bens_detrend(time, flux, n=5):
-    """iterative simple harmonics detrending"""
-
-    def rotation_period(time, flux):
-        """rotation period based on LS periodogram"""
-
-        ls = LombScargle(time, flux)
-        frequency, power = ls.autopower(
-            minimum_frequency=1 / 10, maximum_frequency=1 / 0.1
-        )
-        period = 1 / frequency[np.argmax(power)]
-        return period
-
-    def subtract_sinusoid(time, flux, period):
-        """subtract a sinusoid with period from flux"""
-        X = np.vstack(
-            [
-                np.ones(len(time)),
-                np.cos(2 * np.pi / period * time),
-                np.sin(2 * np.pi / period * time),
-            ]
-        ).T
-
-        w = np.linalg.lstsq(X, flux, rcond=None)[0]
-        model = np.dot(X, w)
-
-        return flux - model, model
-
-    _flux = flux.copy()
-    model = np.zeros_like(flux)
-
-    for _ in range(n):
-        period = rotation_period(time, _flux)
-        _flux, _model = subtract_sinusoid(time, _flux, period)
-        model += _model
-
-    return _flux + 1.0, model

### 2023-12-02 workflows/benchmark/scripts/plot.py

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import yaml
 
-plt.figure(figsize=(9, 2.5))
+plt.figure(figsize=(8, 2.5))
 
 nuance = np.array([yaml.safe_load(open(f))["linear"] for f in snakemake.input.nuance])
 bls = np.array([yaml.safe_load(open(f))["biweight"] for f in snakemake.input.bls])

### 2023-12-03 latex/paper.tex
chore: clean

@@
 \begin{abstract}
-    We present \texttt{nuance}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
+    We present \nuancemethod{}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
 To deal with these nuisance signals, a common approach consists in cleaning a light curve from correlated noise before searching for transits.
@@
 As this degradation depends on the correlated noise characteristics, we explore the parameter space for which transits are altered, and quantify this effect on a wide variety of cases.
-We show that \nuance{} outperforms the detection capabilities of commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth.
+We show that \nuancemethod{} outperforms the detection capabilities of commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth.
 We perform our tests by injecting transits on synthetic light curves, and on known TESS candidates light curves to assess the performance of our algorithm on realistic multi-planetary systems datasets.
@@
 However, it does not explore the light-curves characteristics for which such a complete modeling becomes necessary.\\\\
-In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, when referring to the \nuancemethod{} method, we typeset \nuancemethod{} in italics and, in reference to the implementation of this model, we typeset \nuancecode{} in sans-serif.}, a method to search for transit signals while modeling correlated noises in a tractable way.
+In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} typeseted in italics reference to the method, while \nuancecode{} in sans-serif reference to its implementation.}, a method to search for transit signals while modeling correlated noises in a tractable way.
 In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on planetary signals detectability.
@@
 
-A strong assumption when using BLS-type algorithms is that the searched dataset only contains transit signals and white noise.
+A strong assumption when using BLS-based algorithms is that the searched dataset only contains transit signals and white noise.
 Hence, two sources of correlated noise particularly justify the need for a detrending step before searching for transits: instrumental signals (such as those induced by telescope pointing errors) and astrophysical signals (such as stellar variability induced by pulsations or starspots).
@@
     \begin{centering}
-        \includegraphics[width=0.6\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
+        \includegraphics[width=0.65\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
         \caption{Illustration of the effect of correlated noise on a single transit SNR.
@@
 \subsection{Detrending methods and their effects}\label{detrending_effect}
-The presence of instrumental correlated noise motivated the development of systematics detrending algorithms, such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}) or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}).
-Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal and can be modeled using contemporaneous measurements (like detector's temperature, pointing error, sky background or airmass time series).
-In opposition, stellar variability and other astrophysical signals cannot be correlated with simultaneous measurements.
-This gave rise to several treatments in order to reconstruct and detrend stellar variability.
-One of them is physically-motivated and makes use of Gaussian processes (e.g.
-\citealt{k2sc}), another is empirical and makes use of filtering algorithms (\citealt{Jenkins2010}, \citealt{wotan}).\\\\
-
-In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a Gaussian Process with an SHO kernel described in \autoref{app_gp}.
-For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same Gaussian Process from which the data has been sampled.
-We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
-\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that the degradation is strongly dependant on the correlated noise characteristics encountered.
-
 \begin{figure}[H]
     \begin{centering}
-        \includegraphics[width=1\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
+        \includegraphics[width=\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
         \caption{(Top left) Simulated observation spanning 3 days, including a transit signal with a depth of 0.8\%, duration of 0.1 day, and measurement error of 1.5\% (using the model from \cite{protopapas} described in \autoref{app_proto}).
@@
 \end{figure}
+The presence of instrumental correlated noise motivated the development of systematics detrending algorithms, such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}) or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}).
+Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal and can be modeled using contemporaneous measurements (like detector's temperature, pointing error, sky background or airmass time series).
+In opposition, stellar variability and other astrophysical signals cannot be correlated with simultaneous measurements.
+This gave rise to several treatments in order to reconstruct and detrend stellar variability.
+One of them is physically-motivated and makes use of Gaussian processes (e.g.
+\citealt{k2sc}), another is empirical and makes use of filtering algorithms (\citealt{Jenkins2010}, \citealt{wotan}).\\\\
+In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a Gaussian Process with an SHO kernel described in \autoref{app_gp}.
+For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same Gaussian Process from which the data has been sampled.
+We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
+\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that the degradation is strongly dependant on the correlated noise characteristics encountered.\\\\
 In order to explore the parameter space for which detrending is the most problematic, we employ the method previously described to simulate 10 000 light curves containing a transit signal and correlated noise in the form of stellar variability.
@@
 
+\newpage
 \section{Results}\label{results}
@@
 that a transit is considered recovered if the absolute difference between the injected and recovered period is less than 0.01 days, including periods found at half or twice the injected ones (aliases) and ignoring the value of the recovered transit epoch.
-
-%% STOPPED HERE
 
@@
     \item \texttt{iterative+BLS} iteratively detrends the light curve with a sinusoidal signal fitted to the data, each time using the dominant period of the residuals found using a lomb-scargle periodogram (5 iterations).
-This detrending is followed by a search with the \texttt{BLS} algorithm.
+This detrending is followed by a search with the \texttt{BLS} algorithm.\\\\
 \end{itemize}
@@
 For simplicity, and to adopt a uniform treatment for all target light curves, a design matrix $\bm{X}$ with a single constant column is adopted, such that the systematic model only consists in a single parameter corresponding to the mean value of the differential flux (expected to be close to 1) solved linearly.
-The initial value for the rotation period of the star is found using a lomb-scargle periodogram \citep{Lomb1976, Scargle1982} implemented in \textsf{astropy}\footlink{https://docs.astropy.org/en/stable/timeseries/lombscargle.html}.\\\\
+The initial value for the rotation period of the star is found using a lomb-scargle periodogram \citep{Lomb1976, Scargle1982} implemented in \textsf{astropy}\footlink{https://docs.astropy.org/en/stable/timeseries/lombscargle.html}.
 
@@
 \autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in).
-Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (10950 with the non-dectable ones), and the lowest number of false negatives (\autoref{fig:allsearch}).\\\\
-While the performances of other methods highly depend on the characteristics of the variability, \nuance{} leads to the highest number of true positives in 79\% of cases, and the lowest number of false positives in 72\% of cases (\autoref{fig:allsearch_relative}).
-When considering both true and false positives simultaneously, \nuance{} can be considered the best technique in only 55\% of cases.
-However, when considering transits with a relative timescale $\tau < 5$ (transits that justified the development of this algorithm), \nuance{} is the best technique in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives.
-
+Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (10950 with the non-dectable ones), and the lowest number of false negatives (\autoref{fig:allsearch}).
 \begin{figure}[H]
@@
 \end{figure}
-
 \begin{figure}[H]
@@
 \end{figure}
+While the performances of other methods highly depend on the characteristics of the variability, \nuance{} leads to the highest number of true positives in 79\% of cases, and the lowest number of false positives in 72\% of cases (\autoref{fig:allsearch_relative}).
+When considering both true and false positives simultaneously, \nuance{} can be considered the best technique in only 55\% of cases.
+However, \textbf{when considering transits with a relative timescale $\tau < 5$ (transits that justified the development of this algorithm), \nuance{} is the best technique in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives}.
 
-\newpage
 \subsection{Comparison for a multi-sector TESS candidate: TOI-540}\label{toi540}
@@
 In this respect, we note the similarity of our \textit{linear search} (cf\;\autoref{linear_search}) to the one presented in \citealt{Berta2012}, that focused on the detection of single eclipses in the MEarth light curves \citep{Irwin2009}.
-Similarly, \nuancemethod{} would highly benefit the search for transiting exoplanets around M-dwarf type stars, such as the ones observed by the SPECULOOS survey \citep{speculoos} whose monitoring suffers from both increased red noise (due to atmospheric and instrumental thermal effects discussed e.g.\;in \citealt{Berta2012} and  \citealt{Pedersen2023}) and enhanced stellar variability \citep{Murray2020}.
+Similarly, \textbf{\nuancemethod{} would highly benefit the search for transiting exoplanets around M-dwarf type stars}, such as the ones observed by the SPECULOOS survey \citep{speculoos} whose monitoring suffers from both increased red noise (due to atmospheric and instrumental thermal effects discussed e.g.\;in \citealt{Berta2012} and  \citealt{Pedersen2023}) and enhanced stellar variability \citep{Murray2020}.
 We reserve this promising application to a future study.
@@
     \begin{centering}
-        \includegraphics[width=0.75\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
+        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
         \caption{Example dataset sampled at $N=2880$ times corresponding to an observation of 4 days with an exposure time of 2 minutes.
@@
 
+\newpage
 \section{Injection-recovery on TIC 1019692}\label{1019692}
 
-\begin{figure}
+\begin{figure}[H]
     \begin{centering}
-        \includegraphics[width=0.8\linewidth]{../workflows/tess_injection_recovery/figures/searched/1019692.pdf}
+        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/searched/1019692.pdf}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{centering}
         \caption{Results of the transits injection-recovery on TIC 1019692 half-sector light curve.

### 2023-12-03 workflows/plot_issues/scripts/plot_issue_2.py

@@ -115,7 +115,7 @@ for i, (flux, var, gp) in enumerate(zip(lightcurves, variabilities, cond_gps)):
 # Plots
 # -----
 offset = 0.05
-plt.figure(figsize=(13,9))
+plt.figure(figsize=(10,8))
 
 wotan_color = "C4"
 gp_color = "C3"
@@ -162,7 +162,7 @@ def text(x, method, color):
     xrange = np.array([0, 2.5]) + x + 0.7
     plt.text(xrange.mean(), OFFSET-0.01, method, ha="center", fontsize=15, c=color)
     plt.plot(xrange + 0.1, (OFFSET-0.015, OFFSET-0.015), "-", c=color, lw=2)
-    plt.text(time[-cut] + x + snr_offset, OFFSET-0.03, "transit SNR", ha="center", fontsize=14, c=txt_color)
+    plt.text(time[-cut] + x + snr_offset, OFFSET-0.03, "SNR", ha="center", fontsize=14, c=txt_color)
     plt.text(time[cut:-cut].mean() + x, OFFSET-0.03, "cleaned light curve", ha="center", fontsize=14, c=txt_color)
 
 text(x_wotan, "a. biweight", wotan_color)

### 2023-12-03 latex/paper.tex
chore: change sizes and add notebook to help tess plot

@@
     \begin{centering}
-        \includegraphics[width=0.65\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
+        \includegraphics[width=0.6\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
         \caption{Illustration of the effect of correlated noise on a single transit SNR.
@@
 For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperforms \wtls{} would only apply to this specific example.
-While more examples are considered in the following \autoref{real}, proper guidelines on when to use \textsf{nuance} instead of more classic detrending techniques are discussed in \autoref{when}.
 \begin{figure}[H]

### 2023-12-03 workflows/plot_issues/scripts/plot_issue_1.py

@@ -57,18 +57,18 @@ plt.figure(figsize=(6,5))
 offset = 0.03
 xoff = time.max()*1.2
 
-plt.text(xoff, 1.01, f"transit SNR", ha="center", fontsize=12, va="center")
+plt.text(xoff, 1.01, f"transit SNR", ha="center", fontsize=14, va="center")
 plt.text(time.mean(), 1.01, f"Light curve", ha="center", fontsize=14, alpha=0.7)
 
 plt.plot(time, fluxes[0], ".", c="0.8")
 plt.plot(time, signal, c="k", alpha=0.4, lw=2)
 plt.text(time.mean(), 1-1.8*depth, "white noise + transit", ha="center", fontsize=14)
-plt.text(xoff, 1, f"{snrs[0]:.2f}", ha="center", fontsize=12, va="center")
+plt.text(xoff, 1, f"{snrs[0]:.2f}", ha="center", fontsize=14, va="center")
 
 plt.plot(time, fluxes[1] - offset, ".", c="0.8")
 plt.plot(time, red_noise + 1 - offset, c="C3", alpha=0.5, lw=2)
 plt.text(time.mean(), 1 - offset -1.8*depth, "+ correlated noise", ha="center", fontsize=14, c="C3")
-plt.text(xoff, 1 - offset, f"{snrs[1]:.2f}", ha="center", fontsize=12, va="center")
+plt.text(xoff, 1 - offset, f"{snrs[1]:.2f}", ha="center", fontsize=14, va="center")
 
 plt.axis("off")
 plt.tight_layout()

### 2023-12-03 latex/paper.tex
fix: add processing times + GP acronym

@@
 
+\keywords{transit search, time series analysis, Gaussian processes}
+
 \begin{abstract}
@@
 To deal with these nuisance signals, a common approach consists in cleaning a light curve from correlated noise before searching for transits.
-However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signals, up to the point of no detection.
-As this degradation depends on the correlated noise characteristics, we explore the parameter space for which transits are altered, and quantify this effect on a wide variety of cases.
-We show that \nuancemethod{} outperforms the detection capabilities of commonly used transit search algorithms, especially for light curves featuring correlated noise with an amplitude three times greater than the searched transit depth.
-We perform our tests by injecting transits on synthetic light curves, and on known TESS candidates light curves to assess the performance of our algorithm on realistic multi-planetary systems datasets.
-Beyond its detection efficiency, we make \nuance{} tractable, \TODO{} times faster than current alternatives based on brute force detrending, and available as a well documented open-source Python package\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}}.
+However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signals down to no detection.
+As this degradation depends on the correlated noise characteristics, we explore the relative parameter space for which transits are altered, and quantify this effect on a set of simulated light curves.
+On this synthetic dataset, we show that \nuancemethod{} largely outperforms the detection capabilities of a commonly used transit search technique, which consists in cleaning light curves with an optimal bi-weight filter before searching for transits using a box-least-square algorithm.
+In order to assess the performance of \nuancemethod{} on real datasets, we inject and recover transits into the TESS light curves of 438 M dwarfs, and compare our method to 5 detrending techniques followed by a box-least-square search.
+For transits with a duration exceeding one fifth of the stellar variability timescale, i.e.
+when both signals start ressembling each other, we show that our algorithm is the most perforant in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives detections.
+Although simultaneously searching for transits while modeling correlated noise is expected to be computationaly expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
+Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of both active stars and their transiting exoplanets, both using space-based and sparse ground-based observations.
 \end{abstract}
@@
 A 1-hour transit signal of depth 1\% is generated on top of white noise (variance of $0.0015^2$) as part of a 24-hours observation with an exposure time of 1 minute (top).
-Then, in the bottom plot, correlated noise is added, generated using a Gaussian Process with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%.
+Then, in the bottom plot, correlated noise is added, generated using a Gaussian process (GP) with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%.
 The SNR on the right of each light curve is computed using \autoref{eq:snr}.
@@
 (Left) Correlated noise in the form of stellar variability with different timescales and amplitudes added to the unique transit signal.
-This signal is then reconstructed using (in purple) Tukey's bi-weight filter \citep{wotan} with an optimal window size of three times the transit duration, and (in red) a Gaussian Process with the same kernel used to simulate the data.
+This signal is then reconstructed using (in purple) Tukey's bi-weight filter \citep{wotan} with an optimal window size of three times the transit duration, and (in red) a GP with the same kernel used to simulate the data.
 In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr} with the transit depth estimated as the minimum in-transit flux.}
@@
 This gave rise to several treatments in order to reconstruct and detrend stellar variability.
-One of them is physically-motivated and makes use of Gaussian processes (e.g.
+One of them is physically-motivated and makes use of GP (e.g.
 \citealt{k2sc}), another is empirical and makes use of filtering algorithms (\citealt{Jenkins2010}, \citealt{wotan}).\\\\
-In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a Gaussian Process with an SHO kernel described in \autoref{app_gp}.
-For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same Gaussian Process from which the data has been sampled.
+In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a GP with an SHO kernel described in \autoref{app_gp}.
+For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same GP from which the data has been sampled.
 We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
@@
 These combined likelihoods yield a transit-search periodogram on which the periodic transit detection is based.
-\textsf{nuance} differs from \cite{foreman2016} and other existing transit search algorithms as it models the covariance of the light curve with a Gaussian Process, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
+\textsf{nuance} differs from \cite{foreman2016} and other existing transit search algorithms as it models the covariance of the light curve with a GP, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
 This way, \textsf{nuance} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR.\\\\
@@
 \\\\
-To account for correlated noise, the light curve $f$ is modeled as being drawn from a Gaussian Process, like the simulated light curves used in \autoref{issues}, such that
+To account for correlated noise, the light curve $f$ is modeled as being drawn from a GP, like the simulated light curves used in \autoref{issues}, such that
 \begin{equation*}
@@
 This motivates a final step to produce the transit search periodogram $\mathcal{Q}$.
-For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, the maximum likelihood parameters
+For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, we compute the maximum likelihood parameters
 \begin{equation}\label{eq:phi0}
@@
 \end{equation}
-are retrieved, and $\mathcal{Q}(P)$ is defined as the SNR of the transit of period $P$, epoch $T_0 = \phi_0 P$, duration $D$ and depth $\Delta$, i.e.
+and defined $\mathcal{Q}(P)$ as being equal to the SNR of the transit of period $P$, epoch $T_0 = \phi_0 P$, duration $D$ and depth $\Delta$, i.e.
 \begin{equation*}
@@
 \subsection{An open-source python package}
-Methods presented in this paper are made accessible through the \textsf{nuance} open-source Python package, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} and released on the Python Package Index (PyPI)\footlink{https://pypi.org/project/nuance/}.
+Methods presented in this paper are made accessible through the \nuancecode{} open-source Python package, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} and released on the Python Package Index (PyPI)\footlink{https://pypi.org/project/nuance/}.
 \\\\
@@
 From this \texttt{search} object, the best transiting candidate parameters can be computed (\textsf{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\textsf{search.Q\_snr}), together with valuable information about the transit search.
-The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \textsf{nuance} being that the \textit{linear search} only needs to be performed ones, and reused for the search of several transiting candidates.
+The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \nuancemethod{} being that the \textit{linear search} only needs to be performed ones and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
 An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.
@@
 
-To start testing \textsf{nuance} against existing methods, a simple light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ (relative flux), observed for a duration of 6 days with a cadence of 2 minutes.
-In this signal, box-shaped transits are injected with periods randomly sampled from 0.3 to 2.5 days, a unique transit duration of 50 minutes, and depths randomly sampled from $2.6 \times 10^{-4}$ to $6.8 \times 10^{-4}$ (using the simple model from \cite{protopapas} presented in \autoref{app_proto} with $c=500$).
+To start testing the \nuancemethod{} approach against existing methods, a simple light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ (relative flux), observed for a duration of 6 days with a cadence of 2 minutes.
+In this signal, box-shaped transits are injected with periods randomly sampled from 0.3 to 2.5 days, a unique transit duration of 50 minutes, and depths randomly sampled to lead to a transit SNR from 4 to 30.
+% (i.e.\;depths from $2.6 \times 10^{-4}$ to $6.8 \times 10^{-4}$, using the simple model from \cite{protopapas} presented in \autoref{app_proto} with $c=500$).
 By design, these injected signals have an SNR ranging from 5 to 30 (with $\sigma_r = 0$ in \autoref{eq:snr}).
-For each light curve, transits are searched using two different methods: \textsf{nuance}, using its implementation from the Python package described in the previous section; and \texttt{BLS}, the box-least-square algorithm \citep{bls}, using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}.
-For \nuance{}\\\\
+For each light curve, transits are searched using two different methods: \textsf{nuance}, using its implementation from the Python package described in the previous section; and \textsf{BLS}, the box-least-square algorithm \citep{bls}, using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}.\\\\
 For both methods, 3000 trial periods from 0.2 to 2.6 days are searched, with a single trial duration fixed to the unique known duration of 50 minutes.
 A transit signal is considered detected if the absolute difference between the injected and the recovered period is less than 0.01 day.
-To ease the detection criteria, transit periods recovered at half or twice the injected period (aka \textit{aliases}) are considered as being detected.
+To ease the detection criteria, orbital periods recovered at half or twice the injected ones (aka \textit{aliases}) are considered as being detected.
 For this reason, detected transit epochs are not considered.
@@
         \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
-        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise (whose characteristics are described in the main text), using \texttt{BLS} and \textsf{nuance}.
+        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise (whose characteristics are described in the main text), using \textsf{BLS} and \textsf{nuance}.
 The color scale indicates the recovery of transits in the corresponding (period, depth) parameters space, white for a full recovery and black for no detection.}
@@
 Here, we evaluate the performance of \textsf{nuance} in the relative parameter space $(\tau, \delta)$ described in \autoref{eq:relative_params} and see when its specific treatment of correlated noise in the transit search becomes necessary.\\\\
-We compare \textsf{nuance} to the approach that involves removing stellar variability from the light curves before performing the search on a detrended dataset.
-Similar to \autoref{detrending_effect}, an optimal bi-weight filter implemented in the \textsf{wõtan} Python package\footlink{https://github.com/hippke/wotan} is used with a window size three times that of the injected transit duration.
+We compare \textsf{nuance} to the approach that involves removing stellar variability from light curves before performing the search on a detrended dataset.
+Similarly to \autoref{detrending_effect}, an optimal bi-weight filter implemented in the \textsf{wõtan} Python package\footlink{https://github.com/hippke/wotan} is used with a window size three times that of the injected transit duration.
 A transit search on the detrended light curve is then performed using the \texttt{BLS} algorithm, like in \autoref{control} using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}.
@@
 When considering both true and false positives simultaneously, \nuance{} can be considered the best technique in only 55\% of cases.
-However, \textbf{when considering transits with a relative timescale $\tau < 5$ (transits that justified the development of this algorithm), \nuance{} is the best technique in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives}.
+However, \textbf{when considering transits with a relative timescale $\tau < 5$ (transits that justified the development of this algorithm), \nuancemethod{} is the best technique in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives}.
 
@@
 The resulting light curve for sector 4 and its mean model is shown in \autoref{fig:toi540_clean}.
-
 \begin{figure}[H]
@@
 This search, using \nuance{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the \texttt{BLS} algorithm (denoted \texttt{biweight+BLS} and described in \autoref{real}) on all sectors combined.
-Since we don't know the transit duration a-priori, we employ a sliding window duration of 5 hours, small enough to encompass the stellar variability of TOI-540 but large enough not to detrend existing transits with durations lower tham 1.5 hours.
+Since we don't know the transit duration a-priori, we perform the detrending and search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
 The results of this comparison are shown in \autoref{fig:toi540_periodograms}.
-
 \begin{figure}[H]
@@
 \end{figure}
-
 After a first periodic search, the times falling in the detected periodic transit signal (the one with an SNR higher than the detection limit set to 7) are masked.
@@
 \noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple non-optimized GP with a squared exponential kernel.
-These are compared to the processing time of \texttt{biweight+BLS} (cf.\;\autoref{simu}) separated into the \textit{biweight} filtering step and the \textsf{BLS} search.
+These are compared to the processing time of \texttt{biweight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the \textsf{BLS} search.
 \begin{figure}[H]
@@
 
-As see in \autoref{fig:blsvsnuance}, most of the computational cost of \nuancecode{} and the \texttt{biweight+BLS} method comes from the \textit{linear search} and the \textit{biweight} filtering step.
+As see in \autoref{fig:blsvsnuance}, most of the computational cost of \nuancecode{} and the \texttt{biweight+BLS} method comes from the \textit{linear search} and the bi-weight filtering step.
 One advantage of \nuancemethod{} is that the \textit{linear search} can be performed separately on different continuous observations, and then combined in the periodic search.
@@
 The second advantage of \nuancecode{} is its \textsf{JAX} implementation, allowing the linear search to be run in parallel.
-As a reference, the 12 cores of an Apple M2 Max chip, searching for TOI-540 b transits with nuance (\autoref{toi540}) took 3 minutes in total, 1 minute for the 5 sectors linear searches, and 2 minutes for the combined periodic search.
-In comparison, the search with \texttt{biweight+BLS} took about 4 seconds.\\\\
+As a reference, searching for TOI-540 b transits with nuance (\autoref{toi540}), on the 12 cores of an Apple M2 Max chip, took 5 minutes 35 seconds in total, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds each), and 3 minutes 45 seconds for the combined periodic search.
+In comparison, the brute force search with \texttt{biweight+BLS}, consisting in trying 15 bi-weight windows, took a total of 1 minute 49 seconds (about only 7 seconds each).\\\\
 % As the periodic search is expected to be improved by lower lever optimizations (using a compiled language), we reserve its benchmark for a future work.
@@
 Because of its computational cost, \nuancecode{} must not be employed in the general case, but rather adopted when light curves contain correlated noise with specific characteristics.
-If employing a \textit{biweight} filter for detrending, these characteristics correspond to the ones discussed in \autoref{issues}.
+If employing a bi-weight filter for detrending, these characteristics correspond to the ones discussed in \autoref{issues}.
 But as these strongly depend on the type of detrending technique employed, we do not provide general guidelines as when \nuancemethod{} should be preferred over a specific detrending technique.
@@
 The mean of this signal consists in a periodic transit signal of period $P=0.7$ days, duration $D=0.05$ days and depth of 2\% ($\mathcal{T}$ in gray) plus instrumental signals ($\mathcal{S}$ in blue).
-Correlated noise in the form of stellar variability is simulated by modeling the covariance matrix of the signal with a Gaussian process ($\mathcal{V}$ in red) including a diagonal variance of $0.001^2$ corresponding to white noise ($\epsilon$ in light grey).
+Correlated noise in the form of stellar variability is simulated by modeling the covariance matrix of the signal with a GP ($\mathcal{V}$ in red) including a diagonal variance of $0.001^2$ corresponding to white noise ($\epsilon$ in light grey).
 This simulated signal is not intended to be physically realistic.}
@@
 i.e.
-that $\bm{f}$ is drawn from a Gaussian process of mean $\bm{\mu}$ and covariance matrix $\bm{C}$.
+that $\bm{f}$ is drawn from a GP of mean $\bm{\mu}$ and covariance matrix $\bm{C}$.
 In this equation, $\bm{\mu}$ is such that its $i$-th element is defined by  $\mu_i = \mathcal{T}(t_i) + \mathcal{S}(t_i)$ where $t_i$ is the $i$-th time of observation, $\mathcal{T}$ is a periodic transit function and $\mathcal{S}$ a function describing the instrumental part of the signal (both described below).

### 2023-12-04 latex/paper.tex
chore: typo fixes and GP acronym

@@
 
-\keywords{transit search, time series analysis, Gaussian processes}
+\keywords{transit search, exoplanets, stellar activity, time series analysis, Gaussian processes}
 
@@
 However, it does not explore the light-curves characteristics for which such a complete modeling becomes necessary.\\\\
-In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} typeseted in italics reference to the method, while \nuancecode{} in sans-serif reference to its implementation.}, a method to search for transit signals while modeling correlated noises in a tractable way.
+In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} typeseted in italics refers to the method, while \nuancecode{} in sans-serif refers to its implementation.}, a method to search for transit signals while modeling correlated noises in a tractable way.
 In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on planetary signals detectability.
@@
 \end{equation} 
-with $\bm{\Sigma}$ the covariance matrix modeled using the Gaussian Process.
+with $\bm{\Sigma}$ the covariance matrix modeled using the GP.
 Hence, $\ln p(\bm{f} \vert I)$ can be computed on a grid of epochs and durations, the transit depth being linearly solved for any $(T, D)$.
@@
 were $\Delta_{i,j}$ is the depth linearly solved for a given $(T_i, D_j)$\footnote{The expression of $\set{\ln\mathcal{L}_{i,j}}_{i, j}$ omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{i,j}$) as it is also linearly solved for any given value of $(T_i, D_j)$ and irrelevant in what follows.}.
-\autoref{fig:linear_search} shows this likelihood grid computed for a simulated dataset, using the same Gaussian Process  and design matrix $\bm{X}$ used to simulate the data.
+\autoref{fig:linear_search} shows this likelihood grid computed for a simulated dataset, using the same GP  and design matrix $\bm{X}$ used to simulate the data.
 
@@
 \subsection{An open-source python package}
-Methods presented in this paper are made accessible through the \nuancecode{} open-source Python package, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} and released on the Python Package Index (PyPI)\footlink{https://pypi.org/project/nuance/}.
+Methods presented in this paper are made accessible through the \nuancecode{} open-source Python package, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} and released on the Python Package Index\footlink{https://pypi.org/project/nuance/}.
 \\\\
@@
 \end{lstlisting}
-where \textsf{gp} is a \textsf{tinygp} Gaussian Process instance and \textsf{X} the design matrix of the linear model.
+where \textsf{gp} is a \textsf{tinygp} GP instance and \textsf{X} the design matrix of the linear model.
 \textsf{nuance} exploits the use of \textsf{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package powered by \textsf{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}, allowing for custom kernels to be built and highly tractable computations.
@@
 \begin{lstlisting}[language=Python]
-# a wide range of periods
+# range of periods
 periods = np.linspace(0.1, 5, 2000)
@@
 From this \texttt{search} object, the best transiting candidate parameters can be computed (\textsf{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\textsf{search.Q\_snr}), together with valuable information about the transit search.
-The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary systems, the advantage of \nuancemethod{} being that the \textit{linear search} only needs to be performed ones and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
+The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary hosts, the advantage of \nuancemethod{} being that the \textit{linear search} only needs to be performed ones and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
 An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.
@@
 
-To start testing the \nuancemethod{} approach against existing methods, a simple light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ (relative flux), observed for a duration of 6 days with a cadence of 2 minutes.
-In this signal, box-shaped transits are injected with periods randomly sampled from 0.3 to 2.5 days, a unique transit duration of 50 minutes, and depths randomly sampled to lead to a transit SNR from 4 to 30.
+To start testing \nuancemethod{} against existing methods, a simple adimensional normalized light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ spanning 6 days with an exposure time of 2 minutes.
+From this signal, we produce 4000 light curves, each containing box-shaped transits with periods randomly sampled from 0.3 to 2.5 days, durations of 50 minutes, and depths randomly sampled to lead to transit SNRs ranging from 4 to 30.
 % (i.e.\;depths from $2.6 \times 10^{-4}$ to $6.8 \times 10^{-4}$, using the simple model from \cite{protopapas} presented in \autoref{app_proto} with $c=500$).
 By design, these injected signals have an SNR ranging from 5 to 30 (with $\sigma_r = 0$ in \autoref{eq:snr}).
-For each light curve, transits are searched using two different methods: \textsf{nuance}, using its implementation from the Python package described in the previous section; and \textsf{BLS}, the box-least-square algorithm \citep{bls}, using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}.\\\\
+For each light curve, transits are searched using two different methods: \textsf{nuance}, using its implementation from the Python package described in the previous section; and \textsf{BLS}, the box-least-square algorithm from \cite{bls} (using \textsf{astropy}'s \textsf{BoxLeastSquares}\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html} implementation).\\\\
 For both methods, 3000 trial periods from 0.2 to 2.6 days are searched, with a single trial duration fixed to the unique known duration of 50 minutes.
@@
 To ease the detection criteria, orbital periods recovered at half or twice the injected ones (aka \textit{aliases}) are considered as being detected.
-For this reason, detected transit epochs are not considered.
+For this reason, detected transit epochs are not considered (although manually vetted).
 Results from this \textit{injection-recovery} procedure are shown in \autoref{fig:control}.
@@
         \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
-        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise (whose characteristics are described in the main text), using \textsf{BLS} and \textsf{nuance}.
-The color scale indicates the recovery of transits in the corresponding (period, depth) parameters space, white for a full recovery and black for no detection.}
+        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using \textsf{BLS} and \nuancecode{}.
+The color scale indicates the recovery of transits in the corresponding (period, depth) parameter space, white for a full recovery and black for no detection.}
         \label{fig:control}
@@
 Similarly to \autoref{detrending_effect}, an optimal bi-weight filter implemented in the \textsf{wõtan} Python package\footlink{https://github.com/hippke/wotan} is used with a window size three times that of the injected transit duration.
-A transit search on the detrended light curve is then performed using the \texttt{BLS} algorithm, like in \autoref{control} using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}.
-This strategy is widely used in the community and is denoted \texttt{biweight+BLS}.
+A transit search on the detrended light curve is then performed using the \texttt{BLS} algorithm, like in \autoref{control} using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}; a strategy commonly used in the literature that we denote \texttt{biweight+BLS}.
 In what follows, the transit detection criteria are the same as the ones used in \autoref{control}, i.e.
@@
 \\\\
-For a given pair of $(\tau, \delta)$, we simulate stellar variability using a Gaussian Process with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params} (except for $Q$, described below), computed with respect to the injected transit parameters $D$ and $\Delta$.
+For a given pair of $(\tau, \delta)$, we simulate stellar variability using a GP with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params} (except for $Q$, described below), computed with respect to the injected transit parameters $D$ and $\Delta$.
 The same kernel is used for the search with \textsf{nuance}, an optimal choice on equal footing with the optimal $3\times D$ window size of the bi-weight filter employed in the \wtls{} search.
-While this kernel is optimal, a comprehensive discussion on the use of non-optimal kernels is discussed in .
+%  While this kernel is optimal, a comprehensive discussion on the use of non-optimal kernels is discussed in . 
  4000 pairs of $(\tau, \delta)$ are generated such that
@@
 \subsubsection*{Results}
-The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against the \wtls{} strategy on transits with relatively low depth compared to the stellar variability amplitude, and a relatively small duration compared to the stellar variability period.
+The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against the \wtls{} strategy on transits with relatively large depth compared to the stellar variability amplitude, and a relatively large duration compared to the stellar variability period.
 This empirical statement only concerns light curves with a given amount of white noise, and may vary depending on the length of the observing window or the number of transits.
@@
 
-\noindent This injection-recovery is done in a particularly optimal setup, on light curves not all physically realistic and using an optimal Gaussian Process kernel, hence demonstrating the performance of \textsf{nuance} only on a purely experimental basis.
+\noindent This injection-recovery is done in a particularly optimal setup, on light curves not all physically realistic and using an optimal GP kernel, hence demonstrating the performance of \textsf{nuance} only on a purely experimental basis.
 In the next section, we perform transits injection-recovery on real space-based light curves.
@@
 Top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
-On this bottom plot, the black line corresponds to the mean of one of the Gaussian Process model (with hyperparameters optimized here on the cleaned light curve).}
+On this bottom plot, the black line corresponds to the mean of one of the GP model (with hyperparameters optimized here on the cleaned light curve).}
     \label{fig:cleaned}
@@
 \subsubsection*{Stellar variability kernel}\label{rotation_kernel}
-One of the particularity of nuance is to model a light curve as a Gaussian Process.
+One of the particularity of nuance is to model a light curve as a GP.
 To model these TESS light curves, we employ a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
@@
         \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/4.pdf}
-        \caption{Sector 4 light curve of TOI 540.
-The clean signal (gray points) has been mask masked for flares (black points), and the black line corresponds to the mean of the Gaussian Process model.}
+        \caption{Sector 4 light curve of TOI 540. The clean signal (gray points) has been mask masked for flares (black points), and the black line corresponds to the mean of the GP model.}
         \label{fig:toi540_clean}
@@
 In practice, this is done by masking the linear search products \\\\
-As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{biweight+BLS} and \texttt{nuance} are very similar.
+As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{biweight+BLS} and \texttt{nuance} are very similar, with the known transiting exoplanet TOI-540 b detected with an orbital period P = 1.24 days.
 This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b (using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ min of TOI-540 b \citep{TOI540}).
@@
 The proper treatment of correlated noise in \nuance{}, as observed in \autoref{real}, makes these peaks non-significant, avoiding a large number of false detections.\\\\
-We note that finding a TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques.
-We reserve the search for such new transit signals to a follow-up publication.
+We note that finding a TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques (see e.g.\;\autoref{fig:simu}).
+Nontheless, we verify that \nuancemethod is capable of finding a large number of already known transiting exoplanet candidates, in light curves featuring various forms of correlated noises, at least as efficiently as with commonly used techniques.
+We reserve the search for new transit signals to a follow-up publication.
 
@@
 \subsection{The choice of kernel}
-While not being discussed in our study, the efficiency of \nuancemethod{} to detect transits in correlated noise is highly dependant on the design of its GP kernel.
+While not being discussed in our study, \textbf{the efficiency of \nuancemethod{} to detect transits in correlated noise is highly dependant on the design of its GP kernel}.
 In the ensemble comparison of \autoref{real}, the goal was to choose a kernel and an optimization strategy suited to most of the studied light curves, leading to few outliers in the results that were indicative of a badly designed and/or optimized kernel.
@@
 \subsection{Stellar variability $\nu$}\label{app_gp}
-As this chapter focuses on stellar variability and its effect on transit detection, We employ a simple physically-motivated Gaussian Process kernel, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) taking the form
+As this chapter focuses on stellar variability and its effect on transit detection, We employ a simple physically-motivated GP kernel, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) taking the form
 \begin{equation}
@@
 where $Q$ is the quality factor of the oscillator, $\omega$ its pulsation and $\sigma$ the amplitude of the kernel function.
-Gaussian Process computations in this paper use the implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package exposing the quasi-separable kernels from \cite{celerite2} and powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
-The stellar variability signal in \autoref{fig:app_principle_dataset} has been sampled from a Gaussian Process with an SHO kernel of parameters $\omega = \pi/6D$ (i.e.
+GP computations in this paper use the implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package exposing the quasi-separable kernels from \cite{celerite2} and powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
+The stellar variability signal in \autoref{fig:app_principle_dataset} has been sampled from a GP with an SHO kernel of parameters $\omega = \pi/6D$ (i.e.
 a period equal to 12 times the duration $D$ of the simulated transit), $Q=45$ and $\sigma=0.02$, the depth of the simulated transit.
@@
         \caption{Results of the transits injection-recovery on TIC 1019692 half-sector light curve.
-Left: cleaned light curve with computed trend overplotted in black (except for \textsf{nuance} where it corresponds to the mean of the Gaussian Process model).
+Left: cleaned light curve with computed trend overplotted in black (except for \textsf{nuance} where it corresponds to the mean of the GP model).
 Right: Results of the transit search where a black square denotes a transit signal not detected, gray a signal detected at an alias period ($P/2$ or $2P$), and  white a signal detected with the correct period.

### 2023-12-05 latex/paper.tex
feat: full review + conclusion draft

@@
 \author{Catriona A. Murray}
-\author{Francisco J. Pozuelos}
 \correspondingauthor{Lionel J. Garcia}
@@
     We present \nuancemethod{}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
-To deal with these nuisance signals, a common approach consists in cleaning a light curve from correlated noise before searching for transits.
-However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signals down to no detection.
+To deal with them, a common approach consists in cleaning a light curve from nuisance signals before searching for transits.
+However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signal-to-noise-ratio down to no detection.
 As this degradation depends on the correlated noise characteristics, we explore the relative parameter space for which transits are altered, and quantify this effect on a set of simulated light curves.
-On this synthetic dataset, we show that \nuancemethod{} largely outperforms the detection capabilities of a commonly used transit search technique, which consists in cleaning light curves with an optimal bi-weight filter before searching for transits using a box-least-square algorithm.
-In order to assess the performance of \nuancemethod{} on real datasets, we inject and recover transits into the TESS light curves of 438 M dwarfs, and compare our method to 5 detrending techniques followed by a box-least-square search.
+On this synthetic dataset, we show that \nuancemethod{} largely outperforms the detection capabilities of a commonly used transit search technique, which consists in cleaning light curves from styellar variability with an optimal bi-weight filter before searching for transits using a box-least-square algorithm.
+Then, in order to assess the performance of \nuancemethod{} on real datasets, we inject and recover transits into the TESS light curves of 438 M dwarfs, and compare our method to 4 detrending techniques followed by a Box-Least-Square search.
 For transits with a duration exceeding one fifth of the stellar variability timescale, i.e.
-when both signals start ressembling each other, we show that our algorithm is the most perforant in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives detections.
-Although simultaneously searching for transits while modeling correlated noise is expected to be computationaly expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
-Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of both active stars and their transiting exoplanets, both using space-based and sparse ground-based observations.
+when both signals start resembling each other, we show that our algorithm is the most performant in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
+Although simultaneously searching for transits while modeling correlated noise is expected to be computationaly expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices (?).
+Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of active stars and their transiting exoplanets, both using space-based and sparse ground-based observations.
 \end{abstract}
@@
 However, discovering transiting exoplanets around active stars holds few promises.
-First, as younger stars are more active\footnote{approximately down to spectral type MV ()} \citep{Skumanich1972}, being able to detect planets transiting active stars will favor the discovery of young planetary systems \citep[e.g.][]{Newton2022}.
+First, as younger stars are more active \citep{Skumanich1972}, being able to detect planets transiting active stars will favor the discovery of young planetary systems \citep[e.g.][]{Newton2022}.
 Second, as stellar variability may originate from surface active regions (such as starpots), the increased discovery of occulting companions will increase our chances to map the photosphere of active stars \citep[e.g.][]{Morris2017}, benefiting both the study of stellar atmospheres and the concerning impact of their non-uniformity on planetary atmosphere retrievals \citep{rackham2018}.
@@
 However, it does not explore the light-curves characteristics for which such a complete modeling becomes necessary.\\\\
-In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} typeseted in italics refers to the method, while \nuancecode{} in sans-serif refers to its implementation.}, a method to search for transit signals while modeling correlated noises in a tractable way.
-In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on planetary signals detectability.
-In \autoref{nuance}, we present \nuance{}.
+In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} written in italics refers to the algorithmic and analytical method, while \nuancecode{} in sans-serif refers to its implementation.}, a method to search for transit signals while modeling correlated noises in a tractable way.
+In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on transit signals detectability.
+In \autoref{nuance}, we present \nuance{}, and the two main steps on which this method is based: the linear search and the periodic search.
 In \autoref{results}, we test the performance of \nuance{} on a wide variety of cases, and compare it to commonly used transit search algorithms.
-Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, and we conclude in \autoref{conclusion}.
+This include transits injected in synthetic datasets, but also in the TESS light curves of 438 rapidly-rotating M dwarfs.
+Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, and we conclude in .
 
@@
 
-A strong assumption when using BLS-based algorithms is that the searched dataset only contains transit signals and white noise.
+A strong assumption when using the BLS algorithm to search for transits is that the searched dataset only contains these signals in addition of white noise.
 Hence, two sources of correlated noise particularly justify the need for a detrending step before searching for transits: instrumental signals (such as those induced by telescope pointing errors) and astrophysical signals (such as stellar variability induced by pulsations or starspots).
@@
 \end{equation}
-where $\Delta$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits ($N_{tr}=1$ here), and $\sigma_w^2$ and $\sigma_c^2$ are the white and correlated noise variances.
+where $\Delta$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits ($N_{tr}=1$ here), and $\sigma_w^2$ and $\sigma_c^2$ are the white and correlated noise variances\footnote{i.e.
+$\sigma_w^2 = \sum_{i=j} C_{{ij}}$ and $\sigma_c^2 = \sum_{i\neq j} C_{{ij}}$ with $C$ the covariance matrix of the data.}.
 \autoref{fig:issue1} shows this metric computed for a unique transit observed in the absence (grey) and presence (red) of correlated noise.
@@
 A 1-hour transit signal of depth 1\% is generated on top of white noise (variance of $0.0015^2$) as part of a 24-hours observation with an exposure time of 1 minute (top).
-Then, in the bottom plot, correlated noise is added, generated using a Gaussian process (GP) with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%.
+Then, in the bottom plot, correlated noise is added, generated using a Gaussian process (GP; \citealt{Rasmussen2005}) with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%.
 The SNR on the right of each light curve is computed using \autoref{eq:snr}.
@@
 \subsection{Detrending methods and their effects}\label{detrending_effect}
+The presence of instrumental correlated noise motivated the development of systematics detrending algorithms, such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}) or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}).
+Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal and can be modeled using contemporaneous measurements (like detector's temperature, pointing error, sky background or airmass time series).
 \begin{figure}[H]
@@
 \end{figure}
-The presence of instrumental correlated noise motivated the development of systematics detrending algorithms, such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}) or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}).
-Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal and can be modeled using contemporaneous measurements (like detector's temperature, pointing error, sky background or airmass time series).
 In opposition, stellar variability and other astrophysical signals cannot be correlated with simultaneous measurements.
 This gave rise to several treatments in order to reconstruct and detrend stellar variability.
-One of them is physically-motivated and makes use of GP (e.g.
-\citealt{k2sc}), another is empirical and makes use of filtering algorithms (\citealt{Jenkins2010}, \citealt{wotan}).\\\\
+Some of them are physically-motivated and make use of GPs (e.g.
+\citealt{k2sc}), others are empirical and make use of filtering and data-driven algorithms (\citealt{Jenkins2010}, \citealt{wotan}).\\\\
 In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a GP with an SHO kernel described in \autoref{app_gp}.
@@
 We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
-\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that the degradation is strongly dependant on the correlated noise characteristics encountered.\\\\
+\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that this degradation due to detrending is strongly dependant on the correlated noise characteristics encountered.\\\\
 In order to explore the parameter space for which detrending is the most problematic, we employ the method previously described to simulate 10 000 light curves containing a transit signal and correlated noise in the form of stellar variability.
-For each light curve, we reconstruct the variability signal and detrend it using an optimal bi-weight filter before estimating the remaining transit SNR.
+For each light curve, we reconstruct the variability signal and detrend it using an optimal bi-weight filter, i.e.
+with a window size three times that of the transit duration \citep{wotan}, before estimating the remaining transit SNR.
 For convenience, the simulated light curves share a common transit added on top of white noise, and a variability signal only defined by two parameters: $\tau$, the relative timescale of the variability with respect to the transit duration; and $\delta$, the relative amplitude of the variability against transit depth, corresponding to an SHO kernel with hyperparameters chosen as:
@@
 
-For $\tau=1$ and $\delta=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a variability signal with a period half that of the transit duration, and a standard deviation two times that of the transit depth, i.e.
+For $\tau=1$ and $\delta=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a variability signal with a period half that of the transit duration, and a correlated noise amplitude comparable to the transit depth, i.e.
 strongly resembling the simulated transit signal.\\\\
 \autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 6$).
-Hence, bi-weight filter detrending makes transit-search blind to many systems.
-While this study should be extended to other detrending techniques, it highlights the need for a more informed transit search algorithm able to deal with correlated noise.
+Hence, bi-weight filter detrending makes transit search blind to many systems.
+While this study should be extended to other detrending techniques, it highlights the need for a more informed transit search algorithm able to deal with correlated noise, at least if present in the form of stellar variability.
+
+% ## STOPPED here
 
@@
         \includegraphics[width=0.9\linewidth]{../workflows/cleaning_snr/figures/result.pdf}
-        \caption{SNR of a unique transit after detrending light curves using an optimal bi-weight filter.
-All light curves correspond to a 2.8 days observation with a cadence of 2 minutes, and contain a unique transit of duration 1 hour and a depth of 1\%, added on top of white noise with a standard deviation of 0.5\% plus stellar variability.
-Light curves at the top and right side of the central plot are shown with their corresponding $\tau$ and $\delta$ values.
-The amplitude of the variability increases with $\delta$ and the timescale of the variability increases with $\tau$.}
+        \caption{SNR of a unique transit after detrending light curves using an optimal bi-weight filter (i.e.\;of window size $3\times D$).
+All light span 2.8 days with exposure times of 2 minutes.
+The 10 000 simulated datasets contain the same transit signal with a duration $D=1$ hour and a depth of 1\%, added on top of white noise with a standard deviation of 0.5\% plus stellar variability.
+Light curves at the top and right side of the central plot are shown with their corresponding $\tau$ and $\delta$ values, which again correspond to the relative timescale of stellar variability against the transit duration, and the relative amplitude of stellar variability against the transit depth.
+Hence, the amplitude of the variability increases with $\delta$ and the timescale of the variability increases with $\tau$, the transit signal remaining fixed.}
         \label{fig:snr_detrend}
@@
 Let assume that the flux $f$ of a star is observed and arranged in the vector $\bm{f}$ of size $N$, associated to the vector of times $\bm{t}$.
-This flux, shown in \autoref{fig:linear_search}, contains instrumental signals, stellar variability and a periodic transit signal that we wish to uncover.
+This flux, shown in \autoref{fig:linear_search}, contains instrumental signals, stellar variability and a periodic transit signal that we wish to detect.
 We assume that a set of $M$ observed measurements (such as the position of the star on the detector or the sky background) taken at the same time as the flux can be treated as explanatory variables for $\bm{f}$.
 These measurements constitute the columns of the $(N\times M)$ \textit{design matrix} $\bm{X}$.\\\\
-
 Ideally, we would detect the periodic transit signal in this flux by sampling the posterior likelihood of this data to a full-fledge model including stellar variability (more generally correlated noise), instrumental systematic signals (modeled with explanatory variables or treated as correlate noise), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$.
 We would then reduce the posterior likelihood to $p(\bm{f}\vert P)$, its marginalized version over all parameters except the period $P$, producing a transit search periodogram $\mathcal{Q}(P)$.
-However, this approach has two issues: It is highly intractable, and it may lead to multimodal distributions that are hard to interpret.
+However, this approach has two issues: It is highly intractable, and it may lead to a multimodal distribution that is hard to interpret.
 \\\\
@@
 \end{equation}
-This is done by adopting the strategy of \cite{foreman2016}, and separate the transit search into two components: the \textit{linear search} and the periodic search.
-During the \textit{linear search}, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
+This is done by adopting the strategy of \cite{foreman2016}, separating the transit search into two components: the linear search and the periodic search.
+During the linear search, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
 Then, the periodic search consists in combining these likelihoods to compute the likelihood of the data given a periodic transit signal for a range of periods.
@@
 This way, \textsf{nuance} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR.\\\\
-We note that the approach employed by \textsf{nuance} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.
+We note that the approach employed by \nuancemethod{} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.
 
@@
 
-During the \textit{linear search}, the goal is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of the data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths.
+The goal of the linear search is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of the data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths.
 \\\\
@@
 Hence, $\ln p(\bm{f} \vert I)$ can be computed on a grid of epochs and durations, the transit depth being linearly solved for any $(T, D)$.
-In this equation.
 This \textit{linear search} leads to the set of likelihoods
@@
         \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
-        \caption{Principle and output of the \textit{linear search}.
+        \caption{Principle and output of the linear search.
 The simulated dataset (top) corresponds to the one shown and described in \autoref{fig:app_principle_dataset}.
@@
 $$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi, D)}_{i,j}$$
-shown for few periods in \autoref{fig:periodic_search} (b).
-In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the \textit{linear search} using $\phi_i$, $D_j$ and $T_0 = 0$.
+shown for few periods in \autoref{fig:periodic_search} (b). In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the linear search using $\phi_i$, $D_j$ and $T_0 = 0$.
 
@@
 
-    \caption{Applied to the dataset shown \autoref{fig:linear_search}, this figure shows how the periodic search works at different periods $P$ (including the true period $P = 0.7$ days).
+    \caption{Applied to the dataset shown in \autoref{fig:linear_search}, this figure shows how the periodic search works at different periods $P$ (including the true period $P = 0.7$ days).
 Given different periods $P$ and $T_0=0$, the likelihood $\ln\mathcal{L}$ shown in (a) is phase-folded and interpolated onto a common grid of phases shown in (b).
-As an example, the white lines in (a) mark the edges of each fold for a period of $P=0.5$ days, and the white crosses show the epochs $\set{T_k}_{k\in\mathbb{K}}$ matching a particular phase in the grid (reported in (a) and (b)) on which the corresponding $\set{\ln \mathcal{L}_k}_{k\in\mathbb{K}}$ are interpolated and combined using \autoref{eq:result}.
+As an example, the white lines in (a) mark the edges of each fold for a period $P=0.5$ days, and the white crosses show the epochs $\set{T_k}_{k\in\mathbb{K}}$ matching a particular phase in the grid (reported in (a) and (b)) on which the corresponding $\set{\ln \mathcal{L}_k}_{k\in\mathbb{K}}$ are interpolated and combined using \autoref{eq:result}.
 To allow the use of efficient matrix computations, this is done for all durations $\set{D_i}_i$ so that $\mathcal{P}$ is computed on the full grid $\set{D_i, \phi_i}_{i, j}$ at once.
@@
 But a final issue emerges, one that is fundamentally linked to our strategy.
-Each likelihood $p(\bm{f} \vert T, D, \Delta)$ estimated during the \textit{linear search} is computed using N measurements.
+Each likelihood $p(\bm{f} \vert T, D, \Delta)$ estimated during the linear search is computed using N measurements.
 Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving $N\times K$ measurements.
@@
 maximizing $\mathcal{Q}$, is adopted as the best candidate, basing the confidence in this signal through its SNR.
-The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (given by \autoref{eq:sigma_delta}).
+The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (given by \autoref{eq:result}).
 
@@
 \textsf{nuance} exploits the use of \textsf{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package powered by \textsf{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}, allowing for custom kernels to be built and highly tractable computations.
-We can then define a set of epochs \textsf{t0s} and durations \textsf{Ds} and run the \textit{linear search} with
+We can then define a set of epochs \textsf{t0s} and durations \textsf{Ds} and run the linear search with
 \begin{lstlisting}[language=Python,linewidth=\linewidth]
@@
 From this \texttt{search} object, the best transiting candidate parameters can be computed (\textsf{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\textsf{search.Q\_snr}), together with valuable information about the transit search.
-The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary hosts, the advantage of \nuancemethod{} being that the \textit{linear search} only needs to be performed ones and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
+The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary hosts, the advantage of \nuancemethod{} being that the linear search only needs to be performed ones and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
 An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.
@@
 By design, these injected signals have an SNR ranging from 5 to 30 (with $\sigma_r = 0$ in \autoref{eq:snr}).
-For each light curve, transits are searched using two different methods: \textsf{nuance}, using its implementation from the Python package described in the previous section; and \textsf{BLS}, the box-least-square algorithm from \cite{bls} (using \textsf{astropy}'s \textsf{BoxLeastSquares}\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html} implementation).\\\\
+For each light curve, transits are searched using two different tools: \textsf{nuance}, using its implementation from the Python package described in the previous section; and BLS, the Box-Least-Square algorithm from \cite{bls} (using \textsf{astropy}'s \textsf{BoxLeastSquares}\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html} implementation).\\\\
 For both methods, 3000 trial periods from 0.2 to 2.6 days are searched, with a single trial duration fixed to the unique known duration of 50 minutes.
@@
         \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
-        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using \textsf{BLS} and \nuancecode{}.
+        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using BLS and \nuancecode{}.
 The color scale indicates the recovery of transits in the corresponding (period, depth) parameter space, white for a full recovery and black for no detection.}
@@
 
-These results demonstrate the qualitative match between the detection capabilities of \textsf{nuance} and \texttt{BLS} on light curves with no correlated noise.
+These results demonstrate the qualitative match between the detection capabilities of \textsf{nuance} and BLS on light curves with no correlated noise.
 Explaining the subtle differences observed between the two methods when only white noise is present is beyond the scope of this paper, and we will assume that any differences observed in the following sections are due to the different treatments of correlated noise.
@@
 \newpage
-\section{Results}\label{results}
+\section{Performances}\label{results}
 \subsection{Comparison with \texttt{biweight+BLS} on simulated light curves}\label{simu}
@@
 Similarly to \autoref{detrending_effect}, an optimal bi-weight filter implemented in the \textsf{wõtan} Python package\footlink{https://github.com/hippke/wotan} is used with a window size three times that of the injected transit duration.
-A transit search on the detrended light curve is then performed using the \texttt{BLS} algorithm, like in \autoref{control} using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}; a strategy commonly used in the literature that we denote \texttt{biweight+BLS}.
+A transit search on the detrended light curve is then performed using the BLS algorithm, like in \autoref{control} using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}; a strategy commonly used in the literature that we denote \texttt{biweight+BLS}.
 In what follows, the transit detection criteria are the same as the ones used in \autoref{control}, i.e.
@@
 Each light curve consists in a 4 days observation with an exposure time of 2 minutes, leading to $N=2880$ data points with a normal error of $0.1\%$.
-As in \autoref{control}, only box-shaped transits are injected, using the model from \cite{protopapas} with $c=500$.
-\\\\
+As in \autoref{control}, only box-shaped transits are injected, using the model from \cite{protopapas} with $c=500$ (cf.\;\autoref{app_proto}).\\\\
 For a given pair of $(\tau, \delta)$, we simulate stellar variability using a GP with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params} (except for $Q$, described below), computed with respect to the injected transit parameters $D$ and $\Delta$.
@@
 This empirical statement only concerns light curves with a given amount of white noise, and may vary depending on the length of the observing window or the number of transits.
-For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperforms \wtls{} would only apply to this specific example.
+For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperforms \wtls{} would only apply to this specific but representative example.
+However, we verify that our conclusions remain qualitatively valid for various simulation setups.
 \begin{figure}[H]
@@
 \newpage
-\subsection{Comparison with various techniques on limited TESS light curves}\label{real}
+\subsection{Comparison with various techniques on rapidly-rotating M dwarfs TESS light curves}\label{real}
 
@@
 
-In order to assess the performance of \textsf{nuance} on real datasets, we inject and recover transits into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}).
-The goal of this transit injection-recovery is to compare \textsf{nuance} to other transit search strategies, and to understand the light curves characteristics for which it becomes beneficial.\\\\
+In order to assess the performance of \nuance{} on real datasets, we inject and recover transits into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}).
+The goal of this transit injection-recovery is to compare \nuance{} to other transit search strategies, and to understand the light curves characteristics for which it becomes beneficial.\\\\
 We focus this proof of concept on light curves from a list of 438 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{}.
 For each of the studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) of a single sector (the first being observed for each target) spanning on average 10 days.\\\\
-In this injection-recovery tests, \textsf{nuance} is compared to other transit search strategies employing the \texttt{BLS} algorithm, for each of them after detrending the light curves with a different technique:
+In this injection-recovery tests, \nuance{} is compared to other transit search strategies employing the BLS algorithm, for each of them after detrending the light curves with a different technique:
 \begin{itemize}
-    \item \texttt{bspline+BLS} employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the \texttt{BLS} algorithm.
-    \item \texttt{biweight+BLS} employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package \citep{wotan} with an optimal window size of $3\times D$ (i.e.\, three times the transit duration), followed by a search with the \texttt{BLS} algorithm.
+    \item \texttt{bspline+BLS} employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the BLS algorithm.
+    \item \texttt{biweight+BLS} employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package \citep{wotan} with an optimal window size of $3\times D$ (i.e.\, three times the transit duration), followed by a search with the BLS algorithm.
     \item \texttt{harmonics+BLS} employs a linear harmonic detrending, where the light curve is modeled as a Fourier series including four harmonics of the stellar rotation period found by \cite{Ramsay2020}, with coefficients found through ordinary least square.
-This detrending is followed by a search with the \texttt{BLS} algorithm.
+This detrending is followed by a search with the BLS algorithm.
     \item \texttt{iterative+BLS} iteratively detrends the light curve with a sinusoidal signal fitted to the data, each time using the dominant period of the residuals found using a lomb-scargle periodogram (5 iterations).
-This detrending is followed by a search with the \texttt{BLS} algorithm.\\\\
+This detrending is followed by a search with the BLS algorithm.\\\\
 \end{itemize}
@@
 
-As some of these techniques can be affected by gaps in the data, we only use continuous measurements from half a TESS sector.
-We assume that all methods (including nuance) are based on an incomplete model of the data that do not account for stellar flares.
+As some of the techniques compare to \nuance{} can be affected by gaps in the data, we only use continuous measurements from half a TESS sector.
+We assume that all methods (including \nuance{}) are based on an incomplete model of the data that do not account for stellar flares.
 For this reason, the light curve of each target is cleaned using an iterative sigma clipping approach.
@@
 This way, large flare signals and their expected ingress and egress are masked, using a total of 3 iterations.
+At each iteration, the GP kernel hyperparameters are re-optimized.
 As PDCSAP light curves often start with a ramp-like signal, the first 300 points (as well as the last 300 points) of each continuous observation are masked.
@@
 Top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
-On this bottom plot, the black line corresponds to the mean of one of the GP model (with hyperparameters optimized here on the cleaned light curve).}
+On this bottom plot, the black line corresponds to the mean of the GP model (with hyperparameters optimized here on the cleaned light curve).}
     \label{fig:cleaned}
 \end{figure}
-For each light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injections and recovery per target.
-Orbital periods $P$ are sampled on a regular grid between 0.4 and 5 days, and planetary radii $R_p$ are sampled on a regular grid designed to lead a minimum transit SNR of 4 and a maximum of 30.
+For each light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injection-recovery per target.
+Orbital periods $P$ are sampled on a regular grid between 0.4 and 5 days, and planetary radii $R_p$ are sampled on a regular grid designed to yield a minimum transit SNR of 2 and a maximum of 30.
 Using \autoref{eq:snr} with $\sigma_r = 0$, the planetary radius leading to a transit with a desired SNR $s$ is given by
@@
 \subsubsection*{Stellar variability kernel}\label{rotation_kernel}
-One of the particularity of nuance is to model a light curve as a GP.
+One of the particularity of \nuance{} is to model a light curve as a GP.
 To model these TESS light curves, we employ a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
@@
 For simplicity, and to adopt a uniform treatment for all target light curves, a design matrix $\bm{X}$ with a single constant column is adopted, such that the systematic model only consists in a single parameter corresponding to the mean value of the differential flux (expected to be close to 1) solved linearly.
-The initial value for the rotation period of the star is found using a lomb-scargle periodogram \citep{Lomb1976, Scargle1982} implemented in \textsf{astropy}\footlink{https://docs.astropy.org/en/stable/timeseries/lombscargle.html}.
+Our motivations to choose this very simplistic baseline, despite the capability of \nuancemethod{} to account for more complex linear models, is discussed in \autoref{systematics}.
 
 \subsubsection*{Search parameters and transit detection criteria}
-For all techniques, we only search for transits with a duration fixed to the duration of the injected transits.
-This is mainly done for computational efficiency and allows for a narrower comparison between \texttt{nuance} and all \texttt{BLS}-based techniques.
+For all techniques, we only search for transits with a duration fixed to the known duration of the injected transits.
+This is mainly done for computational efficiency and allows for a narrower comparison between \nuancemethod{} and all BLS-based techniques.
 Finally, the search is done over 4000 trial periods linearly sampled from 0.3 to 6 days.
-Realistic transit searches on a wider parameter space (e.g.
-multiple trial durations) are discussed in \autoref{}.\\\\
+A realistic transit search on a wider parameter space (e.g.
+multiple trial durations) is discussed in the next section.\\\\
 Like in previous sections, we consider the detection of period aliases and ignore the match between the injected and recovered transit epochs (although we visually vet that the found epochs were consistent with the ones injected for cases where the planet was detected).
 We consider a transit detectable if its original SNR is greater than 6.
-Hence we define \textit{true positives} has detectable transits recovered with the correct period (or an alias) and a measured SNR greater or equal to 6, and \textit{false positives} as non-detectable transits recovered with a measured SNR greater or equal to 6.
+Hence we define \textit{true positives} has detectable transits recovered with the correct period (or an alias) and a measured SNR greater or equal to 6, and \textit{false positives} as non-detectable transits recovered with a measured SNR greater or equal to 6.\\\\As we noticed that few methods were still affected by the remaining stellar variability after detrending, we remove from the grid of trial periods all periods close to the stellar rotation period and its aliases.
+In practice this is done by removing all periods $P$ such that $dP = \frac{P}{P_*}$ is less than 2\% from an integer value, i.e.
+$\vert  dP - \lceil dP \rceil \vert < 0.02$.
 
@@
 An example of the transit injection-recovery and its result is shown in \autoref{1019692} for the target TIC 1019692.
-\autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in).
-Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (10950 with the non-dectable ones), and the lowest number of false negatives (\autoref{fig:allsearch}).
+\autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in \autoref{eq:relative_params}).
+Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits and the lowest number of false negatives (\autoref{fig:allsearch}).
 \begin{figure}[H]
@@
 
-In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b.
-We download the 2 min cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
+In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b \citep{TOI540}.
+We download the 2 minutes cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
 Like in the previous section, we use a Lomb-Scargle periodogram and identify the 0.72 days rotation period of the star, that we use as an initial value to optimize the kernel described in \autoref{rotation_kernel} on each sector independently.
 Here again, we employ an upper-sigma-clipping to mask flares out of the data.
-The resulting light curve for sector 4 and its mean model is shown in \autoref{fig:toi540_clean}.
+The resulting light curve for sector 4 and its mean model are shown in \autoref{fig:toi540_clean}.
 \begin{figure}[H]
@@
         \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/4.pdf}
-        \caption{Sector 4 light curve of TOI 540. The clean signal (gray points) has been mask masked for flares (black points), and the black line corresponds to the mean of the GP model.}
+        \caption{Sector 4 light curve of TOI 540. The cleaned signal (gray points) has been masked for flares (black points), and the black line corresponds to the mean of the GP model.}
         \label{fig:toi540_clean}
@@
 
-For each sector, we perform the \textit{linear search} of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
+For each sector, we perform the linear search of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
 We then perform the periodic search on all sectors combined, using a concatenation of all linear searches.
-By adopting this by-sector GP modeling of the light-curve, the \textit{linear search} of \nuance{} scales linearly with the number of sectors being processed.
+By adopting this by-sector GP modeling of the light-curve, the linear search of \nuance{} scales linearly with the number of sectors being processed.
 This approach is adopted for efficiency but also to encapsulate the changing properties of stellar variability from one sector to another, often separated by year-long gaps.
 The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days.
-This search, using \nuance{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the \texttt{BLS} algorithm (denoted \texttt{biweight+BLS} and described in \autoref{real}) on all sectors combined.
-Since we don't know the transit duration a-priori, we perform the detrending and search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
+This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{biweight+BLS} and described in \autoref{real}) on all sectors combined.
+Since we don't know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
 The results of this comparison are shown in \autoref{fig:toi540_periodograms}.
@@
 \end{figure}
-After a first periodic search, the times falling in the detected periodic transit signal (the one with an SNR higher than the detection limit set to 7) are masked.
-In practice, this is done by masking the linear search products \\\\
+After a first periodic search, trial epochs in windows of widths $2\times D$ centered on the detected periodic transits are masked.
+In practice, this is done by masking the linear search products $\set{\ln\mathcal{L}_{i,j}}_{i, j}$, $\set{\Delta_{i,j}}_{i, j}$ and $\set{\sigma_{i,j}}_{i, j}$ (defined in \autoref{linear_search}).\\\\
 As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{biweight+BLS} and \texttt{nuance} are very similar, with the known transiting exoplanet TOI-540 b detected with an orbital period P = 1.24 days.
-This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b (using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ min of TOI-540 b \citep{TOI540}).
-Hence, as $\tau < 7$, the use of \nuance{} on TOI-540 is not expected to be very beneficial (see \autoref{fig:snr_detrend}).
+This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b \footnote{using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ min of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial beneficial (see \autoref{fig:snr_detrend}).
 Nonetheless, the \nuance{} periodogram of TOI 540 features less spurious SNR peaks, largely due to the penalty naturally occurring when single transits with different depths are periodically combined.
@@
 We note that finding a TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques (see e.g.\;\autoref{fig:simu}).
-Nontheless, we verify that \nuancemethod is capable of finding a large number of already known transiting exoplanet candidates, in light curves featuring various forms of correlated noises, at least as efficiently as with commonly used techniques.
+Nontheless, we verify that \nuancemethod{} is capable of finding a large number of already known transiting exoplanet candidates, in light curves featuring various forms of correlated noises, at least as efficiently as with commonly used techniques.
 We reserve the search for new transit signals to a follow-up publication.
@@
 \noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple non-optimized GP with a squared exponential kernel.
-These are compared to the processing time of \texttt{biweight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the \textsf{BLS} search.
+These are compared to the processing time of \texttt{biweight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the BLS search.
 \begin{figure}[H]
@@
 The gray curve shows the performance of \textsf{nuance} \textit{linear search} when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together.
-This study is performed on both a single core of an Apple M2 Max chip (red line) and all of its 12 cores (black line).
-While not being shown, we verify that both the \textsf{BLS} algorithm and \nuancemethod{}processing times scale linearly with the number of probed transit durations and orbital periods.}
+This study is performed on the 12 cores of an Apple M2 Max chip, exploiting the parallelization offered by our implementation.
+While not being shown, we verify that both the BLS algorithm and \nuancemethod{} processing times scale linearly with the number of probed transit durations and orbital periods.}
         \label{fig:blsvsnuance}
@@
 
-As see in \autoref{fig:blsvsnuance}, most of the computational cost of \nuancecode{} and the \texttt{biweight+BLS} method comes from the \textit{linear search} and the bi-weight filtering step.
-One advantage of \nuancemethod{} is that the \textit{linear search} can be performed separately on different continuous observations, and then combined in the periodic search.
+As see in \autoref{fig:blsvsnuance}, most of the computational costs of \nuancecode{} and the \texttt{biweight+BLS} method in this particular example come from the linear search and the bi-weight filtering step.
+This is not always true and depends on the size of the trial durations and periods grids.
+One advantage of \nuancemethod{} is that the linear search can be performed separately on different continuous observations, and then combined in the periodic search.
 Hence, if searching for transits in separate observations with approximately similar durations, such as different TESS sectors or different ground-based nightly observations, the computational cost of \nuancecode{} grows linearly with the number of observations (see grey line in \autoref{fig:blsvsnuance}).
-Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform about one order of magnitude slower than more conventional techniques associated with \textsf{BLS}}.
+Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform about one order of magnitude slower than more conventional techniques associated with BLS}.
 The second advantage of \nuancecode{} is its \textsf{JAX} implementation, allowing the linear search to be run in parallel.
@@
 But as these strongly depend on the type of detrending technique employed, we do not provide general guidelines as when \nuancemethod{} should be preferred over a specific detrending technique.
-Nevertheless, extensive benchmarks and guidelines are reserved to future developments and will be progressively shared on \nuancecode{}'s online documentation\footlink{https://nuance.readthedocs.io/en/latest/}.
+To aid users in making an informed choice of algorithm, extensive benchmarks and guidelines are reserved to future developments and will be progressively shared on \nuancecode{}'s online documentation\footlink{https://nuance.readthedocs.io/en/latest/}.
 
-\subsection{Systematics modeling}
+\subsection{Systematics modeling}\label{systematics}
 Throughout the paper, a single column design matrix $\bm{X}$, corresponding to the mean differential flux (ideally unitary), was employed, hence assuming that the instrumental systematics signals were non-existent.
@@
 To account for stellar variability while fitting the linear systematics model to the data, a step further would be to use a GP, such as done in the EVEREST \citep{everest2} pipeline.
-However this would also involve some potential degradation of the transit signals (see e.g.\;\autoref{fig:issue2}), with a hardly distinguishable origin.
-For these reasons, and to keep our comparisons as targeted as possible, we do not compare commonly used systematics detrending approaches and decided to focus our study on the comparison with stellar variability detrending techniques (although these two aspects often overlap in the literature).\\\\
+However this would also involve some potential degradation of the transit signals (see e.g.\;\autoref{fig:issue2}), with an hardly distinguishable origin.
+For these reasons, and to keep our comparisons as targeted as possible, we do not compare commonly used systematics detrending approaches and decided to focus our comparisons solely on stellar variability detrending techniques (although these two aspects often overlap in the literature).\\\\
 Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsly is extremely promising for the search of transiting exoplanets in ground-based data, that usually suffer from daily interruptions.
@@
 An alternative, recommended for more realistic blind searches, is to perform model comparison on well selected kernels, and to adapt the optimization strategy to each dataset.\\\\
-When using \nuancecode{} on TESS light curves for example, it must be noted that the observed light curve variability might originate from a contaminated photometric aperture, so that a physically-interpretable GP kernel representing a single star activity is not necessarily optimal.
+When using \nuancecode{} on TESS light curves for example, it must be noted that the observed light curve variability might originate from a contaminated photometric aperture, so that a physically-interpretable GP kernel representing a single star activity is not necessarily appropriate.
+On the other hand, a single squared exponential GP kernel might also be sufficient for some applications.
+
+\subsection{Prospects}
+The present implementation of \nuancemethod{} has the potential to be extended to be used beyond the search of periodic transits.
+Here are ideas of possible use and extensions, from the most straightforward to the most ambitious:
+\begin{enumerate}
+    \item In order to compare \nuancecode{} to BLS-based methods, we injected and retrieved only box-shaped transits.
+However, similarly to the Transit-Least-Square algorithm from \cite{tls}, \textbf{limb-darkened transits can serve as base model} in the linear search and are expected to improve the transit search in the same way TLS provided an improvement over BLS.
+    \item The linear search of \nuancemethod{} is a single-event detection algorithm that can be used to search for single transit events, but also \textbf{detect transiting exocomets and flares}, by simply adapting the base astrophysical model in the last colum of the design matrix $\bf{X}$.
+While not being tested and benchmarked, the \nuancecode{} implementation already integrates this feature (make a tutorial!).
+    \item During the periodic search, no prior about the transit duration related to the orbital period of the planet was used.
+This was done in order to allow the detection of grazing transiting exoplanets that would produce shorter timescale transits compared to what is expected from a circular orbit.
+However, adding such prior, \textbf{using the distribution of known exoplanets eccentricities}, might produce less spurious periodogram peaks and be very beneficial for the automatic search of transits in large datasets.
+Another idea, similar to the one employed in \cite{foreman2016}, is to leverage \textbf{model comparison in order to reject single transits that are better described by the GP model alone}.
+Both ideas come at no cost given our modeling approach.
+    \item Finally, the formalism of \nuancemethod{} could be adapted and used to \textbf{search for exoplanets featuring transit time variations (TTV)}.
+Indeed, this application only requires a modification of the periodic search, as maximum likelihood peaks close to linearly predicted transit epochs may be considered.
+This could be done either with a special nearest-neighbor algorithm or with a convolution of the computed likelihood grid with a Gaussian kernel.
+To maximize efficiency and interpretability, we would recommend these approaches to be explored analytically, rather than using a data-driven treatment of the linear search products.
+\end{enumerate}
 
 \section{Conclusion}
-
+This paper presents \nuancemethod{}, an algorithm designed to detect planetary transits in light curves featuring correlated noise in the form of instrumental signals and stellar variability.
+In this context, a conventional approach involves detrending a light curve before searching for transits using a Box-Least-Square algorithm.
+However, we show that the incomplete modeling of a light curve for its detrending, one that ignores the presence of transits, degrades their signal-to-noise-ratio down to the point of not being detectable.
+Adopting a widely used detrending stratgey, we explore the extend of this degradation on simulated light curves, and its dependance on the stellar variability characteristics.\\\\
+The effectiveness of \nuancemethod{} is tested using a synthetic dataset and further validated on real TESS light curves of 438 M dwarfs.
+These injection-recovery tests reveal that \nuancemethod{} consistently outperforms conventional transit search techniques, especially for transits with duration exceeding one-fifth of the stellar variability timescale.
+In these cases, \nuancemethod{} not only identifies a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
+To make \nuancemethod{} accessible for wider use, we make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package.
+This package leverages the power of JAX, allowing for efficient computations and compatibility with distributed computing environments and GPU devices.\\\\
+We acknowledge the limitations of \nuancemethod{} and its increased computational cost.
+In this regard, \nuancemethod{} should be used as an alternative to more traditional techniques only in the presence of substantial correlated noise.
+As guidelines for choosing \nuancemethod{} over other techniques are highly dependant on the type of detrending algorithm employed, we reserve this study to a future work.\\\\
+Finally, we suggest future improvements and extensions of the algorithm, including its application for detecting single transits, exocomets, flares, and transit time variations (TTV), underscoring its versatility and potential for broader impact in astronomical research.
+\\\\
+\vfill{}
 We would like to thank Julien De Wit, Prajwal Niraula and Francisco J.
-Pozuelos for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
+Pozuelos for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.x
 

### 2023-12-06 latex/paper.tex
feat: exocomets

@@
 In \autoref{nuance}, we present \nuance{}, and the two main steps on which this method is based: the linear search and the periodic search.
-In \autoref{results}, we test the performance of \nuance{} on a wide variety of cases, and compare it to commonly used transit search algorithms.
+In \autoref{results}, we test the performance of \nuance{} on a wide variety of cases, and compare our method to commonly used transit search algorithms.
 This include transits injected in synthetic datasets, but also in the TESS light curves of 438 rapidly-rotating M dwarfs.
@@
 
-\subsection{The choice of kernel}
+\subsection{The choice of kernel and model}
 While not being discussed in our study, \textbf{the efficiency of \nuancemethod{} to detect transits in correlated noise is highly dependant on the design of its GP kernel}.
@@
 When using \nuancecode{} on TESS light curves for example, it must be noted that the observed light curve variability might originate from a contaminated photometric aperture, so that a physically-interpretable GP kernel representing a single star activity is not necessarily appropriate.
-On the other hand, a single squared exponential GP kernel might also be sufficient for some applications.
+On the other hand, a single squared exponential GP kernel might also be sufficient for some applications.\\\\
+Something to emphasize is that \nuancemethod{} cannot be used to produce reliable detrended light curves.
+Most of the time, the GP is flexible enough to partially model transits and can never be used to model the data without them.
+The idea behind \nuancemethod{} is to compute the likelihood of data against a model containing both transits and correlated noise, without ever trying to disentangle both signals.
+In practice, it means that it can be very hard to actually verify the presence of transits found by \nuancemethod{} visually, so that transits may be detected but remain hidden in correlated noise.
+This is particularly true for stars displaying very high-frequency photometric pulsations.
 
@@
     \item The linear search of \nuancemethod{} is a single-event detection algorithm that can be used to search for single transit events, but also \textbf{detect transiting exocomets and flares}, by simply adapting the base astrophysical model in the last colum of the design matrix $\bf{X}$.
-While not being tested and benchmarked, the \nuancecode{} implementation already integrates this feature (make a tutorial!).
+While not being tested and benchmarked, the \nuancecode{} implementation already integrates this feature and we show an example of a search of the known exocomets of Beta Pictoris in \autoref{fig:exocomets}.
     \item During the periodic search, no prior about the transit duration related to the orbital period of the planet was used.
-This was done in order to allow the detection of grazing transiting exoplanets that would produce shorter timescale transits compared to what is expected from a circular orbit.
-However, adding such prior, \textbf{using the distribution of known exoplanets eccentricities}, might produce less spurious periodogram peaks and be very beneficial for the automatic search of transits in large datasets.
+This was done in order to allow the detection of grazing transiting exoplanets that would produce shorter timescale transits compared to what is expected from a circular orbit with a null impact parameter.
+However, adding such priors might produce less spurious periodogram peaks and be very beneficial for the automatic search of transits in large datasets.
 Another idea, similar to the one employed in \cite{foreman2016}, is to leverage \textbf{model comparison in order to reject single transits that are better described by the GP model alone}.
@@
 
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{exocomet.pdf}
+        \caption{Demonstration of the search of transiting exocomets in the TESS light curves of the star $\beta$ Pictoris (2 minutes PDCSAP data from sector 33).
+This star is known to display rapid $\delta$ Scuti type photometric variations with a period of about 30 minutes \citep{Lecavelier2022}.
+Here, we simply used the linear search of \nuancecode{} with a different base model, one that mimics the shape of transiting exocomets, to compute the SNR time-series of the signal over one sector.
+The maximum-SNR events are displayed in the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.
+The GP kernel choice and its hyperparameters optimization are similar to the one presented in \autoref{real}.}
+        \end{centering}
+        \label{fig:exocomets} 
+\end{figure}
+
 \section{Conclusion}
@@
 However, we show that the incomplete modeling of a light curve for its detrending, one that ignores the presence of transits, degrades their signal-to-noise-ratio down to the point of not being detectable.
-Adopting a widely used detrending stratgey, we explore the extend of this degradation on simulated light curves, and its dependance on the stellar variability characteristics.\\\\
-The effectiveness of \nuancemethod{} is tested using a synthetic dataset and further validated on real TESS light curves of 438 M dwarfs.
+Adopting a widely used detrending stratgey, we explore the extend of this degradation on simulated light curves, and its dependance on the stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
+The effectiveness of \nuancemethod{} is tested using a synthetic dataset and further validated on real TESS light curves of 438 rapidly-rotating M dwarfs.
 These injection-recovery tests reveal that \nuancemethod{} consistently outperforms conventional transit search techniques, especially for transits with duration exceeding one-fifth of the stellar variability timescale.
 In these cases, \nuancemethod{} not only identifies a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
-To make \nuancemethod{} accessible for wider use, we make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package.
-This package leverages the power of JAX, allowing for efficient computations and compatibility with distributed computing environments and GPU devices.\\\\
-We acknowledge the limitations of \nuancemethod{} and its increased computational cost.
+We make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package.
+This package leverages the power of JAX, allowing for efficient computations and compatibility with distributed computing environments and GPU devices.
+Nonetheless we acknowledge the limitations of \nuancemethod{} and its increased computational cost compared to more tr.
 In this regard, \nuancemethod{} should be used as an alternative to more traditional techniques only in the presence of substantial correlated noise.

### 2023-12-06 latex/paper.tex
chore: redo plot and new intermediate results benchmark

@@
 \author{Catriona A. Murray}
+\author{Francisco J. Pozuelos}
 \correspondingauthor{Lionel J. Garcia}
@@
 when both signals start resembling each other, we show that our algorithm is the most performant in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
-Although simultaneously searching for transits while modeling correlated noise is expected to be computationaly expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices (?).
+Although simultaneously searching for transits while modeling correlated noise is expected to be computationaly expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
 Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of active stars and their transiting exoplanets, both using space-based and sparse ground-based observations.
@@
 As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{biweight+BLS} and \texttt{nuance} are very similar, with the known transiting exoplanet TOI-540 b detected with an orbital period P = 1.24 days.
-This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b \footnote{using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ min of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial beneficial (see \autoref{fig:snr_detrend}).
+This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b \footnote{using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ minutes of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial beneficial (see \autoref{fig:snr_detrend}).
 Nonetheless, the \nuance{} periodogram of TOI 540 features less spurious SNR peaks, largely due to the penalty naturally occurring when single transits with different depths are periodically combined.
@@
 
-\subsection{The choice of kernel and model}
+\subsection{The GP kernel}
 While not being discussed in our study, \textbf{the efficiency of \nuancemethod{} to detect transits in correlated noise is highly dependant on the design of its GP kernel}.
@@
 On the other hand, a single squared exponential GP kernel might also be sufficient for some applications.\\\\
-Something to emphasize is that \nuancemethod{} cannot be used to produce reliable detrended light curves.
-Most of the time, the GP is flexible enough to partially model transits and can never be used to model the data without them.
-The idea behind \nuancemethod{} is to compute the likelihood of data against a model containing both transits and correlated noise, without ever trying to disentangle both signals.
+Something to emphasize is that \nuancemethod{} cannot be used to produce reliable detrended light curves, as an optimal GP is often flexible enough to partially model transits (see \autoref{fig:issue2}).
+In contrast, the idea behind \nuancemethod{} is rather to compute the likelihood of data against a model containing both transits and correlated noise, without ever trying to disentangle both signals.
 In practice, it means that it can be very hard to actually verify the presence of transits found by \nuancemethod{} visually, so that transits may be detected but remain hidden in correlated noise.
-This is particularly true for stars displaying very high-frequency photometric pulsations.
+This is particularly true for stars displaying very high-frequency photometric pulsations (see the example in \autoref{fig:exocomets} where the detrended light curves dismissively follow the base model).
 
 \subsection{Prospects}
-The present implementation of \nuancemethod{} has the potential to be extended to be used beyond the search of periodic transits.
+The present implementation of \nuancemethod{} has the potential to be extended to be used beyond the search of periodic box-shaped transits.
 Here are ideas of possible use and extensions, from the most straightforward to the most ambitious:
@@
 However, similarly to the Transit-Least-Square algorithm from \cite{tls}, \textbf{limb-darkened transits can serve as base model} in the linear search and are expected to improve the transit search in the same way TLS provided an improvement over BLS.
-    \item The linear search of \nuancemethod{} is a single-event detection algorithm that can be used to search for single transit events, but also \textbf{detect transiting exocomets and flares}, by simply adapting the base astrophysical model in the last colum of the design matrix $\bf{X}$.
-While not being tested and benchmarked, the \nuancecode{} implementation already integrates this feature and we show an example of a search of the known exocomets of Beta Pictoris in \autoref{fig:exocomets}.
+    \item The linear search of \nuancemethod{} is a single-event detection algorithm that can be used to search for single transit events, but also \textbf{detect transiting exocomets and flares}, by simply changing the base astrophysical model in the last colum of the design matrix $\bf{X}$.
+While not being tested and benchmarked, \nuancecode{} already integrates this feature.
+An example of the detection of known transiting exocomets in $\beta$ Pictoris TESS light curves is shown in \autoref{fig:exocomets}.
     \item During the periodic search, no prior about the transit duration related to the orbital period of the planet was used.
@@
 However, adding such priors might produce less spurious periodogram peaks and be very beneficial for the automatic search of transits in large datasets.
-Another idea, similar to the one employed in \cite{foreman2016}, is to leverage \textbf{model comparison in order to reject single transits that are better described by the GP model alone}.
+Another idea, similar to the one employed in \cite{foreman2016}, is to leverage \textbf{model comparison in order to reject transits that are better described by the GP model alone}.
 Both ideas come at no cost given our modeling approach.
@@
 This star is known to display rapid $\delta$ Scuti type photometric variations with a period of about 30 minutes \citep{Lecavelier2022}.
-Here, we simply used the linear search of \nuancecode{} with a different base model, one that mimics the shape of transiting exocomets, to compute the SNR time-series of the signal over one sector.
-The maximum-SNR events are displayed in the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.
-The GP kernel choice and its hyperparameters optimization are similar to the one presented in \autoref{real}.}
+The GP kernel and its hyperparameters are chosen and optimized as in \autoref{real}.Here, we simply used the linear search of \nuancecode{} with a different base model, one that mimics the shape of transiting exocomets (with 20 trial durations), to compute the SNR time-series of the signal over one sector.
+The maximum-SNR events are displayed in the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.}
         \end{centering}
@@
 However, we show that the incomplete modeling of a light curve for its detrending, one that ignores the presence of transits, degrades their signal-to-noise-ratio down to the point of not being detectable.
-Adopting a widely used detrending stratgey, we explore the extend of this degradation on simulated light curves, and its dependance on the stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
+Adopting a widely used detrending stratgey, we explore the extent of this degradation on simulated light curves, and its dependance on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
 The effectiveness of \nuancemethod{} is tested using a synthetic dataset and further validated on real TESS light curves of 438 rapidly-rotating M dwarfs.
-These injection-recovery tests reveal that \nuancemethod{} consistently outperforms conventional transit search techniques, especially for transits with duration exceeding one-fifth of the stellar variability timescale.
-In these cases, \nuancemethod{} not only identifies a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
-We make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package.
-This package leverages the power of JAX, allowing for efficient computations and compatibility with distributed computing environments and GPU devices.
-Nonetheless we acknowledge the limitations of \nuancemethod{} and its increased computational cost compared to more tr.
-In this regard, \nuancemethod{} should be used as an alternative to more traditional techniques only in the presence of substantial correlated noise.
-As guidelines for choosing \nuancemethod{} over other techniques are highly dependant on the type of detrending algorithm employed, we reserve this study to a future work.\\\\
-Finally, we suggest future improvements and extensions of the algorithm, including its application for detecting single transits, exocomets, flares, and transit time variations (TTV), underscoring its versatility and potential for broader impact in astronomical research.
+These injection-recovery tests reveal that \nuancemethod{} consistently outperforms conventional transit search techniques, especially for transits with a duration exceeding one-fifth of the stellar variability timescale.
+In these cases, \nuancemethod{} not only lead to a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
+We make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package, developped with \textsf{JAX} to allow its  use on distributed computing environments and GPU devices.
+Overall, we acknowledge the limitations of \nuancemethod{} and its increased computational cost compared to more conventional techniques.
+Hence, \nuancemethod{} should be used as an alternative to more traditional techniques only in the presence of substantial correlated noise.
+As guidelines for choosing our method over other techniques are highly dependant on the type of detrending algorithm employed, we reserve this study to a future work.\\\\
+Finally, we suggest future improvements and extensions of the algorithm, including its application for detecting single transits, exocomets, flares, and exoplanets featuring transit time variations (TTV), underscoring its versatility and potential for broader impact in astronomical research.
 \\\\
 \vfill{}
-We would like to thank Julien De Wit, Prajwal Niraula and Francisco J.
-Pozuelos for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.x
+We would like to thank Julien De Wit, Prajwal Niraula and Jon Jenkins for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
 

### 2023-12-06 workflows/benchmark/scripts/results.py

@@ -1,0 +1,17 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import yaml
+
+nuance = np.array([yaml.safe_load(open(f)) for f in snakemake.input.nuance])
+bls = np.array([yaml.safe_load(open(f))for f in snakemake.input.bls])
+
+result = {
+    "points": np.array(snakemake.params.points).tolist(),
+    "nuance_linear": [n["linear"] for n in nuance],
+    "nuance_periodic": [b["biweight"] for b in bls],
+    "biweight": [n["all"] - n["linear"] for n in nuance],
+    "bls": [b["bls"] - b["biweight"] for b in bls]
+}
+
+with open(snakemake.output[0], "w") as f:
+    yaml.dump(result, f)

### 2023-12-06 latex/paper.tex
chore: new plots  + typos

@@
 
-The hyperparameters of this kernel are optimized on trimmed and cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize}\footlink{https://jax.readthedocs.io/en/latest/_autosummary/jax.scipy.optimize.minimize.html} wrapper provided by the \textsf{jaxopt} Python package, and taking advantage of the \textsf{JAX} implementation of \textsf{tinygp}.
+The hyperparameters of this kernel are optimized on trimmed and cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize}\footlink{https://jax.readthedocs.io/en/latest/_autosummary/jax.scipy.optimize.minimize.html} wrapper provided by the \textsf{jaxopt} Python package, and taking advantage of the \textsf{JAX} implementation of \textsf{tinygp} and its quasiseparable kernels\footlink{https://tinygp.readthedocs.io/en/latest/api/kernels.quasisep.html} \citep{celerite}.
 As correlated noise is expected to affect the light curve uncertainty estimates performed by SPOC, the diagonal of the full covariance matrix of the data (i.e.\,their uncertainty, assuming homoscedasticity) is held free, increasing the number of optimized parameters to 9.
@@
 \autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in \autoref{eq:relative_params}).
-Compared to other methods, we find that \texttt{nuance} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits and the lowest number of false negatives (\autoref{fig:allsearch}).
+These figures are a synthesis of \autoref{fig:allsearchim} that shows the rate of true and false positives binned in the full parameter space $(\tau, \delta)$.
 \begin{figure}[H]
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/true_false_positives.pdf}
+        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_bars.pdf}
+        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals.}
+        \label{fig:allsearch}
+    \end{centering}
+\end{figure}
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp.pdf}
         \caption{Rate of true and false positives of \nuance{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
@@
 \end{figure}
+Compared to other techniques, we find that \nuancecode{} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits (\autoref{fig:allsearch}).
+While the performances of other methods highly depend on the characteristics of the variability, \textbf{\nuancecode{} is the best technique in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positives} (\autoref{fig:allsearch}).
 \begin{figure}[H]
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/true_and_false_positives_bars.pdf}
-        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals.}
-        \label{fig:allsearch}
+        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_ims.pdf}
+        \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods.}
+        \label{fig:allsearchim}
     \end{centering}
 \end{figure}
-While the performances of other methods highly depend on the characteristics of the variability, \nuance{} leads to the highest number of true positives in 79\% of cases, and the lowest number of false positives in 72\% of cases (\autoref{fig:allsearch_relative}).
-When considering both true and false positives simultaneously, \nuance{} can be considered the best technique in only 55\% of cases.
-However, \textbf{when considering transits with a relative timescale $\tau < 5$ (transits that justified the development of this algorithm), \nuancemethod{} is the best technique in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positives}.
 
 \subsection{Comparison for a multi-sector TESS candidate: TOI-540}\label{toi540}
-
 In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b \citep{TOI540}.
@@
 Here again, we employ an upper-sigma-clipping to mask flares out of the data.
-The resulting light curve for sector 4 and its mean model are shown in \autoref{fig:toi540_clean}.
+The resulting light curve for sector 4 and its mean model are shown in \autoref{fig:toi540_clean}.\\\\
+For each sector, we perform the linear search of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
+We then perform the periodic search on all sectors combined, using a concatenation of all linear searches.
+By adopting this by-sector GP modeling of the light-curve, the linear search of \nuance{} scales linearly with the number of sectors being processed.
 \begin{figure}[H]
@@
 \end{figure}
-
-For each sector, we perform the linear search of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
-We then perform the periodic search on all sectors combined, using a concatenation of all linear searches.
-By adopting this by-sector GP modeling of the light-curve, the linear search of \nuance{} scales linearly with the number of sectors being processed.
-This approach is adopted for efficiency but also to encapsulate the changing properties of stellar variability from one sector to another, often separated by year-long gaps.
-The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days.
-This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{biweight+BLS} and described in \autoref{real}) on all sectors combined.
-Since we don't know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
-The results of this comparison are shown in \autoref{fig:toi540_periodograms}.
 \begin{figure}[H]
@@
 \end{figure}
+This approach is adopted for efficiency but also to encapsulate the changing properties of stellar variability from one sector to another, often separated by year-long gaps.
+The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days.
+This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{biweight+BLS} and described in \autoref{real}) on all sectors combined.
+Since we don't know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
+The results of this comparison are shown in \autoref{fig:toi540_periodograms}.\\\\
 After a first periodic search, trial epochs in windows of widths $2\times D$ centered on the detected periodic transits are masked.
@@
 
+\newpage
 \section{Discussion}\label{discussion}
@@
 
+\newpage
 \section{Conclusion}
@@
 However, we show that the incomplete modeling of a light curve for its detrending, one that ignores the presence of transits, degrades their signal-to-noise-ratio down to the point of not being detectable.
-Adopting a widely used detrending stratgey, we explore the extent of this degradation on simulated light curves, and its dependance on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
+Adopting a widely used detrending strategy, we explore the extent of this degradation on simulated light curves, and its dependance on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
 The effectiveness of \nuancemethod{} is tested using a synthetic dataset and further validated on real TESS light curves of 438 rapidly-rotating M dwarfs.
 These injection-recovery tests reveal that \nuancemethod{} consistently outperforms conventional transit search techniques, especially for transits with a duration exceeding one-fifth of the stellar variability timescale.
-In these cases, \nuancemethod{} not only lead to a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
+In all cases, \nuancemethod{} not only lead to a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
 We make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package, developped with \textsf{JAX} to allow its  use on distributed computing environments and GPU devices.
@@
 \vfill{}
-We would like to thank Julien De Wit, Prajwal Niraula and Jon Jenkins for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
+We would like to thank Julien De Wit, Prajwal Niraula for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
 

### 2024-01-25 latex/paper.tex
fix: update paper

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\let\subsectionautorefname\sectionautorefname
+\let\subsubsectionautorefname\sectionautorefname
 
@@
 
-\title{\texttt{nuance}: Detection of planetary transits in the presence of correlated noise}
+\title{\texttt{nuance}: Efficient detection of planets transiting active stars}
 
@@
     We present \nuancemethod{}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
-To deal with them, a common approach consists in cleaning a light curve from nuisance signals before searching for transits.
-However, we show that this approach, based on the prior assumption that transits are not present, strongly degrades their signal-to-noise-ratio down to no detection.
-As this degradation depends on the correlated noise characteristics, we explore the relative parameter space for which transits are altered, and quantify this effect on a set of simulated light curves.
-On this synthetic dataset, we show that \nuancemethod{} largely outperforms the detection capabilities of a commonly used transit search technique, which consists in cleaning light curves from styellar variability with an optimal bi-weight filter before searching for transits using a box-least-square algorithm.
+To deal with this noise, a common approach consists in cleaning a light curve from nuisance signals before searching for transits.
+However, we show that this approach limit the detection efficiency of traditional techniques.
+As this effect depends on the correlated noise characteristics, we map the relevant parameter space for which transits are altered, and quantify the detection efficiency limits on a set of simulated light curves.
+On this synthetic dataset, we show that \nuancemethod{} largely outperforms the detection capabilities of a commonly used transit search technique, which consists in cleaning light curves from stellar variability with an optimal bi-weight filter before searching for transits using a box-least-square algorithm.
 Then, in order to assess the performance of \nuancemethod{} on real datasets, we inject and recover transits into the TESS light curves of 438 M dwarfs, and compare our method to 4 detrending techniques followed by a Box-Least-Square search.
-For transits with a duration exceeding one fifth of the stellar variability timescale, i.e.
-when both signals start resembling each other, we show that our algorithm is the most performant in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
-Although simultaneously searching for transits while modeling correlated noise is expected to be computationaly expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
+For transits with a duration exceeding one fifth of the stellar variability timescale, we show that \nuancemethod is the most performant in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
+Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
 Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of active stars and their transiting exoplanets, both using space-based and sparse ground-based observations.
@@
 For this reason, known transiting exoplanets tend to be found around quieter stars, or belong to the population of close-in giants whose transit signals dominate over stellar rotational variability \citep{Simpson2023}.
-However, discovering transiting exoplanets around active stars holds few promises.
+However, transiting exoplanets around active stars are discoveries with significant scientific value.
 First, as younger stars are more active \citep{Skumanich1972}, being able to detect planets transiting active stars will favor the discovery of young planetary systems \citep[e.g.][]{Newton2022}.
-Second, as stellar variability may originate from surface active regions (such as starpots), the increased discovery of occulting companions will increase our chances to map the photosphere of active stars \citep[e.g.][]{Morris2017}, benefiting both the study of stellar atmospheres and the concerning impact of their non-uniformity on planetary atmosphere retrievals \citep{rackham2018}.
+Second, as stellar variability may originate from surface active regions (such as starpots), transiting exoplanets can be used to map the photosphere of active stars \citep[e.g.][]{Morris2017}, benefiting both the study of stellar atmospheres and the concerning impact of their non-uniformity on planetary atmosphere retrievals \citep{rackham2018}.
 Overall, enabling the detection of transits in light curves with high levels of correlated noises will greatly benefit the study of terrestrial exoplanets around late M-dwarfs, usually observed at lower SNR and more likely to display photometric variability (e.g.
@@
 However, when correlated noise starts resembling transits, this cleaning step (often referred to as \textit{detrending}) is believed to degrade their detectability \cite[see subsection 4.3 of][]{wotan}.
-In this case, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noise, and to compute the likelihood of the data to the transit model on a wide parameter space (an approach largely avoided due to its intractable nature).
+In this case, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noise, and to compute the likelihood of the data to the transit model on a wide parameter space, an approach largely avoided due to its intractable nature.
 Nonetheless, \cite{kovacs2016} ask: \textit{Periodic transit and variability search with simultaneous systematics filtering: Is it worth it?}.
-The latter study explores a handful of cases and generally discards the benefit of using a full-fledged approach.
-However, it does not explore the light-curves characteristics for which such a complete modeling becomes necessary.\\\\
-In this paper, we explore the light curve characteristics for which a full-fledge transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} written in italics refers to the algorithmic and analytical method, while \nuancecode{} in sans-serif refers to its implementation.}, a method to search for transit signals while modeling correlated noises in a tractable way.
+The latter study explores a handful of cases and generally discards the benefit of using a full-fledged approach.\\\\
+In this paper, we identify regions of light curves morphological parameter space for which a full-fledged transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} written in italics refers to the algorithmic and analytical method, while \nuancecode{} in sans-serif refers to its implementation.}, a method to search for transit signals while simultaneously modeling correlated noises in a tractable way.
 In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on transit signals detectability.
@@
 This include transits injected in synthetic datasets, but also in the TESS light curves of 438 rapidly-rotating M dwarfs.
-Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, and we conclude in .
+Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, and we conclude in \autoref{conclusion}.
 
 \newpage
-\section{The issue}\label{issues}
+\section{Motivation}\label{issues}
 
-A strong assumption when using the BLS algorithm to search for transits is that the searched dataset only contains these signals in addition of white noise.
-Hence, two sources of correlated noise particularly justify the need for a detrending step before searching for transits: instrumental signals (such as those induced by telescope pointing errors) and astrophysical signals (such as stellar variability induced by pulsations or starspots).
-In this section, we qualitatively study the effect of correlated noise and its detrending on transits detectability.
+A strong assumption when using the BLS algorithm to search for transits is that the searched dataset only contains transit signals and white noise, justifying the need for detrending.
+In this section, we explore this effect by simulating light curves containing a transit signal and correlated noise in the form of stellar variability, and study how detrending the latter affects the transit signal detectability depending on the light curve morphological characteristics.
+Hence, we first present how transit light curves are simulated, using a stochastic model of stellar variability, and describe the detection metric we employ to quantify transit detectability in the presence of correlated noise.
 
-\subsection{Transit detectability}
+\subsection{Light curve simulations}\label{light_curves_simulations}
+To perform this study, we want to simulate realistic light curves that contain instrumental signals, transit signals and correlated noise with characteristics that can be controlled using a set of interpretable parameters.
+To this end, we model light curves as realizations of a Gaussian Process (GP; \citealt{Rasmussen2005}) with a mean containing the instrumental and transit signals, and a kernel allowing to model different forms of correlated noise controlled by its hyperparameters.\\\\
+Let $f$ be the simulated differential flux of a star sampled and arranged in the vector $\bm{f}$ associated to the vector of times $\bm{t}$, such that
+\begin{equation*}
+    \bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{C}),
+\end{equation*}
+i.e. that $\bm{f}$ is drawn from a GP of mean $\bm{\mu}$ and covariance matrix $\bm{C}$. For practical reasons, we model $\bm{\mu}$ as a linear combination of $M$ explanatory variables, such that
+\begin{equation}\label{eq:linear_model}
+    \bm{\mu} = \bm{X w}
+\end{equation}
+where the first $M-1$ columns of $\bm{X}$ are contemporaneous instrumental time series measurements, such as the position of the star on the detector or the sky background (e.g.\;\cite{}), and its last column is a box-shaped transit signal with a fixed epoch, duration and period.
+This way, the transit signal is part of the mean linear model, meaning that once the design matrix $\bm{X}$ is constructed the transit is only parametrized by its depth $\Delta$.\\\\
+As we are interested in correlated noise in the form of stellar variability, we choose a physically-motivated GP kernel to model the covariance matrix $\bm{C}$, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) parametrized by its quality factor $Q$, its pulsation $\omega$ and the amplitude of the kernel function $\sigma$ (the full expression of the kernel function is provided in \autoref{app_gp}).
+This choice of mean and kernel function completely defines the GP, from which light curves with different levels of correlated noise can be drawn (see an example in \autoref{fig:app_principle_dataset}).
 
-One way to study the detectability of a unique transit signal is to compute its signal-to-noise-ratio (SNR, \citealt{Pont2006}, Equation 12):
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
+        \caption{Flux times series drawn from the GP model described in \autoref{light_curves_simulations}.
+This dataset corresponds to an observation of 4 days with an exposure time of 2 minutes.
+The mean of this signal consists in a periodic transit signal of period $P=0.7$ days, duration $D=1.2$ hours and depth $\Delta=0.02$ (dark gray) plus an instrumental signal (blue).
+Correlated noise in the form of stellar variability is simulated with an SHO kernel of hyperparameters $\omega = \pi / 6 D$, $Q=45$ and $\sigma = \Delta$.
+Finally, white noise with a standard deviation of $0.001$ is added to the diagonal of the covariance matrix.}
+        \label{fig:app_principle_dataset}
+    \end{centering}
+\end{figure}
+
+\subsection{Transit detectability}\label{transit_detectability}
+
+One way to quantify the detectability of a unique transit signal is to compute its signal-to-noise-ratio expressed as 
+\begin{equation*}
+  SNR= \frac{\Delta}{\sigma}\sqrt{n},
+\end{equation*}
+where $\Delta$ is the transit depth, $\sigma$ the measurements uncertainty (assuming homoscedasticity), and $n$ the number of points within transit.
+Although this metric is useful to assess the strength of the transit signal given a certain photometric precision, it does not account for the presence of correlated noise.
+However, instrumental and other astrophysical signals will necessarily affect the detectability of transits in realistic light curves \citep{Pont2006}.
+For this reason, the combined differential photometric precision metric was developed (CDPP; \citealt{Jenkins2010}), and used in the context of the Kepler mission to assess the level of correlated noise in light curves, affecting the detectability of transits with a given duration.
+% Ultimately, this metric was essential to estimate the yield of the survey, and helped to achieve one of the primary goal of the mission: to measure the occurrence rate of Earth-size planets orbiting in the habitable zones of sun-like stars \citep{}.
+The CDPP is computed by decomposing the data in the time-frequency domain using wavelets, and measures the significance of a distorted transit signal in the whitened data.
+For a given transit duration, the CDPP is then a measure of the noise remaining after filtering the light curve in each frequency band, taking into account the presence of non-stationary correlated noise of a given timescale (see \citealt{Jenkins2010} for more details).
+Hence, we can estimate the significance of the transit signal accounting for the presence of correlated noise as
 \begin{equation}\label{eq:snr}
-  SNR= \frac{\Delta}{\sqrt{\frac{\sigma_w^2}{n} + \frac{\sigma_c^2}{N_{tr}}}}
+  SNR= \frac{\hat{\Delta}}{CDPP_D},
 \end{equation}
-where $\Delta$ is the relative transit depth, $n$ is the number of points within transit, $N_{tr}$ the number of transits ($N_{tr}=1$ here), and $\sigma_w^2$ and $\sigma_c^2$ are the white and correlated noise variances\footnote{i.e.
-$\sigma_w^2 = \sum_{i=j} C_{{ij}}$ and $\sigma_c^2 = \sum_{i\neq j} C_{{ij}}$ with $C$ the covariance matrix of the data.}.
-\autoref{fig:issue1} shows this metric computed for a unique transit observed in the absence (grey) and presence (red) of correlated noise.
-
+where $CDPP_D$ is the CDPP computed for a given transit duration $D$ and $\bar{\Delta}$ is the transit depth after detrending.
+For simplicity, the CDPP is computed using the method from \cite{Gilliland2011} implemented in the \textsf{lightkurve} Python package \citep{Lightkurve} and $\hat{\Delta}$ is the depth obtained by solving \autoref{eq:linear_model} with the last column of $\bm{X}$ containing a normalized transit model.
+\autoref{fig:issue1} shows the SNR from \autoref{eq:snr} computed for a unique transit observed in the absence (grey) and presence of correlated noise (red).
 \begin{figure}[H]
@@
         \caption{Illustration of the effect of correlated noise on a single transit SNR.
-A 1-hour transit signal of depth 1\% is generated on top of white noise (variance of $0.0015^2$) as part of a 24-hours observation with an exposure time of 1 minute (top).
-Then, in the bottom plot, correlated noise is added, generated using a Gaussian process (GP; \citealt{Rasmussen2005}) with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%.
+An hour-long transit of depth 0.01 is generated on top of white noise (standard deviation of 0.0015) as part of a 24-hours observation with an exposure time of 1 minute (top).
+Then, in the bottom plot, correlated noise is added, generated using a GP with an SHO kernel of hyperparameters $\omega=60$, $Q=0.5$ and $\sigma=\Delta/4$ .
 The SNR on the right of each light curve is computed using \autoref{eq:snr}.
@@
 The presence of instrumental correlated noise motivated the development of systematics detrending algorithms, such as the Trend Filtering Algorithm (\textsc{TFA}, \citealt{tfa}, in its primary use case), \textsc{SysRem} (\citealt{sysrem}) or Pixel Level Decorrelation (\textsc{PLD}, \citealt{pld}; see also \textsc{Everest} from \citealt{everest1, everest2}).
-Most of these methods rely on the shared nature of instrumental signals among light-curves (or neighboring pixels) such that the correction applied should not degrade the transit signal and can be modeled using contemporaneous measurements (like detector's temperature, pointing error, sky background or airmass time series).
+Most of these methods rely on the shared nature of instrumental signals among light curves (or neighboring pixels) such that the correction applied should not degrade the transit signal and can be modeled using contemporaneous measurements (e.g.\;detector's temperature, pointing error, sky background or airmass time series).
+But even after instrumental signals have been removed, stellar variability and other astrophysical signals remain, which gave rise to several approaches.
+Some of them are physically-motivated and make use of GPs (e.g.\;\citealt{k2sc}), others are empirical and make use of filtering and data-driven algorithms (\citealt{Jenkins2010}, \citealt{wotan}).
+In this section, we show how these techniques impact transits detectability, depending on the morphological characteristics of light curves.\\\\
+In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different timescales, sampled from a GP with an SHO kernel described in \autoref{app_gp}.
+For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same GP from which the data has been sampled.
+We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
+\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that this degradation due to detrending is strongly dependent on the correlated noise characteristics encountered.
 \begin{figure}[H]
@@
         \includegraphics[width=\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
-        \caption{(Top left) Simulated observation spanning 3 days, including a transit signal with a depth of 0.8\%, duration of 0.1 day, and measurement error of 1.5\% (using the model from \cite{protopapas} described in \autoref{app_proto}).
-(Left) Correlated noise in the form of stellar variability with different timescales and amplitudes added to the unique transit signal.
-This signal is then reconstructed using (in purple) Tukey's bi-weight filter \citep{wotan} with an optimal window size of three times the transit duration, and (in red) a GP with the same kernel used to simulate the data.
-In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr} with the transit depth estimated as the minimum in-transit flux.}
+        \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.08$, duration 0.1 day, and white noise with a standard deviation of 0.015 (top left).
+The hyperparameters of the kernel are fixed to $Q=10$ and $\sigma=0.008$, with the pulsation $\omega$ increasing from top to bottom.
+The resulting stellar variability signals are then reconstructed using (in purple) Tukey's bi-weight filter with an optimal window size of $3\times D$  (see \citealt{wotan}), and (in red) a GP with the same kernel used to simulate the data.
+In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr}.}
         \label{fig:issue2}
@@
 \end{figure}
-In opposition, stellar variability and other astrophysical signals cannot be correlated with simultaneous measurements.
-This gave rise to several treatments in order to reconstruct and detrend stellar variability.
-Some of them are physically-motivated and make use of GPs (e.g.
-\citealt{k2sc}), others are empirical and make use of filtering and data-driven algorithms (\citealt{Jenkins2010}, \citealt{wotan}).\\\\
-In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different amplitudes and timescales, sampled from a GP with an SHO kernel described in \autoref{app_gp}.
-For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same GP from which the data has been sampled.
-We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
-\autoref{fig:issue2} clearly shows the effect of both detrending techniques on transits SNR, and intuitively suggests that this degradation due to detrending is strongly dependant on the correlated noise characteristics encountered.\\\\
-In order to explore the parameter space for which detrending is the most problematic, we employ the method previously described to simulate 10 000 light curves containing a transit signal and correlated noise in the form of stellar variability.
-For each light curve, we reconstruct the variability signal and detrend it using an optimal bi-weight filter, i.e.
-with a window size three times that of the transit duration \citep{wotan}, before estimating the remaining transit SNR.
-For convenience, the simulated light curves share a common transit added on top of white noise, and a variability signal only defined by two parameters: $\tau$, the relative timescale of the variability with respect to the transit duration; and $\delta$, the relative amplitude of the variability against transit depth, corresponding to an SHO kernel with hyperparameters chosen as:
+To explore the parameter space for which detrending is the most problematic, we employ the model described in \autoref{signals_simulations} to simulate 10 000 differential light curves with different morphological characteristics, and compute the remaining transit SNR after detrending.
+In order to place the stellar variability hyperparameters on a relative scale with the transit signal parameters, we reparametrize the GP kernel with
 \begin{equation}\label{eq:relative_params}
-    \omega = \frac{\pi}{\tau D}, \hspace{0.5cm} 
-    \sigma = \delta \frac{\Delta}{2} \hspace{0.5cm}  \text{and}  \hspace{0.5cm}  
+    \tau = \frac{\pi}{\omega D}, \hspace{0.5cm} 
+    \delta = \frac{2\sigma}{\Delta} \hspace{0.5cm}  \text{and}  \hspace{0.5cm}  
     Q = 10,
 \end{equation}
-with $\Delta=1\%$ and $D=0.04$ days, the depth and duration of the simulated transit. 
-% We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability signals.
-
-For $\tau=1$ and $\delta=1$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a variability signal with a period half that of the transit duration, and a correlated noise amplitude comparable to the transit depth, i.e.
-strongly resembling the simulated transit signal.\\\\
-\autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 6$).
-Hence, bi-weight filter detrending makes transit search blind to many systems.
-While this study should be extended to other detrending techniques, it highlights the need for a more informed transit search algorithm able to deal with correlated noise, at least if present in the form of stellar variability.
-
-% ## STOPPED here
-
+where $\tau$ is the relative timescale of the variability with respect to the transit duration and $\delta$ the relative amplitude of the variability against the transit depth, both being adimensional.
+Hence, for $(\tau, \delta)=(1, 1)$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a variability signal with a period half that of the transit duration, and a correlated noise amplitude comparable to the transit depth, i.e.
+strongly resembling the simulated transit signal.
 \begin{figure}[H]
     \begin{centering}
-        \includegraphics[width=0.9\linewidth]{../workflows/cleaning_snr/figures/result.pdf}
-        \caption{SNR of a unique transit after detrending light curves using an optimal bi-weight filter (i.e.\;of window size $3\times D$).
-All light span 2.8 days with exposure times of 2 minutes.
-The 10 000 simulated datasets contain the same transit signal with a duration $D=1$ hour and a depth of 1\%, added on top of white noise with a standard deviation of 0.5\% plus stellar variability.
+        \includegraphics[width=\linewidth]{../workflows/cleaning_snr/figures/result.pdf}
+        \caption{Transit SNR computed for 10 000 differential light curves with varying morphological characteristics.
+All light curves span 2.8 days with an exposure time of 2 minutes and contain the same transit signal of  duration $D=1$ hour, depth of 0.005 and white noise with a standard deviation of 0.001.
 Light curves at the top and right side of the central plot are shown with their corresponding $\tau$ and $\delta$ values, which again correspond to the relative timescale of stellar variability against the transit duration, and the relative amplitude of stellar variability against the transit depth.
-Hence, the amplitude of the variability increases with $\delta$ and the timescale of the variability increases with $\tau$, the transit signal remaining fixed.}
+The two central plots show the transit SNR after detrending the light curves with one of two methods.
+On the left, light curves are detrended using Tukey's bi-weight filter with an optimal window size of $3\times D$ (see \citealt{wotan}), while on the right, light curves are detrended using a GP with the same kernel used to simulate the data.}
         \label{fig:snr_detrend}
@@
 \end{figure}
+For each light curve, we separately reconstruct and detrend  the variability signal using the two techniques employed in \autoref{fig:issue2}, i.e.\;one using an optimal bi-weight filter with a window size three times that of the transit duration \citep{wotan} and the other using a GP with an optimal kernel.
+We then compute the detrended transit SNR using \autoref{eq:snr}.\\\\
+\autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 5$).
+Although being more robust, the same effect is observed when detrending with an optimal GP.
+Hence, detrending makes transit search blind to many systems, especially when the widely adopted bi-weight filter is used.
+While this study should be extended to other detrending techniques, it highlights the need for a more informed transit search algorithm able to deal with correlated noise, at least if present in the form of stellar variability.
 
@@
 
-\textit{nuance} is an algorithm capable of searching for planetary transits in light curves containing correlated noise, such as instrumental signals and photometric stellar variability.
-\\\\
-Let assume that the flux $f$ of a star is observed and arranged in the vector $\bm{f}$ of size $N$, associated to the vector of times $\bm{t}$.
-This flux, shown in \autoref{fig:linear_search}, contains instrumental signals, stellar variability and a periodic transit signal that we wish to detect.
-We assume that a set of $M$ observed measurements (such as the position of the star on the detector or the sky background) taken at the same time as the flux can be treated as explanatory variables for $\bm{f}$.
-These measurements constitute the columns of the $(N\times M)$ \textit{design matrix} $\bm{X}$.\\\\
-Ideally, we would detect the periodic transit signal in this flux by sampling the posterior likelihood of this data to a full-fledge model including stellar variability (more generally correlated noise), instrumental systematic signals (modeled with explanatory variables or treated as correlate noise), and a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$.
-We would then reduce the posterior likelihood to $p(\bm{f}\vert P)$, its marginalized version over all parameters except the period $P$, producing a transit search periodogram $\mathcal{Q}(P)$.
-However, this approach has two issues: It is highly intractable, and it may lead to a multimodal distribution that is hard to interpret.
-\\\\
-Given a period $P$, we instead want to compute the likelihood of a periodic transit signal at the maximum likelihood parameters $\hat T_0$, $\hat D$ and $\hat \Delta$, i.e the periodogram
+In this section, we present \textit{nuance}, an algorithm capable of searching for planetary transits in light curves containing correlated noise such as instrumental signals and photometric stellar variability.\\\\
+When searching for exoplanets using an indirect method, such as using transits or radial velocities, the initial detection is often as important as the follow-up observations, ultimately leading to the confirmation of the planetary candidate.
+For that reason, the final product of most transit search algorithms consists in a \textit{periodogram}: a 1D detection metric given a set of trial periods that enables to identify the presence of a potential candidate, and the period and epoch at which it should be followed up.\\\\
+If we assume that a transit is defined by its period $P$, epoch $T$, duration $D$ and depth $\Delta$, we then wish to compute the likelihood of such a transit being present in the for a set of periods $P$, leading to an interpretable periodogram.
+As we are interested in following up transits with specific parameters, such as a well-defined epoch, the likelihood we wish to compute must not be marginalized over all parameters other than $P$ but rather specifically computed at the maximum-likelihood parameters $\hat T$, $\hat D$ and $\hat \Delta$, leading to the \textit{profile likelihood}
 \begin{equation}\label{eq:periodogram}
-        \mathcal{Q}(P) = p(\bm{f} \vert P, \hat T_0 ,\hat D, \hat \Delta)
+       p(\bm{f} \vert P, \hat T ,\hat D, \hat \Delta),
 \end{equation} 
-This is done by adopting the strategy of \cite{foreman2016}, separating the transit search into two components: the linear search and the periodic search.
-During the linear search, the likelihood of a single non-periodic transit is computed for a grid of epochs, durations and depths.
-Then, the periodic search consists in combining these likelihoods to compute the likelihood of the data given a periodic transit signal for a range of periods.
+where $\bm{f}$ is the data. As in \autoref{light_curves_simulations} let assume that $\bm{f}$ is a vector of size $N$ containing the flux of a star observed at times $\bm{t}$ such that
+\begin{equation*}
+    \bm{f} \sim \mathcal{N}(\bm{X w}, \bm{\Sigma}),
+\end{equation*}
+i.e.\;that $\bm{f}$ is drawn from a GP of mean $\bm{Xw}$ and covariance $\bm{\Sigma}$.
+Again, the first $M-1$ columns of the $(N\times M)$ design matrix $\bm{X}$ are contemporaneous instrumental time series measurements and its last column consists in a normalized box-shaped transit of period $P$, epoch $T$ and duration $D$.
+This way, the transit signal is part of the linear model and its depth $\Delta$ can be solved linearly.
+Under this assumption, the log-likelihood of the data given the presence of a periodic transit signal of period $P$, epoch $T$ and duration $D$ is \citep{Rasmussen2005}
+\begin{equation} \label{eq:linear_search_ll}
+    \ln p(\bm{f} \vert P, T, D, \Delta) = -\frac{1}{2}(\bm{f}-\bm{Xw})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{Xw}) -  \frac{1}{2}\ln\vert\bm{\Sigma}\vert - \frac{N}{2}\ln 2\pi.
+\end{equation}
+In this expression, the vector of parameters $\bm{w}$ and their errors $\bm{\sigma}$ are computed using the generalized least-square solution
+\begin{equation}\label{eq:ls}
+    \bm{w} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}\bm{X}^T\bm{\Sigma}^{-1}\bm{f} \hspace{0.5cm} \text{and} \hspace{0.5cm} \bm{\sigma} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1},
+\end{equation}
+so that the depth of the transit corresponds to $\hat{\Delta} = \bm{w}_M$.\\\\
+Hence, given a period $P$, computing $Q(P)$ boils down to a non-linear optimization involving several evaluations of the likelihood given in \autoref{eq:linear_search_ll}.
+While this is tractable for few periods, it becomes highly intractable for the large set of trial periods required to properly sample a transit search periodogram.\\\\
+To be tractable, \nuancemethod{} employs the strategy of \cite{foreman2016}, separating the transit search into two components: the linear search and the periodic search.
+During the linear search, the likelihood of a single non-periodic transit is computed for a grid of epochs and durations, requiring the evaluation of $p(\bm{f} \vert T, D)$ for a fixed number of times.
+Then, the periodic search consists in combining these likelihoods to compute the likelihood $p(\bm{f} \vert P, T, D)$ of the data given a periodic transit signal for a range of periods $P$.
 These combined likelihoods yield a transit-search periodogram on which the periodic transit detection is based.
 \textsf{nuance} differs from \cite{foreman2016} and other existing transit search algorithms as it models the covariance of the light curve with a GP, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
-This way, \textsf{nuance} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR.\\\\
+This way, \textsf{nuance} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR (see \autoref{issues}).\\\\
 We note that the approach employed by \nuancemethod{} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.
@@
 
-The goal of the linear search is to compute the likelihood $p(\bm{f} \vert T , D, \Delta)$ of the data given a single non-periodic transit signal of epoch $T$, duration $D$ and depth $\Delta$, for a grid of epochs, durations and depths.
-\\\\
-To account for correlated noise, the light curve $f$ is modeled as being drawn from a GP, like the simulated light curves used in \autoref{issues}, such that
+The goal of the linear search is to compute the likelihood $p(\bm{f} \vert T , D)$ for a grid of epochs $\set{T_i}_i$ and durations $\set{D_j}_j$.
+For each pair $(T_i, D_j)$ the transit depth $\Delta_{i, j}$ is linearly solved and follow a Gaussian distribution of mean $\hat{\Delta}_{i, j}$ and standard deviation $\sigma_{i, j}$, computed using \autoref{eq:ls}.
+Hence, the linear search lead to the set of maximum-likelihoods
 \begin{equation*}
-    \bm{f} \sim \mathcal{N}(\bm{X w}, \bm{\Sigma}),
+    \set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \hat{\Delta}_{i,j})}_{i, j}.
 \end{equation*}
-with mean $\bm{Xw}$ (i.e.
-a linear model of the M explanatory variables with coefficients $\bm{w}$) and covariance $\bm{\Sigma}$.
-To account for the presence of a single non-periodic transit of epoch $T$ and duration $D$, this signal is computed and appended as the last column of the design matrix $\bm{X}$, using the simple transit model from \cite{protopapas} with a unitary depth (\autoref{eq:protopapas}).
-This way, the transit signal is part of the linear model and its depth $\Delta$ can be solved linearly.
-Under this assumption, the log-likelihood of the data given a single non-periodic transit is \citep{Rasmussen2005}
-\begin{equation} \label{eq:linear_search_ll}
-    \ln p(\bm{f} \vert I) = -\frac{1}{2}(\bm{f}-\bm{Xw})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{Xw}) -  \frac{1}{2}\ln\vert\bm{\Sigma}\vert - \frac{N}{2}\ln 2\pi,
-\end{equation}
-where the parameters vector $\bm{w}$ and their errors $\bm{\sigma}$ are computed using the generalized least-square solution
-\begin{equation}\label{eq:ls}
-    \bm{w} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}\bm{X}^T\bm{\Sigma}^{-1}\bm{f} \hspace{0.5cm} \text{and} \hspace{0.5cm} \bm{\sigma} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1},
-\end{equation} 
-with $\bm{\Sigma}$ the covariance matrix modeled using the GP.
-Hence, $\ln p(\bm{f} \vert I)$ can be computed on a grid of epochs and durations, the transit depth being linearly solved for any $(T, D)$.
-This \textit{linear search} leads to the set of likelihoods
-\begin{equation*}
-    \set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})}_{i, j},
-\end{equation*}
-were $\Delta_{i,j}$ is the depth linearly solved for a given $(T_i, D_j)$\footnote{The expression of $\set{\ln\mathcal{L}_{i,j}}_{i, j}$ omits the vector $\bm{w}_{i,j}$ (except its last value $\Delta_{i,j}$) as it is also linearly solved for any given value of $(T_i, D_j)$ and irrelevant in what follows.}.
-\autoref{fig:linear_search} shows this likelihood grid computed for a simulated dataset, using the same GP  and design matrix $\bm{X}$ used to simulate the data.
+An example of such a grid of likelihoods is shown in \autoref{fig:linear_search}.
 
@@
 \begin{equation*}
-    p(\bm{f} \vert P, T_0 , D, \Delta),
+    p(\bm{f} \vert P, T , D, \Delta),
 \end{equation*}
 i.e.\,the probability of a periodic transit of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$ given the data $\bm{f}$.
-For a given transit duration $D$, any combination of $(P, T_0)$ leads to K transits, for which it is tempting to write
+For a given transit duration $D$, any combination of $(P, T)$ leads to K transits, for which it is tempting to write
 \begin{equation}\label{eq:attempt}
-    p(\bm{f} \vert P, T_0 ,D, \Delta) = \prod_k^K p(\bm{f} \vert T_k, D, \Delta_k),
+    p(\bm{f} \vert P, T ,D, \Delta) = \prod_k^K p(\bm{f} \vert T_k, D, \Delta_k),
 \end{equation}
@@
 \begin{equation*}
-    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) = \sum_k^K \ln \mathcal{L}_k.
+    \ln p(\bm{f} \vert P, T ,D, \Delta) = \sum_k^K \ln \mathcal{L}_k.
 \end{equation*}
@@
 \newpage
-\section{Conclusion}
+\section{Conclusion}\label{conclusion}
 This paper presents \nuancemethod{}, an algorithm designed to detect planetary transits in light curves featuring correlated noise in the form of instrumental signals and stellar variability.

### 2024-01-26 latex/paper.tex
fix: wrapping up nuance section

@@
 \begin{equation*}
-    \bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{C}),
+    \bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{\Sigma}),
 \end{equation*}
-i.e. that $\bm{f}$ is drawn from a GP of mean $\bm{\mu}$ and covariance matrix $\bm{C}$. For practical reasons, we model $\bm{\mu}$ as a linear combination of $M$ explanatory variables, such that
+i.e. that $\bm{f}$ is drawn from a GP of mean $\bm{\mu}$ and covariance matrix $\bm{\Sigma}$. For practical reasons, we model $\bm{\mu}$ as a linear combination of $M$ explanatory variables, such that
 \begin{equation}\label{eq:linear_model}
@@
 This way, the transit signal is part of the mean linear model, meaning that once the design matrix $\bm{X}$ is constructed the transit is only parametrized by its depth $\Delta$.\\\\
-As we are interested in correlated noise in the form of stellar variability, we choose a physically-motivated GP kernel to model the covariance matrix $\bm{C}$, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) parametrized by its quality factor $Q$, its pulsation $\omega$ and the amplitude of the kernel function $\sigma$ (the full expression of the kernel function is provided in \autoref{app_gp}).
+As we are interested in correlated noise in the form of stellar variability, we choose a physically-motivated GP kernel to model the covariance matrix $\bm{\Sigma}$, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) parametrized by its quality factor $Q$, its pulsation $\omega$ and the amplitude of the kernel function $\sigma$ (the full expression of the kernel function is provided in \autoref{app_gp}).
 This choice of mean and kernel function completely defines the GP, from which light curves with different levels of correlated noise can be drawn (see an example in \autoref{fig:app_principle_dataset}).
@@
 \section{\textsf{nuance}}\label{nuance}
+When searching for exoplanets using an indirect method, such as using transits or radial velocities, the initial detection is often as important as the follow-up observations, ultimately leading to the confirmation of the planetary candidate.
+For that reason, the final product of most transit search algorithms consists in a \textit{periodogram}: a 1D detection metric given a set of trial periods that allows to identify the presence of a potential candidate, and the period and epoch at which it should be followed up.\\\\
+If we assume that a transit is defined by its period $P$, epoch $T_0$, duration $D$ and depth $\Delta$, we then wish to compute the likelihood of such a transit being present in the data for a set of periods $P$, leading to an interpretable periodogram.
+As we are interested in following up transits with specific parameters, such as a well-defined epoch, the likelihood we wish to compute must not be marginalized over all parameters other than $P$ but rather specifically computed at the maximum-likelihood parameters $\hat T_0$, $\hat D$ and $\hat \Delta$, leading to the \textit{profile likelihood}
+\begin{equation}\label{eq:periodogram}
+       \mathcal{Q}(P) = p(\bm{f} \vert P, \hat T_0 ,\hat D, \hat \Delta),
+\end{equation} 
+where $\bm{f}$ is the data.
+In this section, we present \textit{nuance}, an algorithm to compute transit-search periodograms in a tractable way, and detect planetary transits in light curves containing correlated noise such as instrumental signals and photometric stellar variability.
+In \autoref{approach}, we explain how our approach requires two separate steps in order to remain tractable, and present these steps in \autoref{linear_search} and \autoref{periodic_search}, leading to the transit-search periodogram described in \autoref{periodogram}.
+Finally, we present the \nuancecode{} Python package in \autoref{package} and perform a control test of our implementation against \textsf{BLS} in \autoref{control}.
 
-In this section, we present \textit{nuance}, an algorithm capable of searching for planetary transits in light curves containing correlated noise such as instrumental signals and photometric stellar variability.\\\\
-When searching for exoplanets using an indirect method, such as using transits or radial velocities, the initial detection is often as important as the follow-up observations, ultimately leading to the confirmation of the planetary candidate.
-For that reason, the final product of most transit search algorithms consists in a \textit{periodogram}: a 1D detection metric given a set of trial periods that enables to identify the presence of a potential candidate, and the period and epoch at which it should be followed up.\\\\
-If we assume that a transit is defined by its period $P$, epoch $T$, duration $D$ and depth $\Delta$, we then wish to compute the likelihood of such a transit being present in the for a set of periods $P$, leading to an interpretable periodogram.
-As we are interested in following up transits with specific parameters, such as a well-defined epoch, the likelihood we wish to compute must not be marginalized over all parameters other than $P$ but rather specifically computed at the maximum-likelihood parameters $\hat T$, $\hat D$ and $\hat \Delta$, leading to the \textit{profile likelihood}
-\begin{equation}\label{eq:periodogram}
-       p(\bm{f} \vert P, \hat T ,\hat D, \hat \Delta),
-\end{equation} 
-where $\bm{f}$ is the data. As in \autoref{light_curves_simulations} let assume that $\bm{f}$ is a vector of size $N$ containing the flux of a star observed at times $\bm{t}$ such that
+\subsection{The approach}\label{approach}
+As in \autoref{light_curves_simulations}, let assume that $\bm{f}$ is a vector of size $N$ containing the flux of a star observed at times $\bm{t}$ such that
 \begin{equation*}
@@
 i.e.\;that $\bm{f}$ is drawn from a GP of mean $\bm{Xw}$ and covariance $\bm{\Sigma}$.
-Again, the first $M-1$ columns of the $(N\times M)$ design matrix $\bm{X}$ are contemporaneous instrumental time series measurements and its last column consists in a normalized box-shaped transit of period $P$, epoch $T$ and duration $D$.
+Again, we set the first $M-1$ columns of the $(N\times M)$ design matrix $\bm{X}$ as contemporaneous measurements and the last column as a normalized box-shaped transit of period $P$, epoch $T_0$ and duration $D$.
 This way, the transit signal is part of the linear model and its depth $\Delta$ can be solved linearly.
-Under this assumption, the log-likelihood of the data given the presence of a periodic transit signal of period $P$, epoch $T$ and duration $D$ is \citep{Rasmussen2005}
+Under this assumption, the log-likelihood of the data given the presence of a periodic transit signal of period $P$, epoch $T_0$, duration $D$, and a mean linear model with coefficients $\bm{w}$ is \citep{Rasmussen2005}
 \begin{equation} \label{eq:linear_search_ll}
-    \ln p(\bm{f} \vert P, T, D, \Delta) = -\frac{1}{2}(\bm{f}-\bm{Xw})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{Xw}) -  \frac{1}{2}\ln\vert\bm{\Sigma}\vert - \frac{N}{2}\ln 2\pi.
+    \ln p(\bm{f} \vert P, T_0, D, \bm{w}) = -\frac{1}{2}(\bm{f}-\bm{Xw})^T\bm{\Sigma}^{-1}(\bm{f}-\bm{Xw}) -  \frac{1}{2}\ln\vert\bm{\Sigma}\vert - \frac{N}{2}\ln 2\pi.
 \end{equation}
-In this expression, the vector of parameters $\bm{w}$ and their errors $\bm{\sigma}$ are computed using the generalized least-square solution
+Given the linearity of the mean model at $P$, $T_0$ and $D$ fixed, this likelihood is maximized for the least-square parameters
 \begin{equation}\label{eq:ls}
-    \bm{w} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}\bm{X}^T\bm{\Sigma}^{-1}\bm{f} \hspace{0.5cm} \text{and} \hspace{0.5cm} \bm{\sigma} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1},
+    \bm{w} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}\bm{X}^T\bm{\Sigma}^{-1}\bm{f} \hspace{0.5cm} \text{with  uncertainties} \hspace{0.5cm} \bm{\sigma} = (\bm{X}^T\bm{\Sigma}^{-1}\bm{X})^{-1}.
 \end{equation}
-so that the depth of the transit corresponds to $\hat{\Delta} = \bm{w}_M$.\\\\
-Hence, given a period $P$, computing $Q(P)$ boils down to a non-linear optimization involving several evaluations of the likelihood given in \autoref{eq:linear_search_ll}.
-While this is tractable for few periods, it becomes highly intractable for the large set of trial periods required to properly sample a transit search periodogram.\\\\
-To be tractable, \nuancemethod{} employs the strategy of \cite{foreman2016}, separating the transit search into two components: the linear search and the periodic search.
-During the linear search, the likelihood of a single non-periodic transit is computed for a grid of epochs and durations, requiring the evaluation of $p(\bm{f} \vert T, D)$ for a fixed number of times.
-Then, the periodic search consists in combining these likelihoods to compute the likelihood $p(\bm{f} \vert P, T, D)$ of the data given a periodic transit signal for a range of periods $P$.
-These combined likelihoods yield a transit-search periodogram on which the periodic transit detection is based.
-\textsf{nuance} differs from \cite{foreman2016} and other existing transit search algorithms as it models the covariance of the light curve with a GP, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
-This way, \textsf{nuance} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR (see \autoref{issues}).\\\\
+% Using this expression, the depth of the transit corresponds to $\hat{\Delta} = \bm{\hat{w}}_M$ with the uncertainty $\hat{\sigma} = \bm{\hat{\sigma}}_{M,M}$, where M denotes the index of the last column of the design matrix $\bm{X}$, corresponding to the normal\ transit signal.
+As we are only interested in the transit depth, we will omit $\bm{w}$ in the reminder of this paper and simply write likelihood expressions depending on its last value $\Delta=\bm{w}_M$, such as $p(\bm{f} \vert P, T_
+0, D, \Delta)$, where all other parameters of $\bm{w}$ are taken at their maximum-likelihood values. \\\\
+Hence, given a period $P$, computing $\mathcal{Q}(P)$ boils down to a non-linear optimization involving several evaluations of the likelihood given in \autoref{eq:linear_search_ll}.
+While this is tractable for few periods, it becomes highly untractable for the large set of trial periods required to properly sample a transit-search periodogram (in the order of tens of thousands).\\\\
+To remain tractable, \nuancemethod{} employs the strategy of \cite{foreman2016}, and separates the transit search into two components: the linear search and the periodic search.
+During the linear search, the likelihood $p(\bm{f} \vert T_0, D, \Delta)$ of a single non-periodic transit is computed for a grid of epochs and durations, each time solving for $\Delta$ linearly.
+Then, the periodic search consists in combining these likelihoods to compute the likelihood $p(\bm{f} \vert P, T_0, D, \Delta)$ of the data given a periodic transit signal for a range of periods $P$.
+These combined likelihoods yield $p(\bm{f} \vert P, \hat T_0 ,\hat D, \hat \Delta)$, a transit-search periodogram on which the periodic transit detection is based.
+\nuancemethod{} differs from \cite{foreman2016} and other existing transit search algorithms as it models the covariance of the light curve with a GP, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
+This way, \nuancemethod{} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR (see \autoref{issues}).\\\\
 We note that the approach employed by \nuancemethod{} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.
@@
 
-The goal of the linear search is to compute the likelihood $p(\bm{f} \vert T , D)$ for a grid of epochs $\set{T_i}_i$ and durations $\set{D_j}_j$.
-For each pair $(T_i, D_j)$ the transit depth $\Delta_{i, j}$ is linearly solved and follow a Gaussian distribution of mean $\hat{\Delta}_{i, j}$ and standard deviation $\sigma_{i, j}$, computed using \autoref{eq:ls}.
-Hence, the linear search lead to the set of maximum-likelihoods
+The goal of the linear search is to compute the likelihood $p(\bm{f} \vert T_0 , D, \Delta)$ for a grid of epochs $\set{T_i}_i$ and durations $\set{D_j}_j$.
+For each pair $(T_i, D_j)$ the transit depth $\Delta_{i, j}$ is linearly solved, which leads to the set of maximum likelihoods
 \begin{equation*}
-    \set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \hat{\Delta}_{i,j})}_{i, j}.
+    \set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i, j})}_{i, j}.
 \end{equation*}
 An example of such a grid of likelihoods is shown in \autoref{fig:linear_search}.
-
+To prepare for the next step, uncertainties on the depths $\sigma_{i,j}$ are also computed using \autoref{eq:ls} and stored.
 \begin{figure}[H]
@@
 
-\noindent To prepare for the next step, the corresponding depths $\Delta_{i,j}$ linearly solved for any $(T_i ,D_j)$ are stored, as well as their associated uncertainties $\sigma_{i,j}$, corresponding to
+\subsection{The periodic search}\label{periodic_search}
+
+We then need to combine the likelihoods computed from the linear search to obtain
 \begin{equation*}
-    \Delta_{i,j} = \bm{w}_M \hspace{0.5cm}\text{and}\hspace{0.5cm} \sigma_{i,j} = \bm{\sigma}_{MM},
-\end{equation*}
-M denoting the index of the last column of the design matrix $\bm{X}$, where the transit signal is contained.
-
-\subsection{The periodic search}
-
-We then need to combine the likelihoods computed from the \textit{linear search} to obtain
-\begin{equation*}
-    p(\bm{f} \vert P, T , D, \Delta),
+    p(\bm{f} \vert P, T_0 , D, \Delta),
 \end{equation*}
 i.e.\,the probability of a periodic transit of period $P$, epoch $T_0$, duration $D$ and depth $\Delta$ given the data $\bm{f}$.
-For a given transit duration $D$, any combination of $(P, T)$ leads to K transits, for which it is tempting to write
+For a given transit duration $D$, any combination of $(P, T_0)$ leads to K transits, for which it is tempting to write
 \begin{equation}\label{eq:attempt}
-    p(\bm{f} \vert P, T ,D, \Delta) = \prod_k^K p(\bm{f} \vert T_k, D, \Delta_k),
+    p(\bm{f} \vert P, T_0 ,D, \Delta) = \prod_k^K p(\bm{f} \vert T_k, D, \Delta_k),
 \end{equation}
@@
 \begin{equation*}
-    \ln p(\bm{f} \vert P, T ,D, \Delta) = \sum_k^K \ln \mathcal{L}_k.
+    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) = \sum_k^K \ln \mathcal{L}_k.
 \end{equation*}
@@
 However, individual transits from a periodic signal cannot be considered independent, and should instead be found periodically and share a common transit depth $\Delta$.
-To this end, it can be shown (see \autoref{proof}) that there is an analytical expression for the joint likelihood of K individual transits with depths and errors $\set{\Delta_k, \sigma_k}_k$ assuming a common depth $\Delta$, corresponding to
+To this end, it can be shown (see \autoref{proof}) that there is a closed form expression for the joint likelihood of K individual transits with depths and errors $\set{\Delta_k, \sigma_k}_k$ assuming a common depth $\Delta$, corresponding to
 \begin{equation}\label{eq:result}
@@
 \end{equation}
-
+In order to compute this joint likelihood, we must assume that the likelihoods computed during the linear search are independent.
+In theory, given that all data points of the light curves are used to compute each likelihood separately, these are not.
+In practice however, this assumption is reasonable and only affect the results of the periodic search for extremely short-period transit signals (where $P$ is less than few durations) that are less likely to be present in physically-realistic datasets.\\\\
 While \autoref{eq:result} takes a closed form, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$.
@@
 
-\subsection{The transit search periodogram}
+\subsection{The transit-search periodogram}\label{periodogram}
 
 Using \autoref{eq:result}, we can now compute $\ln\mathcal{P}$ for a range of periods and build a transit search periodogram using \autoref{eq:periodogram}.
-But a final issue emerges, one that is fundamentally linked to our strategy.
-Each likelihood $p(\bm{f} \vert T, D, \Delta)$ estimated during the linear search is computed using N measurements.
+But a final issue emerges, one that is fundamentally linked to our approach.
+Each likelihood $p(\bm{f} \vert T_0, D, \Delta)$ estimated during the linear search is computed using N measurements.
 Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving $N\times K$ measurements.
@@
 \end{equation}
-and defined $\mathcal{Q}(P)$ as being equal to the SNR of the transit of period $P$, epoch $T_0 = \phi_0 P$, duration $D$ and depth $\Delta$, i.e.
+and defined $\mathcal{Q}(P)$ as being equal to the SNR of the transit of period $P$, epoch $T_0 = \phi_0 \times P$, duration $D$ and depth $\Delta$, i.e.
 \begin{equation*}
@@
 
-\subsection{An open-source python package}
+\subsection{An open-source python package}\label{package}
 Methods presented in this paper are made accessible through the \nuancecode{} open-source Python package, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} and released on the Python Package Index\footlink{https://pypi.org/project/nuance/}.
@@
 
-To start testing \nuancemethod{} against existing methods, a simple adimensional normalized light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ spanning 6 days with an exposure time of 2 minutes.
+To start testing \nuancecode{} against existing methods, a simple adimensional normalized light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ spanning 6 days with an exposure time of 2 minutes.
 From this signal, we produce 4000 light curves, each containing box-shaped transits with periods randomly sampled from 0.3 to 2.5 days, durations of 50 minutes, and depths randomly sampled to lead to transit SNRs ranging from 4 to 30.
@@
         \caption{Study of \textsf{nuance} (black line) processing time against \texttt{biweight+BLS} (cf.\;\autoref{simu}, blue line).
-The gray curve shows the performance of \textsf{nuance} \textit{linear search} when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together.
+The gray curve shows the performance of \textsf{nuance} linear search when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together.
 This study is performed on the 12 cores of an Apple M2 Max chip, exploiting the parallelization offered by our implementation.
@@
 Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsly is extremely promising for the search of transiting exoplanets in ground-based data, that usually suffer from daily interruptions.
-In this respect, we note the similarity of our \textit{linear search} (cf\;\autoref{linear_search}) to the one presented in \citealt{Berta2012}, that focused on the detection of single eclipses in the MEarth light curves \citep{Irwin2009}.
+In this respect, we note the similarity of our linear search (cf\;\autoref{linear_search}) to the one presented in \citealt{Berta2012}, that focused on the detection of single eclipses in the MEarth light curves \citep{Irwin2009}.
 Similarly, \textbf{\nuancemethod{} would highly benefit the search for transiting exoplanets around M-dwarf type stars}, such as the ones observed by the SPECULOOS survey \citep{speculoos} whose monitoring suffers from both increased red noise (due to atmospheric and instrumental thermal effects discussed e.g.\;in \citealt{Berta2012} and  \citealt{Pedersen2023}) and enhanced stellar variability \citep{Murray2020}.
@@
 \appendix
-\section{Light curves simulations}\label{signals_simulations}
-
-In order to study the effect of correlated noise on transit search, this paper relies on transit light curve simulations including realistic effects of stellar variability and instrumental signals.
-The following describes how such signals are modeled.
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
-        \caption{Example dataset sampled at $N=2880$ times corresponding to an observation of 4 days with an exposure time of 2 minutes.
-The mean of this signal consists in a periodic transit signal of period $P=0.7$ days, duration $D=0.05$ days and depth of 2\% ($\mathcal{T}$ in gray) plus instrumental signals ($\mathcal{S}$ in blue).
-Correlated noise in the form of stellar variability is simulated by modeling the covariance matrix of the signal with a GP ($\mathcal{V}$ in red) including a diagonal variance of $0.001^2$ corresponding to white noise ($\epsilon$ in light grey).
-This simulated signal is not intended to be physically realistic.}
-        \label{fig:app_principle_dataset}
-    \end{centering}
-\end{figure}
-\noindent Let $f$ be the simulated flux of a star sampled and arranged in the vector $\bm{f}$ associated to the vector of times $\bm{t}$, such that
-\begin{equation*}
-    \bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{C}),
-\end{equation*}
-i.e.
-that $\bm{f}$ is drawn from a GP of mean $\bm{\mu}$ and covariance matrix $\bm{C}$.
-In this equation, $\bm{\mu}$ is such that its $i$-th element is defined by  $\mu_i = \mathcal{T}(t_i) + \mathcal{S}(t_i)$ where $t_i$ is the $i$-th time of observation, $\mathcal{T}$ is a periodic transit function and $\mathcal{S}$ a function describing the instrumental part of the signal (both described below).
-The covariance matrix $\bm{C}$ is built such that $C_{i, j} = k(t_i, t_j)$ where $k$ is a covariance function (or \textit{kernel}) accounting for correlated noise in the form of stellar variability (with added white noise).
-An example of such signal is simulated and shown in \autoref{fig:app_principle_dataset}.
-
-\subsection{Transit signal $\mathcal{T}$}\label{app_proto}
-The periodic transit signal $\mathcal{T}$ is simulated using the simple model described in \cite{protopapas}, where a transit of period $P$, epoch $T_0$, duration $D$ and unitary depth observed at time $t$ is given by
-\begin{equation}\label{eq:protopapas}
-    \begin{gathered}
-        \mathcal{T}_c(t, P, T_0, D) = \frac{1}{2}\tanh\left(c\left[\theta - \frac{1}{2}\right]\right) - \frac{1}{2}\tanh\left(c\left[\theta + \frac{1}{2}\right]\right), \\
-        \text{with}\quad\theta = \frac{P}{\pi  D}\sin\left(\frac{\pi(t-T_0)}{P}\right),
-    \end{gathered}
-\end{equation}
-where the dimensionless parameter $c$ controls the roundness of the transit depth ($c\gg1$ corresponding to a box-shaped transit as shown in \autoref{fig:protopapas}).
-This analytical model is fully empirical but easily differentiable.
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/principle/figures/protopapas.pdf}
-        \caption{Simulations of a single transit signal (\autoref{eq:protopapas}) shown for different values of $c$.}
-        \label{fig:protopapas}
-    \end{centering}
-\end{figure}
-\noindent In this paper, unless specified, all transits are simulated with $c=12$, a value arbitrarily chosen that can be fined-tuned in real applications using the limb-darkening coefficients of a given star.
-The periodic transit signal $\mathcal{T}$ seen in \autoref{fig:app_principle_dataset} corresponds to $\mathcal{T} = 0.02 \times \mathcal{T}_{c=12}(\bm{t}, P=0.7, T_0=0.2, D=0.05)$, all parameters in unit of days.
-
-\subsection{Instrumental signals $\mathcal{S}$}
-Instrumental signals are simulated as a linear model of $M$ explanatory variables arranged in the $(N\times M)$ design matrix $\bm{X}$. Hence,
-$$
-    \mathcal{S} = \bm{X w},
-$$
-where the vector $\bm{w}$ contains the linear coefficients of the model.
-The simulated flux shown in \autoref{fig:app_principle_dataset} contains a linear model where the $M=4$ columns of the design matrix $\bm{X}$ are given by $\bm{X}_{i} = \bm{t}^i$ (i.e.
-$\bm{X}$ is the Vandermonde matrix order $3$ of time $t$) and $\bm{w} = [1.0\quad0.0005\quad\text{-}0.0002\quad\text{-}0.0005].$
-
-\subsection{Stellar variability $\nu$}\label{app_gp}
+\section{SHO kernel}\label{app_gp}
 As this chapter focuses on stellar variability and its effect on transit detection, We employ a simple physically-motivated GP kernel, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) taking the form
@@
 GP computations in this paper use the implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package exposing the quasi-separable kernels from \cite{celerite2} and powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
-The stellar variability signal in \autoref{fig:app_principle_dataset} has been sampled from a GP with an SHO kernel of parameters $\omega = \pi/6D$ (i.e.
-a period equal to 12 times the duration $D$ of the simulated transit), $Q=45$ and $\sigma=0.02$, the depth of the simulated transit.
-An extra term $\sigma_f^2=0.001^2$ is added to the diagonal of the covariance matrix, corresponding to the variance of the simulated measurement $f$ and leading to the white noise observed in \autoref{fig:app_principle_dataset}.
 
@@
 \newcommand{\sumTk}{i\neq k}
-From the \textit{linear search} presented in \autoref{linear_search}, we retain and index by $k$ the parameters of the $K$ individual transits whose epochs $\{T_k\}_k$ are compatible with a periodic signal of period $P$ and epoch $T_0$.
+From the linear search presented in \autoref{linear_search}, we retain and index by $k$ the parameters of the $K$ individual transits whose epochs $\{T_k\}_k$ are compatible with a periodic signal of period $P$ and epoch $T_0$.
 From the likelihoods of these transits (computed in \autoref{linear_search}), we want an expression for

### 2024-01-30 latex/paper.tex
fix: adressing more minor comments

@@
 During the linear search, the likelihood $p(\bm{f} \vert T_0, D, \Delta)$ of a single non-periodic transit is computed for a grid of epochs and durations, each time solving for $\Delta$ linearly.
-Then, the periodic search consists in combining these likelihoods to compute the likelihood $p(\bm{f} \vert P, T_0, D, \Delta)$ of the data given a periodic transit signal for a range of periods $P$.
+Then, the periodic search consists of combining these likelihoods to compute the likelihood $p(\bm{f} \vert P, T_0, D, \Delta)$ of the data given a periodic transit signal for a range of periods $P$.
 These combined likelihoods yield $p(\bm{f} \vert P, \hat T_0 ,\hat D, \hat \Delta)$, a transit-search periodogram on which the periodic transit detection is based.
 \nuancemethod{} differs from \cite{foreman2016} and other existing transit search algorithms as it models the covariance of the light curve with a GP, accounting for correlated noise (especially in the form of stellar variability) while keeping the model linear and tractable.
-This way, \nuancemethod{} searches for transits while, at the same time, modeling correlated noise, avoiding the detrending step that degrades transit signals SNR (see \autoref{issues}).\\\\
+This way, \nuancemethod{} searches for transits while, at the same time, modeling correlated noise, avoiding the separate detrending step that degrades transit signals SNR (see \autoref{issues}).\\\\
 We note that the approach employed by \nuancemethod{} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.
@@
 In theory, given that all data points of the light curves are used to compute each likelihood separately, these are not.
-In practice however, this assumption is reasonable and only affect the results of the periodic search for extremely short-period transit signals (where $P$ is less than few durations) that are less likely to be present in physically-realistic datasets.\\\\
+In practice however, this assumption is reasonable and only affect the results of the periodic search for extremely short-period transit signals (where $P$ is less than few times the duration $D$) that are less likely to be present in physically-realistic datasets.\\\\
 While \autoref{eq:result} takes a closed form, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$.
@@
 Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving $N\times K$ measurements.
-This lead to a normalization issue when trying to compare the joint log-likelihoods $\mathcal{P}(P)$ from one period to another, as the number of observed transits differs from one period to another.
+This leads to a normalization issue when comparing the joint log-likelihoods $\mathcal{P}(P)$ from one period to another, as the number of observed transits differs from one period to another.
 This motivates a final step to produce the transit search periodogram $\mathcal{Q}$.
@@
 From this \texttt{search} object, the best transiting candidate parameters can be computed (\textsf{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\textsf{search.Q\_snr}), together with valuable information about the transit search.
-The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary hosts, the advantage of \nuancemethod{} being that the linear search only needs to be performed ones and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
+The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary hosts, the advantage of \nuancemethod{} being that the linear search only needs to be performed once and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
 An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.
@@
 
-To start testing \nuancecode{} against existing methods, a simple adimensional normalized light curve is simulated, consisting in pure white noise with a standard deviation of $5\times 10^{-4}$ spanning 6 days with an exposure time of 2 minutes.
+To start testing \nuancecode{} against existing methods, a simple adimensional normalized light curve is simulated, consisting of pure white noise with a standard deviation of $5\times 10^{-4}$ spanning 6 days with an exposure time of 2 minutes.
 From this signal, we produce 4000 light curves, each containing box-shaped transits with periods randomly sampled from 0.3 to 2.5 days, durations of 50 minutes, and depths randomly sampled to lead to transit SNRs ranging from 4 to 30.
@@
 \newpage
-\section{Performances}\label{results}
+\section{Performance}\label{results}
 \subsection{Comparison with \texttt{biweight+BLS} on simulated light curves}\label{simu}

### 2024-01-30 latex/paper.tex
fix: update paper and add reviews

@@
 \usepackage{float}
+\usepackage{soul}
 
@@
 \newcommand{\footlink}[1]{\footnote{\url{#1}}}
-\newcommand{\wtls}{\texttt{biweight+BLS}}
+\newcommand{\wtls}{\texttt{bi-weight+BLS}}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 \affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
+\author{Catriona A. Murray}
+\affiliation{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, CO 80309, USA}
+\affiliation{Center for Astrophysics and Space Astronomy, University of Colorado, 593 UCB, Boulder, CO 80309}
+\author{Suzanne Aigrain}
+\affiliation{Sub-department of Astrophysics, Department of Physics, University of Oxford, Oxford, OX1 3RH, UK}
 \author{Dax L. Feliz}
-\author{Catriona A. Murray}
+\affiliation{American Museum of Natural History, 200 Central Park West, Manhattan, NY 10024, USA}
 \author{Francisco J. Pozuelos}
+\affiliation{Instituto de Astrofísica de Andalucía (IAA-CSIC), Glorieta de la
+Astronomía s/n, 18008 Granada, Spain}
 \correspondingauthor{Lionel J. Garcia}
@@
 This dataset corresponds to an observation of 4 days with an exposure time of 2 minutes.
-The mean of this signal consists in a periodic transit signal of period $P=0.7$ days, duration $D=1.2$ hours and depth $\Delta=0.02$ (dark gray) plus an instrumental signal (blue).
+The mean of this signal consists in a periodic transit signal of period $P=0.7$ days, duration $D=1.2$ hours and depth $\Delta=2\%$ (dark gray) plus an instrumental signal (blue).
 Correlated noise in the form of stellar variability is simulated with an SHO kernel of hyperparameters $\omega = \pi / 6 D$, $Q=45$ and $\sigma = \Delta$.
-Finally, white noise with a standard deviation of $0.001$ is added to the diagonal of the covariance matrix.}
+Finally, white noise with a standard deviation of $0.1\%$ is added to the diagonal of the covariance matrix.}
         \label{fig:app_principle_dataset}
@@
         \caption{Illustration of the effect of correlated noise on a single transit SNR.
-An hour-long transit of depth 0.01 is generated on top of white noise (standard deviation of 0.0015) as part of a 24-hours observation with an exposure time of 1 minute (top).
+An hour-long transit of depth 0.5\% is generated on top of white noise (standard deviation of 0.15\%) as part of a 24-hours observation with an exposure time of 1 minute (top).
 Then, in the bottom plot, correlated noise is added, generated using a GP with an SHO kernel of hyperparameters $\omega=60$, $Q=0.5$ and $\sigma=\Delta/4$ .
 The SNR on the right of each light curve is computed using \autoref{eq:snr}.
-Models used to simulate these data are provided in \autoref{signals_simulations}.}
+Models used to simulate these data are provided in \autoref{light_curves_simulations}.}
         \label{fig:issue1}
@@
         \includegraphics[width=\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
-        \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.08$, duration 0.1 day, and white noise with a standard deviation of 0.015 (top left).
-The hyperparameters of the kernel are fixed to $Q=10$ and $\sigma=0.008$, with the pulsation $\omega$ increasing from top to bottom.
+        \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.8\%$, duration 0.1 day, and white noise with a standard deviation of 0.15\% (top left).
+The hyperparameters of the kernel are fixed to $Q=10$ and $\sigma=0.8\%$, with the pulsation $\omega$ increasing from top to bottom.
 The resulting stellar variability signals are then reconstructed using (in purple) Tukey's bi-weight filter with an optimal window size of $3\times D$  (see \citealt{wotan}), and (in red) a GP with the same kernel used to simulate the data.
@@
 \end{figure}
-To explore the parameter space for which detrending is the most problematic, we employ the model described in \autoref{signals_simulations} to simulate 10 000 differential light curves with different morphological characteristics, and compute the remaining transit SNR after detrending.
+To explore the parameter space for which detrending is the most problematic, we employ the model described in \autoref{light_curves_simulations} to simulate 10 000 differential light curves with different morphological characteristics, and compute the remaining transit SNR after detrending.
 In order to place the stellar variability hyperparameters on a relative scale with the transit signal parameters, we reparametrize the GP kernel with
@@
 \end{equation}
-where $\set{T_k}_k$ are the epochs matching $(T_0, P)$ and $\set{\Delta_k}_k$ the corresponding depths. So that
+where $\set{T_k}_k$ are the epochs matching $(T_0, P)$ and $\set{\Delta_k}_k$ the corresponding depths, so that
 \begin{equation*}
@@
 \section{Performance}\label{results}
-\subsection{Comparison with \texttt{biweight+BLS} on simulated light curves}\label{simu}
 \autoref{fig:snr_detrend} shows that \textsf{nuance}'s full-fledged modeling capabilities may not always be necessary and may only be beneficial for certain noise characteristics, relative to the searched transit parameters.
-Here, we evaluate the performance of \textsf{nuance} in the relative parameter space $(\tau, \delta)$ described in \autoref{eq:relative_params} and see when its specific treatment of correlated noise in the transit search becomes necessary.\\\\
-We compare \textsf{nuance} to the approach that involves removing stellar variability from light curves before performing the search on a detrended dataset.
-Similarly to \autoref{detrending_effect}, an optimal bi-weight filter implemented in the \textsf{wõtan} Python package\footlink{https://github.com/hippke/wotan} is used with a window size three times that of the injected transit duration.
-A transit search on the detrended light curve is then performed using the BLS algorithm, like in \autoref{control} using the \textsf{astropy} \textsf{BoxLeastSquares} implementation\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html}; a strategy commonly used in the literature that we denote \texttt{biweight+BLS}.
-In what follows, the transit detection criteria are the same as the ones used in \autoref{control}, i.e.
-that a transit is considered recovered if the absolute difference between the injected and recovered period is less than 0.01 days, including periods found at half or twice the injected ones (aliases) and ignoring the value of the recovered transit epoch.
-
-\subsubsection*{Dataset}\label{sim_dataset}
-The dataset consists in 4000 light curves simulated using the model fully described in \autoref{signals_simulations}.
+Here, we evaluate the performance of \textsf{nuance} in the relative parameter space $(\tau, \delta)$ described in \autoref{eq:relative_params} and probe when its specific treatment of correlated noise in the transit search becomes necessary.\\\\
+We perform this study by comparing \textsf{nuance} to the approach that involves removing stellar variability from light curves before performing the search on a detrended dataset.
+The following detrending strategies, each followed by a search with the BLS algorithm, are compared:\\\\
+\ul{\texttt{bi-weight+BLS}}: employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package with an optimal window size of $3\times D$, i.e.\, three times the transit duration (e.g.\;\citealt{wotan} and \citealt{Dransfield2024}).\\\\
+\ul{\texttt{GP+BLS}}: employs a GP conditioned on the data (e.g.\;\citealt{Lienhard2020}). The kernel of the GP and its optimization is described on a case-by-case basis.\\\\
+\ul{\texttt{Bspline+BLS}}: employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html} (e.g.\;\citealt{wotan} and \citealt{Canocchi2023}).\\\\
+\ul{\texttt{harmonics+BLS}}: employs a linear harmonic detrending, where the light curve is modeled as a Fourier series including four harmonics of the stellar rotation period with coefficients found through ordinary least square.\\\\
+\ul{\texttt{iterative+BLS}} iteratively detrends the light curve with a sinusoidal signal fitted to the data, each time using the dominant period of the residuals found using a lomb-scargle periodogram (5 iterations).\\\\
+Like in previous sections, we consider the planet detected if the recovered period is within 0.01 day of the true period or a direct alias (such as two times or half the true period).
+Again, we ignore the exact match between the injected and recovered transit epochs, although we visually vet that the found epochs are consistent with the ones injected.
+We consider a transit detectable if its original SNR is greater than 6.
+Hence we define \textit{true positives} has detectable transits recovered with the correct period (or an alias) and a measured SNR greater or equal to 6, and \textit{false positives} as non-detectable transits recovered with a measured SNR greater or equal to 6.\\\\As we noticed that few methods were still affected by the remaining stellar variability after detrending, the grid of orbital periods being searched for does not contain the stellar rotation period $P_*$ and its aliases.
+In practice this is done by removing all orbital periods $P$ from the search grid such that $dP = \frac{P}{P_*}$ is less than 2\% from an integer value, i.e.
+$\vert  dP - \lceil dP \rceil \vert < 0.02$.\\\\
+\subsection{Comparisons on simulated light curves}\label{simu}
+Our first comparison dataset consists in 4000 light curves simulated using the model described in \autoref{light_curves_simulations}.
 We simulate a common periodic transit added to all light curves, of period $P=1.1$ days, epoch $T_0=0.2$ days, duration $D=0.04$ days and depth $\Delta=1\%$.
-Each light curve consists in a 4 days observation with an exposure time of 2 minutes, leading to $N=2880$ data points with a normal error of $0.1\%$.
-As in \autoref{control}, only box-shaped transits are injected, using the model from \cite{protopapas} with $c=500$ (cf.\;\autoref{app_proto}).\\\\
-For a given pair of $(\tau, \delta)$, we simulate stellar variability using a GP with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params} (except for $Q$, described below), computed with respect to the injected transit parameters $D$ and $\Delta$.
-The same kernel is used for the search with \textsf{nuance}, an optimal choice on equal footing with the optimal $3\times D$ window size of the bi-weight filter employed in the \wtls{} search.
+Each light curve consists in a 4 days observation with an exposure time of 2 minutes, leading to $N=2880$ data points with a normal error of $0.1\%$.\\\\
+For a given pair of $(\tau, \delta)$, we simulate stellar variability using a GP with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params} computed with respect to the injected transit parameters $D$ and $\Delta$.
+The same kernel is used for the search with \textsf{nuance} and with the \texttt{GP+BLS} method, an optimal choice on equal footing with the optimal $3\times D$ window size of the bi-weight filter employed in the \wtls{} search.
 %  While this kernel is optimal, a comprehensive discussion on the use of non-optimal kernels is discussed in . 
- 4000 pairs of $(\tau, \delta)$ are generated such that
+The 4000 pairs of $(\tau, \delta)$ are drawn from
 \begin{equation*}
@@
 \end{equation*}
-where $\mathcal{U}(a, b)$ denotes a uniform distribution of lower bound $a$ and upper bound $b$.
-
+where $\mathcal{U}(a, b)$ denotes a uniform distribution of lower bound $a$ and upper bound $b$.\\\\
 \subsubsection*{Results}
-The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against the \wtls{} strategy on transits with relatively large depth compared to the stellar variability amplitude, and a relatively large duration compared to the stellar variability period.
-This empirical statement only concerns light curves with a given amount of white noise, and may vary depending on the length of the observing window or the number of transits.
-For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperforms \wtls{} would only apply to this specific but representative example.
+The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against most other methods.
+Except for the \texttt{GP+BLS} approach, \nuancecode{} lead to a much higher rate of true positives for transits with relatively small depths compared to the stellar variability amplitude, and a relatively small duration compared to the stellar variability period.
+On the other hand, the \texttt{GP+BLS} strategy seems almost as performant as \nuancecode, recovering most of the injected transits and only lacking detection in a relatively comparable portion of the $(\tau, \delta)$ parameter space.
+We note that these empirical statements only concerns simulated light curves with a given amount of white noise, and may vary depending on the length of the observing window or the number of transits.
+For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperforms these methods would only apply to this specific but representative example.
 However, we verify that our conclusions remain qualitatively valid for various simulation setups.
@@
         \includegraphics[width=\linewidth]{../workflows/synthetic-injection-recovery/figures/synthetic_ir.pdf}
-        \caption{Injection-recovery results on 4000 simulated light curves described in \autoref{sim_dataset}.
-The color scale represents the fraction of recovered transits, white if all injected transits are recovered in a given portion of the $(\tau, \delta)$ parameter space, black if none are recovered.}
+        \caption{Rate of true positives for the 4000 simulated light curves.
+The color scale represents the fraction of true positives, white if all injected transits are recovered in a given portion of the $(\tau, \delta)$ parameter space, black if none are recovered.}
         \label{fig:simu}
@@
 
-\noindent This injection-recovery is done in a particularly optimal setup, on light curves not all physically realistic and using an optimal GP kernel, hence demonstrating the performance of \textsf{nuance} only on a purely experimental basis.
-In the next section, we perform transits injection-recovery on real space-based light curves.
-
-\newpage
-\subsection{Comparison with various techniques on rapidly-rotating M dwarfs TESS light curves}\label{real}
+\noindent This injection-recovery is done in a particularly optimal setup, on simulated light curves not all physically realistic and using an optimal GP kernel, hence demonstrating the performance of \textsf{nuance} only on a purely experimental basis.
+In the next section, we perform transits injection-recovery on real space-based light curves.\\\\
+\subsection{Comparisons on rapidly-rotating M dwarfs TESS light curves}\label{real}
 
@@
 In order to assess the performance of \nuance{} on real datasets, we inject and recover transits into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}).
-The goal of this transit injection-recovery is to compare \nuance{} to other transit search strategies, and to understand the light curves characteristics for which it becomes beneficial.\\\\
 We focus this proof of concept on light curves from a list of 438 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{}.
-For each of the studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) of a single sector (the first being observed for each target) spanning on average 10 days.\\\\
-In this injection-recovery tests, \nuance{} is compared to other transit search strategies employing the BLS algorithm, for each of them after detrending the light curves with a different technique:
-\begin{itemize}
-    \item \texttt{bspline+BLS} employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html}, followed by a search with the BLS algorithm.
-    \item \texttt{biweight+BLS} employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package \citep{wotan} with an optimal window size of $3\times D$ (i.e.\, three times the transit duration), followed by a search with the BLS algorithm.
-    \item \texttt{harmonics+BLS} employs a linear harmonic detrending, where the light curve is modeled as a Fourier series including four harmonics of the stellar rotation period found by \cite{Ramsay2020}, with coefficients found through ordinary least square.
-This detrending is followed by a search with the BLS algorithm.
-    \item \texttt{iterative+BLS} iteratively detrends the light curve with a sinusoidal signal fitted to the data, each time using the dominant period of the residuals found using a lomb-scargle periodogram (5 iterations).
-This detrending is followed by a search with the BLS algorithm.\\\\
-\end{itemize}
-
+For each studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) of a single sector (the first being observed for each target) spanning on average 10 days.
+To our knowledge, none of these 438 targets have been searched for planetary transits before.
+However, we note that the presence of existing transit signals in these light curves before the injection of simulated ones is possible, but will only affect our results on a relative basis.\\\\
 \subsubsection*{Light curve cleaning and transits injection}
-
-As some of the techniques compare to \nuance{} can be affected by gaps in the data, we only use continuous measurements from half a TESS sector.
-We assume that all methods (including \nuance{}) are based on an incomplete model of the data that do not account for stellar flares.
+As some of the techniques compared to \nuance{} can be affected by gaps in the data, we only use continuous measurements from half a TESS sector.
+We assume that all methods (including \nuance{}) are based on an incomplete model of the data that does not account for stellar flares.
 For this reason, the light curve of each target is cleaned using an iterative sigma clipping approach.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
+    \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692.
+The top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
+On this bottom plot, the black line corresponds to the mean of the GP model (with hyperparameters optimized here on the cleaned light curve).}
+    \label{fig:cleaned}
+\end{figure}
 For each iteration, points 3 times above the standard deviation of the full light curve (previously subtracted by its median) are identified.
 Then, the 30 adjacent points on each side of the found outliers are masked.
-This way, large flare signals and their expected ingress and egress are masked, using a total of 3 iterations.
+This way, large flare signals are masked, using a total of 3 iterations.
 At each iteration, the GP kernel hyperparameters are re-optimized.
@@
 We note that the gaps left after sigma clipping may be problematic for some of the detrending techniques (such as \texttt{bspline+BLS}).
-However, adopting this flare cleaning step and analyzing light curves with few small gaps is a practice commonly found in the literature.
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
-    \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692.
-Top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
-On this bottom plot, the black line corresponds to the mean of the GP model (with hyperparameters optimized here on the cleaned light curve).}
-    \label{fig:cleaned}
-\end{figure}
+However, adopting this flare cleaning step and analyzing light curves with few small gaps is a practice commonly found in the literature.\\\\
 For each light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injection-recovery per target.
@@
 \end{equation*}
-with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.
-
+with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.\\\\
 \subsubsection*{Stellar variability kernel}\label{rotation_kernel}
-One of the particularity of \nuance{} is to model a light curve as a GP.
-To model these TESS light curves, we employ a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
+In the \texttt{GP+BLS} approach and the \nuancecode{} algorithm, we model these TESS light curves using a GP with a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
 In order to account for additional corelated noises, we complement this kernel with a short and a long-timescale exponential term, so that the full kernel can be expressed as
@@
 In total, the rotation kernel $k$ has 8 hyperparameters.\\\\
-
 The hyperparameters of this kernel are optimized on trimmed and cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize}\footlink{https://jax.readthedocs.io/en/latest/_autosummary/jax.scipy.optimize.minimize.html} wrapper provided by the \textsf{jaxopt} Python package, and taking advantage of the \textsf{JAX} implementation of \textsf{tinygp} and its quasiseparable kernels\footlink{https://tinygp.readthedocs.io/en/latest/api/kernels.quasisep.html} \citep{celerite}.
@@
 For simplicity, and to adopt a uniform treatment for all target light curves, a design matrix $\bm{X}$ with a single constant column is adopted, such that the systematic model only consists in a single parameter corresponding to the mean value of the differential flux (expected to be close to 1) solved linearly.
-Our motivations to choose this very simplistic baseline, despite the capability of \nuancemethod{} to account for more complex linear models, is discussed in \autoref{systematics}.
-
+Our motivations to choose this very simplistic baseline, despite the capability of \nuancemethod{} to account for more complex linear models, is discussed in \autoref{systematics}.\\\\
 \subsubsection*{Search parameters and transit detection criteria}
@@
 multiple trial durations) is discussed in the next section.\\\\
-Like in previous sections, we consider the detection of period aliases and ignore the match between the injected and recovered transit epochs (although we visually vet that the found epochs were consistent with the ones injected for cases where the planet was detected).
-We consider a transit detectable if its original SNR is greater than 6.
-Hence we define \textit{true positives} has detectable transits recovered with the correct period (or an alias) and a measured SNR greater or equal to 6, and \textit{false positives} as non-detectable transits recovered with a measured SNR greater or equal to 6.\\\\As we noticed that few methods were still affected by the remaining stellar variability after detrending, we remove from the grid of trial periods all periods close to the stellar rotation period and its aliases.
-In practice this is done by removing all periods $P$ such that $dP = \frac{P}{P_*}$ is less than 2\% from an integer value, i.e.
-$\vert  dP - \lceil dP \rceil \vert < 0.02$.
-
-\newpage
 \subsubsection*{Results}
@@
 \end{figure}
-\begin{figure}[H]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp.pdf}
-        \caption{Rate of true and false positives of \nuance{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
-The solid red line corresponds to the maximum of true positives among all methods on the top panel, and the minimum of false positives on the bottom panel.}
-        \label{fig:allsearch_relative}
-    \end{centering}
-\end{figure}
 Compared to other techniques, we find that \nuancecode{} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits (\autoref{fig:allsearch}).
@@
     \begin{centering}
+        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp.pdf}
+        \caption{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
+The solid blue line corresponds to the maximum of true positives among all methods on the top panel (dominated by the \texttt{iterative+BLS} method), and the minimum of false positives on the bottom panel (waned by the \texttt{GP+BLS} method, all other methods being above \nuancecode{}).}
+        \label{fig:allsearch_relative}
+    \end{centering}
+\end{figure}
+\begin{figure}[H]
+    \begin{centering}
         \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_ims.pdf}
-        \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods.}
+        \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods and all targets.}
         \label{fig:allsearchim}
@@
 Here again, we employ an upper-sigma-clipping to mask flares out of the data.
-The resulting light curve for sector 4 and its mean model are shown in \autoref{fig:toi540_clean}.\\\\
-For each sector, we perform the linear search of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
-We then perform the periodic search on all sectors combined, using a concatenation of all linear searches.
-By adopting this by-sector GP modeling of the light-curve, the linear search of \nuance{} scales linearly with the number of sectors being processed.
+The resulting light curve for sector 4 and its mean model are shown in \autoref{fig:toi540_clean}.
 \begin{figure}[H]
@@
 \end{figure}
+For each sector, we perform the linear search of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
+We then perform the periodic search on all sectors combined, using a concatenation of all linear searches.
+By adopting this by-sector GP modeling of the light-curve, the linear search of \nuance{} scales linearly with the number of sectors being processed.
 \begin{figure}[H]
@@
         \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/periodograms.pdf} 
-        \caption{Transit search SNR periodograms of TOI-540 using \texttt{biweight+BLS} and \nuance{}.
+        \caption{Transit search SNR periodograms of TOI-540 using \texttt{bi-weight+BLS} and \nuance{}.
 After a first periodic search (left panel), the epochs corresponding to the maximum-SNR transit are masked before the second search is performed (right panel).}
@@
 The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days.
-This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{biweight+BLS} and described in \autoref{real}) on all sectors combined.
+This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{bi-weight+BLS} and described in \autoref{real}) on all sectors combined.
 Since we don't know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
@@
 In practice, this is done by masking the linear search products $\set{\ln\mathcal{L}_{i,j}}_{i, j}$, $\set{\Delta_{i,j}}_{i, j}$ and $\set{\sigma_{i,j}}_{i, j}$ (defined in \autoref{linear_search}).\\\\
-As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{biweight+BLS} and \texttt{nuance} are very similar, with the known transiting exoplanet TOI-540 b detected with an orbital period P = 1.24 days.
-This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b \footnote{using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ minutes of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial beneficial (see \autoref{fig:snr_detrend}).
+As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{bi-weight+BLS} and \texttt{nuance} are very similar, with the known transiting exoplanet TOI-540 b detected with an orbital period P = 1.24 days.
+This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b\footnote{using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ minutes of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial (see \autoref{fig:snr_detrend}).
 Nonetheless, the \nuance{} periodogram of TOI 540 features less spurious SNR peaks, largely due to the penalty naturally occurring when single transits with different depths are periodically combined.
-In the second search (right panel) of \autoref{fig:toi540_periodograms}, we also notice a higher number of peaks that would lead to false positive detections of transits in the \texttt{biweight+BLS} case.
+In the second search (right panel) of \autoref{fig:toi540_periodograms}, we also notice a higher number of peaks that would lead to false positive detections of transits in the \texttt{bi-weight+BLS} case.
 The proper treatment of correlated noise in \nuance{}, as observed in \autoref{real}, makes these peaks non-significant, avoiding a large number of false detections.\\\\
@@
 \noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple non-optimized GP with a squared exponential kernel.
-These are compared to the processing time of \texttt{biweight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the BLS search.
+These are compared to the processing time of \texttt{bi-weight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the BLS search.
 \begin{figure}[H]
@@
         \includegraphics[width=\linewidth]{../workflows/benchmark/figures/nuance_vs_bls.pdf}
-        \caption{Study of \textsf{nuance} (black line) processing time against \texttt{biweight+BLS} (cf.\;\autoref{simu}, blue line).
+        \caption{Study of \textsf{nuance} (black line) processing time against \texttt{bi-weight+BLS} (cf.\;\autoref{simu}, blue line).
 The gray curve shows the performance of \textsf{nuance} linear search when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together.
-This study is performed on the 12 cores of an Apple M2 Max chip, exploiting the parallelization offered by our implementation.
+This study is performed on a single CPU core of an Apple M2 Max chip.
 While not being shown, we verify that both the BLS algorithm and \nuancemethod{} processing times scale linearly with the number of probed transit durations and orbital periods.}
@@
 
-As see in \autoref{fig:blsvsnuance}, most of the computational costs of \nuancecode{} and the \texttt{biweight+BLS} method in this particular example come from the linear search and the bi-weight filtering step.
+As see in \autoref{fig:blsvsnuance}, most of the computational costs of \nuancecode{} and the \texttt{bi-weight+BLS} method in this particular example come from the linear search and the bi-weight filtering step.
 This is not always true and depends on the size of the trial durations and periods grids.
@@
 Hence, if searching for transits in separate observations with approximately similar durations, such as different TESS sectors or different ground-based nightly observations, the computational cost of \nuancecode{} grows linearly with the number of observations (see grey line in \autoref{fig:blsvsnuance}).
-Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform about one order of magnitude slower than more conventional techniques associated with BLS}.
-The second advantage of \nuancecode{} is its \textsf{JAX} implementation, allowing the linear search to be run in parallel.
-As a reference, searching for TOI-540 b transits with nuance (\autoref{toi540}), on the 12 cores of an Apple M2 Max chip, took 5 minutes 35 seconds in total, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds each), and 3 minutes 45 seconds for the combined periodic search.
-In comparison, the brute force search with \texttt{biweight+BLS}, consisting in trying 15 bi-weight windows, took a total of 1 minute 49 seconds (about only 7 seconds each).\\\\
+Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform about two order of magnitude slower than more conventional techniques associated with BLS}.\\\\
+As a reference, searching for TOI-540 b transits with nuance (\autoref{toi540}), on the 12 cores of an Apple M2 Max chip, took 5 minutes 35 seconds in total, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds per TESS sector), and 3 minutes 45 seconds for the combined periodic search.
+In comparison, the brute force search with \texttt{bi-weight+BLS}, consisting in trying 15 bi-weight windows, took a total of 1 minute 49 seconds (about only 7 seconds each).\\\\
 % As the periodic search is expected to be improved by lower lever optimizations (using a compiled language), we reserve its benchmark for a future work.
 Thanks to its \textsf{JAX} implementation, we also note the possibility of running \nuancecode{} on GPUs, which is out of the scope of the present comparison.\\\\
-Because of its computational cost, \nuancecode{} must not be employed in the general case, but rather adopted when light curves contain correlated noise with specific characteristics.
+Because of its computational cost, we do not recommend using \nuancecode{} in the general case, but rather adopted when light curves contain correlated noise with specific characteristics.
 If employing a bi-weight filter for detrending, these characteristics correspond to the ones discussed in \autoref{issues}.
-But as these strongly depend on the type of detrending technique employed, we do not provide general guidelines as when \nuancemethod{} should be preferred over a specific detrending technique.
+But as these strongly depend on the type of detrending technique employed (see \autoref{fig:simu}), we do not provide general guidelines as when \nuancemethod{} should be preferred over a specific detrending technique.
 To aid users in making an informed choice of algorithm, extensive benchmarks and guidelines are reserved to future developments and will be progressively shared on \nuancecode{}'s online documentation\footlink{https://nuance.readthedocs.io/en/latest/}.
@@
 Comparisons involving more complex design matrices would also be sensitive to the choice of linear components, and would have unwanted repercussions on their results.\\\\
-As an illustration, the NEMESIS pipeline \citep{nemesis} starts processing the differential light curves by employing a linear systematics detrending using a least square fit of the data with a reduced PLD basis, before smoothing the signal from stellar variability using an approach similar to the one employed in the \texttt{biweight+BLS} approach (cf.\;\autoref{real}), hence detrending the systematics with an incomplete model that does not account for stellar variability.
+As an illustration, the NEMESIS pipeline \citep{nemesis} starts processing the differential light curves by employing a linear systematics detrending using a least square fit of the data with a reduced PLD basis, before smoothing the signal from stellar variability using an approach similar to the one employed in the \texttt{bi-weight+BLS} approach (cf.\;\autoref{real}), hence detrending the systematics with an incomplete model that does not account for stellar variability.
 To account for stellar variability while fitting the linear systematics model to the data, a step further would be to use a GP, such as done in the EVEREST \citep{everest2} pipeline.
@@
 An alternative, recommended for more realistic blind searches, is to perform model comparison on well selected kernels, and to adapt the optimization strategy to each dataset.\\\\
-When using \nuancecode{} on TESS light curves for example, it must be noted that the observed light curve variability might originate from a contaminated photometric aperture, so that a physically-interpretable GP kernel representing a single star activity is not necessarily appropriate.
+When using \nuancecode{} on TESS light curves for example, it must be noted that the observed light curve variability might originate from contamination due to several nearby blended stars, so that a physically-interpretable GP kernel representing a single star activity is not necessarily appropriate.
 On the other hand, a single squared exponential GP kernel might also be sufficient for some applications.\\\\
@@
 In practice, it means that it can be very hard to actually verify the presence of transits found by \nuancemethod{} visually, so that transits may be detected but remain hidden in correlated noise.
-This is particularly true for stars displaying very high-frequency photometric pulsations (see the example in \autoref{fig:exocomets} where the detrended light curves dismissively follow the base model).
+This is particularly true for stars displaying very high-frequency photometric pulsations (see the example in \autoref{fig:exoco}).
 
@@
 While not being tested and benchmarked, \nuancecode{} already integrates this feature.
-An example of the detection of known transiting exocomets in $\beta$ Pictoris TESS light curves is shown in \autoref{fig:exocomets}.
+An example of the detection of known transiting exocomets in $\beta$ Pictoris TESS light curves is shown in \autoref{fig:exoco}.
     \item During the periodic search, no prior about the transit duration related to the orbital period of the planet was used.
@@
 \begin{figure}[H]    
-    \begin{centering}
     \includegraphics[width=\linewidth]{exocomet.pdf}
@@
 The maximum-SNR events are displayed in the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.}
-        \end{centering}
-        \label{fig:exocomets} 
+    \label{fig:exoco} 
 \end{figure}
 
-\newpage
 \section{Conclusion}\label{conclusion}
@@
 In this context, a conventional approach involves detrending a light curve before searching for transits using a Box-Least-Square algorithm.
-However, we show that the incomplete modeling of a light curve for its detrending, one that ignores the presence of transits, degrades their signal-to-noise-ratio down to the point of not being detectable.
+However, we show that this approach degrades transit signal-to-noise-ratios down to the point of not being detectable.
 Adopting a widely used detrending strategy, we explore the extent of this degradation on simulated light curves, and its dependance on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
@@
 \section{SHO kernel}\label{app_gp}
-As this chapter focuses on stellar variability and its effect on transit detection, We employ a simple physically-motivated GP kernel, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) taking the form
+In order to model stellar variability and its effect on transit detection, we employ a simple physically-motivated GP kernel, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) taking the form
 \begin{equation}
@@
 where $Q$ is the quality factor of the oscillator, $\omega$ its pulsation and $\sigma$ the amplitude of the kernel function.
-GP computations in this paper use the implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package exposing the quasi-separable kernels from \cite{celerite2} and powered by \texttt{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
+GP computations in this paper use the implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package exposing the quasi-separable kernels from \cite{celerite2} and powered by \textsf{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
 

### 2024-01-31 latex/paper.tex
fix: few comments solved

@@
     We present \nuancemethod{}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
-To deal with this noise, a common approach consists in cleaning a light curve from nuisance signals before searching for transits.
+To deal with this noise, a common approach consists of cleaning a light curve from nuisance signals before searching for transits.
 However, we show that this approach limit the detection efficiency of traditional techniques.
 As this effect depends on the correlated noise characteristics, we map the relevant parameter space for which transits are altered, and quantify the detection efficiency limits on a set of simulated light curves.
-On this synthetic dataset, we show that \nuancemethod{} largely outperforms the detection capabilities of a commonly used transit search technique, which consists in cleaning light curves from stellar variability with an optimal bi-weight filter before searching for transits using a box-least-square algorithm.
-Then, in order to assess the performance of \nuancemethod{} on real datasets, we inject and recover transits into the TESS light curves of 438 M dwarfs, and compare our method to 4 detrending techniques followed by a Box-Least-Square search.
-For transits with a duration exceeding one fifth of the stellar variability timescale, we show that \nuancemethod is the most performant in 94\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
+Then, in order to assess the performance of \nuancemethod{} on real datasets, we inject and recover transits into the TESS light curves of 438 M dwarfs, and compare our method to 5 detrending techniques followed by a Box-Least-Square search.
+Overall, we demonstrate that \nuancemethod{} is the most performant method in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
 Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
-Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of active stars and their transiting exoplanets, both using space-based and sparse ground-based observations.
+Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of active stars by increasing the yield of transit search surveys, both using space-based and sparse ground-based observations.
 \end{abstract}
@@
 This strategy is widely adopted by the community, both using physically-motivated systematic models like \cite{everest1, everest2}, or filtering techniques (\citealt{Jenkins2010}, \citealt{wotan}).
-However, when correlated noise starts resembling transits, this cleaning step (often referred to as \textit{detrending}) is believed to degrade their detectability \cite[see subsection 4.3 of][]{wotan}.
+However, when correlated noise starts resembling transits, this cleaning step, often called \textit{detrending}, is believed to degrade their detectability \cite[see subsection 4.3 of][]{wotan}.
 In this case, the only alternative to search for transits is to perform a full-fledged modeling of the light curve, including both transits and correlated noise, and to compute the likelihood of the data to the transit model on a wide parameter space, an approach largely avoided due to its intractable nature.
@@
 This include transits injected in synthetic datasets, but also in the TESS light curves of 438 rapidly-rotating M dwarfs.
-Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, and we conclude in \autoref{conclusion}.
-
+Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, before concluding in \autoref{conclusion}.
 \newpage
 \section{Motivation}\label{issues}
-
-A strong assumption when using the BLS algorithm to search for transits is that the searched dataset only contains transit signals and white noise, justifying the need for detrending.
+A strong assumption when using the \textsf{BLS} algorithm to search for transits is that the searched dataset only contains transit signals and white noise, justifying the need for detrending.
 In this section, we explore this effect by simulating light curves containing a transit signal and correlated noise in the form of stellar variability, and study how detrending the latter affects the transit signal detectability depending on the light curve morphological characteristics.
 Hence, we first present how transit light curves are simulated, using a stochastic model of stellar variability, and describe the detection metric we employ to quantify transit detectability in the presence of correlated noise.
-
 \subsection{Light curve simulations}\label{light_curves_simulations}
@@
 \end{equation*}
-i.e. that $\bm{f}$ is drawn from a GP of mean $\bm{\mu}$ and covariance matrix $\bm{\Sigma}$. For practical reasons, we model $\bm{\mu}$ as a linear combination of $M$ explanatory variables, such that
+i.e.\;that $\bm{f}$ is drawn from a GP of mean $\bm{\mu}$ and covariance matrix $\bm{\Sigma}$.
+For practical reasons, we model $\bm{\mu}$ as a linear combination of $M$ explanatory variables, such that
 \begin{equation}\label{eq:linear_model}
@@
 \end{equation}
-where the first $M-1$ columns of $\bm{X}$ are contemporaneous instrumental time series measurements, such as the position of the star on the detector or the sky background (e.g.\;\cite{}), and its last column is a box-shaped transit signal with a fixed epoch, duration and period.
+where the first $M-1$ columns of $\bm{X}$ are contemporaneous instrumental time series measurements, such as the position of the star on the detector or the sky background, and its last column is a box-shaped transit signal with a fixed epoch, duration and period.
 This way, the transit signal is part of the mean linear model, meaning that once the design matrix $\bm{X}$ is constructed the transit is only parametrized by its depth $\Delta$.\\\\
-As we are interested in correlated noise in the form of stellar variability, we choose a physically-motivated GP kernel to model the covariance matrix $\bm{\Sigma}$, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) parametrized by its quality factor $Q$, its pulsation $\omega$ and the amplitude of the kernel function $\sigma$ (the full expression of the kernel function is provided in \autoref{app_gp}).
+As we are interested in active stars whose fluxes feature correlated noise in the form of stellar variability, we choose a physically-motivated GP kernel to model the covariance matrix $\bm{\Sigma}$, describing stellar variability through the covariance of a stochastically-driven damped harmonic oscillator (SHO, \citealt{celerite, celerite2}) parametrized by its quality factor $Q$, its pulsation $\omega$ and the amplitude of the kernel function $\sigma$ (the full expression of the kernel function is provided in \autoref{app_gp}).
 This choice of mean and kernel function completely defines the GP, from which light curves with different levels of correlated noise can be drawn (see an example in \autoref{fig:app_principle_dataset}).
-
 \begin{figure}[H]
@@
         \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
-        \caption{Flux times series drawn from the GP model described in \autoref{light_curves_simulations}.
+        \caption{The top plot shows a flux time series of an active star drawn from the GP model described in \autoref{light_curves_simulations}.
 This dataset corresponds to an observation of 4 days with an exposure time of 2 minutes.
@@
 \end{figure}
-
 \subsection{Transit detectability}\label{transit_detectability}
@@
         \includegraphics[width=\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
-        \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.8\%$, duration 0.1 day, and white noise with a standard deviation of 0.15\% (top left).
+        \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.8\%$, duration $D=0.1$ day, and white noise with a standard deviation of 0.15\% (top left).
 The hyperparameters of the kernel are fixed to $Q=10$ and $\sigma=0.8\%$, with the pulsation $\omega$ increasing from top to bottom.
@@
 To explore the parameter space for which detrending is the most problematic, we employ the model described in \autoref{light_curves_simulations} to simulate 10 000 differential light curves with different morphological characteristics, and compute the remaining transit SNR after detrending.
-In order to place the stellar variability hyperparameters on a relative scale with the transit signal parameters, we reparametrize the GP kernel with
+In order to place the stellar variability hyperparameters on a relative scale with the transit signal parameters, we reparametrize the SHO kernel with
 \begin{equation}\label{eq:relative_params}
@@
 Hence, for $(\tau, \delta)=(1, 1)$, the expressions of $\omega$ and $\sigma$ given in \autoref{eq:relative_params} correspond to a variability signal with a period half that of the transit duration, and a correlated noise amplitude comparable to the transit depth, i.e.
-strongly resembling the simulated transit signal.
+strongly resembling the simulated transit signal.\\\\
+For each of the 10 000 light curve generated, we separately reconstruct and detrend  the variability signal using the two techniques employed in \autoref{fig:issue2}, i.e.\;one using an optimal bi-weight filter with a window size three times that of the transit duration \citep{wotan} and the other using a GP with an optimal kernel.
+We then compute the detrended transit SNR using \autoref{eq:snr}.
 \begin{figure}[H]
@@
 \end{figure}
-For each light curve, we separately reconstruct and detrend  the variability signal using the two techniques employed in \autoref{fig:issue2}, i.e.\;one using an optimal bi-weight filter with a window size three times that of the transit duration \citep{wotan} and the other using a GP with an optimal kernel.
-We then compute the detrended transit SNR using \autoref{eq:snr}.\\\\
 \autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 5$).
@@
 Hence, detrending makes transit search blind to many systems, especially when the widely adopted bi-weight filter is used.
-While this study should be extended to other detrending techniques, it highlights the need for a more informed transit search algorithm able to deal with correlated noise, at least if present in the form of stellar variability.
+We note that, outside this problematic parameter space, both techniques perform relatively well and such a degradation of the transit SNR should not be expected.
+Although this study should be extended to other detrending techniques, it highlights the need for a more informed transit search algorithm able to deal with correlated noise, at least if present in the form of stellar variability.
 
@@
 While this is tractable for few periods, it becomes highly untractable for the large set of trial periods required to properly sample a transit-search periodogram (in the order of tens of thousands).\\\\
-To remain tractable, \nuancemethod{} employs the strategy of \cite{foreman2016}, and separates the transit search into two components: the linear search and the periodic search.
+To remain tractable, \nuancemethod{} employs the strategy of \cite{foreman2016}\footnote{Also introduced in \cite{Aigrain2004}.}, and separates the transit search into two components: the linear search and the periodic search.
 During the linear search, the likelihood $p(\bm{f} \vert T_0, D, \Delta)$ of a single non-periodic transit is computed for a grid of epochs and durations, each time solving for $\Delta$ linearly.
@@
 In \cite{foreman2016}, a similar issue is solved by using the nearest neighbors in the epochs grid.
-Instead, to allow the efficient matrix computation of \autoref{eq:result}, the likelihood grid is linearly interpolated from $\set{T_i}_i$ to a common grid of transit phases $\set{\phi_i}_i$, leading to the periodic search log-likelihood
-$$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi, D)}_{i,j}$$
+Instead, to allow the efficient matrix computation of \autoref{eq:result}, we interpolate the likelihood grid from $\set{T_i}_i$ to a common grid of transit phases $\set{\phi_i}_i$, leading to the periodic search log-likelihood
+$$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i,j}$$
 shown for few periods in \autoref{fig:periodic_search} (b). In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the linear search using $\phi_i$, $D_j$ and $T_0 = 0$.
@@
         \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
-        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using BLS and \nuancecode{}.
+        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using \textsf{BLS} and \nuancecode{}.
 The color scale indicates the recovery of transits in the corresponding (period, depth) parameter space, white for a full recovery and black for no detection.}
@@
 
-These results demonstrate the qualitative match between the detection capabilities of \textsf{nuance} and BLS on light curves with no correlated noise.
+These results demonstrate the qualitative match between the detection capabilities of \textsf{nuance} and \textsf{BLS} on light curves with no correlated noise, where the \textsf{BLS} algorithm should be optimal.
 Explaining the subtle differences observed between the two methods when only white noise is present is beyond the scope of this paper, and we will assume that any differences observed in the following sections are due to the different treatments of correlated noise.
@@
 We perform this study by comparing \textsf{nuance} to the approach that involves removing stellar variability from light curves before performing the search on a detrended dataset.
-The following detrending strategies, each followed by a search with the BLS algorithm, are compared:\\\\
+The following detrending strategies, each followed by a search with the \textsf{BLS} algorithm, are compared:\\\\
 \ul{\texttt{bi-weight+BLS}}: employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package with an optimal window size of $3\times D$, i.e.\, three times the transit duration (e.g.\;\citealt{wotan} and \citealt{Dransfield2024}).\\\\
@@
 The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against most other methods.
-Except for the \texttt{GP+BLS} approach, \nuancecode{} lead to a much higher rate of true positives for transits with relatively small depths compared to the stellar variability amplitude, and a relatively small duration compared to the stellar variability period.
+Except for the \texttt{GP+BLS} approach, \nuancecode{} lead to a much higher rate of true positives for transits with relatively small depths compared to the stellar variability amplitude (i.e.\;$\delta>2$), and a duration comparable to the stellar variability period (i.e.\;$\tau<2$).
 On the other hand, the \texttt{GP+BLS} strategy seems almost as performant as \nuancecode, recovering most of the injected transits and only lacking detection in a relatively comparable portion of the $(\tau, \delta)$ parameter space.
@@
 To our knowledge, none of these 438 targets have been searched for planetary transits before.
-However, we note that the presence of existing transit signals in these light curves before the injection of simulated ones is possible, but will only affect our results on a relative basis.\\\\
+However, we note that the presence of existing transit signals in these light curves before the injection of simulated ones is possible, but will not affect the relative comparison of one method to another.\\\\
 \subsubsection*{Light curve cleaning and transits injection}
@@
 \subsubsection*{Stellar variability kernel}\label{rotation_kernel}
-In the \texttt{GP+BLS} approach and the \nuancecode{} algorithm, we model these TESS light curves using a GP with a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
+In the \texttt{GP+BLS} and \nuancecode{} methods we model light curves using a GP with a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
 In order to account for additional corelated noises, we complement this kernel with a short and a long-timescale exponential term, so that the full kernel can be expressed as
@@
 In total, the rotation kernel $k$ has 8 hyperparameters.\\\\
-The hyperparameters of this kernel are optimized on trimmed and cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize}\footlink{https://jax.readthedocs.io/en/latest/_autosummary/jax.scipy.optimize.minimize.html} wrapper provided by the \textsf{jaxopt} Python package, and taking advantage of the \textsf{JAX} implementation of \textsf{tinygp} and its quasiseparable kernels\footlink{https://tinygp.readthedocs.io/en/latest/api/kernels.quasisep.html} \citep{celerite}.
+The hyperparameters of this kernel are optimized on trimmed and cleaned light curves containing the injected transits, using the \textsf{scipy.optimize.minimize} wrapper provided by the \textsf{jaxopt} Python package\footlink{https://jaxopt.github.io/stable/_autosummary/jaxopt.ScipyMinimize.html}, and taking advantage of the \textsf{JAX} implementation of \textsf{tinygp} and its quasiseparable kernels\footlink{https://tinygp.readthedocs.io/en/latest/api/kernels.quasisep.html} \citep{celerite}.
 As correlated noise is expected to affect the light curve uncertainty estimates performed by SPOC, the diagonal of the full covariance matrix of the data (i.e.\,their uncertainty, assuming homoscedasticity) is held free, increasing the number of optimized parameters to 9.
@@
 A realistic transit search on a wider parameter space (e.g.
-multiple trial durations) is discussed in the next section.\\\\
+multiple trial durations) is discussed in the next section.
 \subsubsection*{Results}
@@
 \end{figure}
-Compared to other techniques, we find that \nuancecode{} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits (\autoref{fig:allsearch}).
+Compared to other techniques, we find that \nuancecode{} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (\autoref{fig:allsearch}).
 While the performances of other methods highly depend on the characteristics of the variability, \textbf{\nuancecode{} is the best technique in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positives} (\autoref{fig:allsearch}).
@@
 \end{figure}
+From \autoref{fig:allsearch}, we note that the performance of the \texttt{GP+BLS} method is much worse that what could be anticipated from the results presented in \autoref{fig:simu}, where the GP kernel was fully optimal (given it was also used to simulate the data).
+In that case, our kernel might not be optimal, whether because of its form or because of the values of its hyperparameters.
+In any case, the fact that the same kernel performs significantly better when used with \nuancecode{} shows that our method is less sensitive to the choice of kernel and its optimization compared to detrending with a GP.
 \begin{figure}[H]
@@
 \end{figure}
-
 \subsection{Comparison for a multi-sector TESS candidate: TOI-540}\label{toi540}
-In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b \citep{TOI540}.
-We download the 2 minutes cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
-Like in the previous section, we use a Lomb-Scargle periodogram and identify the 0.72 days rotation period of the star, that we use as an initial value to optimize the kernel described in \autoref{rotation_kernel} on each sector independently.
-Here again, we employ an upper-sigma-clipping to mask flares out of the data.
-The resulting light curve for sector 4 and its mean model are shown in \autoref{fig:toi540_clean}.
 \begin{figure}[H]
@@
 \end{figure}
+In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b \citep{TOI540}.
+We download the 2 minutes cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
+Like in the previous section, we use a Lomb-Scargle periodogram and identify the 0.72 days rotation period of the star, that we use as an initial value to optimize the kernel described in \autoref{rotation_kernel} on each sector independently.
+Here again, we employ an upper-sigma-clipping to mask flares out of the data.
+The resulting light curve for sector 4 and its mean model are shown in \autoref{fig:toi540_clean}.\\\\
 For each sector, we perform the linear search of nuance on the cleaned light curve, using the original times as the trial epochs and 10 trial transit durations linearly sampled from 15 minutes to 1.5 hours.
@@
 This approach is adopted for efficiency but also to encapsulate the changing properties of stellar variability from one sector to another, often separated by year-long gaps.
-The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days.
-This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{bi-weight+BLS} and described in \autoref{real}) on all sectors combined.
-Since we don't know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
+The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days\footnote{We acknowledge that the grid of trial periods and durations used in this study is non-optimal for a real transit search.
+However, this simple choice is sufficient for our periodogram comparison against BLS.
+Optimal grids of parameters for realistic transit searches are discussed in \cite{tls}.}.
+This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the \textsf{BLS} algorithm (denoted \texttt{bi-weight+BLS} and described in \autoref{real}) on all sectors combined.
+Since we don't know the transit duration a-priori, we perform the detrending and \textsf{BLS} search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
 The results of this comparison are shown in \autoref{fig:toi540_periodograms}.\\\\
@@
 \noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple non-optimized GP with a squared exponential kernel.
-These are compared to the processing time of \texttt{bi-weight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the BLS search.
+These are compared to the processing time of \texttt{bi-weight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the \textsf{BLS} search.
 \begin{figure}[H]
@@
 This study is performed on a single CPU core of an Apple M2 Max chip.
-While not being shown, we verify that both the BLS algorithm and \nuancemethod{} processing times scale linearly with the number of probed transit durations and orbital periods.}
+While not being shown, we verify that both the \textsf{BLS} algorithm and \nuancemethod{} processing times scale linearly with the number of trial transit durations and orbital periods.}
         \label{fig:blsvsnuance}
@@
 
-As see in \autoref{fig:blsvsnuance}, most of the computational costs of \nuancecode{} and the \texttt{bi-weight+BLS} method in this particular example come from the linear search and the bi-weight filtering step.
+As see in \autoref{fig:blsvsnuance}, most of the computational costs of \nuancecode{} and the \texttt{bi-weight+BLS} method come from the linear search and the bi-weight filtering step.
 This is not always true and depends on the size of the trial durations and periods grids.
@@
 Hence, if searching for transits in separate observations with approximately similar durations, such as different TESS sectors or different ground-based nightly observations, the computational cost of \nuancecode{} grows linearly with the number of observations (see grey line in \autoref{fig:blsvsnuance}).
-Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform about two order of magnitude slower than more conventional techniques associated with BLS}.\\\\
-As a reference, searching for TOI-540 b transits with nuance (\autoref{toi540}), on the 12 cores of an Apple M2 Max chip, took 5 minutes 35 seconds in total, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds per TESS sector), and 3 minutes 45 seconds for the combined periodic search.
-In comparison, the brute force search with \texttt{bi-weight+BLS}, consisting in trying 15 bi-weight windows, took a total of 1 minute 49 seconds (about only 7 seconds each).\\\\
+Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform between one and two order of magnitude slower than more conventional techniques associated with BLS}.\\\\
+As a reference, searching for TOI-540 b transits with the parallelized implementation of \nuancecode{} (see \autoref{toi540}) on the 12 cores of an Apple M2 Max chip, took 5 minutes 35 seconds in total, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds per TESS sector), and 3 minutes 45 seconds for the combined periodic search.
+In comparison, the brute force search with \texttt{bi-weight+BLS}, consisting in trying 15 bi-weight windows, took a total of 1 minute 49 seconds (only 7 seconds each).\\\\
 % As the periodic search is expected to be improved by lower lever optimizations (using a compiled language), we reserve its benchmark for a future work.
 Thanks to its \textsf{JAX} implementation, we also note the possibility of running \nuancecode{} on GPUs, which is out of the scope of the present comparison.\\\\
-Because of its computational cost, we do not recommend using \nuancecode{} in the general case, but rather adopted when light curves contain correlated noise with specific characteristics.
+Because of its computational cost, we do not recommend using \nuancecode{} in the general case, but rather when light curves contain correlated noise with specific characteristics.
 If employing a bi-weight filter for detrending, these characteristics correspond to the ones discussed in \autoref{issues}.
@@
 However this would also involve some potential degradation of the transit signals (see e.g.\;\autoref{fig:issue2}), with an hardly distinguishable origin.
-For these reasons, and to keep our comparisons as targeted as possible, we do not compare commonly used systematics detrending approaches and decided to focus our comparisons solely on stellar variability detrending techniques (although these two aspects often overlap in the literature).\\\\
+For these reasons, and to keep our comparisons as targeted as possible, we do not compare commonly used systematics detrending approaches and decided to focus our comparisons solely on stellar variability detrending techniques (although these two aspects often overlap in the literature, e.g.
+in \citealt{everest1}).\\\\
 Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsly is extremely promising for the search of transiting exoplanets in ground-based data, that usually suffer from daily interruptions.
@@
 \subsection{The GP kernel}
-While not being discussed in our study, \textbf{the efficiency of \nuancemethod{} to detect transits in correlated noise is highly dependant on the design of its GP kernel}.
+While not being discussed in our study, \textbf{the efficiency of \nuancemethod{} to detect transits in correlated noise might be dependant on the design of its GP kernel}.
 In the ensemble comparison of \autoref{real}, the goal was to choose a kernel and an optimization strategy suited to most of the studied light curves, leading to few outliers in the results that were indicative of a badly designed and/or optimized kernel.
@@
 When using \nuancecode{} on TESS light curves for example, it must be noted that the observed light curve variability might originate from contamination due to several nearby blended stars, so that a physically-interpretable GP kernel representing a single star activity is not necessarily appropriate.
-On the other hand, a single squared exponential GP kernel might also be sufficient for some applications.\\\\
+On the other hand, a single squared exponential GP kernel might also be sufficient for some applications; an aspect we intend to explore in future applications.\\\\
 Something to emphasize is that \nuancemethod{} cannot be used to produce reliable detrended light curves, as an optimal GP is often flexible enough to partially model transits (see \autoref{fig:issue2}).
@@
 
+\newpage
 \section{Conclusion}\label{conclusion}
@@
 However, we show that this approach degrades transit signal-to-noise-ratios down to the point of not being detectable.
-Adopting a widely used detrending strategy, we explore the extent of this degradation on simulated light curves, and its dependance on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
+Adopting commonly used detrending strategies, we explore the extent of this degradation on simulated light curves, and its dependance on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
 The effectiveness of \nuancemethod{} is tested using a synthetic dataset and further validated on real TESS light curves of 438 rapidly-rotating M dwarfs.
-These injection-recovery tests reveal that \nuancemethod{} consistently outperforms conventional transit search techniques, especially for transits with a duration exceeding one-fifth of the stellar variability timescale.
-In all cases, \nuancemethod{} not only lead to a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
+These injection-recovery tests reveal that \nuancemethod{} consistently outperforms commonly used transit search techniques, especially when the timescale of stellar variability is less than twice that of the transit duration.
+In all cases, \nuancemethod{} not only leads to a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
 We make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package, developped with \textsf{JAX} to allow its  use on distributed computing environments and GPU devices.

### 2024-02-09 latex/paper.tex
feat: fist last review comments

@@
 \author{Catriona A. Murray}
-\affiliation{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, CO 80309, USA}
-\affiliation{Center for Astrophysics and Space Astronomy, University of Colorado, 593 UCB, Boulder, CO 80309}
+\affiliation{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, USA}
 \author{Suzanne Aigrain}
-\affiliation{Sub-department of Astrophysics, Department of Physics, University of Oxford, Oxford, OX1 3RH, UK}
+\affiliation{Sub-department of Astrophysics, Department of Physics, University of Oxford, Oxford, UK}
 \author{Dax L. Feliz}
-\affiliation{American Museum of Natural History, 200 Central Park West, Manhattan, NY 10024, USA}
+\affiliation{American Museum of Natural History, New York, NY, USA}
 \author{Francisco J. Pozuelos}
-\affiliation{Instituto de Astrofísica de Andalucía (IAA-CSIC), Glorieta de la
-Astronomía s/n, 18008 Granada, Spain}
+\affiliation{Instituto de Astrofísica de Andalucía, Granada, Spain}
 \correspondingauthor{Lionel J. Garcia}
@@
 
-\keywords{transit search, exoplanets, stellar activity, time series analysis, Gaussian processes}
+\keywords{transit search, exoplanets, active stars, stellar variability, time series analysis, Gaussian process, TESS}
 
 \begin{abstract}
-    We present \nuancemethod{}, an algorithm to search for planetary transits in light curves featuring correlated noise, such as instrumental signals and stellar photometric variability.
-To deal with this noise, a common approach consists of cleaning a light curve from nuisance signals before searching for transits.
-However, we show that this approach limit the detection efficiency of traditional techniques.
-As this effect depends on the correlated noise characteristics, we map the relevant parameter space for which transits are altered, and quantify the detection efficiency limits on a set of simulated light curves.
-Then, in order to assess the performance of \nuancemethod{} on real datasets, we inject and recover transits into the TESS light curves of 438 M dwarfs, and compare our method to 5 detrending techniques followed by a Box-Least-Square search.
+    The detection of planetary transits in the light curves of active stars, featuring correlated noise in the form of stellar variability, remains a challenge.
+Depending on the noise characteristics, we show that the traditional technique that consists of detrending a light curve before searching for transits alters their signal-to-noise ratio, and hinders our capability to discover exoplanets transiting rapidly-rotating active stars.
+We present \nuancemethod{}, an algorithm to search for transits in light curves while simultaneously accounting for the presence of correlated noise, such as stellar variability and instrumental signals.
+We assess the performance of \nuancemethod{} on simulated light curves as well as on the TESS light curves of 438 rapidly-rotating M dwarfs.
+For each dataset, we compare our method to 5 commonly-used detrending techniques followed by a search with the Box-Least-Square algorithm.
 Overall, we demonstrate that \nuancemethod{} is the most performant method in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
-Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
-Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of active stars by increasing the yield of transit search surveys, both using space-based and sparse ground-based observations.
+Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and availa ble as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
+Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of planetary systems around active stars, both using space-based surveys and sparse ground-based observations.
 \end{abstract}
@@
 \section*{Introduction}
-Transiting exoplanets are keystone objects for the field of exoplanetary science, but detecting transits in light curves featuring stellar variability and instrumental signals remains a challenge \citep{Pont2006,Howell2016}.
+Transiting exoplanets are keystone objects for the field of exoplanetary science, but detecting transits in light curves featuring stellar variability and instrumental signals remains a challenge (e.g.
+\citealt{Pont2006}, \citealt{Howell2016} or \citealt{Yaptangco2024}).
 For this reason, known transiting exoplanets tend to be found around quieter stars, or belong to the population of close-in giants whose transit signals dominate over stellar rotational variability \citep{Simpson2023}.
@@
 Overall, enabling the detection of transits in light curves with high levels of correlated noises will greatly benefit the study of terrestrial exoplanets around late M-dwarfs, usually observed at lower SNR and more likely to display photometric variability (e.g.
-\citealt{Murray2020}).
+\citealt{Murray2020} or \citealt{Petrucci2024}).
 \\\\
@@
 The latter study explores a handful of cases and generally discards the benefit of using a full-fledged approach.\\\\
-In this paper, we identify regions of light curves morphological parameter space for which a full-fledged transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} written in italics refers to the algorithmic and analytical method, while \nuancecode{} in sans-serif refers to its implementation.}, a method to search for transit signals while simultaneously modeling correlated noises in a tractable way.
+\newpage
+In this paper, we identify regions of light curves morphological parameter space for which a full-fledged transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} written in italics refers to the algorithm, while \nuancecode{} in sans-serif refers to its implementation published as a Python package in its version 0.5.2 (see \autoref{package}).}, a method to search for transit signals while simultaneously modeling correlated noises in a tractable way.
 In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on transit signals detectability.
@@
 \section{Motivation}\label{issues}
-A strong assumption when using the \textsf{BLS} algorithm to search for transits is that the searched dataset only contains transit signals and white noise, justifying the need for detrending.
+A strong assumption when using the BLS algorithm to search for transits is that the searched dataset only contains transit signals and white noise, justifying the need for detrending.
 In this section, we explore this effect by simulating light curves containing a transit signal and correlated noise in the form of stellar variability, and study how detrending the latter affects the transit signal detectability depending on the light curve morphological characteristics.
@@
 To perform this study, we want to simulate realistic light curves that contain instrumental signals, transit signals and correlated noise with characteristics that can be controlled using a set of interpretable parameters.
-To this end, we model light curves as realizations of a Gaussian Process (GP; \citealt{Rasmussen2005}) with a mean containing the instrumental and transit signals, and a kernel allowing to model different forms of correlated noise controlled by its hyperparameters.\\\\
+To this end, we model light curves as realizations of a Gaussian process (GP; \citealt{Rasmussen2005}, \citealt{Aigrain2023}) with a mean containing the instrumental and transit signals, and a kernel allowing to model different forms of correlated noise controlled by its hyperparameters.\\\\
 Let $f$ be the simulated differential flux of a star sampled and arranged in the vector $\bm{f}$ associated to the vector of times $\bm{t}$, such that
@@
 \end{equation}
-where the first $M-1$ columns of $\bm{X}$ are contemporaneous instrumental time series measurements, such as the position of the star on the detector or the sky background, and its last column is a box-shaped transit signal with a fixed epoch, duration and period.
+where the first $M-1$ columns of $\bm{X}$ are contemporaneous instrumental time series measurements, such as the position of the star on the detector, the sky background, co-trending basis vectors, or any other explanatory variables.
+The last column of $\bm{X}$ is a box-shaped transit signal with a fixed epoch, duration and period.
 This way, the transit signal is part of the mean linear model, meaning that once the design matrix $\bm{X}$ is constructed the transit is only parametrized by its depth $\Delta$.\\\\
@@
         \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
-        \caption{The top plot shows a flux time series of an active star drawn from the GP model described in \autoref{light_curves_simulations}.
+        \caption{The top plot shows the simulated flux time series of an active star drawn from the GP model described in \autoref{light_curves_simulations}.
 This dataset corresponds to an observation of 4 days with an exposure time of 2 minutes.
@@
 However, instrumental and other astrophysical signals will necessarily affect the detectability of transits in realistic light curves \citep{Pont2006}.
-For this reason, the combined differential photometric precision metric was developed (CDPP; \citealt{Jenkins2010}), and used in the context of the Kepler mission to assess the level of correlated noise in light curves, affecting the detectability of transits with a given duration.
+For this reason, the combined differential photometric precision (CDPP; \citealt{Jenkins2010}) metric was developed, and used in the context of the Kepler mission to assess the level of correlated noise in light curves, affecting the detectability of transits with a given duration.
 % Ultimately, this metric was essential to estimate the yield of the survey, and helped to achieve one of the primary goal of the mission: to measure the occurrence rate of Earth-size planets orbiting in the habitable zones of sun-like stars \citep{}.
@@
 For simplicity, the CDPP is computed using the method from \cite{Gilliland2011} implemented in the \textsf{lightkurve} Python package \citep{Lightkurve} and $\hat{\Delta}$ is the depth obtained by solving \autoref{eq:linear_model} with the last column of $\bm{X}$ containing a normalized transit model.
-\autoref{fig:issue1} shows the SNR from \autoref{eq:snr} computed for a unique transit observed in the absence (grey) and presence of correlated noise (red).
 \begin{figure}[H]
@@
 \end{figure}
-
-As illustrated in \autoref{fig:issue1}, the presence of correlated noise strongly decreases the transit signal SNR, which would ultimately limit its detectability.
+\autoref{fig:issue1} shows the SNR from \autoref{eq:snr} computed for a unique transit observed in the absence (grey) and presence of correlated noise (red).
+As illustrated in this figure, the presence of correlated noise strongly decreases the transit signal SNR, which would ultimately limit its detectability.
 
@@
 strongly resembling the simulated transit signal.\\\\
-For each of the 10 000 light curve generated, we separately reconstruct and detrend  the variability signal using the two techniques employed in \autoref{fig:issue2}, i.e.\;one using an optimal bi-weight filter with a window size three times that of the transit duration \citep{wotan} and the other using a GP with an optimal kernel.
+For each of the 10 000 light curves generated, we separately reconstruct and detrend the variability signal using the two techniques employed in \autoref{fig:issue2}, i.e.\;one using an optimal bi-weight filter with a window size three times that of the transit duration \citep{wotan} and the other using a GP with an optimal kernel.
 We then compute the detrended transit SNR using \autoref{eq:snr}.
@@
         \caption{Transit SNR computed for 10 000 differential light curves with varying morphological characteristics.
-All light curves span 2.8 days with an exposure time of 2 minutes and contain the same transit signal of  duration $D=1$ hour, depth of 0.005 and white noise with a standard deviation of 0.001.
+All light curves span 2.8 days with an exposure time of 2 minutes and contain the same transit signal of  duration $D=1$ hour, depth of 0.5\% and white noise with a standard deviation of 0.1\%.
 Light curves at the top and right side of the central plot are shown with their corresponding $\tau$ and $\delta$ values, which again correspond to the relative timescale of stellar variability against the transit duration, and the relative amplitude of stellar variability against the transit depth.
@@
 When searching for exoplanets using an indirect method, such as using transits or radial velocities, the initial detection is often as important as the follow-up observations, ultimately leading to the confirmation of the planetary candidate.
-For that reason, the final product of most transit search algorithms consists in a \textit{periodogram}: a 1D detection metric given a set of trial periods that allows to identify the presence of a potential candidate, and the period and epoch at which it should be followed up.\\\\
+For that reason, the final product of most transit search algorithms consists of a \textit{periodogram}: a 1D detection metric given a set of trial periods that allows to identify the presence of a potential candidate, and the period and epoch at which it should be followed up.\\\\
 If we assume that a transit is defined by its period $P$, epoch $T_0$, duration $D$ and depth $\Delta$, we then wish to compute the likelihood of such a transit being present in the data for a set of periods $P$, leading to an interpretable periodogram.
@@
 In \autoref{approach}, we explain how our approach requires two separate steps in order to remain tractable, and present these steps in \autoref{linear_search} and \autoref{periodic_search}, leading to the transit-search periodogram described in \autoref{periodogram}.
-Finally, we present the \nuancecode{} Python package in \autoref{package} and perform a control test of our implementation against \textsf{BLS} in \autoref{control}.
+Finally, we present the \nuancecode{} Python package in \autoref{package} and perform a control test of our implementation against BLS in \autoref{control}.
 
@@
 While this is tractable for few periods, it becomes highly untractable for the large set of trial periods required to properly sample a transit-search periodogram (in the order of tens of thousands).\\\\
-To remain tractable, \nuancemethod{} employs the strategy of \cite{foreman2016}\footnote{Also introduced in \cite{Aigrain2004}.}, and separates the transit search into two components: the linear search and the periodic search.
+To remain tractable, \nuancemethod{} employs an extension of the strategy used by \cite{foreman2016}, which has significant intellectual overlap with the methods used by \cite{Aigrain2004} and \cite{Jenkins2010}.
+This approach separates the transit search into two components: the linear search and the periodic search.
 During the linear search, the likelihood $p(\bm{f} \vert T_0, D, \Delta)$ of a single non-periodic transit is computed for a grid of epochs and durations, each time solving for $\Delta$ linearly.
@@
 This way, \nuancemethod{} searches for transits while, at the same time, modeling correlated noise, avoiding the separate detrending step that degrades transit signals SNR (see \autoref{issues}).\\\\
-We note that the approach employed by \nuancemethod{} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.
+% We note that the approach employed by \nuancemethod{} (using the two steps proposed by \citealt{foreman2016}), shares similarities with the approach of \cite{Jenkins2010}, where a single event statistic is computed and combined into a multiple event statistics.
 
@@
 In order to compute this joint likelihood, we must assume that the likelihoods computed during the linear search are independent.
-In theory, given that all data points of the light curves are used to compute each likelihood separately, these are not.
-In practice however, this assumption is reasonable and only affect the results of the periodic search for extremely short-period transit signals (where $P$ is less than few times the duration $D$) that are less likely to be present in physically-realistic datasets.\\\\
+However, the individual transits combined in the periodic search are not independent.
+Indeed, by using a GP, we assume that points belonging to separate transits may have a non-zero covariance.
+In practice, we notice that this covariance is small enough to consider each transit as independent, a reasonable assumption for most physically-realistic datasets.\\\\
 While \autoref{eq:result} takes a closed form, the individual epochs matching $T_0$ and $P$ are not necessarily available in the grid of epochs $\set{T_k}_k$.
@@
 $$\ln\mathcal{P}(P) = \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i,j}$$
-shown for few periods in \autoref{fig:periodic_search} (b). In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the linear search using $\phi_i$, $D_j$ and $T_0 = 0$.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_periodic_0.pdf}\\
-    {(a) $\ln \mathcal{L}$}
-\end{figure}
+shown for few periods in \autoref{fig:periodic_search}. In the latter equation, $\Delta_{i,j}$ is omitted since being interpolated from the linear search using $\phi_i$, $D_j$ and $T_0 = 0$.
 
@@
     \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
-    {(b) $\mathcal{P}(P)$}
-
-    \caption{Applied to the dataset shown in \autoref{fig:linear_search}, this figure shows how the periodic search works at different periods $P$ (including the true period $P = 0.7$ days).
-Given different periods $P$ and $T_0=0$, the likelihood $\ln\mathcal{L}$ shown in (a) is phase-folded and interpolated onto a common grid of phases shown in (b).
-As an example, the white lines in (a) mark the edges of each fold for a period $P=0.5$ days, and the white crosses show the epochs $\set{T_k}_{k\in\mathbb{K}}$ matching a particular phase in the grid (reported in (a) and (b)) on which the corresponding $\set{\ln \mathcal{L}_k}_{k\in\mathbb{K}}$ are interpolated and combined using \autoref{eq:result}.
-To allow the use of efficient matrix computations, this is done for all durations $\set{D_i}_i$ so that $\mathcal{P}$ is computed on the full grid $\set{D_i, \phi_i}_{i, j}$ at once.
-We understand from the folded likelihoods plots in (b) that a different choice of epoch $T_0$ may only shift the results in phase but do not affect the values of $\mathcal{P}$.
-For this reason, computing $\mathcal{P}$ for $T_0=0$ is sufficient.
-We also notice how the maximum value of $\mathcal{P}(P=0.7/2)$ (left plot of (b)) is lower than for $P=0.7$ days, a result of combining the log-likelihoods using \autoref{eq:result} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
+    \caption{Periodic search likelihood $\mathcal{P}(P)$ computed for different trial periods $P$.
+Notice how the maximum value of $\mathcal{P}$ for the alias period $P=0.35$ day (left plot) is lower than for $P=2\times 0.35=0.7$ day, a result of combining the log-likelihoods using \autoref{eq:result} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
     \label{fig:periodic_search}
@@
 
-Using \autoref{eq:result}, we can now compute $\ln\mathcal{P}$ for a range of periods and build a transit search periodogram using \autoref{eq:periodogram}.
-But a final issue emerges, one that is fundamentally linked to our approach.
-Each likelihood $p(\bm{f} \vert T_0, D, \Delta)$ estimated during the linear search is computed using N measurements.
-Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving $N\times K$ measurements.
-This leads to a normalization issue when comparing the joint log-likelihoods $\mathcal{P}(P)$ from one period to another, as the number of observed transits differs from one period to another.
+Using \autoref{eq:result}, we can now compute $\ln\mathcal{P}$ for a range of periods and phases, and build a transit search periodogram using \autoref{eq:periodogram}.
+This has two disadvantages: First, each likelihood $p(\bm{f} \vert T_0, D, \Delta)$ estimated during the linear search is computed using N measurements.
+Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving up to $N\times K$ measurements.
+For this reason, one has to normalize each likelihood $\ln\mathcal{P}$, by keeping track of the number of points used to compute each of them, which differs from one phase to another.
+Second, the maximum value of the likelihood $\ln\mathcal{P}$ is relative to a given dataset, so that a more intuitive and absolute metric must be used to relate to the transit signal detection (such as the Signal Detection Efficiency in \citealt{bls}).
 This motivates a final step to produce the transit search periodogram $\mathcal{Q}$.
-For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, we compute the maximum likelihood parameters
-\begin{equation}\label{eq:phi0}
-    (\phi_0 ,D) = \argmax_{\phi_i, D_j} \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i, j}
-\end{equation}
-and defined $\mathcal{Q}(P)$ as being equal to the SNR of the transit of period $P$, epoch $T_0 = \phi_0 \times P$, duration $D$ and depth $\Delta$, i.e.
-\begin{equation*}
-    \mathcal{Q}(P) = \frac{\Delta}{\sigma},
-\end{equation*}
-where $\Delta$ and $\sigma$ are obtained using \autoref{eq:sigma_delta} with the last column of $X$ containing a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $1$.
-This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.
 \begin{figure}[H]
@@
 \end{figure}
-The periodic transit of period $P$ with the maximum SNR, i.e.
+For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, we compute the maximum likelihood parameters
+\begin{equation}\label{eq:phi0}
+    (\phi_0 ,D) = \argmax_{\phi_i, D_j} \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i, j}
+\end{equation}
+and define $\mathcal{Q}(P)$ as being equal to the SNR of the transit of period $P$, epoch $T_0 = \phi_0 \times P$, duration $D$ and depth $\Delta$, i.e.
+\begin{equation*}
+    \mathcal{Q}(P) = \frac{\Delta}{\sigma},
+\end{equation*}
+where $\Delta$ and $\sigma$ are obtained using \autoref{eq:result} with the last column of $X$ containing a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $1$.
+This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.\\\\
+Hence, periodic transit of period $P$ with the maximum SNR, i.e.
 maximizing $\mathcal{Q}$, is adopted as the best candidate, basing the confidence in this signal through its SNR.
-The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (given by \autoref{eq:result}).
-
+The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (given by \autoref{eq:result}).\\\\
 \subsection{An open-source python package}\label{package}
-Methods presented in this paper are made accessible through the \nuancecode{} open-source Python package, hosted on Github\footnote{\href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}} and released on the Python Package Index\footlink{https://pypi.org/project/nuance/}.
-\\\\
+Methods presented in this paper are made available through the \nuancecode{} open-source Python package hosted at \href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}, released on the Python Package Index\footlink{https://pypi.org/project/nuance/} and with documentation and tutorials hosted at \href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}.
+All following mentions of \nuancecode{} refers to its version 0.5.2.\\\\
 To instantiate a search, a user can start by creating a \textsf{Nuance} object with
@@
 For this reason, detected transit epochs are not considered (although manually vetted).
-Results from this \textit{injection-recovery} procedure are shown in \autoref{fig:control}.
+Results from this \textit{injection-recovery} procedure are shown in \autoref{fig:control}.\\\\
+These results demonstrate the qualitative match between the detection capabilities of \textsf{nuance} and \textsf{BLS} on light curves with no correlated noise, where the BLS algorithm should be optimal.
+Explaining the subtle differences observed between the two methods when only white noise is present is beyond the scope of this paper, and we will assume that any differences observed in the following sections are due to the different treatments of correlated noise.
 
@@
 
-These results demonstrate the qualitative match between the detection capabilities of \textsf{nuance} and \textsf{BLS} on light curves with no correlated noise, where the \textsf{BLS} algorithm should be optimal.
-Explaining the subtle differences observed between the two methods when only white noise is present is beyond the scope of this paper, and we will assume that any differences observed in the following sections are due to the different treatments of correlated noise.
-
 \newpage
@@
 We perform this study by comparing \textsf{nuance} to the approach that involves removing stellar variability from light curves before performing the search on a detrended dataset.
-The following detrending strategies, each followed by a search with the \textsf{BLS} algorithm, are compared:\\\\
+The following detrending strategies, each followed by a search with the BLS algorithm, are compared:\\\\
 \ul{\texttt{bi-weight+BLS}}: employs an optimal bi-weight filter implemented in the \textsf{wõtan} Python package with an optimal window size of $3\times D$, i.e.\, three times the transit duration (e.g.\;\citealt{wotan} and \citealt{Dransfield2024}).\\\\
@@
 \subsubsection*{Stellar variability kernel}\label{rotation_kernel}
-In the \texttt{GP+BLS} and \nuancecode{} methods we model light curves using a GP with a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ (with $P_\star$ the rotation period of the star), a model representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python}.
+In the \texttt{GP+BLS} and \nuancecode{} methods we model light curves using a GP with a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ where $P_\star$ is the rotation period of the star.
+This model is representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python} \citep[e.g.][]{David2019, Gillen2020}.
 In order to account for additional corelated noises, we complement this kernel with a short and a long-timescale exponential term, so that the full kernel can be expressed as
@@
 Compared to other techniques, we find that \nuancecode{} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (\autoref{fig:allsearch}).
-While the performances of other methods highly depend on the characteristics of the variability, \textbf{\nuancecode{} is the best technique in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positives} (\autoref{fig:allsearch}).
+While the performances of other methods strongly depend on the characteristics of the variability, \textbf{\nuancecode{} is the best technique in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positives} (\autoref{fig:allsearch}).
 \begin{figure}[H]
@@
 \end{figure}
-From \autoref{fig:allsearch}, we note that the performance of the \texttt{GP+BLS} method is much worse that what could be anticipated from the results presented in \autoref{fig:simu}, where the GP kernel was fully optimal (given it was also used to simulate the data).
+From \autoref{fig:allsearch}, we note that the number of true positives of the \texttt{GP+BLS} method is much worse that what could be anticipated from the results presented in \autoref{fig:simu}, where the GP kernel was fully optimal (given it was also used to simulate the data).
 In that case, our kernel might not be optimal, whether because of its form or because of the values of its hyperparameters.
-In any case, the fact that the same kernel performs significantly better when used with \nuancecode{} shows that our method is less sensitive to the choice of kernel and its optimization compared to detrending with a GP.
+In any case, the fact that the same kernel performs significantly better when used with \nuancecode{} shows that \textbf{our method is less sensitive to the choice of kernel and its optimization compared to detrending with a GP}.
 \begin{figure}[H]
@@
 Optimal grids of parameters for realistic transit searches are discussed in \cite{tls}.}.
-This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the \textsf{BLS} algorithm (denoted \texttt{bi-weight+BLS} and described in \autoref{real}) on all sectors combined.
-Since we don't know the transit duration a-priori, we perform the detrending and \textsf{BLS} search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
+This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{bi-weight+BLS} and described in \autoref{real}) on all sectors combined.
+Since we don't know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
 The results of this comparison are shown in \autoref{fig:toi540_periodograms}.\\\\
@@
 The proper treatment of correlated noise in \nuance{}, as observed in \autoref{real}, makes these peaks non-significant, avoiding a large number of false detections.\\\\
-We note that finding a TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques (see e.g.\;\autoref{fig:simu}).
+We note that finding a known TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques (see e.g.\;\autoref{fig:simu}).
 Nontheless, we verify that \nuancemethod{} is capable of finding a large number of already known transiting exoplanet candidates, in light curves featuring various forms of correlated noises, at least as efficiently as with commonly used techniques.
@@
 \noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple non-optimized GP with a squared exponential kernel.
-These are compared to the processing time of \texttt{bi-weight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the \textsf{BLS} search.
+These are compared to the processing time of \texttt{bi-weight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the BLS search.
 \begin{figure}[H]
@@
 This study is performed on a single CPU core of an Apple M2 Max chip.
-While not being shown, we verify that both the \textsf{BLS} algorithm and \nuancemethod{} processing times scale linearly with the number of trial transit durations and orbital periods.}
+While not being shown, we verify that both the BLS algorithm and \nuancemethod{} processing times scale linearly with the number of trial transit durations and orbital periods.}
         \label{fig:blsvsnuance}
@@
 Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform between one and two order of magnitude slower than more conventional techniques associated with BLS}.\\\\
-As a reference, searching for TOI-540 b transits with the parallelized implementation of \nuancecode{} (see \autoref{toi540}) on the 12 cores of an Apple M2 Max chip, took 5 minutes 35 seconds in total, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds per TESS sector), and 3 minutes 45 seconds for the combined periodic search.
+Nonetheless, searching for TOI-540 b transits with the parallelized implementation of \nuancecode{} (see \autoref{toi540}) on the 12 cores of an Apple M2 Max chip, took only 5 minutes and 35 seconds, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds per TESS sector), and 3 minutes 45 seconds for the combined periodic search.
 In comparison, the brute force search with \texttt{bi-weight+BLS}, consisting in trying 15 bi-weight windows, took a total of 1 minute 49 seconds (only 7 seconds each).\\\\
@@
 in \citealt{everest1}).\\\\
-Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsly is extremely promising for the search of transiting exoplanets in ground-based data, that usually suffer from daily interruptions.
+Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsly is extremely promising for the search of transiting exoplanets, including for ground-based observations that usually suffer from daily interruptions.
 In this respect, we note the similarity of our linear search (cf\;\autoref{linear_search}) to the one presented in \citealt{Berta2012}, that focused on the detection of single eclipses in the MEarth light curves \citep{Irwin2009}.
@@
 \\\\
-\vfill{}
-We would like to thank Julien De Wit, Prajwal Niraula for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
+The software presented in this work is open source under the MIT License and is
+available at \href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}, with documentation and tutorials hosted at \href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}.
+
+\newpage
+We would like to thank Julien De Wit and Prajwal Niraula for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, the member of the Exotic group at the University of Liège, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.\\\\
+
+\software{{numpy} \citep{harris2020array}, {scipy} \citep{2020SciPy-NMeth}, {matplotlib} \citep{Hunter:2007}, {jax} \citep{jax2018github}, {jaxopt} \citep{jaxopt_implicit_diff}, {tinygp} \citep{foreman_mackey_2024_10463641}, {lightkurve} \citep{2018ascl.soft12013L}, {astropy} \citep{astropy:2022}, {pandas} \citep{mckinney-proc-scipy-2010, reback2020pandas} and {wotan} \citep{wotan}.}
 

### 2024-02-09 latex/paper.tex
chore: acknowledgment + minor fixes

@@
 Second, the maximum value of the likelihood $\ln\mathcal{P}$ is relative to a given dataset, so that a more intuitive and absolute metric must be used to relate to the transit signal detection (such as the Signal Detection Efficiency in \citealt{bls}).
-This motivates a final step to produce the transit search periodogram $\mathcal{Q}$.
+This motivates a final step to produce an interpretable transit search periodogram $\mathcal{Q}$.
 \begin{figure}[H]
@@
 \newpage
-We would like to thank Julien De Wit and Prajwal Niraula for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, the member of the Exotic group at the University of Liège, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.\\\\
+We would like to thank Julien De Wit and Prajwal Niraula for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, the member of the Exotic group at the University of Liège, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
+This publication benefits from the support of the French Community of Belgium in the context of the FRIA Doctoral Grant awarded to L.J.G.
+F.J.P acknowledges financial support from the grant CEX2021-001131-S funded by MCIN/AEI/ 10.13039/501100011033 and through projects PID2019-109522GB-C52 and PID2022-137241NB-C4.
+S.A.
+acknowledges funding from the European Research Council (ERC) under grant agreement 865624 (GPRV) and from the UK Science and Technology Facilities Council (STFC) under grants ST/S000488/1 and ST/R004846/1.
+This paper includes data collected by the TESS mission.
+Funding for the TESS mission is provided by the NASA's Science Mission Directorate and NASA Explorer Program.
+We acknowledge the use of public TESS data from pipelines at the TESS Science Office and at the TESS Science Processing Operations Center.
+Resources supporting this work were provided by the NASA High-End Computing (HEC) Program through the NASA Advanced Supercomputing (NAS) Division at Ames Research Center for the production of the SPOC data products.
+TESS data were obtained from the MAST data archive at the Space Telescope Science Institute (STScI).
+STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5–26555.
+The Center for Computational Astrophysics at the Flatiron Institute is supported by the Simons Foundation.\\\\
 

### 2024-03-29 referee_answer.txt
feat: account for referee report

In what follows, all references correspond to lines in the revised manuscript, where changes have been highlighted in red. We thank the referee for an extremely meaningful report, we believe the paper will truly benefit from their experience on the subject.

Referee:
The selection of BLS (instead of TLS) together with simultanous "best" detrending seems odd at first glance. Why not combine both? The reason is clear however (and should be stated explicitely): Planetary transit detection around active stars is hindered almost entirely by stellar noise, not by the small change caused by planetary limb-darkening. The authors correctly note a (small) possible improvement by TLS (section 4.4) and this is fine. But the neglect of TLS-type methods should be explained early on as well.

Authors:
We agree that an early justification would be better. We added a sentence in line 374.

Referee:
I am not at all convinced by the results of the synthetic tests. It is too easy to create synthetic data that are particularly useful for a given method (and vice-versa). Specifically, the periodic GP kernel (while physically motivated) and the (multi)periodic data obviously work together well. This test does not show, however, that the method is useful for real-world photometry. I recommend the authors make this clearer in the synthetic section, also giving a stronger motivation for the real-world test.

Authors:
We made that point a bit more explicit in line 473, highlighting the "synthetic" value of these tests.

Referee:
For the real-world tests, I'm a bit confused. The text states injection experiments "from a list of 438 targets" - how many different light curves were chosen for injections? All 438 or a subset of these? How many injections were made per LC, and how many in total? I would guess 438x100, but I'm not sure. It is certainly necessary to test a large randomly chosen variety of active star light curves for such injection-retrieval experiments. This way, over-tuning and selection bias can be reduced. The section on top of page 22 should be written more clearly.

Authors:
We did use the full dataset of 438 light curves (one half-sector for each target). We made this point much clearer throughout the manuscript, in lines 482, 485, 492, 514 and 524.

Referee:
In addition, it would be useful to state an average (or median) planetary radius (in physical units, not just in SNR) of the injections to gain some insights into the "real" success of this new method: What planet discoveries can be expected? By "how much" is the new method better, e.g. as measured in planetary radius for an e.g. SNR=10 detection?

Authors:
We thank the referee for this very meaningful suggestion. Line 524 gives some first order statistics about the injected planets radii. We also added an extensive study of the kind of planets for which nuance tends to be beneficial (line 601 and figure 14). Although we give an explanation of why nuance is better at finding larger planets here, we warn the readers that this result is very specific to the studied dataset. Nonetheless, it is nice to show that nuance is at least as performant as other methods in finding planets with all kinds of radii.

Referee:
Finally, your readers are wondering: You have searched 438 (?) previously untouched light curves of active stars. Have planets been found in these real data with the new method (before or after injection)? Of course this is a paper focusing on a new tool and not planet discovery, but a note seems to be in order of either a companion paper that is being worked on (or not). That way, the community can avoid double work.

Authors:
We did not make such a search, and the community should be encouraged to do it. We make that explicit in line 614.

Referee:
Lastly, one item is missing for section 4.4 "prospects". The "nuance" periodogram (Figure 7), just as the "BLS periodogram", suffers from an increase of SNR with larger periods P. This increase is an algorithmic artifact and can be (approximately) removed (for a clearer trend visualization see Figure 3a and section 2.2 in Hippke 2019). The underyling effect was first explained well by Ofir 2014. I think it is not strictly required for "nuance" to consider this effect, but it should be added to the list of suggested future improvements (especially because being almost trivial to implement).

Authors:
After investigation it seems that, since our periodogram is made of the actual SNR of the transit, this effect might not be an algorithmic artifact. It makes sense that the longer period model, corresponding to a unique transit when the period is large enough (See figure 7) have a higher SNR than a shorter-period signal where multiple transits with a given period are forced to fit the light curve. Ofir 2014 (section 5.1) suggests that "a progressively smaller number of points are actually required to be in-transit, and so the probability that the mean value of smaller random groups of points will be different from the global mean will increase, driving both the mean value and the scatter of the raw BLS spectrum up". Even if this point was valid in our case, an increase of the transit depth uncertainty could also lead to a smaller SNR. That being said, we acknowledge that there is a fundamental parallel in all other periodograms constructs (such as the Lomb-Scargle). For now, as we do not have a theoretical justification to detrend the SNR periodogram, we prefer not to mention this point and defer this addition to an eventual upgrade of the nuance package and documentation.

Referee:
Two real-world examples are dicussed: TOI-540 (Figure 14) and TIC 1019692 (Figure 10). Are these "typical examples" or why were they chosen?

Authors:
These example are indeed typical, which we highlight in lines 578 and 623.

Other changes:
- An extra past affiliation has been added to the author list (line 5)
- A mistake about the second-best method has been fixed in the caption of Figure 12.
- Fix "Nonetheless" line 661
- Added "shorter-timescale" line 801
- Addition to the acknowledgements in line 858

----

Data Editor:
Per our software policy [1], we recommend that authors with living code in a GitHub repository place a "frozen" version on Zenodo (or other 3rd party repository that issues DOIs) [2,3] and then cite the deposited version in the article [4] and it in the reference list

Authors:
A Zenodo record was created for version 0.6.0 and cited in the paper (line 332 and 884).

Data Editor:
Note that the astropy collaboration requests that all 3 of their articles be cited

Authors:
All references for astropy were added in line 882.

Again, many thanks to both the referee and data editor!

### 2024-03-29 latex/paper.tex
chore: add line numbers for review
feat: account for referee report

@@
-\documentclass[modern]{aastex631}
+\documentclass[modern,linenumbers]{aastex631}
 \usepackage{amsmath}
@@
 \newcommand{\TODO}{\texttt{todo}}
+\newcommand{\review}[1]{\textcolor{red}{#1}\textbf}
+% \newcommand{\review}[1]{#1} 
 \newcommand{\set}[1]{\{\,#1\,\}}
@@
 \author{Lionel J. Garcia}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
+\affiliation{\review{Astrobiology Research Unit, University of Liège, Liège, Belgium}{}}
 \author{Daniel Foreman-Mackey}
@@
 \affiliation{Instituto de Astrofísica de Andalucía, Granada, Spain}
+\affiliation{Astrobiology Research Unit, University of Liège, Liège, Belgium}
 \correspondingauthor{Lionel J. Garcia}
@@
 
-\keywords{transit search, exoplanets, active stars, stellar variability, time series analysis, Gaussian process, TESS}
+\keywords{exoplanet detection methods, stellar activity, time series analysis, gaussian processes regression, computational methods, GPU computing}
 
@@
 Overall, we demonstrate that \nuancemethod{} is the most performant method in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
-Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and availa ble as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
+Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices.
 Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of planetary systems around active stars, both using space-based surveys and sparse ground-based observations.
@@
 Methods presented in this paper are made available through the \nuancecode{} open-source Python package hosted at \href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}, released on the Python Package Index\footlink{https://pypi.org/project/nuance/} and with documentation and tutorials hosted at \href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}.
-All following mentions of \nuancecode{} refers to its version 0.5.2.\\\\
+All following mentions of \nuancecode{} refers to its \review{version 0.6.0 \citep{zenodo}.}{} \\\\
 To instantiate a search, a user can start by creating a \textsf{Nuance} object with
@@
 By design, these injected signals have an SNR ranging from 5 to 30 (with $\sigma_r = 0$ in \autoref{eq:snr}).
-For each light curve, transits are searched using two different tools: \textsf{nuance}, using its implementation from the Python package described in the previous section; and BLS, the Box-Least-Square algorithm from \cite{bls} (using \textsf{astropy}'s \textsf{BoxLeastSquares}\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html} implementation).\\\\
+For each light curve, transits are searched using two different tools: \textsf{nuance}, using its implementation from the Python package described in the previous section; and BLS, the Box-Least-Square algorithm from \cite{bls} (using \textsf{astropy}'s \textsf{BoxLeastSquares}\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html} implementation).
+\review{We note that the Transit-Least-Square algorithm \citep{tls}, that includes the effect of limb-darkening in the base transit template, could also be used here.
+However, as this effect have a negligible impact on the transit detection compared to the effect of correlated noise, we choose to make our comparisons with BLS only.}\\\\
 For both methods, 3000 trial periods from 0.2 to 2.6 days are searched, with a single trial duration fixed to the unique known duration of 50 minutes.
@@
 
-\noindent This injection-recovery is done in a particularly optimal setup, on simulated light curves not all physically realistic and using an optimal GP kernel, hence demonstrating the performance of \textsf{nuance} only on a purely experimental basis.
+\noindent This injection-recovery is done in a particularly optimal setup, on simulated light curves not all physically realistic and using an optimal GP kernel, hence demonstrating the performance of \textsf{nuance} only on a purely \review{synthetic}{} basis.
+\review{Hence, we strongly emphasize that these tests do not reflect well the performance of each method on real datasets}.
 In the next section, we perform transits injection-recovery on real space-based light curves.\\\\
@@
 In order to assess the performance of \nuance{} on real datasets, we inject and recover transits into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}).
-We focus this proof of concept on light curves from a list of 438 M-dwarfs found to have detectable rotation signals with periods lower than one day \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{}.
-For each studied target, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) of a single sector (the first being observed for each target) spanning on average 10 days.
+We focus this proof of concept \review{on the light curves of 438 M-dwarfs found to have detectable rotation signals with periods lower than one day}{} \citep{Ramsay2020}, which lead to a parameter space justifying the use of \nuance{}.
+For each \review{of the 438}{} targets, transits are injected and recovered in the TESS 2 min cadence SPOC Simple Aperture Photometry and Pre-search Data Conditioning light curves (PDCSAP, \citealt{spoc}) of a single sector (the first being observed for each target) spanning on average 10 days.
 To our knowledge, none of these 438 targets have been searched for planetary transits before.
-However, we note that the presence of existing transit signals in these light curves before the injection of simulated ones is possible, but will not affect the relative comparison of one method to another.\\\\
+However, we note that the presence of existing transit signals in these light curves before the injection of simulated ones is possible, but will not affect the relative  comparison of one method to another.
+\review{As described in the following sections, our experiment considers a total of $438 \times 100$ light curves, where transit signals are injected and recovered.}\\\\
 \subsubsection*{Light curve cleaning and transits injection}
@@
 However, adopting this flare cleaning step and analyzing light curves with few small gaps is a practice commonly found in the literature.\\\\
-For each light curve, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injection-recovery per target.
+For each \review{of the 438 light curves considered}, transits of planets with 10 different orbital periods combined with 10 planetary radii are individually considered, for a total of 100 periodic transits injection-recovery per target.
 Orbital periods $P$ are sampled on a regular grid between 0.4 and 5 days, and planetary radii $R_p$ are sampled on a regular grid designed to yield a minimum transit SNR of 2 and a maximum of 30.
@@
 \end{equation*}
-with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.\\\\
+with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.
+\review{In total, $438 \times 100$ light curves are produced and searched for transits, corresponding to planets radii ranging from $0.46 R_\oplus$ to $12.29 R_\oplus$ and a median radius of $2.40 R_\oplus$.}\\\\
 \subsubsection*{Stellar variability kernel}\label{rotation_kernel}
@@
 multiple trial durations) is discussed in the next section.
+\newpage
 \subsubsection*{Results}
-An example of the transit injection-recovery and its result is shown in \autoref{1019692} for the target TIC 1019692.
+An example of the transit injection-recovery and its result is shown in \autoref{1019692} for the target TIC 1019692, \review{a representative example of the kind of stellar variability encountered in our dataset}.
 \autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in \autoref{eq:relative_params}).
@@
 \end{figure}
-Compared to other techniques, we find that \nuancecode{} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (\autoref{fig:allsearch}).
-While the performances of other methods strongly depend on the characteristics of the variability, \textbf{\nuancecode{} is the best technique in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positives} (\autoref{fig:allsearch}).
 \begin{figure}[H]
@@
         \caption{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
-The solid blue line corresponds to the maximum of true positives among all methods on the top panel (dominated by the \texttt{iterative+BLS} method), and the minimum of false positives on the bottom panel (waned by the \texttt{GP+BLS} method, all other methods being above \nuancecode{}).}
+The solid blue line corresponds to the maximum of true positives among all methods on the top panel (dominated \review{by the \texttt{harmonics+BLS} method}), and the minimum of false positives on the bottom panel (waned by the \texttt{GP+BLS} method, all other methods being above \nuancecode{}).}
         \label{fig:allsearch_relative}
@@
 \end{figure}
-From \autoref{fig:allsearch}, we note that the number of true positives of the \texttt{GP+BLS} method is much worse that what could be anticipated from the results presented in \autoref{fig:simu}, where the GP kernel was fully optimal (given it was also used to simulate the data).
-In that case, our kernel might not be optimal, whether because of its form or because of the values of its hyperparameters.
-In any case, the fact that the same kernel performs significantly better when used with \nuancecode{} shows that \textbf{our method is less sensitive to the choice of kernel and its optimization compared to detrending with a GP}.
+Compared to other techniques, we find that \nuancecode{} leads to the highest number of true positives, with a successful detection of 76\% of the 7008 detectable transits injected (\autoref{fig:allsearch}).
+While the performances of other methods strongly depend on the characteristics of the variability, \textbf{\nuancecode{} is the best technique in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positives} (\autoref{fig:allsearch}).\\\\
 \begin{figure}[H]
@@
 \end{figure}
+From \autoref{fig:allsearch}, we note that the number of true positives of the \texttt{GP+BLS} method is much worse that what could be anticipated from the results presented in \autoref{fig:simu}, where the GP kernel was fully optimal (given it was also used to simulate the data).
+In that case, our kernel might not be optimal, whether because of its form or because of the values of its hyperparameters.
+In any case, the fact that the same kernel performs significantly better when used with \nuancecode{} shows that \textbf{our method is less sensitive to the choice of kernel and its optimization compared to detrending with a GP}.\\\\
+\review{Finally, in \autoref{fig:tpfp_radius}, we observe a dependence between the rate of true positives and the injected planets radii.
+Overall \texttt{nuance} leads to significantly higher true positives in all radii bins compared to other methods, except for \texttt{BLS+harmonics} (which dominates other methods) that shows a similar yield for planets with radii less than 2$R_\oplus$.
+Hence, we note that the advantage of \texttt{nuance} is particularly significant for planets larger than 2$R_\oplus$.
+This can be explained by the fact that, for a transit signal to be comparable to a high-amplitude stellar variability signal, i.e.
+when \texttt{nuance} shows an advantage, a larger transit depth, hence a larger planetary radius, is required.
+Although this particular result highly depends on the variability present in the studied dataset, we conclude that \textbf{the performance of \nuance{} is not sensitive to the searched planets' radii.}}\\\\
+\begin{figure}[H]
+    \begin{centering}
+        \includegraphics[width=0.6\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_radius.pdf}
+        \caption{\review{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the injected planet's radius.
+The solid blue line corresponds to the maximum of true positives among all other methods than \nuance{} (dominated by the \texttt{harmonics+BLS} method).}{}}
+        \label{fig:tpfp_radius}
+    \end{centering}
+\end{figure}
+\review{When analyzing the light curves of the 438 targets from \citealt{Ramsay2020}, our search solely focused on the injected transit signals.
+Although \nuancecode{} seems particularly well suited to search for real transit signals in this dataset, we did not conduct such study.
+To be done properly, this task would simply require a search on a wider and finer trial epochs and durations grid, using all available TESS sectors; a project that we highly encourage.}{}
+
 \subsection{Comparison for a multi-sector TESS candidate: TOI-540}\label{toi540}
@@
 In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b \citep{TOI540}.
+\review{As we will see, \nuancemethod{} does not have significantly superior performance for this candidate.
+However, its light curves feature a very typical stellar variability signal (see \autoref{fig:toi540_clean}), that would commonly be detrended before searching for planetary transits}.\\\\
 We download the 2 minutes cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
@@
 This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b\footnote{using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ minutes of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial (see \autoref{fig:snr_detrend}).
-Nonetheless, the \nuance{} periodogram of TOI 540 features less spurious SNR peaks, largely due to the penalty naturally occurring when single transits with different depths are periodically combined.
+\review{Nonetheless}, the \nuance{} periodogram of TOI 540 features less spurious SNR peaks, largely due to the penalty naturally occurring when single transits with different depths are periodically combined.
 In the second search (right panel) of \autoref{fig:toi540_periodograms}, we also notice a higher number of peaks that would lead to false positive detections of transits in the \texttt{bi-weight+BLS} case.
@@
 We note that finding a known TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques (see e.g.\;\autoref{fig:simu}).
-Nontheless, we verify that \nuancemethod{} is capable of finding a large number of already known transiting exoplanet candidates, in light curves featuring various forms of correlated noises, at least as efficiently as with commonly used techniques.
+Nonetheless, we verify that \nuancemethod{} is capable of finding a large number of already known transiting exoplanet candidates, in light curves featuring various forms of correlated noises, at least as efficiently as with commonly used techniques.
 We reserve the search for new transit signals to a follow-up publication.
@@
     \item During the periodic search, no prior about the transit duration related to the orbital period of the planet was used.
-This was done in order to allow the detection of grazing transiting exoplanets that would produce shorter timescale transits compared to what is expected from a circular orbit with a null impact parameter.
+This was done in order to allow the detection of grazing transiting exoplanets that would produce \review{shorter-timescale}{} transits compared to what is expected from a circular orbit with a null impact parameter.
 However, adding such priors might produce less spurious periodogram peaks and be very beneficial for the automatic search of transits in large datasets.
@@
 Both ideas come at no cost given our modeling approach.
+    % \item \review{The transit-search periodogram was intentionally reduced to a one-dimensional metric, mainly for practical reasons.
+However, we believe that the profile likelihood (or SNR) over orbital periods \textit{and} transit durations (see \autoref{snrperiod}) holds valuable information that could be exploited to further direct the search, especially when transits with different durations are present.}{}
     \item Finally, the formalism of \nuancemethod{} could be adapted and used to \textbf{search for exoplanets featuring transit time variations (TTV)}.
@@
 \end{enumerate}
-
 \begin{figure}[H]    
@@
 \end{figure}
+% \begin{figure}[H]
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_snr_period.pdf}
+%         \caption{}
+%         \label{fig:snrperiod}
+%     \end{centering}
+% \end{figure}
 
@@
 We would like to thank Julien De Wit and Prajwal Niraula for meaningful discussions at the beginning of this project, Germain Garcia for his useful insights about numerical optimization, Michaël Gillon for his overall support, the member of the Exotic group at the University of Liège, and the member of the Astronomical Data Group at the Center for Computational Astrophysics for many enriching discussions and feedback.
-This publication benefits from the support of the French Community of Belgium in the context of the FRIA Doctoral Grant awarded to L.J.G.
+\review{We would also like to thank the journal referee for their thorough and encouraging report, bringing some very meaningful additions to the paper.}{} This publication benefits from the support of the French Community of Belgium in the context of the FRIA Doctoral Grant awarded to L.J.G.
 F.J.P acknowledges financial support from the grant CEX2021-001131-S funded by MCIN/AEI/ 10.13039/501100011033 and through projects PID2019-109522GB-C52 and PID2022-137241NB-C4.
@@
 
-\software{{numpy} \citep{harris2020array}, {scipy} \citep{2020SciPy-NMeth}, {matplotlib} \citep{Hunter:2007}, {jax} \citep{jax2018github}, {jaxopt} \citep{jaxopt_implicit_diff}, {tinygp} \citep{foreman_mackey_2024_10463641}, {lightkurve} \citep{2018ascl.soft12013L}, {astropy} \citep{astropy:2022}, {pandas} \citep{mckinney-proc-scipy-2010, reback2020pandas} and {wotan} \citep{wotan}.}
+\software{{numpy} \citep{harris2020array}, {scipy} \citep{2020SciPy-NMeth}, {matplotlib} \citep{Hunter:2007}, {jax} \citep{jax2018github}, {jaxopt} \citep{jaxopt_implicit_diff}, {tinygp} \citep{foreman_mackey_2024_10463641}, {lightkurve} \citep{2018ascl.soft12013L}, \review{{astropy} (\citealt{astropy:2013,astropy:2018,astropy:2022})}, {pandas} \citep{mckinney-proc-scipy-2010, reback2020pandas}, {wotan} \citep{wotan} \review{and nuance version 0.6.0 \citep{zenodo}}.}{}
 

### 2024-05-17 latex/paper.tex
fix: add proof reading corrections

@@
 \newcommand{\TODO}{\texttt{todo}}
-\newcommand{\review}[1]{\textcolor{red}{#1}\textbf}
+\newcommand{\review}[1]{#1}
 % \newcommand{\review}[1]{#1} 
@@
 We assess the performance of \nuancemethod{} on simulated light curves as well as on the TESS light curves of 438 rapidly-rotating M dwarfs.
-For each dataset, we compare our method to 5 commonly-used detrending techniques followed by a search with the Box-Least-Square algorithm.
+For each dataset, we compare our method to 5 commonly-used detrending techniques followed by a search with the Box-Least-Squares algorithm.
 Overall, we demonstrate that \nuancemethod{} is the most performant method in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections.
@@
 First, as younger stars are more active \citep{Skumanich1972}, being able to detect planets transiting active stars will favor the discovery of young planetary systems \citep[e.g.][]{Newton2022}.
-Second, as stellar variability may originate from surface active regions (such as starpots), transiting exoplanets can be used to map the photosphere of active stars \citep[e.g.][]{Morris2017}, benefiting both the study of stellar atmospheres and the concerning impact of their non-uniformity on planetary atmosphere retrievals \citep{rackham2018}.
+Second, as stellar variability may originate from surface active regions (such as starspots), transiting exoplanets can be used to map the photosphere of active stars \citep[e.g.][]{Morris2017}, benefiting both the study of stellar atmospheres and the concerning impact of their nonuniformity on planetary atmosphere retrievals \citep{rackham2018}.
 Overall, enabling the detection of transits in light curves with high levels of correlated noises will greatly benefit the study of terrestrial exoplanets around late M-dwarfs, usually observed at lower SNR and more likely to display photometric variability (e.g.
@@
 \\\\
-Commonly used transit-search algorithms, such as the Box-Least-Square algorithm \citep[BLS,][]{bls} are capable of detecting transits in light curves containing only transit signals and white noise.
+Commonly used transit-search algorithms, such as the Box-Least-Squares algorithm \citep[BLS,][]{bls} are capable of detecting transits in light curves containing only transit signals and white noise.
 Using this method, the simplest way to detect transits in a light curve featuring correlated noise (either astrophysical or instrumental), is to first clean it from nuisance signals before performing the search.
@@
 This dataset corresponds to an observation of 4 days with an exposure time of 2 minutes.
-The mean of this signal consists in a periodic transit signal of period $P=0.7$ days, duration $D=1.2$ hours and depth $\Delta=2\%$ (dark gray) plus an instrumental signal (blue).
+The mean of this signal consists of a periodic transit signal of period $P=0.7$ days, duration $D=1.2$ hours and depth $\Delta=2\%$ (dark gray) plus an instrumental signal (blue).
 Correlated noise in the form of stellar variability is simulated with an SHO kernel of hyperparameters $\omega = \pi / 6 D$, $Q=45$ and $\sigma = \Delta$.
@@
 The CDPP is computed by decomposing the data in the time-frequency domain using wavelets, and measures the significance of a distorted transit signal in the whitened data.
-For a given transit duration, the CDPP is then a measure of the noise remaining after filtering the light curve in each frequency band, taking into account the presence of non-stationary correlated noise of a given timescale (see \citealt{Jenkins2010} for more details).
+For a given transit duration, the CDPP is then a measure of the noise remaining after filtering the light curve in each frequency band, taking into account the presence of nonstationary correlated noise of a given timescale (see \citealt{Jenkins2010} for more details).
 Hence, we can estimate the significance of the transit signal accounting for the presence of correlated noise as
@@
 \end{equation}
-where $CDPP_D$ is the CDPP computed for a given transit duration $D$ and $\bar{\Delta}$ is the transit depth after detrending.
+where $CDPP_D$ is the CDPP computed for a given transit duration $D$ and $\hat{\Delta}$ is the transit depth after detrending.
 For simplicity, the CDPP is computed using the method from \cite{Gilliland2011} implemented in the \textsf{lightkurve} Python package \citep{Lightkurve} and $\hat{\Delta}$ is the depth obtained by solving \autoref{eq:linear_model} with the last column of $\bm{X}$ containing a normalized transit model.
@@
 \end{figure}
-\autoref{fig:issue1} shows the SNR from \autoref{eq:snr} computed for a unique transit observed in the absence (grey) and presence of correlated noise (red).
+\autoref{fig:issue1} shows the SNR from \autoref{eq:snr} computed for a unique transit observed in the absence (gray) and presence of correlated noise (red).
 As illustrated in this figure, the presence of correlated noise strongly decreases the transit signal SNR, which would ultimately limit its detectability.
@@
 In \autoref{fig:issue2}, we simulate a transit signal on top of which we add photometric stellar variability with different timescales, sampled from a GP with an SHO kernel described in \autoref{app_gp}.
-For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same GP from which the data has been sampled.
+For each light curve, we reconstruct and detrend stellar variability in two ways: one using the widely-adopted Tukey's bi-weight filter, presented in \cite{tukey} and using the implementation from \textsf{wõtan}\footnote{\href{https://github.com/hippke/wotan}{https://github.com/hippke/wotan}} \citep{wotan}; the other using the same GP from which the data have been sampled.
 We then estimate the resulting transit depth and compute the remaining transit SNR using \autoref{eq:snr}.
@@
 \end{figure}
-\autoref{fig:snr_detrend} shows that it exists an entire region of the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 5$).
+\autoref{fig:snr_detrend} shows that there exists an entire region in the $(\tau, \delta)$ parameter space for which the bi-weight detrending degrades transit SNR to the point of no detection ($SNR < 5$).
 Although being more robust, the same effect is observed when detrending with an optimal GP.
@@
 \end{equation} 
-where $\bm{f}$ is the data.
+where $\bm{f}$ denotes the data.
 In this section, we present \textit{nuance}, an algorithm to compute transit-search periodograms in a tractable way, and detect planetary transits in light curves containing correlated noise such as instrumental signals and photometric stellar variability.
@@
 \end{equation}
-% Using this expression, the depth of the transit corresponds to $\hat{\Delta} = \bm{\hat{w}}_M$ with the uncertainty $\hat{\sigma} = \bm{\hat{\sigma}}_{M,M}$, where M denotes the index of the last column of the design matrix $\bm{X}$, corresponding to the normal\ transit signal.
-As we are only interested in the transit depth, we will omit $\bm{w}$ in the reminder of this paper and simply write likelihood expressions depending on its last value $\Delta=\bm{w}_M$, such as $p(\bm{f} \vert P, T_
+As we are only interested in the transit depth, we will omit $\bm{w}$ in the remainder of this paper and simply write likelihood expressions depending on its last value $\Delta=\bm{w}_M$, such as $p(\bm{f} \vert P, T_
 0, D, \Delta)$, where all other parameters of $\bm{w}$ are taken at their maximum-likelihood values. \\\\
-Hence, given a period $P$, computing $\mathcal{Q}(P)$ boils down to a non-linear optimization involving several evaluations of the likelihood given in \autoref{eq:linear_search_ll}.
+Hence, given a period $P$, computing $\mathcal{Q}(P)$ boils down to a nonlinear optimization involving several evaluations of the likelihood given in \autoref{eq:linear_search_ll}.
 While this is tractable for few periods, it becomes highly untractable for the large set of trial periods required to properly sample a transit-search periodogram (in the order of tens of thousands).\\\\
@@
 This approach separates the transit search into two components: the linear search and the periodic search.
-During the linear search, the likelihood $p(\bm{f} \vert T_0, D, \Delta)$ of a single non-periodic transit is computed for a grid of epochs and durations, each time solving for $\Delta$ linearly.
+During the linear search, the likelihood $p(\bm{f} \vert T_0, D, \Delta)$ of a single nonperiodic transit is computed for a grid of epochs and durations, each time solving for $\Delta$ linearly.
 Then, the periodic search consists of combining these likelihoods to compute the likelihood $p(\bm{f} \vert P, T_0, D, \Delta)$ of the data given a periodic transit signal for a range of periods $P$.
@@
 However, the individual transits combined in the periodic search are not independent.
-Indeed, by using a GP, we assume that points belonging to separate transits may have a non-zero covariance.
+Indeed, by using a GP, we assume that points belonging to separate transits may have a nonzero covariance.
 In practice, we notice that this covariance is small enough to consider each transit as independent, a reasonable assumption for most physically-realistic datasets.\\\\
@@
 maximizing $\mathcal{Q}$, is adopted as the best candidate, basing the confidence in this signal through its SNR.
-The parameters of this transit are the period $P$, the epoch $T_0 = \phi_0 P$ and duration $D$ (\autoref{eq:phi0}), and the depth $\Delta$ with error $\sigma$ (given by \autoref{eq:result}).\\\\
+The parameters of this transit are the period $P$, epoch $T_0 = \phi_0 P$, duration $D$ (\autoref{eq:phi0}), and depth $\Delta$ with error $\sigma$ (given by \autoref{eq:result}).\\\\
 \subsection{An open-source python package}\label{package}
-Methods presented in this paper are made available through the \nuancecode{} open-source Python package hosted at \href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}, released on the Python Package Index\footlink{https://pypi.org/project/nuance/} and with documentation and tutorials hosted at \href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}.
-All following mentions of \nuancecode{} refers to its \review{version 0.6.0 \citep{zenodo}.}{} \\\\
+The methods presented in this paper are made available through the \nuancecode{} open-source Python package hosted at \href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}, released on the Python Package Index\footlink{https://pypi.org/project/nuance/} and with documentation and tutorials hosted at \href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}.
+All following mentions of \nuancecode{} refer to its \review{version 0.6.0 \citep{zenodo}.}{} \\\\
 To instantiate a search, a user can start by creating a \textsf{Nuance} object with
@@
 From this \texttt{search} object, the best transiting candidate parameters can be computed (\textsf{search.best}), or the $\mathcal{Q}$ periodogram retrieved (\textsf{search.Q\_snr}), together with valuable information about the transit search.
-The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary hosts, the advantage of \nuancemethod{} being that the linear search only needs to be performed once and reused for the search of several transiting candidates (cf.\;\autoref{toi540}).
+The \textsf{Nuance} object also provides methods to perform transit search on light curves from multi-planetary hosts, the advantage of \nuancemethod{} being that the linear search only needs to be performed once and reused for the search of several transiting candidates (see \autoref{toi540}).
 An extensive and maintained online documentation is provided at \href{https://nuance.readthedocs.io}{\texttt{nuance.readthedocs.io}}.
@@
 By design, these injected signals have an SNR ranging from 5 to 30 (with $\sigma_r = 0$ in \autoref{eq:snr}).
-For each light curve, transits are searched using two different tools: \textsf{nuance}, using its implementation from the Python package described in the previous section; and BLS, the Box-Least-Square algorithm from \cite{bls} (using \textsf{astropy}'s \textsf{BoxLeastSquares}\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html} implementation).
-\review{We note that the Transit-Least-Square algorithm \citep{tls}, that includes the effect of limb-darkening in the base transit template, could also be used here.
-However, as this effect have a negligible impact on the transit detection compared to the effect of correlated noise, we choose to make our comparisons with BLS only.}\\\\
+For each light curve, transits are searched using two different tools: \textsf{nuance}, using its implementation from the Python package described in the previous section; and BLS, the Box-Least-Squares algorithm from \cite{bls} (using \textsf{astropy}'s \textsf{BoxLeastSquares}\footlink{https://docs.astropy.org/en/stable/api/astropy.timeseries.BoxLeastSquares.html} implementation).
+\review{We note that the Transit-Least-Squares algorithm \citep{tls}, that includes the effect of limb-darkening in the base transit template, could also be used here.
+However, as this effect has a negligible impact on transit detection compared to the effect of correlated noise, we choose to make our comparisons with BLS only.}\\\\
 For both methods, 3000 trial periods from 0.2 to 2.6 days are searched, with a single trial duration fixed to the unique known duration of 50 minutes.
@@
 \ul{\texttt{Bspline+BLS}}: employs a B-spline\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.BSpline.html} for detrending, fitted using the \textsf{scipy.interpolate.splrep} function\footlink{https://docs.scipy.org/doc/scipy/reference/generated/scipy.interpolate.splrep.html} (e.g.\;\citealt{wotan} and \citealt{Canocchi2023}).\\\\
-\ul{\texttt{harmonics+BLS}}: employs a linear harmonic detrending, where the light curve is modeled as a Fourier series including four harmonics of the stellar rotation period with coefficients found through ordinary least square.\\\\
+\ul{\texttt{harmonics+BLS}}: employs a linear harmonic detrending, where the light curve is modeled as a Fourier series including four harmonics of the stellar rotation period with coefficients found through ordinary least squares.\\\\
 \ul{\texttt{iterative+BLS}} iteratively detrends the light curve with a sinusoidal signal fitted to the data, each time using the dominant period of the residuals found using a lomb-scargle periodogram (5 iterations).\\\\
@@
 We consider a transit detectable if its original SNR is greater than 6.
-Hence we define \textit{true positives} has detectable transits recovered with the correct period (or an alias) and a measured SNR greater or equal to 6, and \textit{false positives} as non-detectable transits recovered with a measured SNR greater or equal to 6.\\\\As we noticed that few methods were still affected by the remaining stellar variability after detrending, the grid of orbital periods being searched for does not contain the stellar rotation period $P_*$ and its aliases.
-In practice this is done by removing all orbital periods $P$ from the search grid such that $dP = \frac{P}{P_*}$ is less than 2\% from an integer value, i.e.
+Hence we define \textit{true positives} as detectable transits recovered with the correct period (or an alias) and a measured SNR greater or equal to 6, and \textit{false positives} as undetectable transits recovered with a measured SNR greater or equal to 6.\\\\As we noticed that few methods were still affected by the remaining stellar variability after detrending, the grid of orbital periods being searched for does not contain the stellar rotation period $P_*$ and its aliases.
+In practice, this is done by removing all orbital periods $P$ from the search grid such that $dP = \frac{P}{P_*}$ is less than 2\% from an integer value, i.e.
 $\vert  dP - \lceil dP \rceil \vert < 0.02$.\\\\
 \subsection{Comparisons on simulated light curves}\label{simu}
-Our first comparison dataset consists in 4000 light curves simulated using the model described in \autoref{light_curves_simulations}.
+Our first comparison dataset consists of 4000 light curves simulated using the model described in \autoref{light_curves_simulations}.
 We simulate a common periodic transit added to all light curves, of period $P=1.1$ days, epoch $T_0=0.2$ days, duration $D=0.04$ days and depth $\Delta=1\%$.
-Each light curve consists in a 4 days observation with an exposure time of 2 minutes, leading to $N=2880$ data points with a normal error of $0.1\%$.\\\\
+Each light curve consists of a 4 days observation with an exposure time of 2 minutes, leading to $N=2880$ data points with a normal error of $0.1\%$.\\\\
 For a given pair of $(\tau, \delta)$, we simulate stellar variability using a GP with an SHO kernel of hyperparameters defined by \autoref{eq:relative_params} computed with respect to the injected transit parameters $D$ and $\Delta$.
 The same kernel is used for the search with \textsf{nuance} and with the \texttt{GP+BLS} method, an optimal choice on equal footing with the optimal $3\times D$ window size of the bi-weight filter employed in the \wtls{} search.
-%  While this kernel is optimal, a comprehensive discussion on the use of non-optimal kernels is discussed in . 
 The 4000 pairs of $(\tau, \delta)$ are drawn from
@@
 The results of this injection-recovery procedure are shown in \autoref{fig:simu} and highlight particularly well the benefit of \textsf{nuance} against most other methods.
-Except for the \texttt{GP+BLS} approach, \nuancecode{} lead to a much higher rate of true positives for transits with relatively small depths compared to the stellar variability amplitude (i.e.\;$\delta>2$), and a duration comparable to the stellar variability period (i.e.\;$\tau<2$).
+Except for the \texttt{GP+BLS} approach, \nuancecode{} leads to a much higher rate of true positives for transits with relatively small depths compared to the stellar variability amplitude (i.e.\;$\delta>2$), and a duration comparable to the stellar variability period (i.e.\;$\tau<2$).
 On the other hand, the \texttt{GP+BLS} strategy seems almost as performant as \nuancecode, recovering most of the injected transits and only lacking detection in a relatively comparable portion of the $(\tau, \delta)$ parameter space.
-We note that these empirical statements only concerns simulated light curves with a given amount of white noise, and may vary depending on the length of the observing window or the number of transits.
-For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperforms these methods would only apply to this specific but representative example.
+We note that these empirical statements only concern simulated light curves with a given amount of white noise, and may vary depending on the length of the observing window or the number of transits.
+For this reason, quantifying for which values of $(\tau, \delta)$ \textsf{nuance} outperform these methods would only apply to this specific but representative example.
 However, we verify that our conclusions remain qualitatively valid for various simulation setups.
@@
 
-\noindent This injection-recovery is done in a particularly optimal setup, on simulated light curves not all physically realistic and using an optimal GP kernel, hence demonstrating the performance of \textsf{nuance} only on a purely \review{synthetic}{} basis.
+\noindent This injection-recovery is done in a particularly optimal setup, on simulated light curves that are not all physically realistic and using an optimal GP kernel, hence demonstrating the performance of \textsf{nuance} only on a purely \review{synthetic}{} basis.
 \review{Hence, we strongly emphasize that these tests do not reflect well the performance of each method on real datasets}.
@@
 \subsection{Comparisons on rapidly-rotating M dwarfs TESS light curves}\label{real}
-
-% In order to validate
-
-% The parameter space ($\tau$, $\delta$) introduced in \autoref{detrending_effect} is very convenient from a signal processing point of view, but lacks physical reality.
-Very low values of $\tau$ and $\delta$ often translate into non-physical exoplanets properties as well as unrealistic stellar variability, with a strong dependence on the host star properties.
-\\\\
-% In order to assess the performance of \textsf{nuance} on real light curves, its transit recovery capabilities are rather evaluated in the planetary (period, radius) parameter space, more often encountered in the literature (e.g.
-Figure 12 of \citealt{Delrez2022}).
-In what follows, transits described by the simple model of \cite{protopapas} (see \autoref{app_proto}) are
-
 In order to assess the performance of \nuance{} on real datasets, we inject and recover transits into light curves from the Transiting Exoplanet Survey Satellite (TESS, \citealt{tess}).
@@
 \end{equation*}
-with $\sigma$ equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ the radius of the star reported by \cite{Ramsay2020} and $n$ the number of points in transit computed using a transit duration assuming a circular orbit.
-\review{In total, $438 \times 100$ light curves are produced and searched for transits, corresponding to planets radii ranging from $0.46 R_\oplus$ to $12.29 R_\oplus$ and a median radius of $2.40 R_\oplus$.}\\\\
+where $\sigma$ is equal to the mean uncertainty estimated by the SPOC pipeline, $R_\star$ is the radius of the star reported by \cite{Ramsay2020} and $n$ is the number of points in transit computed using a transit duration assuming a circular orbit.
+\review{In total, $438 \times 100$ light curves are produced and searched for transits, corresponding to planet radii ranging from $0.46 R_\oplus$ to $12.29 R_\oplus$ and a median radius of $2.40 R_\oplus$.}\\\\
 \subsubsection*{Stellar variability kernel}\label{rotation_kernel}
-In the \texttt{GP+BLS} and \nuancecode{} methods we model light curves using a GP with a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ where $P_\star$ is the rotation period of the star.
+In the \texttt{GP+BLS} and \nuancecode{} methods, we model light curves using a GP with a mixture of two SHO kernels of period $P_\star$ and $P_\star/2$ where $P_\star$ is the rotation period of the star.
 This model is representative of a wide range of stochastic variability in stellar time series\footlink{https://celerite2.readthedocs.io/en/latest/api/python} \citep[e.g.][]{David2019, Gillen2020}.
-In order to account for additional corelated noises, we complement this kernel with a short and a long-timescale exponential term, so that the full kernel can be expressed as
+In order to account for additional correlated noises, we complement this kernel with a short and a long-timescale exponential term, so that the full kernel can be expressed as
 \begin{equation*}
@@
 \end{equation*}
-with $\ell$ and $\sigma$ the scale and standard deviation of the process.
-These are meant to model short and long-timescale non-periodic correlated noise.
+with $\ell$ and $\sigma$ the scale and standard deviation of the process, respectively.
+These are meant to model short and long-timescale nonperiodic correlated noise.
 In total, the rotation kernel $k$ has 8 hyperparameters.\\\\
@@
 As correlated noise is expected to affect the light curve uncertainty estimates performed by SPOC, the diagonal of the full covariance matrix of the data (i.e.\,their uncertainty, assuming homoscedasticity) is held free, increasing the number of optimized parameters to 9.
-The optimization is performed using the \textsf{BFGS} algorithm \citep{Fletcher1987}, minimizing the negative log-likelihood of the data as expressed in \autoref{eq:linear_search_ll} (without transit), i.e.\,accounting for a linear systematic model of the data in addition of stellar variability.
-For simplicity, and to adopt a uniform treatment for all target light curves, a design matrix $\bm{X}$ with a single constant column is adopted, such that the systematic model only consists in a single parameter corresponding to the mean value of the differential flux (expected to be close to 1) solved linearly.
-Our motivations to choose this very simplistic baseline, despite the capability of \nuancemethod{} to account for more complex linear models, is discussed in \autoref{systematics}.\\\\
+The optimization is performed using the \textsf{BFGS} algorithm \citep{Fletcher1987}, minimizing the negative log-likelihood of the data as expressed in \autoref{eq:linear_search_ll} (without transit), i.e.\,accounting for a linear systematic model of the data in addition to stellar variability.
+For simplicity, and to adopt a uniform treatment for all target light curves, a design matrix $\bm{X}$ with a single constant column is adopted, such that the systematic model only consists of a single parameter corresponding to the mean value of the differential flux (expected to be close to 1) solved linearly.
+Our motivations for choosing this very simplistic baseline, despite the capability of \nuancemethod{} to account for more complex linear models, is discussed in \autoref{systematics}.\\\\
 \subsubsection*{Search parameters and transit detection criteria}
@@
 \autoref{fig:allsearch} shows the global results of the injection-recovery for all targets, while \autoref{fig:allsearch_relative} shows the true and false positives plotted against the relative parameters $\tau$ and $\delta$ (defined in \autoref{eq:relative_params}).
-These figures are a synthesis of \autoref{fig:allsearchim} that shows the rate of true and false positives binned in the full parameter space $(\tau, \delta)$.
+These figures are a synthesis of \autoref{fig:allsearchim}, which shows the rate of true and false positives binned in the full parameter space $(\tau, \delta)$.
 \begin{figure}[H]
@@
 \end{figure}
-From \autoref{fig:allsearch}, we note that the number of true positives of the \texttt{GP+BLS} method is much worse that what could be anticipated from the results presented in \autoref{fig:simu}, where the GP kernel was fully optimal (given it was also used to simulate the data).
+From \autoref{fig:allsearch}, we note that the number of true positives of the \texttt{GP+BLS} method is much worse than what could be anticipated from the results presented in \autoref{fig:simu}, where the GP kernel was fully optimal (given it was also used to simulate the data).
 In that case, our kernel might not be optimal, whether because of its form or because of the values of its hyperparameters.
@@
 \review{Finally, in \autoref{fig:tpfp_radius}, we observe a dependence between the rate of true positives and the injected planets radii.
-Overall \texttt{nuance} leads to significantly higher true positives in all radii bins compared to other methods, except for \texttt{BLS+harmonics} (which dominates other methods) that shows a similar yield for planets with radii less than 2$R_\oplus$.
+Overall \texttt{nuance} leads to significantly higher true positives in all radii bins compared to other methods, except for \texttt{BLS+harmonics} (which dominates other methods), which shows a similar yield for planets with radii less than 2$R_\oplus$.
 Hence, we note that the advantage of \texttt{nuance} is particularly significant for planets larger than 2$R_\oplus$.
@@
 \end{figure}
-In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its earth-like companion TOI-540 b \citep{TOI540}.
+In order to further validate nuance on a realistic dataset, we focus this section on the multi-sector TESS light curves of TOI-540, and the search for its Earth-like companion TOI-540 b \citep{TOI540}.
 \review{As we will see, \nuancemethod{} does not have significantly superior performance for this candidate.
 However, its light curves feature a very typical stellar variability signal (see \autoref{fig:toi540_clean}), that would commonly be detrended before searching for planetary transits}.\\\\
-We download the 2 minutes cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
-Like in the previous section, we use a Lomb-Scargle periodogram and identify the 0.72 days rotation period of the star, that we use as an initial value to optimize the kernel described in \autoref{rotation_kernel} on each sector independently.
+We downloaded the 2 minute cadence SPOC PDCSAP light curves of TOI-540 observed in 5 sectors (4, 5, 6, 31 and 32).
+Like in the previous section, we use a Lomb-Scargle periodogram and identify the 0.72 day rotation period of the star, which we use as an initial value to optimize the kernel described in \autoref{rotation_kernel} on each sector independently.
 Here again, we employ an upper-sigma-clipping to mask flares out of the data.
@@
 This approach is adopted for efficiency but also to encapsulate the changing properties of stellar variability from one sector to another, often separated by year-long gaps.
-The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days\footnote{We acknowledge that the grid of trial periods and durations used in this study is non-optimal for a real transit search.
+The periodic search is done on 20 000 trial periods ranging from 0.5 to 10 days\footnote{We acknowledge that the grid of trial periods and durations used in this study is nonoptimal for a real transit search.
 However, this simple choice is sufficient for our periodogram comparison against BLS.
 Optimal grids of parameters for realistic transit searches are discussed in \cite{tls}.}.
-This search, using \nuancecode{}, is compared to the more traditional approach that consists in detrending each sector with a bi-weight filter and then search for transits with the BLS algorithm (denoted \texttt{bi-weight+BLS} and described in \autoref{real}) on all sectors combined.
-Since we don't know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
+This search, using \nuancecode{}, is compared to the more traditional approach that consists of detrending each sector with a bi-weight filter and then searching for transits with the BLS algorithm (denoted \texttt{bi-weight+BLS} and described in \autoref{real}) on all sectors combined.
+Since we do not know the transit duration a-priori, we perform the detrending and BLS search using 15 filtering window sizes sampled from 30 minutes to 5 hours, and retain the search that leads to the highest transit SNR peak in the periodogram (as done e.g.\;in the SHERLOCK transit search pipeline described in \citealt{Pozuelos2020}).
 The results of this comparison are shown in \autoref{fig:toi540_periodograms}.\\\\
@@
 As seen in \autoref{fig:toi540_periodograms}, the SNR periodogram using \texttt{bi-weight+BLS} and \texttt{nuance} are very similar, with the known transiting exoplanet TOI-540 b detected with an orbital period P = 1.24 days.
-This is well expected as the relative parameter $\tau$ equal 13 for TOI-540 b\footnote{using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ minutes of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial (see \autoref{fig:snr_detrend}).
+This is well expected as the relative parameter $\tau$ equals 13 for TOI-540 b\footnote{Using \autoref{eq:relative_params} with the stellar rotation $P=0.72$ days and the known transit duration $D=29.5$ minutes of TOI-540 b \citep{TOI540}}, which lies outside the range where \nuancemethod{} is expected to be beneficial (see \autoref{fig:snr_detrend}).
 \review{Nonetheless}, the \nuance{} periodogram of TOI 540 features less spurious SNR peaks, largely due to the penalty naturally occurring when single transits with different depths are periodically combined.
 In the second search (right panel) of \autoref{fig:toi540_periodograms}, we also notice a higher number of peaks that would lead to false positive detections of transits in the \texttt{bi-weight+BLS} case.
-The proper treatment of correlated noise in \nuance{}, as observed in \autoref{real}, makes these peaks non-significant, avoiding a large number of false detections.\\\\
-We note that finding a known TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques (see e.g.\;\autoref{fig:simu}).
+The proper treatment of correlated noise in \nuance{}, as observed in \autoref{real}, makes these peaks nonsignificant, avoiding a large number of false detections.\\\\
+We note that finding a known TESS candidate that displays characteristics for which \nuance{} is expected to be significantly beneficial proved to be very challenging during the writing of this paper, as transits with such characteristics are expected to be missed by current state-of-the-art techniques (see, e.g., \autoref{fig:simu}).
 Nonetheless, we verify that \nuancemethod{} is capable of finding a large number of already known transiting exoplanet candidates, in light curves featuring various forms of correlated noises, at least as efficiently as with commonly used techniques.
-We reserve the search for new transit signals to a follow-up publication.
+We reserve the search for new transit signals for a follow-up publication.
 
@@
 
-In the previous section we demonstrated the capability of \nuance{} to search for synthetic or known transit signals, in simulated or real datasets.
+In the previous section, we demonstrated the capability of \nuance{} to search for synthetic or known transit signals, in simulated or real datasets.
 Here, we discuss the caveats of this algorithm, the advantages and limitations of the \nuancecode{} implementation, and future prospects for its extension.
@@
 
-\noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple non-optimized GP with a squared exponential kernel.
-These are compared to the processing time of \texttt{bi-weight+BLS} (cf.\;\autoref{simu}) separated into the bi-weight filtering step and the BLS search.
+\noindent In \autoref{fig:blsvsnuance}, the processing time of \nuancecode{} linear and periodic search are recorded against the number of points in a simulated light curve, assuming a simple un-optimized GP with a squared exponential kernel.
+These are compared to the processing time of \texttt{bi-weight+BLS} (see \autoref{simu}) separated into the bi-weight filtering step and the BLS search.
 \begin{figure}[H]
@@
         \includegraphics[width=\linewidth]{../workflows/benchmark/figures/nuance_vs_bls.pdf}
-        \caption{Study of \textsf{nuance} (black line) processing time against \texttt{bi-weight+BLS} (cf.\;\autoref{simu}, blue line).
+        \caption{Study of \textsf{nuance} (black line) processing time against \texttt{bi-weight+BLS} (see \autoref{simu}, blue line).
 The gray curve shows the performance of \textsf{nuance} linear search when applied to chunks of 10 000 points continuous observations, instead of considering these observations all together.
@@
 
-As see in \autoref{fig:blsvsnuance}, most of the computational costs of \nuancecode{} and the \texttt{bi-weight+BLS} method come from the linear search and the bi-weight filtering step.
-This is not always true and depends on the size of the trial durations and periods grids.
+As seen in \autoref{fig:blsvsnuance}, most of the computational costs of \nuancecode{} and the \texttt{bi-weight+BLS} method come from the linear search and the bi-weight filtering step.
+This is not always true and depends on the size of the trial durations and period grids.
 One advantage of \nuancemethod{} is that the linear search can be performed separately on different continuous observations, and then combined in the periodic search.
-Hence, if searching for transits in separate observations with approximately similar durations, such as different TESS sectors or different ground-based nightly observations, the computational cost of \nuancecode{} grows linearly with the number of observations (see grey line in \autoref{fig:blsvsnuance}).
+Hence, if searching for transits in separate observations with approximately similar durations, such as different TESS sectors or different ground-based nightly observations, the computational cost of \nuancecode{} grows linearly with the number of observations (see gray line in \autoref{fig:blsvsnuance}).
 Nonetheless, considering a variety of other detrending algorithms, \textbf{\nuancecode{} is expected to perform between one and two order of magnitude slower than more conventional techniques associated with BLS}.\\\\
 Nonetheless, searching for TOI-540 b transits with the parallelized implementation of \nuancecode{} (see \autoref{toi540}) on the 12 cores of an Apple M2 Max chip, took only 5 minutes and 35 seconds, 1 minute 50 seconds for the 5 sectors linear searches (around 22 seconds per TESS sector), and 3 minutes 45 seconds for the combined periodic search.
-In comparison, the brute force search with \texttt{bi-weight+BLS}, consisting in trying 15 bi-weight windows, took a total of 1 minute 49 seconds (only 7 seconds each).\\\\
+In comparison, the brute force search with \texttt{bi-weight+BLS}, consisting of trying 15 bi-weight windows, took a total of 1 minute 49 seconds (only 7 seconds each).\\\\
 % As the periodic search is expected to be improved by lower lever optimizations (using a compiled language), we reserve its benchmark for a future work.
@@
 But as these strongly depend on the type of detrending technique employed (see \autoref{fig:simu}), we do not provide general guidelines as when \nuancemethod{} should be preferred over a specific detrending technique.
-To aid users in making an informed choice of algorithm, extensive benchmarks and guidelines are reserved to future developments and will be progressively shared on \nuancecode{}'s online documentation\footlink{https://nuance.readthedocs.io/en/latest/}.
+To aid users in making an informed choice of algorithm, extensive benchmarks, and guidelines are reserved for future developments and will be progressively shared on \nuancecode{}'s online documentation\footlink{https://nuance.readthedocs.io/en/latest/}.
 
 \subsection{Systematics modeling}\label{systematics}
-Throughout the paper, a single column design matrix $\bm{X}$, corresponding to the mean differential flux (ideally unitary), was employed, hence assuming that the instrumental systematics signals were non-existent.
+Throughout the paper, a single-column design matrix $\bm{X}$, corresponding to the mean differential flux (ideally unitary), was employed, hence assuming that the instrumental systematics signals were nonexistent.
 In practice, \nuancemethod{} has been developed to linearly model systematic signals through more complex design matrices (as in \citealt{foreman2016}), in addition with its capability to model correlated noise while searching for transits.
@@
 Comparisons involving more complex design matrices would also be sensitive to the choice of linear components, and would have unwanted repercussions on their results.\\\\
-As an illustration, the NEMESIS pipeline \citep{nemesis} starts processing the differential light curves by employing a linear systematics detrending using a least square fit of the data with a reduced PLD basis, before smoothing the signal from stellar variability using an approach similar to the one employed in the \texttt{bi-weight+BLS} approach (cf.\;\autoref{real}), hence detrending the systematics with an incomplete model that does not account for stellar variability.
+As an illustration, the NEMESIS pipeline \citep{nemesis} starts processing the differential light curves by employing a linear systematics detrending using a least squares fit of the data with a reduced PLD basis, before smoothing the signal from stellar variability using an approach similar to the one employed in the \texttt{bi-weight+BLS} approach (see \autoref{real}), hence detrending the systematics with an incomplete model that does not account for stellar variability.
 To account for stellar variability while fitting the linear systematics model to the data, a step further would be to use a GP, such as done in the EVEREST \citep{everest2} pipeline.
@@
 in \citealt{everest1}).\\\\
-Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsly is extremely promising for the search of transiting exoplanets, including for ground-based observations that usually suffer from daily interruptions.
+Although not being demonstrated here, modeling systematics signals while searching for transits on data acquired sparsely is extremely promising for the search of transiting exoplanets, including for ground-based observations that usually suffer from daily interruptions.
 In this respect, we note the similarity of our linear search (cf\;\autoref{linear_search}) to the one presented in \citealt{Berta2012}, that focused on the detection of single eclipses in the MEarth light curves \citep{Irwin2009}.
 Similarly, \textbf{\nuancemethod{} would highly benefit the search for transiting exoplanets around M-dwarf type stars}, such as the ones observed by the SPECULOOS survey \citep{speculoos} whose monitoring suffers from both increased red noise (due to atmospheric and instrumental thermal effects discussed e.g.\;in \citealt{Berta2012} and  \citealt{Pedersen2023}) and enhanced stellar variability \citep{Murray2020}.
-We reserve this promising application to a future study.
+We reserve this promising application for a future study.
 
 \subsection{The GP kernel}
-While not being discussed in our study, \textbf{the efficiency of \nuancemethod{} to detect transits in correlated noise might be dependant on the design of its GP kernel}.
-In the ensemble comparison of \autoref{real}, the goal was to choose a kernel and an optimization strategy suited to most of the studied light curves, leading to few outliers in the results that were indicative of a badly designed and/or optimized kernel.
+While not being discussed in our study, \textbf{the efficiency of \nuancemethod{} to detect transits in correlated noise might be dependent on the design of its GP kernel}.
+In the ensemble comparison of \autoref{real}, the goal was to choose a kernel and an optimization strategy suited to most of the studied light curves, leading to few outliers in the results, which were indicative of a badly designed and/or optimized kernel.
 An alternative, recommended for more realistic blind searches, is to perform model comparison on well selected kernels, and to adapt the optimization strategy to each dataset.\\\\
@@
     \item In order to compare \nuancecode{} to BLS-based methods, we injected and retrieved only box-shaped transits.
-However, similarly to the Transit-Least-Square algorithm from \cite{tls}, \textbf{limb-darkened transits can serve as base model} in the linear search and are expected to improve the transit search in the same way TLS provided an improvement over BLS.
-    \item The linear search of \nuancemethod{} is a single-event detection algorithm that can be used to search for single transit events, but also \textbf{detect transiting exocomets and flares}, by simply changing the base astrophysical model in the last colum of the design matrix $\bf{X}$.
+However, similarly to the Transit-Least-Squares algorithm from \cite{tls}, \textbf{limb-darkened transits can serve as base model} in the linear search and are expected to improve the transit search in the same way TLS provided an improvement over BLS.
+    \item The linear search of \nuancemethod{} is a single-event detection algorithm that can be used to search for single transit events, but also \textbf{detect transiting exocomets and flares}, by simply changing the base astrophysical model in the last column of the design matrix $\bf{X}$.
 While not being tested and benchmarked, \nuancecode{} already integrates this feature.
@@
 This was done in order to allow the detection of grazing transiting exoplanets that would produce \review{shorter-timescale}{} transits compared to what is expected from a circular orbit with a null impact parameter.
-However, adding such priors might produce less spurious periodogram peaks and be very beneficial for the automatic search of transits in large datasets.
+However, adding such priors might produce fewer spurious periodogram peaks and be very beneficial for the automatic search of transits in large datasets.
 Another idea, similar to the one employed in \cite{foreman2016}, is to leverage \textbf{model comparison in order to reject transits that are better described by the GP model alone}.
@@
 However, we believe that the profile likelihood (or SNR) over orbital periods \textit{and} transit durations (see \autoref{snrperiod}) holds valuable information that could be exploited to further direct the search, especially when transits with different durations are present.}{}
-    \item Finally, the formalism of \nuancemethod{} could be adapted and used to \textbf{search for exoplanets featuring transit time variations (TTV)}.
-Indeed, this application only requires a modification of the periodic search, as maximum likelihood peaks close to linearly predicted transit epochs may be considered.
+    \item Finally, the formalism of \nuancemethod{} could be adapted and used to \textbf{search for exoplanets featuring transit time variations (TTVs)}.
+Indeed, this application only requires a modification of the periodic search, as the maximum likelihood peaks close to linearly predicted transit epochs may be considered.
 This could be done either with a special nearest-neighbor algorithm or with a convolution of the computed likelihood grid with a Gaussian kernel.
@@
 The GP kernel and its hyperparameters are chosen and optimized as in \autoref{real}.Here, we simply used the linear search of \nuancecode{} with a different base model, one that mimics the shape of transiting exocomets (with 20 trial durations), to compute the SNR time-series of the signal over one sector.
-The maximum-SNR events are displayed in the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.}
+The maximum-SNR events are displayed at the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.}
     \label{fig:exoco} 
@@
 This paper presents \nuancemethod{}, an algorithm designed to detect planetary transits in light curves featuring correlated noise in the form of instrumental signals and stellar variability.
-In this context, a conventional approach involves detrending a light curve before searching for transits using a Box-Least-Square algorithm.
+In this context, a conventional approach involves detrending a light curve before searching for transits using a Box-Least-Squares algorithm.
 However, we show that this approach degrades transit signal-to-noise-ratios down to the point of not being detectable.
-Adopting commonly used detrending strategies, we explore the extent of this degradation on simulated light curves, and its dependance on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
+Adopting commonly used detrending strategies, we explore the extent of this degradation on simulated light curves, and its dependence on the photometric stellar variability characteristics, showing the need for a full-fledged transit search method like \nuancemethod{}.\\\\
 The effectiveness of \nuancemethod{} is tested using a synthetic dataset and further validated on real TESS light curves of 438 rapidly-rotating M dwarfs.
@@
 In all cases, \nuancemethod{} not only leads to a higher number of true positive detections but also minimizes false positives, demonstrating its robustness and reliability.\\\\
-We make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package, developped with \textsf{JAX} to allow its  use on distributed computing environments and GPU devices.
+We make \nuancemethod{} publicly available through the \nuancecode{} open-source Python package, developed with \textsf{JAX} to allow its  use on distributed computing environments and GPU devices.
 Overall, we acknowledge the limitations of \nuancemethod{} and its increased computational cost compared to more conventional techniques.
 Hence, \nuancemethod{} should be used as an alternative to more traditional techniques only in the presence of substantial correlated noise.
-As guidelines for choosing our method over other techniques are highly dependant on the type of detrending algorithm employed, we reserve this study to a future work.\\\\
+As guidelines for choosing our method over other techniques are highly dependant on the type of detrending algorithm employed, we reserve this study for a future work.\\\\
 Finally, we suggest future improvements and extensions of the algorithm, including its application for detecting single transits, exocomets, flares, and exoplanets featuring transit time variations (TTV), underscoring its versatility and potential for broader impact in astronomical research.
@@
 \review{We would also like to thank the journal referee for their thorough and encouraging report, bringing some very meaningful additions to the paper.}{} This publication benefits from the support of the French Community of Belgium in the context of the FRIA Doctoral Grant awarded to L.J.G.
-F.J.P acknowledges financial support from the grant CEX2021-001131-S funded by MCIN/AEI/ 10.13039/501100011033 and through projects PID2019-109522GB-C52 and PID2022-137241NB-C4.
+F.J.P acknowledges financial support from grant CEX2021-001131-S funded by MCIN/AEI/ 10.13039/501100011033 and through projects PID2019-109522GB-C52 and PID2022-137241NB-C4.
 S.A.
@@
 This paper includes data collected by the TESS mission.
-Funding for the TESS mission is provided by the NASA's Science Mission Directorate and NASA Explorer Program.
+Funding for the TESS mission is provided by NASA's Science Mission Directorate and NASA Explorer Program.
 We acknowledge the use of public TESS data from pipelines at the TESS Science Office and at the TESS Science Processing Operations Center.
@@
 \end{equation}
-where $Q$ is the quality factor of the oscillator, $\omega$ its pulsation and $\sigma$ the amplitude of the kernel function.
+where $Q$ is the quality factor of the oscillator, $\omega$ its pulsation, and $\sigma$ the amplitude of the kernel function.
 GP computations in this paper use the implementation from \texttt{tinygp}\footnote{\href{https://github.com/dfm/tinygp}{https://github.com/dfm/tinygp}}, a Python package exposing the quasi-separable kernels from \cite{celerite2} and powered by \textsf{JAX}\footnote{\href{https://github.com/google/jax}{https://github.com/google/jax}}.
@@
 where $p(\bm{f} \vert T_k, D, \tilde\Delta)$ is the probability of the $k$-th transit to have a depth $\tilde\Delta$ and $p(\tilde\Delta | \Delta)$ the probability to observe the depth $\tilde\Delta$ knowing the existence of a common depth $\Delta$.
-In other words, \autoref{eq:non_part_of_per} involves the likelihood of the non-periodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
+In other words, \autoref{eq:non_part_of_per} involves the likelihood of the nonperiodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
 \\\\
-Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$ and an amplitude $\mathcal{L}_k$, leading to the likelihood function
+Since each depth $\Delta_k$ is found through generalized least squares, each follows a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$ and an amplitude $\mathcal{L}_k$, leading to the likelihood function
 \begin{equation*}

### 2024-07-19 latex/paper.tex
add all figures to same folder + snakemake ref + icons for online ressources and source code

@@
-\documentclass[modern,linenumbers]{aastex631}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{bm}
-\usepackage{listings}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{soul}
+\documentclass[modern]{aastex631}
 
-% commands
-\newcommand{\nuancemethod}{\textit{nuance}}
-\newcommand{\nuance}{\nuancemethod{}}
-\newcommand{\nuancecode}{\textsf{nuance}}
-
-\newcommand{\TODO}{\texttt{todo}}
-\newcommand{\review}[1]{#1}
-% \newcommand{\review}[1]{#1} 
-\newcommand{\set}[1]{\{\,#1\,\}}
-\newcommand{\footlink}[1]{\footnote{\url{#1}}}
-\newcommand{\wtls}{\texttt{bi-weight+BLS}}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-\let\subsectionautorefname\sectionautorefname
-\let\subsubsectionautorefname\sectionautorefname
-
-% no indent
-\setlength\parindent{0pt}
-
-% code style
-\lstdefinestyle{mystyle}{
-    backgroundcolor=\color{white},
-    commentstyle=\color{gray!70},
-    keywordstyle=\color{Bittersweet},
-    stringstyle=\color{RoyalBlue},
-    basicstyle=\fontsize{8.5}{13}\fontfamily{DejaVuSansMono-TLF}\selectfont,
-    breakatwhitespace=false,         
-    breaklines=true,
-    rulecolor=\color{black!15},
-    numbers=none,
-    numberstyle=\fontsize{7}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont\color{gray!50},
-    framerule=0pt,
-    breakindent=5pt,
-    resetmargins=true,
-    numbersep=10pt,
-    frame=single,
-    aboveskip=1em,
-    belowskip=1em,
-    xleftmargin=6pt,
-    framexleftmargin=4pt
-}
-\lstset{style=mystyle}
+\include{stylez}
 
@@
 % author info
-\author{Lionel J. Garcia}
+\author[0000-0002-4296-2246]{Lionel J. Garcia}
 \affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
 \affiliation{\review{Astrobiology Research Unit, University of Liège, Liège, Belgium}{}}
-\author{Daniel Foreman-Mackey}
+\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
 \affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
-\author{Catriona A. Murray}
+\author[0000-0001-8504-5862]{Catriona A. Murray}
 \affiliation{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, USA}
-\author{Suzanne Aigrain}
+\author[0000-0003-1453-0574]{Suzanne Aigrain}
 \affiliation{Sub-department of Astrophysics, Department of Physics, University of Oxford, Oxford, UK}
-\author{Dax L. Feliz}
+\author[0000-0002-2457-7889]{Dax L. Feliz}
 \affiliation{American Museum of Natural History, New York, NY, USA}
-\author{Francisco J. Pozuelos}
+\author[0000-0003-1572-7707]{Francisco J. Pozuelos}
 \affiliation{Instituto de Astrofísica de Andalucía, Granada, Spain}
@@
 Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of planetary systems around active stars, both using space-based surveys and sparse ground-based observations.
+\href{https://github.com/lgrcia/nuance}{\color{linkcolor}\faCode}\,\,\href{https://github.com/lgrcia/paper-nuance}{\color{linkcolor}\faFileTextO}\,\,\href{https://nuance.readthedocs.io}{\color{linkcolor}\faBook}
 \end{abstract}
@@
 This include transits injected in synthetic datasets, but also in the TESS light curves of 438 rapidly-rotating M dwarfs.
-Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, before concluding in \autoref{conclusion}.
+Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, before concluding in \autoref{conclusion}.\\\\
+This paper was compiled from source code available in a git-versioned repository (\href{https://github.com/lgrcia/paper-nuance}{\color{linkcolor}\faFileTextO}) and used the \textsf{snakemake}\footlink{https://snakemake.readthedocs.io} workflow management tool.
+Source code of the \textsf{nuance} \textsf{Python} package is also available in a git-versioned repository (\href{https://github.com/lgrcia/nuance}{\color{linkcolor}\faCode}) together with an online documentation (\href{https://nuance.readthedocs.io}{\color{linkcolor}\faBook}).
+Finally, we provide links to the source code used to produce each figure (inidcated by the {\color{linkcolor}\codeicon} icon).
+
 \newpage
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
+        \includegraphics[width=\linewidth]{./principle_dataset_decomposed.pdf}
         \caption{The top plot shows the simulated flux time series of an active star drawn from the GP model described in \autoref{light_curves_simulations}.
@@
 Correlated noise in the form of stellar variability is simulated with an SHO kernel of hyperparameters $\omega = \pi / 6 D$, $Q=45$ and $\sigma = \Delta$.
-Finally, white noise with a standard deviation of $0.1\%$ is added to the diagonal of the covariance matrix.}
+Finally, white noise with a standard deviation of $0.1\%$ is added to the diagonal of the covariance matrix.
+\codelink{workflows/principle/principle.ipynb}}
         \label{fig:app_principle_dataset}
@@
 The SNR on the right of each light curve is computed using \autoref{eq:snr}.
-Models used to simulate these data are provided in \autoref{light_curves_simulations}.}
+Models used to simulate these data are provided in \autoref{light_curves_simulations}.
+\codelink{workflows/plot_issues/scripts/plot_issue_1.py}}
         \label{fig:issue1}
@@
 The resulting stellar variability signals are then reconstructed using (in purple) Tukey's bi-weight filter with an optimal window size of $3\times D$  (see \citealt{wotan}), and (in red) a GP with the same kernel used to simulate the data.
-In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr}.}
+In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr}.
+\codelink{workflows/plot_issues/scripts/plot_issue_2.py}}
         \label{fig:issue2}
@@
 The two central plots show the transit SNR after detrending the light curves with one of two methods.
-On the left, light curves are detrended using Tukey's bi-weight filter with an optimal window size of $3\times D$ (see \citealt{wotan}), while on the right, light curves are detrended using a GP with the same kernel used to simulate the data.}
+On the left, light curves are detrended using Tukey's bi-weight filter with an optimal window size of $3\times D$ (see \citealt{wotan}), while on the right, light curves are detrended using a GP with the same kernel used to simulate the data.
+\codelink{workflows/cleaning_snr/scripts/plot.py}}
         \label{fig:snr_detrend}
@@
 For each pair of indices $(i,j)$, the likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}.
-This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly using \autoref{eq:ls}.}
+This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly using \autoref{eq:ls}.
+\codelink{workflows/principle/principle.ipynb}}
         \label{fig:linear_search}
@@
     \caption{Periodic search likelihood $\mathcal{P}(P)$ computed for different trial periods $P$.
-Notice how the maximum value of $\mathcal{P}$ for the alias period $P=0.35$ day (left plot) is lower than for $P=2\times 0.35=0.7$ day, a result of combining the log-likelihoods using \autoref{eq:result} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
+Notice how the maximum value of $\mathcal{P}$ for the alias period $P=0.35$ day (left plot) is lower than for $P=2\times 0.35=0.7$ day, a result of combining the log-likelihoods using \autoref{eq:result} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.
+\codelink{workflows/principle/principle.ipynb}}
     \label{fig:periodic_search}
@@
         \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_Q.pdf}
-        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
+        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.
+\codelink{workflows/principle/principle.ipynb}}
         \label{fig:periodogram}
@@
         \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using \textsf{BLS} and \nuancecode{}.
-The color scale indicates the recovery of transits in the corresponding (period, depth) parameter space, white for a full recovery and black for no detection.}
+The color scale indicates the recovery of transits in the corresponding (period, depth) parameter space, white for a full recovery and black for no detection.
+\codelink{workflows/control_test_bls/scripts/plot.py}}
         \label{fig:control}
@@
         \caption{Rate of true positives for the 4000 simulated light curves.
-The color scale represents the fraction of true positives, white if all injected transits are recovered in a given portion of the $(\tau, \delta)$ parameter space, black if none are recovered.}
+The color scale represents the fraction of true positives, white if all injected transits are recovered in a given portion of the $(\tau, \delta)$ parameter space, black if none are recovered.
+\codelink{workflows/synthetic-injection-recovery/scripts/plot.py}}
         \label{fig:simu}
@@
     \centering
-    \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
+    \includegraphics[width=\linewidth]{cleaned_1019692.pdf}
     \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692.
 The top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot.
-On this bottom plot, the black line corresponds to the mean of the GP model (with hyperparameters optimized here on the cleaned light curve).}
+On this bottom plot, the black line corresponds to the mean of the GP model (with hyperparameters optimized here on the cleaned light curve).
+\codelink{workflows/tess_injection_recovery/scripts/plot_lc.py}}
     \label{fig:cleaned}
@@
         \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_bars.pdf}
-        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals.}
+        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals. \codelink{workflows/tess_injection_recovery/scripts/plot_tpfp.py}}
         \label{fig:allsearch}
@@
         \caption{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
-The solid blue line corresponds to the maximum of true positives among all methods on the top panel (dominated \review{by the \texttt{harmonics+BLS} method}), and the minimum of false positives on the bottom panel (waned by the \texttt{GP+BLS} method, all other methods being above \nuancecode{}).}
+The solid blue line corresponds to the maximum of true positives among all methods on the top panel (dominated \review{by the \texttt{harmonics+BLS} method}), and the minimum of false positives on the bottom panel (waned by the \texttt{GP+BLS} method, all other methods being above \nuancecode{}).
+\codelink{workflows/tess_injection_recovery/scripts/plot_tpfp.py}}
         \label{fig:allsearch_relative}
@@
         \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_ims.pdf}
-        \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods and all targets.}
+        \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods and all targets.
+\codelink{workflows/tess_injection_recovery/scripts/plot_tpfp.py}}
         \label{fig:allsearchim}
@@
         \caption{\review{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the injected planet's radius.
-The solid blue line corresponds to the maximum of true positives among all other methods than \nuance{} (dominated by the \texttt{harmonics+BLS} method).}{}}
+The solid blue line corresponds to the maximum of true positives among all other methods than \nuance{} (dominated by the \texttt{harmonics+BLS} method).}{} \codelink{workflows/tess_injection_recovery/radii.ipynb}}
         \label{fig:tpfp_radius}
@@
         \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/4.pdf}
-        \caption{Sector 4 light curve of TOI 540. The cleaned signal (gray points) has been masked for flares (black points), and the black line corresponds to the mean of the GP model.}
+        \caption{Sector 4 light curve of TOI 540.
+The cleaned signal (gray points) has been masked for flares (black points), and the black line corresponds to the mean of the GP model.
+\codelink{workflows/comparison_toi/scripts/plot_light_curve.py}}
         \label{fig:toi540_clean}
@@
         \caption{Transit search SNR periodograms of TOI-540 using \texttt{bi-weight+BLS} and \nuance{}.
-After a first periodic search (left panel), the epochs corresponding to the maximum-SNR transit are masked before the second search is performed (right panel).}
+After a first periodic search (left panel), the epochs corresponding to the maximum-SNR transit are masked before the second search is performed (right panel).
+\codelink{workflows/comparison_toi/scripts/plot_periodograms.py}}
         \label{fig:toi540_periodograms}
@@
 This study is performed on a single CPU core of an Apple M2 Max chip.
-While not being shown, we verify that both the BLS algorithm and \nuancemethod{} processing times scale linearly with the number of trial transit durations and orbital periods.}
+While not being shown, we verify that both the BLS algorithm and \nuancemethod{} processing times scale linearly with the number of trial transit durations and orbital periods.
+\codelink{workflows/benchmark/scripts/plot.py}}
         \label{fig:blsvsnuance}
@@
 The GP kernel and its hyperparameters are chosen and optimized as in \autoref{real}.Here, we simply used the linear search of \nuancecode{} with a different base model, one that mimics the shape of transiting exocomets (with 20 trial durations), to compute the SNR time-series of the signal over one sector.
-The maximum-SNR events are displayed at the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.}
+The maximum-SNR events are displayed at the bottom of the figure and match with the ones found by \cite{Lecavelier2022}.
+\outcodelink{https://nuance.readthedocs.io/en/latest/notebooks/tutorials/exocomet.html}}
     \label{fig:exoco} 
@@
 As guidelines for choosing our method over other techniques are highly dependant on the type of detrending algorithm employed, we reserve this study for a future work.\\\\
-Finally, we suggest future improvements and extensions of the algorithm, including its application for detecting single transits, exocomets, flares, and exoplanets featuring transit time variations (TTV), underscoring its versatility and potential for broader impact in astronomical research.
+Finally, we suggest future improvements and extensions of the algorithm, including its application for detecting single transits, exocomets, flares, and exoplanets featuring transit time variations (TTVs), underscoring its versatility and potential for broader impact in astronomical research.
 \\\\
@@
 
-\software{{numpy} \citep{harris2020array}, {scipy} \citep{2020SciPy-NMeth}, {matplotlib} \citep{Hunter:2007}, {jax} \citep{jax2018github}, {jaxopt} \citep{jaxopt_implicit_diff}, {tinygp} \citep{foreman_mackey_2024_10463641}, {lightkurve} \citep{2018ascl.soft12013L}, \review{{astropy} (\citealt{astropy:2013,astropy:2018,astropy:2022})}, {pandas} \citep{mckinney-proc-scipy-2010, reback2020pandas}, {wotan} \citep{wotan} \review{and nuance version 0.6.0 \citep{zenodo}}.}{}
+\software{{numpy} \citep{harris2020array}, {scipy} \citep{2020SciPy-NMeth}, {matplotlib} \citep{Hunter:2007}, {jax} \citep{jax2018github}, {jaxopt} \citep{jaxopt_implicit_diff}, {tinygp} \citep{foreman_mackey_2024_10463641}, {lightkurve} \citep{2018ascl.soft12013L}, \review{{astropy} (\citealt{astropy:2013,astropy:2018,astropy:2022})}, {pandas} \citep{mckinney-proc-scipy-2010, reback2020pandas}, {wotan} \citep{wotan}, {snakemake} \citep{Snakemake} \review{and nuance version 0.6.0 \citep{zenodo}}.}{}
 
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/searched/1019692.pdf}
+        \includegraphics[width=\linewidth]{searched_1019692.pdf}
     \end{centering}
@@
 For each method, the upper right legend on the left plot indicates its ranking based on the percent of recovered transit signals (where a transit with an aliased period counts as being detected).
+\codelink{workflows/tess_injection_recovery/scripts/plot_comparison.py}
         }

### 2024-07-19 copy_figures.sh

@@ -1,0 +1,5 @@
+#/bin/bash
+
+cp workflows/tess_injection_recovery/_figures/searched/1019692.pdf figures/searched_1019692.pdf
+cp workflows/tess_injection_recovery/_figures/cleaned/1019692.pdf figures/cleaned_1019692.pdf
+find workflows -type d -maxdepth 2 -name 'figures' -exec sh -c 'cp -r "$0"/* figures' "{}" \;

### 2024-07-19 latex/paper.tex
move figures and add arxiv submission script

@@
 % figure options
-\graphicspath{{../figures}}
+\graphicspath{{./figures}}
 
@@
 Source code of the \textsf{nuance} \textsf{Python} package is also available in a git-versioned repository (\href{https://github.com/lgrcia/nuance}{\color{linkcolor}\faCode}) together with an online documentation (\href{https://nuance.readthedocs.io}{\color{linkcolor}\faBook}).
-Finally, we provide links to the source code used to produce each figure (inidcated by the {\color{linkcolor}\codeicon} icon).
+Finally, we provide links to the source codes used to produce each figure (indicated by {\color{linkcolor}\codeicon}).
 
@@
     \begin{centering}
-        \includegraphics[width=0.6\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
+        \includegraphics[width=0.6\linewidth]{./issue1.pdf}
         \caption{Illustration of the effect of correlated noise on a single transit SNR.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
+        \includegraphics[width=\linewidth]{./issue2.pdf}
         \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.8\%$, duration $D=0.1$ day, and white noise with a standard deviation of 0.15\% (top left).
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/cleaning_snr/figures/result.pdf}
+        \includegraphics[width=\linewidth]{./result.pdf}
         \caption{Transit SNR computed for 10 000 differential light curves with varying morphological characteristics.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
+        \includegraphics[width=\linewidth]{./principle_linear_search.pdf}
         \caption{Principle and output of the linear search.
@@
     \centering
-    \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
+    \includegraphics[width=\linewidth]{./principle_periodic_1.pdf}\\
     \caption{Periodic search likelihood $\mathcal{P}(P)$ computed for different trial periods $P$.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_Q.pdf}
+        \includegraphics[width=\linewidth]{./principle_Q.pdf}
         \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
+        \includegraphics[width=\linewidth]{./control_test.pdf}
         \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using \textsf{BLS} and \nuancecode{}.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/synthetic-injection-recovery/figures/synthetic_ir.pdf}
+        \includegraphics[width=\linewidth]{./synthetic_ir.pdf}
         \caption{Rate of true positives for the 4000 simulated light curves.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_bars.pdf}
+        \includegraphics[width=\linewidth]{./tpfp_bars.pdf}
         \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals. \codelink{workflows/tess_injection_recovery/scripts/plot_tpfp.py}}
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp.pdf}
+        \includegraphics[width=\linewidth]{./tpfp.pdf}
         \caption{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_ims.pdf}
+        \includegraphics[width=\linewidth]{./tpfp_ims.pdf}
         \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods and all targets.
@@
     \begin{centering}
-        \includegraphics[width=0.6\linewidth]{../workflows/tess_injection_recovery/figures/tpfp_radius.pdf}
+        \includegraphics[width=0.6\linewidth]{./tpfp_radius.pdf}
         \caption{\review{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the injected planet's radius.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/4.pdf}
+        \includegraphics[width=\linewidth]{./TOI_540/4.pdf}
         \caption{Sector 4 light curve of TOI 540.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/comparison_toi/figures/TOI 540/periodograms.pdf} 
+        \includegraphics[width=\linewidth]{./TOI_540/periodograms.pdf} 
         \caption{Transit search SNR periodograms of TOI-540 using \texttt{bi-weight+BLS} and \nuance{}.
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{../workflows/benchmark/figures/nuance_vs_bls.pdf}
+        \includegraphics[width=\linewidth]{./nuance_vs_bls.pdf}
         \caption{Study of \textsf{nuance} (black line) processing time against \texttt{bi-weight+BLS} (see \autoref{simu}, blue line).

### 2024-07-19 copy_figures.sh

@@ -1,5 +1,6 @@
 #/bin/bash
 
-cp workflows/tess_injection_recovery/_figures/searched/1019692.pdf figures/searched_1019692.pdf
-cp workflows/tess_injection_recovery/_figures/cleaned/1019692.pdf figures/cleaned_1019692.pdf
-find workflows -type d -maxdepth 2 -name 'figures' -exec sh -c 'cp -r "$0"/* figures' "{}" \;
+cp workflows/tess_injection_recovery/_figures/searched/1019692.pdf latex/figures/searched_1019692.pdf
+cp workflows/tess_injection_recovery/_figures/cleaned/1019692.pdf latex/figures/cleaned_1019692.pdf
+find workflows -type d -maxdepth 2 -name 'figures' -exec sh -c 'cp -r "$0"/* latex/figures' "{}" \;
+cp -r "latex/figures/TOI_540" "latex/figures/TOI 540"
